### 2016-06-19 doc/draft1.tex
Initial Commit

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{a4paper}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{hyperref} % verlinkungen
\usepackage{textcomp}
\usepackage{flafter}
\usepackage{booktabs}
\usepackage{array}
\usepackage{paralist}
\usepackage{dsfont}
\usepackage{color}
\usepackage{bbold}
\usepackage{float}
\usepackage[font=small,labelfont=bf]{caption}
\usepackage{subcaption}
\usepackage{fancyhdr}
\setlength{\headheight}{15.2pt}
\pagestyle{fancy}
\usepackage{listings} 
\usepackage{pict2e}
\usepackage{xfrac}
\usepackage{german}
\usepackage{mathtools}
\usepackage{graphicx}

%%%%%%%%         EIGENEBEFEHLE  %%%%%%%%%
\newcommand{\ddt}{\frac{\partial}{\partial{t}}}
\newcommand{\dnach}[1]{\frac{\partial}{\partial{#1}}}
\newcommand{\ddnach}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\dddnach}[3]{\frac{\partial^2{#1}}{\partial{#2} \partial{#3}}}
\newcommand{\dznach}[2]{\frac{\partial^2{#1}}{\partial{#2}^2}}
\newcommand{\vnabla}{\vec{\nabla}}
\newcommand{\evec}[1]{\mathbf{\hat{#1}}}
\newcommand{\bra}[1]{\langle{#1}|}
\newcommand{\ket}[1]{|{#1}\rangle}
\newcommand{\bracket}[2]{\langle{#1}|{#2}\rangle}
\newcommand{\up}{\uparrow}
\newcommand{\down}{\downarrow}
\newcommand{\updown}{\uparrow\downarrow}
\newcommand{\downup}{\downarrow\uparrow}
\newcommand{\upup}{\uparrow\uparrow}
\newcommand{\downdown}{\downarrow\downarrow}
\newcommand{\sandwich}[2]{\bra{#1} #2 \ket{#1}}
\newcommand{\fsqrt}[2]{\sqrt{\frac{#1}{#2}}}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{%
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  basicstyle=\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  %deletekeywords={...},            % if you want to delete keywords from the given language
  %escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,                    % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  language=C,                 % the language of the code
  %morekeywords={*,...},            % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=2,                       % sets default tabsize to 2 spaces
  %title=\lstname                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}

\lhead[\ ]{\ }
\chead[\ ]{\ }
\rhead[\ ]{\ }

\lfoot[\ ]{\ }
\cfoot[\ ]{\ }
\rfoot[\ ]{\ }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}
\title{Monte-Carlo Simulation von TIP5P-Wasser}
\author{Mathias H"old, Theodor Adaktylos}
\date{Wintersemester 2015/2016}
\maketitle
\thispagestyle{empty}
\newpage
\section{Introduction}

\end{document}

### 2016-06-19 include/classes.hpp

@@ -1,0 +1,29 @@
+#ifndef _INCLUDE_CLASSES_H_
+#define _INCLUDE_CLASSES_H_
+
+#include "globals.hpp"
+
+//class definitions
+class Particle
+{
+    public:
+     double *r;
+     double *v;
+     double *a;
+     double *rnew;
+     double *vnew;
+     double *vhalf;
+     double* r0;
+     double* v0;
+     int type;
+     int ID;
+     std::string name;
+     bool surface;
+
+     Particle();
+     Particle(std::string name);
+     Particle(double x,double y, double z, double v1, double v2, double v3);
+     Particle(double x,double y, double z, double v1, double v2, double v3, std::string Name);
+};
+
+#endif

### 2016-06-19 misc/list_parser.cpp

@@ -1,0 +1,55 @@
+#include <iostream>
+#include <cstdio>
+#include <string>
+#include <list>
+#include <iterator>
+
+
+void ManipulateList(std::list<int>& list)
+{
+    for(std::list<int>::iterator i1 = list.begin(); i1!=list.end();i1++)
+    {
+        (*i1) *= 2;
+    }
+}
+
+void DontManipulateList(std::list<int> list)
+{
+    for(std::list<int>::iterator i1 = list.begin(); i1!=list.end();i1++)
+    {
+        (*i1) *= 2;
+    }
+}
+
+
+void PrintList(std::list<int> list)
+{
+    for(std::list<int>::iterator i1 = list.begin(); i1!=list.end();i1++)
+    {
+        std::cout << (*i1) << std::endl;
+    }
+}
+
+
+int main()
+{
+    std::list<int> list1;
+    for(int i=0;i<20;i++)
+    {
+        list1.push_back(i);
+    }
+
+
+
+    PrintList(list1);
+
+    DontManipulateList(list1);
+
+    PrintList(list1);
+
+    ManipulateList(list1);
+
+    PrintList(list1);
+
+    return 0;
+}

### 2016-06-19 src/functions.cpp

@@ -1,0 +1,1108 @@
+#include "../include/functions.hpp"
+
+void InitPositions(Particle* particle)
+{
+    unsigned int M=1;
+    unsigned int n,x,y,z,k;
+    n=0;
+    
+    while(4*M*M*M < N)
+        ++M;
+
+    double a = L/M;
+    double xCell[4] = {0.25,0.75,0.75,0.25};
+    double yCell[4] = {0.25,0.75,0.25,0.75};
+    double zCell[4] = {0.25,0.25,0.75,0.75};
+    
+    for(x=0;x<M;x++)
+        for(y=0;y<M;y++)
+            for(z=0;z<M;z++)
+                for(k=0;k<4;k++)
+                    if(n<N)
+                    {
+                        particle[n].r[0] = (x + xCell[k])*a;
+                        particle[n].r[1] = (y + yCell[k])*a;
+                        particle[n].r[2] = (z + zCell[k])*a;
+                        particle[n].type = 1;
+                        particle[n].name = "Glass";
+						if((x==0 && (k == 3||k==0))|| (x==M-1 && (k==1 || k==2))|| (y==0 && (k==0 || k==2)) || (y==M-1 && (k==1 || k==3)) || (z==0 && (k==0 || k==1)) || (z==M-1 && (k==2 || k==3)))
+							particle[n].surface=true;
+                        n++;
+                    }
+}
+
+unsigned int NumberOfParticles()
+{
+        double mu;
+        //double P=0.0024;
+        //This is the calculated atmospheric pressure with LJ units
+        double P=5;
+        double k=1.0;
+        double T=0.9;
+        mu = 0.01*L*L*P*pow((1/(2*M_PI*k*T)),0.5);
+        //std::cout << mu << std::endl;
+        return gsl_ran_poisson(r,mu);
+}
+
+void ComputeAccelerations(Particle* particle) 
+{
+    double rij[3];
+    double rSqd = 0;
+    for (unsigned int i = 0; i < N; i++)         // set all accelerations to zero
+        for (unsigned int k = 0; k < 3; k++)
+            particle[i].a[k] = 0;
+    
+    for(unsigned int i=0;i<N-1;i++)
+    {
+        for(unsigned int j=i+1;j<N;j++)
+        {
+            rSqd = 0;
+            for(unsigned int m=0;m<3;m++)
+            {
+                rij[m] = particle[i].r[m]-particle[j].r[m];
[…]

### 2016-06-19 src/main.cpp

@@ -1,0 +1,107 @@
+#include "../include/globals.hpp"
+#include "../include/classes.hpp"
+#include "../include/functions.hpp"
+
+int main(int argc,char** argv)
+{
+    system("rm output/combined/*");
+    int a;
+    int run=0;
+    gsl_rng_set(r,98);
+    FILE* output = fopen("output/box_test.xyz","w");
+    FILE* pressure = fopen("output/pressure.dat","w");
+    FILE* baroOut = fopen("output/barostat.dat","w");
+    FILE* combinedOut = fopen("output/combined.xyz","w");
+    FILE* particledata = fopen("output/particledata.dat","w");
+    FILE* tempout = fopen("output/temperature.dat","w");
+    FILE* cmData = fopen("output/cmdata.dat","w");
+    FILE* equiData = fopen("output/equidata.dat","w");
+    FILE* readData = fopen("output/equidata.dat","r");
+    FILE* totErgData = fopen("output/energy.dat","w");
+    FILE* equidata = fopen("output/readwritetest.dat","w");
+    FILE* equidata2 = fopen("output/readwritetest2.dat","w");
+    FILE* surfacedata = fopen("output/surfacedata.xyz","w");
+    FILE* particleTracker = fopen("output/combined/id_1400.xyz","w");
+    fclose(combinedOut);
+    Particle *cube = new Particle[N];
+    std::list<Particle*> gas;
+    double* rCM = new double[3];
+    double* vCM = new double[3];
+    double energy = 0;
+    InitPositions(cube);
+    fclose(surfacedata);
+    calcCM(cube,rCMStart,vCM);
+    calcCM(cube,rCM,vCM);
+    fprintf(output,"%d\nCube\n",N);
+    InitVelocities(cube);
+    fprintf(pressure,"%lf\n",Pressure(cube));
+    ComputeAccelerations(cube);
+/*
+ *    for(run=0;run<10000;run++)
+ *    {
+ *            if(run%200==0)
+ *                printf("Zeitschritt %d\n",run);
+ *            VelocityVerlet(cube,0,output);
+ *            if(run%100 == 0)
+ *            {
+ *                calcTemp(cube,tempout); 
+ *                //energy = totErg(cube);
+ *                //fprintf(totErgData,"%d\t%lf\n",run,energy);
+ *            }
+ *    }
[…]

### 2016-06-19 test/barotest.cpp

@@ -1,0 +1,52 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../include/globals.hpp"
+#include "../include/classes.hpp"
+#include "../include/functions.hpp"
+
+int main(int argc, char** argv)
+{
+    system("rm output/combined/*");
+    FILE* tempdata = fopen("tempdata.dat","w");
+    int run=0;
+    gsl_rng_set(r,98);
+    Particle *cube = new Particle[N];
+    std::list<Particle*> gas;
+    double* rCM = new double[3];
+    double* vCM = new double[3];
+    InitPositions(cube);
+    calcCM(cube,rCMStart,vCM);
+    calcCM(cube,rCM,vCM);
+    InitVelocities(cube);
+    ComputeAccelerations(cube);
+
+    for(run=0;run<10000;run++)
+    {
+            if(run%200==0)
+                printf("Zeitschritt %d\n",run);
+            VelocityVerlet(cube,0,NULL);
+            if(run%100 == 0)
+                calcTemp(cube,tempdata); 
+    }
+
+    ComputeSoftSphere(gas,cube);
+
+    for(run = 0;run<20000;run++)
+    {
+        if(run%100==0)
+            printf("Zeitschritt %d\n",run);
+        VelocityVerlet(cube,0,NULL);
+        Barostat(cube,gas);
+        harmonicTrap(rCM,vCM,rCMStart,cube);
+        if(run%10==0)
+            GenerateOutput(cube,gas,run);
+        if(run%10==0)
+            calcTemp(cube,tempdata); 
+    }
+    return 0;
+}

### 2016-06-19 test/harmonic.cpp

@@ -1,0 +1,42 @@
+#include "../include/globals.hpp"
+#include "../include/classes.hpp"
+#include "../include/functions.hpp"
+
+void computeHarmonic(Particle* particle1, double* pos)
+{
+    double rij[3];
+    double rSqd = 0;
+    const double k = 15.;
+    for (int k = 0; k < 3; k++)
+        particle1->a[k] = 0;
+    
+    for(int i=0;i<3;i++)
+        rij[i] = pos[i] - particle1->r[i];
+    rSqd = rij[0]*rij[0]+rij[0]*rij[0]+rij[0]*rij[0];
+
+    for(int m=0;m<3;m++)
+        particle1->a[m] += k*rij[m];
+}
+
+void velocityHarmonic(Particle* particle1,double* pos)
+{
+    int i,m;
+
+    for(m=0;m<3;m++)
+    {
+        particle1->r[m] += particle1->v[m]*dt + 0.5 * particle1->a[m] * dt * dt;
+        particle1->v[m] += 0.5 * particle1->a[m] * dt;
+    }
+     
+    computeHarmonic(particle1,pos);
+    for(m=0;m<3;m++)
+        particle1->v[m] += 0.5 * particle1->a[m] * dt;
+}
+
+int main()
+{
+    Particle *particles = new Particle[N];
+    InitPositions(particles);
+    FILE* positions = fopen("comtestpos.xyz","w"); 
+
+}

### 2016-06-19 test/linetest.cpp

@@ -1,0 +1,163 @@
+#include <iostream>
+#include "../include/classes.hpp"
+#include "../include/globals.hpp"
+#include "../include/functions.hpp"
+
+void oneParticleVerlet(Particle* particle);
+void computeOneAcceleration(Particle* particle);
+//bool checkIfOnLine(Particle* particle);
+
+int main()
+{
+    std::ofstream outfile;
+    outfile.open("oneparticle.xyz");
+    Particle* p1 = new Particle();
+    for(int m=0;m<3;m++)
+        p1->r[m] = 0;
+    p1 -> v[0] = 0.5;
+    p1 -> v[1] = 0.5;
+    p1 -> v[2] = 0.5;
+    for(int m=0;m<3;m++)
+    {
+        p1->r0[m] = p1->r[m];
+        p1->v0[m] = p1->v[m];
+    }
+
+    /*
+     *outfile << "1\nParticleTest\n";
+     *outfile << "1\t";
+     *for(int m=0;m<3;m++)
+     *    outfile << p1->r[m] << " ";
+     *outfile << std::endl;
+     */
+    
+    for(int i=0;i<1000;i++)
+    {
+        outfile << "1\nParticleTest\n";
+        outfile << "1\t";
+        for(int m=0;m<3;m++)
+            outfile << p1->r[m] << " ";
+        outfile << std::endl;
+
+        oneParticleVerlet(p1);
+        if(!checkIfOnLine(p1))
+        {
+            std::cout << "Linie wurde verlassen! (" << i << ")" << std::endl;
+            break;
+        }
+    }
+    p1->v[0] = 0.1;
+    for(int i=1000;i<2000;i++)
+    {
+        outfile << "1\nParticleTest\n";
+        outfile << "1\t";
+        for(int m=0;m<3;m++)
+            outfile << p1->r[m] << " ";
+        outfile << std::endl;
+
+        oneParticleVerlet(p1);
+        if(!checkIfOnLine(p1))
+        {
+            std::cout << "Linie wurde verlassen! (" << i << ")" << std::endl;
+            break;
+        }
+    }
+
+    return 0;
+}
+
+void oneParticleVerlet(Particle* particle)
+{
+    int i,m;
+
+    for(m=0;m<3;m++)
+    {
+        particle->r[m] += particle->v[m]*dt + 0.5 * particle->a[m] * dt * dt;
+        particle->v[m] += 0.5 * particle->a[m] * dt;
[…]

### 2016-06-19 test/old/classes_test.cpp

@@ -1,0 +1,317 @@
+#include <iostream>
+#include <cmath>
+#include <gsl/gsl_rng.h>
+#include <gsl/gsl_randist.h>
+
+using namespace std;
+int ID=1;
+
+
+const gsl_rng_type *T = gsl_rng_taus;
+gsl_rng *r = gsl_rng_alloc(T);
+
+double mean = 5;
+double sigma = 0.5;
+
+class Particle
+{
+	public:
+	 double x,y,z;
+	 double vx,vy,vz;
+	 Particle();
+};
+
+Particle::Particle()
+{
+				x = 0;
+				y = 0;
+				z = 0;
+				vx = 0;
+				vy = 0;
+				vz = 0;
+}
+
+struct Node
+{
+				Particle *part;
+				int id;
+				Node *next;
+};
+
+typedef struct Node s_Node;
+
+Particle *testparticle = new Particle[20];
+
+//initialization 
+void initNode(struct Node *head)
+{
+				head -> part = new Particle();
+				head -> id = ID;
+				ID+=1;
+				head -> part -> x = gsl_rng_uniform(r);
+				head -> part -> y = gsl_rng_uniform(r);
+				head -> part -> z = gsl_rng_uniform(r);
+				head -> next = NULL;
+}
+
+//appending
+void addNode(struct Node *head)
+{
+				Node *newNode = new Node;
+				newNode -> part = new Particle();
+				newNode -> id = ID;
+				ID+=1;
+				newNode -> next = NULL;
+
+				Node *current = head;
+				while(current)
+				{
+								if(current -> next == NULL)
+								{
+												current -> next = newNode;
+												return;
+								}
+								current = current -> next;
+				}
+}
+
+bool deleteNode(struct Node  **head, Node *toDelete)
+{
+				Node *current = *head;
+				if(toDelete == *head)
+				{
+								cout << "Cannot delete Head" << endl;
+								return false;
+				}
+				//if(toDele6te == *head)
+				//{
+								//*head = current -> next;
+								//delete toDelete;
+								//return true;
+				//}
+
+				while(current)
+				{
+								if(current -> next == toDelete)
+								{
+												current -> next = toDelete -> next;
+												delete toDelete;
+												return true;
+								}
+								current = current -> next;
+				}
+				return false;
+}
+
+void printList(struct Node *head)
+{
+				Node *current = head;
+				while(current)
+				{
+								cout << "Particle " << current -> id << 
[…]

### 2016-06-19 test/old/erasetest.cpp

@@ -1,0 +1,46 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <list>
+#include <iterator>
+#include <vector>
+
+int main()
+{
+    std::list<int> list1;
+    for(int i=0;i<100;i++)
+    {
+        list1.push_back(i);
+    }
+
+    std::list<int>::iterator iter1;
+    std::list<int>::iterator iter2;
+    for(iter1 = list1.begin();iter1 != list1.end();iter1++)
+    {
+        std::cout << *iter1 << ", ";
+        //if(*iter1==50)
+            //iter1=list1.erase(iter1);
+    }
+    std::cout << std::endl;
+
+    for(iter1 = list1.begin();iter1 != list1.end();iter1++)
+    {
+        if(*iter1==50)
+            iter1=list1.erase(iter1);
+    }
+
+    for(iter1 = list1.begin();iter1 != list1.end();iter1++)
+    {
+        std::cout << *iter1 << ", ";
+    }
+    std::cout << std::endl; 
+
+    int a;
+    int b = 5;
+    
+    for(a=10;a<b;a++)
+    {
+        std::cout << a << std::endl;
+    }
+    return 0;
+}

### 2016-06-19 test/readwritetest.cpp

@@ -1,0 +1,86 @@
+#include "../include/globals.hpp"
+#include "../include/classes.hpp"
+#include "../include/functions.hpp"
+
+int main()
+{
+    //Particle* p1 = new Particle[N];
+    //for(int i=0;i<N;i++)
+        //p1[i] = new Particle();
+/*
+ *    int M=1;
+ *    int n,x,y,z,k;
+ *    n=0;
+ *    
+ *    while(4*M*M*M < N)
+ *        ++M;
+ *
+ *    double a = L/M;
+ *    double xCell[4] = {0.25,0.75,0.75,0.25};
+ *    double yCell[4] = {0.25,0.75,0.25,0.75};
+ *    double zCell[4] = {0.25,0.25,0.75,0.75};
+ *    
+ *    for(x=0;x<M;x++)
+ *        for(y=0;y<M;y++)
+ *            for(z=0;z<M;z++)
+ *                for(k=0;k<4;k++)
+ *                    if(n<N)
+ *                    {
+ *                        p1[n]->r[0] = (x + xCell[k])*a;
+ *                        p1[n]->r[1] = (y + yCell[k])*a;
+ *                        p1[n]->r[2] = (z + zCell[k])*a;
+ *                        n++;
+ *                    }
+ */
+    /*
+     *InitPositions(p1);
+     *FILE* p1data = fopen("p1data.dat","wb");
+     *for(int i=0;i<N;i++)
+     *{
+     *    for(int j=0;j<3;j++)
+     *        fwrite(&p1[i].r[j],sizeof(double),1,p1data);
+     *    for(int j=0;j<3;j++)
+     *        fwrite(&p1[i].v[j],sizeof(double),1,p1data);
+     *    for(int j=0;j<3;j++)
+     *        fwrite(&p1[i].a[j],sizeof(double),1,p1data);
+     *}
+     *FILE* testout1 = fopen("testread1.dat","w");
+     *fclose(p1data);
+     *
+     *for(int n=0;n<N;n++)
+     *{
+     *    fprintf(testout1,"%lf\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\n",p1[n].r[0],p1[n].r[1],p1[n].r[2],p1[n].v[0],p1[n].v[1],p1[n].v[2],p1[n].a[0],p1[n].a[1],p1[n].a[2]);
+     *}
+     */
+
+    /*
[…]

### 2016-06-19 test/vector_erase_test.cpp

@@ -1,0 +1,24 @@
+#include <iostream>
+#include <vector>
+#include <string>
+
+int main()
+{
+    std::vector<int> numbers;
+    for(int i=0;i<20;i++)
+        numbers.push_back(i);
+    std::cout << std::string(20,'-') << std::endl;
+    std::cout << "Before erase" << std::endl; 
+    std::cout << std::string(20,'-') << std::endl;
+    for(int i=0;i<numbers.size();i++)
+        std::cout << "numbers[" << i << "]: " << numbers[i] << std::endl;
+    for(int i=0;i<numbers.size();i++)
+        if(numbers[i] == 5)
+            numbers.erase(numbers.begin()+i);
+    std::cout << std::string(20,'-') << std::endl;
+    std::cout << "After erase" << std::endl; 
+    std::cout << std::string(20,'-') << std::endl;
+    for(int i=0;i<numbers.size();i++)
+        std::cout << "numbers[" << i << "]: " << numbers[i] << std::endl;
+    return 0;
+}

### 2016-06-19 test/velocitytest.cpp

@@ -1,0 +1,56 @@
+#include "../include/globals.hpp"
+#include "../include/classes.hpp"
+#include "../include/functions.hpp"
+
+int main()
+{
+    Particle *particles = new Particle[N];
+    InitPositions(particles);
+    FILE* positions = fopen("velocitytest.xyz","w"); 
+    FILE* brainfart = NULL;
+    fprintf(positions,"%d\nComTest\n",N);
+    for(int i=0;i<N;i++)
+    {
+        fprintf(positions,"Ar\t");
+        for(int j=0;j<3;j++)
+            fprintf(positions,"%lf\t",(particles[i].r[j])*10);
+        fprintf(positions,"\n");
+    }
+    InitVelocities(particles);
+    for(int i=0;i<1000;i++)
+    {
+        VelocityVerlet(particles,brainfart);
+        fprintf(positions,"%d\nComTest\n",N);
+        for(int i=0;i<N;i++)
+        {
+            fprintf(positions,"Ar\t");
+            for(int j=0;j<3;j++)
+                fprintf(positions,"%lf\t",(particles[i].r[j])*10);
+            fprintf(positions,"\n");
+        }
+    }
+    double* vCM = new double[3];
+    double* rCM = new double[3];
+    calcCM(particles,rCM,vCM);
+    std::cout << "rCM: (" << rCM[0] << ", " << rCM[1] << ", " << rCM[2] << ") " << std::endl;
+    std::cout << "vCM: (" << vCM[0] << ", " << vCM[1] << ", " << vCM[2] << ") " << std::endl;
+    for(int i=0;i<N;i++)
+        for(int k=0;k<3;k++)
+            particles[i].v[k] += 0.1;
+    for(int i=0;i<1000;i++)
+    {
+        VelocityVerlet(particles,brainfart);
+        fprintf(positions,"%d\nComTest\n",N);
+        for(int i=0;i<N;i++)
+        {
+            fprintf(positions,"Ar\t");
+            for(int j=0;j<3;j++)
+                fprintf(positions,"%lf\t",(particles[i].r[j])*10);
+            fprintf(positions,"\n");
+        }
+    }
+    calcCM(particles,rCM,vCM);
+    std::cout << "rCM: (" << rCM[0] << ", " << rCM[1] << ", " << rCM[2] << ") " << std::endl;
+    std::cout << "vCM: (" << vCM[0] << ", " << vCM[1] << ", " << vCM[2] << ") " << std::endl;
+    return 0;
+}

### 2016-06-21 src/classes.cpp
good things going on
**changed the particle tracker, so it prints out energies as well
**changed particle class, so mass and potential energy is included
**changed calculation of forces, so that potential energy is included

**added bib file for references
**outlined structure of thesis

@@ -12,6 +12,8 @@ Particle::Particle()
     vhalf = new double[3];
     r0 = new double[3];
     v0 = new double[3];
+    m = 1.;
+    potE = 0;
     
 
     for(int i=0;i<3;i++)
@@ -53,6 +55,8 @@ Particle::Particle(double x,double y, double z, double v1, double v2, double v3)
     ID = g_ID;
     type=2;
     g_ID++;
+    m = 1.;
+    potE = 0;
 }
 
 Particle::Particle(double x,double y, double z, double v1, double v2, double v3,std::string Name)
@@ -80,6 +84,11 @@ Particle::Particle(double x,double y, double z, double v1, double v2, double v3,
     type=2;
     g_ID++;
     name = Name;
+    if(strcmp(name.c_str(),"GasIn")==0 || strcmp(name.c_str(),"GasOut")==0)
+        m = 0.1;
+    else
+        m = 1.;
+    potE = 0;
 }
 
 Particle::Particle(std::string Name)
@@ -104,5 +113,10 @@ Particle::Particle(std::string Name)
     type = 0;
     g_ID++;
     name = Name;
+    if(strcmp(name.c_str(),"GasIn")==0 || strcmp(name.c_str(),"GasOut")==0)
+        m = 0.1;
+    else
+        m = 1.;
+    potE = 0;
 }
 

### 2016-06-21 test/barotest.cpp

@@ -13,6 +13,19 @@ int main(int argc, char** argv)
 {
     system("rm output/combined/*");
     FILE* tempdata = fopen("tempdata.dat","w");
+    FILE* trackdata[500];
+    std::string filename = "output/id/";
+    std::string number;
+    std::string filetype = ".dat";
+    std::string fileopen;
+    for(int i=1000;i<1500;i++)
+    {
+        number = numberToString(i);
+        fileopen = filename+number+filetype;
+        trackdata[i-1000] = fopen(fileopen.c_str(),"w");
+    }
+
+    FILE* trackdata_1000 = fopen("output/id_1000.dat","w");
     int run=0;
     gsl_rng_set(r,98);
     Particle *cube = new Particle[N];
@@ -36,15 +49,20 @@ int main(int argc, char** argv)
 
     ComputeSoftSphere(gas,cube);
 
-    for(run = 0;run<20000;run++)
+    for(run = 0;run<10000;run++)
     {
         if(run%100==0)
             printf("Zeitschritt %d\n",run);
         VelocityVerlet(cube,0,NULL);
         Barostat(cube,gas);
         harmonicTrap(rCM,vCM,rCMStart,cube);
+        for(int k=1000;k<1500;k++)
+            trackParticle(cube,gas,k,trackdata[k-1000]);
         if(run%10==0)
+        {
             GenerateOutput(cube,gas,run);
+            //trackParticle(cube,gas,1000, trackdata_1000);
+        }
         if(run%10==0)
             calcTemp(cube,tempdata); 
     }

### 2016-06-21 test/output/id/findmax.py

@@ -1,0 +1,72 @@
+import numpy as np
+import scipy as sp
+import re
+import matplotlib.pyplot as plot
+
+def findmaxvalue(number):
+    if(number < 10):
+        filename = "100"+str(number)+".dat"
+    elif(number >= 10 and number < 100):
+        filename = "10"+str(number)+".dat"
+    elif(number >= 100):
+        filename = "1"+str(number)+".dat"
+
+    data_array = []
+    with open(filename,"r") as infile:
+        line = infile.read()
+        data = line.split("\n")
+        for i in data:
+            if i != "test" and i != "1":
+                data_array.append(i)
+
+    for i in range(0,len(data_array)):
+        data_array[i] = data_array[i].split("\t")
+
+
+    for i in range(0,len(data_array)-1):
+        data_array[i].remove("GasIn")
+
+    data_array.remove(data_array[len(data_array)-1])
+
+
+    for i in range(0,len(data_array)):
+        for j in range(0,len(data_array[i])):
+            data_array[i][j] = float(data_array[i][j])
+    x_data  = []
+    y_data  = []
+    z_data  = []
+    vx_data = []
+    vy_data = []
+    vz_data = []
+    v_data  = []
+    ax_data = []
+    ay_data = []
+    az_data = []
+    a_data  = []
+
+    for i in range(0,len(data_array)):
+        x_data.append(data_array[i][0])
+        y_data.append(data_array[i][1])
+        z_data.append(data_array[i][2])
+        vx_data.append(data_array[i][3])
+        vy_data.append(data_array[i][4])
+        vz_data.append(data_array[i][5])
+        v_data.append(data_array[i][6])
+        ax_data.append(data_array[i][7])
+        ay_data.append(data_array[i][8])
+        az_data.append(data_array[i][9])
+        a_data.append(data_array[i][10])
+
+    max_a= np.amax(a_data)
+    max_a_index = np.argmax(a_data)
+    return([max_a_index,max_a])
+
+
+vals = []
+for i in range(0,500):
+    vals.append(findmaxvalue(i))
+for i in range(0,len(vals)):
+    print(str(vals[i][0]) + "   " + str(vals[i][1]))
+# print(vals)
+# plot.scatter(vals)
+# plot.show()

### 2016-07-14 src/main.cpp
Added tests for single particle investigation

@@ -36,22 +36,20 @@ int main(int argc,char** argv)
     InitVelocities(cube);
     fprintf(pressure,"%lf\n",Pressure(cube));
     ComputeAccelerations(cube);
-/*
- *    for(run=0;run<10000;run++)
- *    {
- *            if(run%200==0)
- *                printf("Zeitschritt %d\n",run);
- *            VelocityVerlet(cube,0,output);
- *            if(run%100 == 0)
- *            {
- *                calcTemp(cube,tempout); 
- *                //energy = totErg(cube);
- *                //fprintf(totErgData,"%d\t%lf\n",run,energy);
- *            }
- *    }
- *
- *    writePositions(cube,"output/states/LJequilibriumSur.dat");
- */
+    for(run=0;run<10000;run++)
+    {
+            if(run%200==0)
+                printf("Zeitschritt %d\n",run);
+            VelocityVerlet(cube,0,output);
+            if(run%100 == 0)
+            {
+                calcTemp(cube,tempout); 
+                //energy = totErg(cube);
+                //fprintf(totErgData,"%d\t%lf\n",run,energy);
+            }
+    }
+
+    writePositions(cube,"output/states/LJequilibriumSur.dat");
 
     /*
      *for(run;run<50000;run++)
@@ -68,7 +66,7 @@ int main(int argc,char** argv)
      *}
      */
     
-    readPositions(cube,"output/states/LJequilibriumSur.dat");
+    //readPositions(cube,"output/states/LJequilibriumSur.dat");
     //readPositions(cube,"output/states/eHEXEquilibrium.dat");
         
     ComputeSoftSphere(gas,cube);

### 2016-07-14 test/output/id/pics/plotscript.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+files="../*.dat"
+#for f in $files
+#do
+    #echo $(basename ${f%.*})
+#done
+for f in $files
+do
+    filename=$(basename ${f%.*})
+    gnuplot -e "\
+        set term png; \
+        set output '$filename.png'; \
+        plot '$f' using 0:12 w l "
+done
+

### 2016-07-29 include/functions.hpp
Radical changes!
preparation for the remodeling of the program, such that
many configurations can be investigated and saved

@@ -24,9 +24,11 @@ void CheckBoundaries(std::vector<Particle*> &particles);
 void Barostat(Particle* cube, std::list<Particle*>& gas);
 void ComputeSoftSphere(std::list<Particle*>& gas, Particle* cube);
 void PrintAllData(Particle* cube, std::list<Particle*> gas,FILE* output);
-void eHEX(Particle* cube,FILE* output);
+//void eHEX(Particle* cube,FILE* output);
+void eHEX(Particle* cube);
 void calcTemp(Particle* cube,FILE* output);
 void calcCM(Particle* particles,double *rCM, double* vCM);
+void calcCM(Particle* particles,double *rCM, FILE* output);
 bool fileExist(const std::string& filename);
 double totErg(Particle* particles);
 void writePositions(Particle* particles, std::string filename);
@@ -44,6 +46,7 @@ std::string numberToString(T number)
 void printDistances(Particle* cube, std::list<Particle*> &particles,int Run);
 //bool checkIfOnLine(Particle* particle);
 bool checkIfOnLine(std::list<Particle*>::iterator iterator);
+std::string DateToString();
 
 
 #endif

### 2016-07-29 src/functions.cpp

@@ -36,7 +36,7 @@ unsigned int NumberOfParticles()
         double mu;
         //double P=0.0024;
         //This is the calculated atmospheric pressure with LJ units
-        double P=5;
+        double P=4;
         double k=1.0;
         double T=0.9;
         mu = 0.01*L*L*P*pow((1/(2*M_PI*k*T)),0.5);
@@ -276,20 +276,20 @@ void InitBarostat(std::list<Particle*>& particles)
         //std::cout << "N = " << N << std::endl;
         for(i=0;i<N;i++)
         {
+                particles.push_back(new Particle(-L/2,
+                            gsl_rng_uniform(r)*L,
+                            gsl_rng_uniform(r)*L,
+                            gsl_ran_rayleigh(r,sigma),
+                            gsl_ran_gaussian(r,sigma),
+                            gsl_ran_gaussian(r,sigma),"GasIn"));
                 /*
-                 *particles.push_back(new Particle(-L/2,
-                 *            gsl_rng_uniform(r)*L,
-                 *            gsl_rng_uniform(r)*L,
+                 *particles.push_back(new Particle(-L/2.,
+                 *            (2*(gsl_rng_uniform(r)-0.5)*L)+0.5*L,
+                 *            (2*(gsl_rng_uniform(r)-0.5)*L)+0.5*L,
                  *            gsl_ran_rayleigh(r,sigma),
                  *            gsl_ran_gaussian(r,sigma),
                  *            gsl_ran_gaussian(r,sigma),"GasIn"));
                  */
-                particles.push_back(new Particle(-L/2.,
-                            (2*(gsl_rng_uniform(r)-0.5)*L)+0.5*L,
-                            (2*(gsl_rng_uniform(r)-0.5)*L)+0.5*L,
-                            gsl_ran_rayleigh(r,sigma),
-                            gsl_ran_gaussian(r,sigma),
-                            gsl_ran_gaussian(r,sigma),"GasIn"));
                 iter++;
                 //printf("%d: (%lf,%lf,%lf)\n",(*iter)->ID,(*iter)->r[0],(*iter)->r[1],(*iter)->r[2]);
                 fprintf(baroOut,"%lf\t%lf\t%lf\n",(*iter)->r[0],(*iter)->r[1],(*iter)->r[2]);
@@ -300,20 +300,20 @@ void InitBarostat(std::list<Particle*>& particles)
         //std::cout << "N = " << N << std::endl;
         for(i=0;i<N;i++)
[…]

### 2016-07-29 src/globals.cpp

@@ -8,14 +8,17 @@ gsl_rng *r1 = gsl_rng_alloc(T);
 int g_ID=1;
 const unsigned int N = 864;
 
-double FG;
 double* rCMStart = new double[3];
-const double rho = 1.1;
-const double L = pow(N/rho,1.0/3);
-const double rCutOff = 2.5;
-const double dt = 0.01;
-const double eps = 0.5;
-const double dQ=0.010;
+double rho = 1.1;
+double Temp;
+double L = pow(N/rho,1.0/3);
+double rCutOff = 2.5;
+double dt = 0.01;
+double eps = 0.5;
+double dQ=0.050;
+char* input;
+//ACHTUNG: FG war bis jetzt nicht gesetzt!!! moegliche Aenderungen hier beginnen nachzuvollziehen
+double FG = dQ/dt;
 
 bool PBC_FLAG = false;
 bool EHEX_FLAG = false;

### 2016-07-29 test/initvalues.cpp

@@ -1,0 +1,28 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../include/globals.hpp"
+#include "../include/classes.hpp"
+#include "../include/functions.hpp"
+
+int main(int argc, char** argv)
+{
+    std::string dir = "test";
+    chdir(dir.c_str());
+    FILE* initvalues = fopen("initvalues.dat","wb");
+    double T;
+    double dQ;
+    std::cout << "\tT = ";
+    std::cin >> T;  
+    std::cout << "\tdQ = ";
+    std::cin >> dQ;  
+    fwrite(&T,sizeof(double),1,initvalues);
+    fwrite(&dQ,sizeof(double),1,initvalues);
+    
+    fclose(initvalues);
+    return 0;
+}

### 2016-07-29 test/newbarotest.cpp

@@ -1,0 +1,79 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../include/globals.hpp"
+#include "../include/classes.hpp"
+#include "../include/functions.hpp"
+
+int main(int argc, char** argv)
+{
+    system("rm output/combined/*");
+    FILE* tempdata = fopen("tempdata.dat","w");
+    FILE* trackdata[500];
+    std::list<FILE*> trackfiles;
+    std::list<FILE*>::iterator t_iter;
+
+    std::string filename = "output/id/";
+    std::string number;
+    std::string filetype = ".dat";
+    std::string fileopen;
+    /*
+     *for(int i=1000;i<1500;i++)
+     *{
+     *    number = numberToString(i);
+     *    fileopen = filename+number+filetype;
+     *    trackdata[i-1000] = fopen(fileopen.c_str(),"w");
+     *}
+     */
+
+    //FILE* trackdata_1000 = fopen("output/id_1000.dat","w");
+    int run=0;
+    gsl_rng_set(r,98);
+    Particle *cube = new Particle[N];
+    std::list<Particle*> gas;
+    double* rCM = new double[3];
+    double* vCM = new double[3];
+    InitPositions(cube);
+    calcCM(cube,rCMStart,vCM);
+    calcCM(cube,rCM,vCM);
+    InitVelocities(cube);
+    ComputeAccelerations(cube);
+
+    for(run=0;run<10000;run++)
+    {
+            if(run%200==0)
+                printf("Zeitschritt %d\n",run);
+            VelocityVerlet(cube,0,NULL);
+            if(run%100 == 0)
+                calcTemp(cube,tempdata); 
+    }
+    writePositions(cube,"output/positions.dat");
+
+    ComputeSoftSphere(gas,cube);
+    int cntr = 0;
+    for(run = 0;run<10000;run++)
+    {
+        if(run%100==0)
+            printf("Zeitschritt %d\n",run);
+        VelocityVerlet(cube,0,NULL);
+        Barostat(cube,gas);
+        harmonicTrap(rCM,vCM,rCMStart,cube);
+    
+        /*
+         *for(int k=1000;k<1500;k++)
+         *    trackParticle(cube,gas,k,trackdata[k-1000]);
+         */
+        if(run%10==0)
+        {
+            GenerateOutput(cube,gas,run);
[…]

### 2016-07-29 test/user_input.cpp

@@ -5,6 +5,8 @@
 #include <string>
 #include <cstring>
 
+char* input;
+
 int main(int argc,char** argv)
 {
     if(argc == 1)
@@ -12,16 +14,30 @@ int main(int argc,char** argv)
     if(argc > 1)
     {
         if(std::strcmp(argv[1],"init") == 0)
+        {
             puts("Initializing system");
+            input = argv[1];
+        }
         else if(std::strcmp(argv[1],"heat") == 0)
+        {
             puts("Heating system");
+            input = argv[1];
+        }
         else if(std::strcmp(argv[1],"baro") == 0)
+        {
             puts("Running barostat");
+            input = argv[1];
+        }
         else
         {
             puts("Input not recognized. Options are 'init', 'heat' and 'baro'");
             return 1;
         }
+    }
+    if(input != NULL)
+    {
+        std::string output(input);
+        std::cout << output << std::endl;
     }
     return 0;
 }

### 2016-08-08 doc/draft1.tex
good things going on
**changed the particle tracker, so it prints out energies as well
**changed particle class, so mass and potential energy is included
**changed calculation of forces, so that potential energy is included

**added bib file for references
**outlined structure of thesis
Added tests for single particle investigation
Radical changes!
preparation for the remodeling of the program, such that
many configurations can be investigated and saved
running through parameters
does not work yet. run gets stuck and does not do much...

@@
 \usepackage{xfrac}
-\usepackage{german}
+\usepackage[english]{babel}
 \usepackage{mathtools}
@@
 \begin{document}
-\title{Monte-Carlo Simulation von TIP5P-Wasser}
-\author{Mathias H"old, Theodor Adaktylos}
-\date{Wintersemester 2015/2016}
+\title{Temperature Measurements in Optical Tweezer Experiments}
+\author{Mathias H\"old, BSc.}
+\date{2016}
 \maketitle
@@
 \section{Introduction}
+%something about computer science in general, optical tweezers and the need to study the
+%systems on computers yada yada
+
+\newpage
+\section{Motivation}
+%introduction of the experiment, the problem and the idea
+
+\newpage
+\section{Simulation}
+%simulation techniques and general concept of the simulation itself, used methods and so on 
+To study the problem on the computer, one needs certain techniques. These techniques and their application will be described in this section.
+
+\subsection{Molecular Dynamics}
+
+\subsection{Reduced Units}
+\subsection{The Glass Nanoparticle}
+The glass particle from the experiment will be represented by a system of particles, interacting via a Lennard-Jones potential of the form
+\begin{equation}
+    U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right]
+\end{equation}
+or, using reduced units:
+\begin{equation}
+    U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right]
+\end{equation}
+
+\newpage
+\section{Results}
+% results with different parameters, graphs, etc.
+
+\newpage
+\section{Conclusion}
+% what does this all tell us? do we need to study more of this?
+
+\newpage
+\bibliography{references}
+\bibliographystyle{plain}
 

### 2016-08-08 src/functions.cpp

@@ -1192,7 +1192,7 @@ std::string DateToString()
     if(1+ltm->tm_hour < 10)
         datestring += "0";
     datestring += numberToString(ltm->tm_hour);
-    if(1+ltm->tm_min < 10)
+    if(1+ltm->tm_min < 10 || 1+ltm->tm_min == 9)
         datestring += "0";
     datestring += numberToString(ltm->tm_min);
     return datestring;

### 2016-08-08 test/test/show.sh

@@ -1,0 +1,9 @@
+#!/bin/bash
+
+directories="*/"
+
+for i in $directories
+do
+    echo $i
+    ls $i
+done

### 2016-08-23 doc/draft1.tex
changes changes changes

@@
 
-\subsection{Reduced Units}
 \subsection{The Glass Nanoparticle}
-The glass particle from the experiment will be represented by a system of particles, interacting via a Lennard-Jones potential of the form
+The glass particle from the experiment will be represented by a system of 864 particles, aligned regularly on a FCC (face centered cubic) lattice.
+This alignment is achieved by defining a minimum cell, containing points:
+\begin{eqnarray*}
+    p_1 &=& \{0,0,0\}\\
+    p_2 &=& \{0.5,0.5,0\}\\
+    p_3 &=& \{0.5,0,0.5\}\\
+    p_4 &=& \{0,0.5,0.5\}
+\end{eqnarray*}
+The whole system is then  created by copying this unit cell.\\
+The interaction between the atoms is modeled by the Lennard-Jones potential. It has the form
 \begin{equation}
-    U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right]
+    U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right].
 \end{equation}
-or, using reduced units:
+To simplify the equation and computation of the potential and other quantities, like the forces or pressure, it is useful to introduce reduced units.
+In general, reduced units BLABLABLA THINK OF GOOD TEXT HEREJKA\\
+The basic units in a system with Lennard-Jones interaction are length ($\sigma$), energy ($\varepsilon$ or $\varepsilon / k_B$) and mass ($m$). Every 
+quantity can now be written in terms of this units, so they become reduced quantities, denotet by an asterisk. The most important ones are:
+\begin{eqnarray*}
+    r^* &=& r/\sigma\\
+    T^* &=& k_B/\varepsilon \ T\\
+    U^* &=& U/\varepsilon \\
+    P^* &=& P \sigma^3/\varepsilon 
+\end{eqnarray*}
+With the above introduced reduced units, the Lennard-Jones potential can be written as
 \begin{equation}
-    U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right]
+    U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right].
 \end{equation}
+Since this is the most practical form, it will be used without the asterisk from here on.\\
+The Lennard-Jones potential is an additive pair-potential, so the total energy of the system can be calculated by summing over all pairs of atoms:
+\begin{equation}
+    U_{\text{tot}} = \sum_{i=1}^{N-1}\sum_{j=i+1}^N4\left[{r_{ij}}^{-12} - {r_{ij}}^{-6}\right]
+\end{equation}
+where $r_{ij}$ denotes the distance between atom $i$ and $j$. \\
+The Velocity-Verlet algorithm uses the forces to calculate the positions and velocities for the next timestep, we need to calculate the derivative of
+the potential energy:
+\begin{eqnarray}
+    F_{x} &=& -\frac{\partial}{\partial x} U(r) \nonumber\\
+                &=& -\frac{\partial}{\partial x} 4\left[{r}^{-12} - {r}^{-6}\right] \nonumber\\
+                &=& -4 \left[(-12){r}^{-13} - (-6){r}^{-7}\right] \frac{\partial r}{\partial x} \nonumber\\
+                &=& 48 \left[r^{-13} - 0.5 \ r^{-7}\right] \frac{x}{r} \nonumber\\
+    \label{eq:ljforce} &=& 48 \left[r^{-14} - 0.5 \ r^{-8}\right] x
+\end{eqnarray}
+This is the component of the force in x-direction -- the other components are calculated analogously.
 

### 2016-08-23 include/functions.hpp

@@ -20,8 +20,10 @@ void GenerateOutput(Particle *cube, std::list<Particle*> gas,FILE *combinedOut);
 void GenerateOutput(Particle *cube, std::vector<Particle*> gas,FILE *combinedOut);
 void CheckBoundaries(std::list<Particle*> &particles);
 void CheckBoundaries(Particle* cube,std::list<Particle*> &particles);
+void CheckBoundariesNew(Particle* cube,std::list<Particle*> &particles);
 void CheckBoundaries(std::vector<Particle*> &particles);
 void Barostat(Particle* cube, std::list<Particle*>& gas);
+void BarostatNew(Particle* cube, std::list<Particle*>& gas);
 void ComputeSoftSphere(std::list<Particle*>& gas, Particle* cube);
 void PrintAllData(Particle* cube, std::list<Particle*> gas,FILE* output);
 //void eHEX(Particle* cube,FILE* output);

### 2016-08-23 src/functions.cpp

@@ -739,6 +739,69 @@ void CheckBoundaries(Particle* cube,std::list<Particle*> &particles)
 
 }
 
+void CheckBoundariesNew(Particle* cube,std::list<Particle*> &particles)
+{
+    std::list<Particle*>::iterator temp;
+    double lowerBound = -(L/2 + eps);
+    double upperBound = L + L/2 + eps;
+    double aTotOld = 0;
+    double aTotNew = 0;
+    //double dist[3] = {0,0,0};
+    //double distSqd = 0.;
+    //double vSqd = 0.;
+
+    for(temp=particles.begin();temp!=particles.end();)
+    {
+        //if(!checkIfOnLine(temp))
+            //(*temp)->name = "GasOut";
+        /*
+         *vSqd = 0.;
+         *for(int m=0;m<3;m++)
+         *    vSqd += (*temp)->v[m];
+         *if(vSqd < 1e-4)
+         *    (*temp)->name = "GasOut";
+         */
+
+
+        /*
+         *for(int i=0;i<N;i++) // iterate over all particles
+         *{
+         *    if(cube[i].surface && std::strcmp(((*temp)->name).c_str(),"GasOut") != 0) // if particle is on the surface
+         *    {
+         *       for(int m=0;m<3;m++)
+         *          dist[m] = (*temp)->r[m] - cube[i].r[m];
+         *      distSqd = dist[0]*dist[0]+dist[1]*dist[1]+dist[2]*dist[2];
+         *      if(distSqd < 1.2)
+         *      {
+         *          (*temp)->name = "GasOut";
+         *          //std::cout << (*temp)->ID << std::endl;
+         *      }
+         *    }
+         *}
+         */
+        aTotOld = sqrt((*temp)->aOld[0]*(*temp)->aOld[0]+(*temp)->aOld[1]*(*temp)->aOld[1]+(*temp)->aOld[2]*(*temp)->aOld[2]);
+        aTotNew = sqrt((*temp)->a[0]*(*temp)->a[0]+(*temp)->a[1]*(*temp)->a[1]+(*temp)->a[2]*(*temp)->a[2]);
+        /*
+         *if(aTotOld > aTotNew)
+         *    (*temp)->name = "GasOut";
+         */
+        if(aTotNew > aTotOld && strcmp((*temp)->name.c_str(),"GasIn") == 0)
+            (*temp)->name = "GasChange";
[…]

### 2016-08-23 src/main.cpp

@@ -20,16 +20,25 @@ int main(int argc,char** argv)
  *double eps = 0.5;
  *double dQ=0.050;
  */
-    double q = 0.06;
-    while(q < 0.15)
+    double q = 0.01;
+    double at = 0.5;
+    while(q < 0.12)
     {
-        setValues(0.2,q,0.5,4,0.9);
-        std::cout << "==================================================" << std::endl;
-        std::cout << "q: " << q << std::endl;
-        std::cout << "==================================================" << std::endl;
-        mainLoop();
+        at = 0.6;
+        while(at < 1.3)
+        {
+            setValues(0.2,q,0.2,4,at);
+            std::cout << "==================================================" << std::endl;
+            std::cout << "q: " << q << std::endl;
+            std::cout << "Ambient Temperature: " << at << std::endl;
+            std::cout << "==================================================" << std::endl;
+            mainLoop();
+            at += 0.1;
+        }
         q += 0.01;
     }
+    //setValues(0.2,0.06,0.2,4,1.1);
+    //mainLoop();
     return 0;
 }
 
@@ -49,10 +58,16 @@ void mainLoop() {
     std::string folderName = "output/runs/";
     folderName += DateToString();
     int dir = mkdir(folderName.c_str(),S_IRWXU | S_IRWXG | S_IROTH | S_IXOTH);
+    int i=1;
     while(dir == -1)
     {
-        folderName += "_1";
-        mkdir(folderName.c_str(),S_IRWXU | S_IRWXG | S_IROTH | S_IXOTH);
+        std::string newFolderName = folderName + "_" + numberToString(i);
+        //folderName += "_";
+        //folderName += numberToString(i);
+        dir = mkdir(newFolderName.c_str(),S_IRWXU | S_IRWXG | S_IROTH | S_IXOTH);
+        if(dir == 0)
+            folderName = newFolderName;
+        i++;
     }
     chdir(folderName.c_str()); 
     system("mkdir combined");
@@ -117,7 +132,7 @@ void mainLoop() {
     ComputeAccelerations(cube);
     for(run=0;run<10000;run++)
     {
-            if(run%200==0)
+            if(run%500==0)
                 printf("(INIT) Zeitschritt %d\n",run);
             VelocityVerlet(cube,0,NULL);
             if(run%100 == 0)
@@ -125,6 +140,8 @@ void mainLoop() {
                 rescaleVelocities(cube);
[…]

### 2016-08-23 test/barotest.cpp

@@ -35,18 +35,22 @@ int main(int argc, char** argv)
     InitPositions(cube);
     calcCM(cube,rCMStart,vCM);
     calcCM(cube,rCM,vCM);
-    InitVelocities(cube);
-    ComputeAccelerations(cube);
-
-    for(run=0;run<10000;run++)
-    {
-            if(run%200==0)
-                printf("Zeitschritt %d\n",run);
-            VelocityVerlet(cube,0,NULL);
-            if(run%100 == 0)
-                calcTemp(cube,tempdata); 
-    }
-    writePositions(cube,"output/positions.dat");
+/*
+ *    InitVelocities(cube);
+ *    ComputeAccelerations(cube);
+ *
+ *    for(run=0;run<10000;run++)
+ *    {
+ *            if(run%200==0)
+ *                printf("Zeitschritt %d\n",run);
+ *            VelocityVerlet(cube,0,NULL);
+ *            if(run%100 == 0)
+ *                calcTemp(cube,tempdata); 
+ *    }
+ */
+    //writePositions(cube,"output/positions.dat");
+    readPositions(cube,"output/positions.dat");
+    chdir("output");
 
     ComputeSoftSphere(gas,cube);
 
@@ -55,7 +59,7 @@ int main(int argc, char** argv)
         if(run%100==0)
             printf("Zeitschritt %d\n",run);
         VelocityVerlet(cube,0,NULL);
-        Barostat(cube,gas);
+        BarostatNew(cube,gas);
         harmonicTrap(rCM,vCM,rCMStart,cube);
         for(int k=1000;k<1500;k++)
             trackParticle(cube,gas,k,trackdata[k-1000]);

### 2016-08-23 test/energytest.cpp

@@ -1,0 +1,73 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <cstdlib>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../include/classes.hpp"
+#include "../include/functions.hpp"
+#include "../include/globals.hpp"
+
+int main(int argc, char** argv)
+{
+    gsl_rng_set(r,98);
+    FILE* tempout = fopen("temperature_internal.dat","w");
+    FILE* comData = fopen("comdata.dat","w");
+    Particle *cube = new Particle[N];
+    std::list<Particle*> gas;
+    double* rCM = new double[3];
+    double* rCMtemp = new double[3];
+    double* vCM = new double[3];
+    double energy = 0;
+    int run;
+    InitPositions(cube);
+    calcCM(cube,rCMStart,vCM);
+    calcCM(cube,rCM,vCM);
+    InitVelocities(cube);
+    ComputeAccelerations(cube);
+    for(run=0;run<10000;run++)
+    {
+            if(run%500==0)
+                printf("(INIT) Zeitschritt %d\n",run);
+            VelocityVerlet(cube,0,NULL);
+            if(run%100 == 0)
+            {
+                rescaleVelocities(cube);
+                calcTemp(cube,tempout); 
+            }
+            if(run%400 == 0)
+                GenerateOutput(cube,gas,run);
+    }
+    ComputeSoftSphere(gas,cube);
+
+    for(run = 0;run<40000;run++)
+    {
+        if(run%500==0)
+            printf("(MEASURE)Zeitschritt %d\n",run);
+        eHEX(cube);
+        BarostatNew(cube,gas);
+        //std::cout << "works!" << std::endl;
+        harmonicTrap(rCM,vCM,rCMStart,cube);
+        //std::cout << "works!" << std::endl;
+        if(run%400==0)
+        {
+            //trackParticle(cube,gas,1400,particleTracker);
+            GenerateOutput(cube,gas,run+10000);
+            //std::cout << "works!" << std::endl;
+            calcCM(cube,rCMtemp,comData);
+        }
+        if(run%100==0)
+            calcTemp(cube,tempout); 
+    }
+    chdir("../../../");
+    delete [] cube;
+    gas.clear(); 
+    g_ID = 1;
+	EHEX_FLAG = false;
+
+    fclose(tempout);
+    fclose(comData);
+    return 0;
+}

### 2016-08-23 test/test/datetest.cpp

@@ -47,3 +47,8 @@ int main(int argc, char** argv)
 
     return 0;
 }
+
+void makeDir()
+{
+
+}

### 2016-09-02 doc/draft1.tex
Added forces and distances array
the idea is to have the forces and distances available in an array, so
they can be accessed at any given moment (-> pressure calculation via
virial)

--- 2016-08-08 doc/draft1.tex
@@
 %simulation techniques and general concept of the simulation itself, used methods and so on 
-To study the problem on the computer, one needs certain techniques. These techniques and their application will be described in this section.
+The problem at hand can be studied on an atomic level with the use of computer simulation. There is a variety of methods for computer simulations
+that are widely used, one of which being Molecular Dynamics (MD) simulations. The following section will give a brief overview of the concepts of this
+method, which is followed by the application to the simulation of the experiment.
 
 \subsection{Molecular Dynamics}
+Molecular Dynamics\cite{Frenkel2001} simulations is a technique for simulating, as the name suggests, the dynamics of a classical many-body system. In this case,
+classical means, that the trajectories of the individual particles are calculated using classical mechanics rather then quantum mechanics. For
+relatively big atoms/molecules this is a very good approximation, whereas for systems consisting of hydrogen or helium the effects of quantum
+mechanics cannot be neglected and other methods (such as ab-initio simulation) has to be used.\\
+The dynamics of the system are obtained by solving Newton's equations of motion for every particle.
 
-\subsection{Reduced Units}
-\subsection{The Glass Nanoparticle}
-The glass particle from the experiment will be represented by a system of particles, interacting via a Lennard-Jones potential of the form
-\begin{equation}
-    U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right]
-\end{equation}
-or, using reduced units:
-\begin{equation}
-    U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right]
-\end{equation}
+%\subsection{The Glass Nanoparticle}
+%The glass particle from the experiment will be represented by a system of 864 particles, aligned regularly on a FCC (face centered cubic) lattice.
+%This alignment is achieved by defining a minimum cell, containing points:
+%\begin{eqnarray*}
+    %p_1 &=& \{0,0,0\}\\
+    %p_2 &=& \{0.5,0.5,0\}\\
+    %p_3 &=& \{0.5,0,0.5\}\\
+    %p_4 &=& \{0,0.5,0.5\}
+%\end{eqnarray*}
+%The whole system is then  created by copying this unit cell.\\
+%The interaction between the atoms is modeled by the Lennard-Jones potential. It has the form
+%\begin{equation}
+    %U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right].
+%\end{equation}
+%To simplify the equation and computation of the potential and other quantities, like the forces or pressure, it is useful to introduce reduced units.
+%In general, reduced units BLABLABLA THINK OF GOOD TEXT HEREJKA\\
+%The basic units in a system with Lennard-Jones interaction are length ($\sigma$), energy ($\varepsilon$ or $\varepsilon / k_B$) and mass ($m$). Every 
+%quantity can now be written in terms of this units, so they become reduced quantities, denotet by an asterisk. The most important ones are:
+%\begin{eqnarray*}
+    %r^* &=& r/\sigma\\
+    %T^* &=& k_B/\varepsilon \ T\\
+    %U^* &=& U/\varepsilon \\
+    %P^* &=& P \sigma^3/\varepsilon 
+%\end{eqnarray*}
+%With the above introduced reduced units, the Lennard-Jones potential can be written as
+%\begin{equation}
+    %U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right].
+%\end{equation}
+%Since this is the most practical form, it will be used without the asterisk from here on.\\
+%The Lennard-Jones potential is an additive pair-potential, so the total energy of the system can be calculated by summing over all pairs of atoms:
+%\begin{equation}
+    %U_{\text{tot}} = \sum_{i=1}^{N-1}\sum_{j=i+1}^N4\left[{r_{ij}}^{-12} - {r_{ij}}^{-6}\right]
+%\end{equation}
+%where $r_{ij}$ denotes the distance between atom $i$ and $j$. \\
+%The Velocity-Verlet algorithm uses the forces to calculate the positions and velocities for the next timestep, we need to calculate the derivative of
+%the potential energy:
+%\begin{eqnarray}
+    %F_{x} &=& -\frac{\partial}{\partial x} U(r) \nonumber\\
+                %&=& -\frac{\partial}{\partial x} 4\left[{r}^{-12} - {r}^{-6}\right] \nonumber\\
+                %&=& -4 \left[(-12){r}^{-13} - (-6){r}^{-7}\right] \frac{\partial r}{\partial x} \nonumber\\
+                %&=& 48 \left[r^{-13} - 0.5 \ r^{-7}\right] \frac{x}{r} \nonumber\\
+    %\label{eq:ljforce} &=& 48 \left[r^{-14} - 0.5 \ r^{-8}\right] x
+%\end{eqnarray}
+%This is the component of the force in x-direction -- the other components are calculated analogously.
 

### 2016-09-02 include/functions.hpp

@@ -28,7 +28,9 @@ void ComputeSoftSphere(std::list<Particle*>& gas, Particle* cube);
 void PrintAllData(Particle* cube, std::list<Particle*> gas,FILE* output);
 //void eHEX(Particle* cube,FILE* output);
 void eHEX(Particle* cube);
+void calcPressure(Particle cube,FILE* output);
 void calcTemp(Particle* cube,FILE* output);
+void calcCOMTemp(double* vCOM,FILE* output);
 void calcCM(Particle* particles,double *rCM, double* vCM);
 void calcCM(Particle* particles,double *rCM, FILE* output);
 bool fileExist(const std::string& filename);
@@ -49,6 +51,7 @@ void printDistances(Particle* cube, std::list<Particle*> &particles,int Run);
 std::string DateToString();
 void rescaleVelocities(Particle* cube);
 void setValues(double temp, double dq, double Eps, double Pressure, double ambienttemp);
+
 
 
 #endif

### 2016-09-02 include/globals.hpp

@@ -40,6 +40,11 @@ extern double L;
 extern double rCutOff;
 extern double dt;
 
+extern double*** Forces;
+extern double*** Distances;
+
+extern std::vector<double> virial;
+
 extern char* input;
 
 extern bool EHEX_FLAG;
@@ -47,5 +52,10 @@ extern bool PBC_FLAG;
 
 enum Cubeface { X0, XL, Y0, YL, Z0, ZL };
 
+/*
+ *extern std::vector<std::vector<double>> energyIn;
+ *extern std::vector<std::vector<double>> energyOut;
+ */
 
 #endif
+

### 2016-09-02 src/main.cpp

@@ -20,14 +20,26 @@ int main(int argc,char** argv)
  *double eps = 0.5;
  *double dQ=0.050;
  */
+    for(unsigned int i=0;i<N;i++)
+    {
+        Forces[i] = new double*[N];
+        Distances[i] = new double*[N];
+    }
+    for(unsigned int i=0;i<N;i++)
+        for(unsigned int j=0;j<N;j++)
+        {
+            Forces[i][j] = new double[3];
+            Distances[i][j] = new double[3];
+        }
+
     double q = 0.01;
     double at = 0.5;
     while(q < 0.12)
     {
-        at = 0.6;
+        at = 0.5;
         while(at < 1.3)
         {
-            setValues(0.2,q,0.2,4,at);
+            setValues(0.4,q,0.2,4,at);
             std::cout << "==================================================" << std::endl;
             std::cout << "q: " << q << std::endl;
             std::cout << "Ambient Temperature: " << at << std::endl;
@@ -118,7 +130,9 @@ void mainLoop() {
      *FILE* particleTracker = fopen("output/combined/id_1400.xyz","w");
      */
     FILE* tempout = fopen("temperature_internal.dat","w");
+    FILE* COMtempout = fopen("temperature_com.dat","w");
     FILE* comData = fopen("comdata.dat","w");
+    FILE* vCOMData = fopen("vCOMData.dat","w");
     Particle *cube = new Particle[N];
     std::list<Particle*> gas;
     double* rCM = new double[3];
@@ -183,8 +197,16 @@ void mainLoop() {
             calcCM(cube,rCMtemp,comData);
         }
         if(run%100==0)
+        {
             calcTemp(cube,tempout); 
+            calcCOMTemp(vCM,COMtempout);
+            fprintf(vCOMData,"%lf\t%lf\t%lf\n",vCM[0],vCM[1],vCM[2]);
+        }
     }
+    FILE* pressure = fopen("inst_pressure.dat","w");
+    for(unsigned int i=0;i<virial.size();i++)
+        fprintf(pressure,"%lf\n",virial[i]);
+    fclose(pressure);
     chdir("../../../");
     delete [] cube;
     gas.clear(); 
@@ -192,4 +214,6 @@ void mainLoop() {
 	EHEX_FLAG = false;
     fclose(tempout);
     fclose(comData);
+    fclose(COMtempout);
+    fclose(vCOMData);
 }

### 2016-10-10 doc/draft1.tex
various changes

@@
 \usepackage{float}
-\usepackage[font=small,labelfont=bf]{caption}
-\usepackage{subcaption}
+\usepackage[font=scriptsize,labelfont=bf]{caption}
+%\usepackage[font=footnotesize,labelfont=bf]{caption}
+%\usepackage{subcaption}
 \usepackage{fancyhdr}
@@
 %introduction of the experiment, the problem and the idea
+\section{The experiment}
+The starting point of this thesis is an experiment conducted by Gieseler et al \cite{Gieseler2014}. It is an optical tweezer experiment, where the
+motion of a glass nanoparticle in a laser trap was used to investigate the fluctuation theorem\cite{Crooks1999}.
+\subsection{Experimental setup}
+In the experiment, a silica nano particle with a radius of about 75 nm and mass of about $3 \times 10^{-18}$kg is trapped in a laser beam within a
+vacuum chamber. The trapping of the silica nano particle (which will be referred to as \textit{glass particle}) is achieved by a gradient force of the
+laser beam acting on the particle. The experimental setup is depicted in fig. \ref{fig:setup}.\\
+The particle fluctuates within the trap in all three spatial directions. These fluctuations can be approximated
+such that they are decoupled, which means that they can be described by a 1-dimensional Langevin equation:
+\begin{equation}
+    \ddot{x} + \Gamma_0 \dot{x} + \Omega^2_0x = \frac 1 m \left(F_\text{fluct} + F_\text{ext}\right)
+\end{equation}
+%The experiment is set up as follows: a glass sphere with radius $r \approx 75$ nm and mass $m \approx 3 \times 10^{-18}$ kg is trapped in a laser beam
+%in a vacuum chamber. The initial state of the glass particle is prepared by modulation of the laser, such that the oscillation of the glass particle
+%is suppressed in every direction. This is done by measuring the oscillation of the particle in every direction (assuming that the oscillation in every
+%direction is decoupled) and using this information as feedback for the laser. This experimental setup can be seen in figure \ref{fig:setup}.
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[scale=0.3]{images/experimental_setup.jpg}
+        \caption{Experimental setup of the optical tweezer experiment. A silica nano particle is trapped in a laser beam via gradient force in a
+        vacuum. The feedback is used to cool down the particle and create a non-equilibrium steady state. In the first part of the experiment, the
+    feedback is turned off and the motion of the particle towards an equilibrated state is observed. In the second part of the experiment, the steady state
+of the particle is modified by a parametric drive. Both the parametric drive and the feedback are turned off and -- as in the first part -- the
+motion of the particle towards an equilibrated state is observerd.}
+        \label{fig:setup}
+    \end{center}
+\end{figure}
 

### 2016-10-10 include/classes.hpp

@@ -14,6 +14,7 @@ class Particle
      double *rnew;
      double *vnew;
      double *vhalf;
+     double *vOld;
      double* r0;
      double* v0;
      double m;

### 2016-10-10 include/globals.hpp

@@ -6,6 +6,7 @@
 #include <cmath>
 #include <gsl/gsl_rng.h>
 #include <gsl/gsl_randist.h>
+#include <gsl/gsl_histogram.h>
 #include <list>
 #include <iterator>
 #include <vector>
@@ -24,6 +25,10 @@
 extern gsl_rng *r;
 extern gsl_rng *r1;
 extern const gsl_rng_type *T;
+
+extern gsl_histogram *gas_in;
+extern gsl_histogram *gas_out;
+extern gsl_histogram *gas_real_in;
 
 extern int g_ID;
 extern const unsigned int N;

### 2016-10-10 src/classes.cpp

@@ -11,6 +11,7 @@ Particle::Particle()
     rnew = new double[3];
     vnew = new double[3];
     vhalf = new double[3];
+    vOld = new double[3];
     r0 = new double[3];
     v0 = new double[3];
     m = 1.;
@@ -28,6 +29,7 @@ Particle::Particle()
         rnew[i]=0;
         vnew[i]=0;
         vhalf[i]=0;
+        vOld[i] = 0;
     }
     ID = g_ID;
     type = 0;
@@ -41,10 +43,12 @@ Particle::Particle(double x,double y, double z, double v1, double v2, double v3)
     v = new double[3];      
     a = new double[3];
     aOld = new double[3];
+    vOld = new double[3];
     for(int i=0;i<3;i++)
     {
         a[i]=0;
         aOld[i]=0;
+        vOld[i]=0;
     }
     r[0]=x;
     r[1]=y;
@@ -73,10 +77,12 @@ Particle::Particle(double x,double y, double z, double v1, double v2, double v3,
     v0 = new double[3];      
     a = new double[3];
     aOld = new double[3];
+    vOld = new double[3];
     for(int i=0;i<3;i++)
     {
         a[i]=0;
         aOld[i]=0;
+        vOld[i] = 0;
     }
     r[0]=x;
     r[1]=y;
@@ -107,6 +113,7 @@ Particle::Particle(std::string Name)
     v = new double[3];      
     a = new double[3];
     aOld = new double[3];
+    vOld = new double[3];
     rnew = new double[3];
     vnew = new double[3];
     vhalf = new double[3];
@@ -120,6 +127,7 @@ Particle::Particle(std::string Name)
         rnew[i]=0;
         vnew[i]=0;
         vhalf[i]=0;
+        vOld[i] = 0;
     }
     ID = g_ID;
     type = 0;
@@ -143,4 +151,5 @@ Particle::~Particle()
     delete [] vhalf;
     delete [] r0;
     delete [] v0;
+    delete [] vOld;
 }

### 2016-10-10 src/globals.cpp

@@ -5,6 +5,10 @@ const gsl_rng_type *T = gsl_rng_taus;
 gsl_rng *r = gsl_rng_alloc(T);
 gsl_rng *r1 = gsl_rng_alloc(T);
 
+gsl_histogram *gas_in = gsl_histogram_alloc(1000);
+gsl_histogram *gas_out = gsl_histogram_alloc(1000);
+gsl_histogram *gas_real_in = gsl_histogram_alloc(1000);
+
 int g_ID=1;
 const unsigned int N = 864;
 

### 2016-10-10 src/main.cpp

@@ -85,6 +85,7 @@ void mainLoop() {
     }
     chdir(folderName.c_str()); 
     system("mkdir combined");
+    system("mkdir gas");
     std::ofstream params;
     params.open("parameters.txt");
     params << "rho = " << rho << std::endl;
@@ -115,6 +116,9 @@ void mainLoop() {
     int a;
     int run=0;
     gsl_rng_set(r,98);
+    gsl_histogram_set_ranges_uniform(gas_in,-1,5);
+    gsl_histogram_set_ranges_uniform(gas_out,-1,5);
+    gsl_histogram_set_ranges_uniform(gas_real_in,-1,5);
     /*
      *FILE* output = fopen("output/box_test.xyz","w");
      *FILE* pressure = fopen("output/pressure.dat","w");
@@ -135,6 +139,10 @@ void mainLoop() {
     FILE* COMtempout = fopen("temperature_com.dat","w");
     FILE* comData = fopen("comdata.dat","w");
     FILE* vCOMData = fopen("vCOMData.dat","w");
+    FILE* gasInData = fopen("histogram_in","w");
+    FILE* gasOutData = fopen("histogram_out","w");
+    FILE* gasRealInData = fopen("histogram_real_in","w");
+    FILE* gasTempData = fopen("gasTempData.dat","w");
     Particle *cube = new Particle[N];
     std::list<Particle*> gas;
     std::list<Particle*> gasHistory;
@@ -183,13 +191,15 @@ void mainLoop() {
         
     ComputeSoftSphere(gas,cube);
 
-    for(run = 0;run<40000;run++)
+    for(run = 0;run<80000;run++)
     {
         if(run%500==0)
+        {
             printf("(MEASURE) Zeitschritt %d - Number of Gas particles: %d\n",run,gas.size());
+        }
         eHEX(cube);
-        //BarostatNew(cube,gas);
-        BarostatNew(cube,gas,gasHistory);
+        BarostatNew(cube,gas);
+        //BarostatNew(cube,gas,gasHistory);
         //std::cout << "works!" << std::endl;
         harmonicTrap(rCM,vCM,rCMStart,cube);
         //std::cout << "works!" << std::endl;
@@ -205,12 +215,30 @@ void mainLoop() {
             calcTemp(cube,tempout); 
             calcCOMTemp(vCM,COMtempout);
[…]

### 2016-10-10 test/particleTest.cpp

@@ -24,6 +24,12 @@ int main(int argc, char** argv)
     p2 = new Particle();
     p3 = new Particle();
 
+    std::list<Particle*> particles;
+    std::list<Particle*> history;
+    InitBarostat(particles,history);
+
+
+
     delete p1;
     delete p2;
     delete p3;

### 2016-11-16 doc/draft1.tex
added tests and updated draft of thesis

@@
 \usepackage{paralist}
-\usepackage{dsfont}
-\usepackage{color}
+\usepackage{dsfont} \usepackage{color}
 \usepackage{bbold}
@@
 \end{figure}
+On the left hand side we have the friction coefficient $\Gamma_0$ and the angular frequency that describes the fluctuation along the chosen axis. On
+the right hand side, there are two forces. The first one is $F_\text{fluct}$, which describes a stochastic force caused by interactions with the gas
+in the vacuum chamber. This force is given by
+\begin{equation}
+    F_\text{fluct} = \sqrt{2m\Gamma_0k_BT_0} \ \xi\left(t\right)
+\end{equation}
+where $T_0$ is the temperature of the heat bath (i.e. the surrounding gas in the vacuum chamber), $k_B$ is the Boltzmann constant and $\xi(t)$ is
+white noise, which obeys the equations $\left\langle\xi(t)\right\rangle= 0$ and $\left\langle\xi(t)\xi(t')\right\rangle= \delta(t-t')$, which means
+that it is a random force. The term $\Gamma_0$ appears in the formulat due to the fluctuation-dissipation theorem, which links the damping rate to the
+stochastic force.\\
+The external force $F_\text{ext}$ is part of the experimental setup, where the frequency of the fluctuation along an axis, $\Omega_0$, is measured and
+used to suppress the motion along said axis. This causes a decrease in the particle fluctuations and thus acts as a cooling mechanism for the particle
+in the trap. This process creates a non-equilibrium steady state $\rho_{ss}(u,\alpha)$, which is not known analytically. This state is the starting
+point of this thesis.\\
 
@@
 The dynamics of the system are obtained by solving Newton's equations of motion for every particle.
+
+\subsection{The Glass Particle}
+The glass particle from the experiment will be modeled as a system of particles interacting via a Lennard-Jones pair potential, 
+\begin{equation}
+    \label{eq:lj}
+    U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right]
+\end{equation}
+where $\varepsilon$ is the depth of the potential well (and thus its unit is energy) and $\sigma$ is the distance at which the potential is zero. 
+The form of the potential and the relation to the parameters is depicted in Fig. \ref{fig:lj}.
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.2]{images/LJ-2.pdf}
+        \caption{The Lennard-Jones 12-6 potential from \eqref{eq:lj}. The x-axis is the particle distance divided by $\sigma$ and  the y-axis is the
+        potential divided by the depth of the potential well.}
+        \label{fig:lj}
+    \end{center}
+\end{figure}
+Since $\varepsilon$ and $\sigma$ are crucial parameters for the simulation and do (usually) not change over time, it is practical to use them to
+define the dimensions of the system. This means that the unit of distance is $\sigma$, the unit of energy is $\varepsilon$ and the unit of mass is 
+the mass of the simulated particle. The so called \textit{reduced units} can be constructed from these three parameters and put into relation to the
+original units. Here are some examples:
+\begin{itemize}
+    \item {distance:} $r^* = r/\sigma$
+    \item {potential energy:} $U^* = U/\varepsilon$
+    \item {temperature:} $T^* = k_B T/\varepsilon$
+    \item {time:} $t^* = t\sqrt{\varepsilon/(m\sigma^2)}$
+    \item {pressure:} $P^* = P\sigma^3/\varepsilon$
+    \item {density:} $\rho^* = \rho \sigma^3$
+\end{itemize}
+One very popular choice for the simulated atoms is Argon because it is an inert gas and the atoms behave 
+approximately like hard spheres which attract each other with weak van der Waals forces, which justifies the use of the Lennard-Jones potential. 
+Argon has a mass of m = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$m and $\varepsilon = 1.65 \times 10^{-21}$J.\\
+With the above introduced reduced units, the Lennard-Jones potential can be written as
+\begin{equation}
+    U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right].
+\end{equation}
+Since the reduced units will be used throughout the rest of this thesis, i will drop the asterisk henceforth.\\
+From the Lennard-Jones potential the corresponding force can be calculated by taking the derivative with respect to the direction of interest:
+\begin{eqnarray}
+    F_{x} &=& -\frac{\partial}{\partial x} U(r) \nonumber\\
+                &=& -\frac{\partial}{\partial x} 4\left[{r}^{-12} - {r}^{-6}\right] \nonumber\\
+                &=& -4 \left[(-12){r}^{-13} - (-6){r}^{-7}\right] \frac{\partial r}{\partial x} \nonumber\\
+                &=& 48 \left[r^{-13} - 0.5 \ r^{-7}\right] \frac{x}{r} \nonumber\\
+    \label{eq:ljforce} &=& 48 \left[r^{-14} - 0.5 \ r^{-8}\right] x
+\end{eqnarray}
+The force in the y and z direction can be calculated analogously.\\
 

### 2016-11-16 output/runs/161012_1708/means.py

@@ -1,0 +1,7 @@
+import numpy as np
+
+data_in = np.loadtxt("gasTempData.dat",usecols=(0,))
+data_out = np.loadtxt("gasTempData.dat",usecols=(1,))
+
+print("In: " + str(np.mean(data_in)) + " +- " + str(np.std(data_in)))
+print("Out: " + str(np.mean(data_out)) + " +- " + str(np.std(data_out)))

### 2016-11-16 src/functions.cpp

@@ -39,7 +39,7 @@ unsigned int NumberOfParticles()
         //double P=4;
         double k=1.0;
         //double T=0.9;
-        mu = dt*L*L*P*pow((1/(2*M_PI*k*AmbientTemp)),0.5);
+        mu = dt*L*L*P*pow((1./(2*M_PI*k*AmbientTemp)),0.5);
         //std::cout << mu << std::endl;
         return gsl_ran_poisson(r,mu);
 }
@@ -88,7 +88,7 @@ void ComputeAccelerations(Particle* particle)
             rSqd = rij[0]*rij[0]+rij[1]*rij[1]+rij[2]*rij[2];
             if(rSqd <= rCutOff*rCutOff)
             {       
-                //double f = 8*48 * ( pow(rSqd,-7.) - 0.5*pow(rSqd,-4.) );
+                //f = 8*48 * ( pow(rSqd,-7.) - 0.5*pow(rSqd,-4.) );
                 f = 48 * ( pow(rSqd,-7) - 0.5*pow(rSqd,-4) );
                 for(unsigned int m=0; m<3; m++)
                 {
@@ -447,6 +447,841 @@ void InitBarostat(std::list<Particle*>& particles)
          *        iter++;
          *}
          */
+}
+
+void InitBarostatFull(std::list<Particle*>& particles)
+{
+        unsigned int i;
+        unsigned int N;
+        //double sigma=1.1;
+        double sigma=AmbientTemp;
+
+        Particle *px0;
+        Particle *pxL;
+        Particle *py0;
+        Particle *pyL;
+        Particle *pz0;
+        Particle *pzL;
+
+        double kinE = 0.;
+
+        //side: x0
+        N = NumberOfParticles();
+        for(i=0;i<N;i++)
+        {
+            
+            px0 = new Particle(-L,
+                            ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
+                            ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
+                            gsl_ran_rayleigh(r,sigma),
+                            gsl_ran_gaussian(r,sigma),
+                            gsl_ran_gaussian(r,sigma),"GasIn");
+            particles.push_back(px0);
+            kinE = px0->m*(px0->v[0]*px0->v[0]+px0->v[1]*px0->v[1]+px0->v[2]*px0->v[2])*0.5;
+            gsl_histogram_increment(gas_in,kinE);
+        }
+
+        //side xL
+        N = NumberOfParticles();
[…]

### 2016-11-16 test/cell_list.cpp

@@ -1,0 +1,76 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <cstdlib>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../include/globals.hpp"
+#include "../include/classes.hpp"
+#include "../include/functions.hpp"
+
+
+struct cell{
+    double xStart;
+    double yStart;
+    double zStart;
+};
+int main(int argc, char** argv)
+{
+    int a;
+    int run=0;
+    gsl_rng_set(r,98);
+    gsl_histogram_set_ranges_uniform(gas_in,-1,5);
+    gsl_histogram_set_ranges_uniform(gas_out,-1,5);
+    gsl_histogram_set_ranges_uniform(gas_real_in,-1,5);
+
+    Particle *cube = new Particle[N];
+    std::list<Particle*> gas;
+    std::list<Particle*> gasHistory;
+    double* rCM = new double[3];
+    double* rCMtemp = new double[3];
+    double* vCM = new double[3];
+    double energy = 0;
+
+    InitPositions(cube);
+    calcCM(cube,rCMStart,vCM);
+    calcCM(cube,rCM,vCM);
+    InitVelocities(cube);
+    ComputeAccelerations(cube);
+    /*
+     *for(run=0;run<10000;run++)
+     *{
+     *        if(run%500==0)
+     *            printf("(INIT) Zeitschritt %d\n",run);
+     *        VelocityVerlet(cube,0,NULL);
+     *        if(run%100 == 0)
+     *            rescaleVelocities(cube);
+     *        if(run%100 == 0)
+     *            GenerateOutput(cube,gas,run);
+     *}
+     */
+
+    gsl_histogram *cells = gsl_histogram_alloc(100);
+    gsl_histogram_set_ranges_uniform(cells,0,50);
+    double lcx = floor(L/rCutOff);
+    double rcx = L/lcx;
+    double c = 0;
+    double ci[3] = {0,0,0};
+    for(int i=0;i<N;i++)
+    {
+        for(int k=0;k<3;k++)
+            ci[k] = floor(cube[i].r[k]/rcx);
+        c = ci[0]*lcx*lcx+ci[1]*lcx+ci[2];
+        std::cout << c << std::endl;
+        gsl_histogram_increment(cells,c);
+    } 
+
+    FILE* celldata = fopen("cell_list_test.dat","w");
[…]

### 2016-11-16 test/compile_cell_list.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+g++  cell_list.cpp ../src/functions.cpp ../src/globals.cpp ../src/classes.cpp -lgsl -lgslcblas  -o  cell_list.o

### 2016-11-16 test/nums.py

@@ -1,0 +1,12 @@
+import numpy as np
+
+data = np.loadtxt("test.dat",dtype=int)
+m = []
+for j in range(0,30):
+    m.append([])
+print(m)
+for i in data:
+    m[i].append(int(i))
+
+for j in range(0,30):
+    print(len(m[j]))

### 2016-11-16 test/temperature_test/compile_gasTemperature_test.sh

@@ -1,0 +1,4 @@
+#!/bin/bash
+
+g++  gasTemperature_test.cpp ../../src/globals.cpp ../../src/functions.cpp  ../../src/classes.cpp -lgsl -lgslcblas -o  gasTemperature_test.o
+

### 2016-11-16 test/temperature_test/gasTemperature_test.cpp

@@ -1,0 +1,104 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <cstdlib>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../../include/functions.hpp"
+#include "../../include/globals.hpp"
+#include "../../include/classes.hpp"
+
+int main(int argc, char** argv)
+{
+    setValues(0.2,0.07,0.2,4,0.1);
+
+    double* rCM = new double[3];
+    double* rCMtemp = new double[3];
+    double* rCMstart = new double[3];
+    double* vCM = new double[3];
+    double* pos = new double[3];
+    //double* energies = new double[3];
+    double energy = 0;
+    FILE* eData = fopen("energies.dat","w");
+    FILE* positions = fopen("gas_test.xyz","w");
+    //FILE* tempout = fopen("temperature.dat","w");
+    FILE* output = fopen("gas_temperature.dat","w");
+
+    for(int i=0;i<3;i++)
+        pos[i] = 0.;
+    int run;
+
+    Particle *cube = new Particle[N];
+    std::list<Particle*> gas;
+    InitPositions(cube);
+    calcCM(cube,rCMStart,vCM);
+    calcCM(cube,rCM,vCM);
+    InitVelocities(cube);
+    ComputeAccelerations(cube);
+
+    for(run=0;run<40000;run++)
+    {
+            if(run%500==0)
+                printf("(INIT) Zeitschritt %d\n",run);
+            fprintf(positions,"%d\nTest\n",N+gas.size());
+            VelocityVerlet(cube,1,positions);
+            if(run%100 == 0 && run < 20000)
+                rescaleVelocities(cube);
+            //energies = calculateEnergies(cube,gas); 
+            //energy = calculateEnergies(cube,gas); 
+            //fprintf(eData,"%lf\t%lf\t%lf\n",energies[0],energies[1],energies[2]);
+            //fprintf(eData,"%lf\n",energy);
+    }
+
+    /*
+     *ComputeSoftSphere(gas,cube);
+     *BarostatNew(cube,gas);
+     *ComputeSoftSphere(gas,cube);
+     *BarostatNew(cube,gas);
+     */
+    //ComputeSoftSphere(gas,cube);
+    //BarostatNew(cube,gas);
+    for(int k=0;k<20;k++)
+        InitBarostatFullNew(gas);
[…]

### 2016-11-16 test/temperature_test/temperature_test.cpp

@@ -1,0 +1,64 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <cstdlib>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../../include/functions.hpp"
+#include "../../include/globals.hpp"
+#include "../../include/classes.hpp"
+
+int main(int argc, char** argv)
+{
+    setValues(0.2,0.07,0.2,4,0.1);
+
+    double* rCM = new double[3];
+    double* rCMtemp = new double[3];
+    double* rCMstart = new double[3];
+    double* vCM = new double[3];
+    double* pos = new double[3];
+    FILE* positions = fopen("temperature_test.xyz","w");
+    FILE* tempout = fopen("temperature.dat","w");
+
+    for(int i=0;i<3;i++)
+        pos[i] = 0.;
+    int run;
+
+    Particle *cube = new Particle[N];
+    std::list<Particle*> gas;
+    InitPositions(cube);
+    calcCM(cube,rCMStart,vCM);
+    calcCM(cube,rCM,vCM);
+    InitVelocities(cube);
+    ComputeAccelerations(cube);
+
+    for(run=0;run<40000;run++)
+    {
+            if(run%500==0)
+                printf("(INIT) Zeitschritt %d\n",run);
+            fprintf(positions,"%d\nTrap\n",N);
+            VelocityVerlet(cube,1,positions);
+            if(run%100 == 0 && run < 20000)
+                rescaleVelocities(cube);
+            calcTemp(cube,tempout); 
+    }
+    for(run = 0;run<40000;run++)
+    {
+        if(run%500==0)
+        {
+            printf("(MEASURE) Zeitschritt %d - Number of Gas particles: %d\n",run,gas.size());
+        }
+        eHEX(cube);
+        BarostatNew(cube,gas);
+        harmonicTrap(rCM,vCM,rCMStart,cube);
+        if(run%400==0)
+            GenerateOutput(cube,gas,run+10000);
+        if(run%100==0)
+        {
+            calcTemp(cube,tempout); 
+        }
+    }
+	return 0;
+}

### 2016-11-18 doc/draft1.tex
More work on the draft, updated gitignore
xyz files are ignored, as well as the histograms
more tests added

--- 2016-10-10 doc/draft1.tex
@@
 %\usepackage[font=footnotesize,labelfont=bf]{caption}
-%\usepackage{subcaption}
+\usepackage{subcaption}
 \usepackage{fancyhdr}
@@
 The dynamics of the system are obtained by solving Newton's equations of motion for every particle.
+
+\subsection{The Glass Particle}
+The glass particle from the experiment will be modeled as a system of particles interacting via a Lennard-Jones pair potential, 
+\begin{equation}
+    \label{eq:lj}
+    U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right]
+\end{equation}
+where $\varepsilon$ is the depth of the potential well (and thus its unit is energy) and $\sigma$ is the distance at which the potential is zero. 
+The form of the potential and the relation to the parameters is depicted in Fig. \ref{fig:lj}.
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.2]{images/LJ-2.pdf}
+        \caption{The Lennard-Jones 12-6 potential from \eqref{eq:lj}. The x-axis is the particle distance divided by $\sigma$ and  the y-axis is the
+        potential divided by the depth of the potential well.}
+        \label{fig:lj}
+    \end{center}
+\end{figure}
+Since $\varepsilon$ and $\sigma$ are crucial parameters for the simulation and do (usually) not change over time, it is practical to use them to
+define the dimensions of the system. This means that the unit of distance is $\sigma$, the unit of energy is $\varepsilon$ and the unit of mass is 
+the mass of the simulated particle. The so called \textit{reduced units} can be constructed from these three parameters and put into relation to the
+original units. Here are some examples:
+\begin{itemize}
+    \item {distance:} $r^* = r/\sigma$
+    \item {potential energy:} $U^* = U/\varepsilon$
+    \item {temperature:} $T^* = k_B T/\varepsilon$
+    \item {time:} $t^* = t\sqrt{\varepsilon/(m\sigma^2)}$
+    \item {pressure:} $P^* = P\sigma^3/\varepsilon$
+    \item {density:} $\rho^* = \rho \sigma^3$
+\end{itemize}
+One very popular choice for the simulated atoms is Argon because it is an inert gas and the atoms behave 
+approximately like hard spheres which attract each other with weak van der Waals forces, which justifies the use of the Lennard-Jones potential. 
+Argon has a mass of m = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$m and $\varepsilon = 1.65 \times 10^{-21}$J.\\
+With the above introduced reduced units, the Lennard-Jones potential can be written as
+\begin{equation}
+    U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right].
+\end{equation}
+Since the reduced units will be used throughout the rest of this thesis, i will drop the asterisk henceforth.\\
+From the Lennard-Jones potential the corresponding force can be calculated by taking the derivative with respect to the direction of interest:
+\begin{eqnarray}
+    F_{x} &=& -\frac{\partial}{\partial x} U(r) \nonumber\\
+                &=& -\frac{\partial}{\partial x} 4\left[{r}^{-12} - {r}^{-6}\right] \nonumber\\
+                &=& -4 \left[(-12){r}^{-13} - (-6){r}^{-7}\right] \frac{\partial r}{\partial x} \nonumber\\
+                &=& 48 \left[r^{-13} - 0.5 \ r^{-7}\right] \frac{x}{r} \nonumber\\
+    \label{eq:ljforce} &=& 48 \left[r^{-14} - 0.5 \ r^{-8}\right] x
+\end{eqnarray}
+The force in the y and z direction can be calculated analogously.\\
+The initial configuration of the particles is a face centered cubic (FCC) lattice. A schematic of the FCC lattice is depicted in fig. \ref{fig:fcc}.
+With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since once FCC cell (as
+depicted) shares its atoms with its next neighbours, the number of atom per unit cell is 4, as depicted in fig. \ref{fig:wiki}. The whole system of
+atoms is then created by repeating this cell structure. One convenient way is to arrange the unit cells in a cubic system, so if there are $M$ FCC
+unit cells on one edge, the whole system consists of $M^3$ cells. Since there are 4 particles per cell, there are ideal or so called \textit{magic
+numbers} for atoms for which this setup works perfectly: $N = 4M^3 = 4,32,108,256,500,864,\ldots$.\\
+\begin{figure}[h]
+    \begin{center}
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.3]{images/fcc-2.pdf}
+            \caption{Schematic figure of a face centered cubic (FCC) lattice}
+            \label{fig:fcc}
+        \end{subfigure} 
+        \
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.15]{images/fcc_unit.png}
+            \caption{Schematic of the number of particles contained in one FCC unit cell. The atoms on the edges are shared by 8 cells and the ones 
+                on the faces by 2 cells\cite{unitcell}}
+            \label{fig:fcccut}
+        \end{subfigure} 
+        \
+        %\hspace{60pt}
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.3]{images/unit_cell.pdf}
+            \caption{Setup for successively building a FCC lattice}
+            \label{fig:unitcell}
+        \end{subfigure}
+    \end{center}
+\end{figure}
+%\begin{figure}
+    %\begin{center}
+        %\includegraphics[scale=0.3]{images/fcc_unit.png}
+        %\caption{Schematic of the number of particles contained in one FCC unit cell. The atoms on the edges are shared by 8 cells and the ones on the
+            %faces by 2 cells. Image taken from
+            %{https://upload.wikimedia.org/wikipedia/commons/2/27/Elementarzelle\_einer\_kubisch\_raumzentrierten\_Elementarzelle.png}}
+    %\end{center}
+%\end{figure}
+% do i need to credit the uni buffalo course? 
+% http://www.physics.buffalo.edu/phy411-506/
+There are several ways to achieve this initial configuration and the one used in this thesis \cite{buffalo} was to create a kind of unit cell 
+consisting of four atoms, as shown in fig. \ref{fig:unitcell}, which can be described by a set of points
+\begin{eqnarray*}
+    p_1 &=& \{0,0,0\}\\
+    p_2 &=& \{0.5,0.5,0\}\\
+    p_3 &=& \{0.5,0,0.5\}\\
+    p_4 &=& \{0,0.5,0.5\}
+\end{eqnarray*}
+From the particle number $N$ and the number of FCC unit cells per edge $M$ the lattice constant $a$ can be calculated
+\begin{equation}
+    a = \frac{L}{M}
+\end{equation}
+where $L$ is the side length of the cube that is the whole system and it is calculated via the density of the system
+\begin{equation}
+    L = \sqrt[3]{\frac{N}{\rho}}
+\end{equation}
+With the lattice constant and the 4 points of the FCC cell, all the particles can be put into place.
+
+\subsection{The Velocity-Verlet Algorithm}
+When we look at the system from a thermodynamical standpoint, we see that it follows some kind of path in the phase space as time progresses. Every
+point in this space corresponds to a set of positions and momenta and the connection between two points corresponds to the evolution of the system
+from one state to another. As mentioned above, this evolution (the dynamics of the system) is a crucial element to Molecular Dynamics. Since the
+equations of motion cannot be solved analytically in general, we need to approximate the solution.\\
+The method used here is called finite difference approach. The trajectory of the system in the phase space is cut into finite pieces of length $\Delta
+t$ and the equations of motion are solved for every segment separately (see fig. \ref{fig:finitedifference}).
+\begin{figure}
+    \begin{center}
+        \includegraphics[scale=0.5]{images/finite_approach.pdf}
+        \caption{Simplified graphical schematic of the finite difference approach. The evolution of the system from a point $(r^N(t),p^N(t))$ in the
+            phase space is approximated by slicing it up into pieces of length $\Delta t$. On every stop after the starting point ($t+\Delta t$, 
+            $t+2\Delta t$, $t+3\Delta t$,$\ldots$) the equations of motion can be solved numerically.}
+        \label{fig:finitedifference} 
+    \end{center}
+\end{figure}
 

### 2016-11-18 include/functions.hpp

@@ -47,6 +47,7 @@ void PrintAllData(Particle* cube, std::list<Particle*> gas,FILE* output);
 void eHEX(Particle* cube);
 void calcPressure(Particle cube,FILE* output);
 void calcTemp(Particle* cube,FILE* output);
+void calcTemp(Particle* cube,FILE* output,int run);
 void calcCOMTemp(double* vCOM,FILE* output);
 
 void calcCM(Particle* particles,double *rCM, double* vCM);

### 2016-11-18 src/functions.cpp

@@ -2334,6 +2334,16 @@ void calcTemp(Particle* cube,FILE* output)
     fprintf(output,"%lf\n",T);
 }
 
+void calcTemp(Particle* cube,FILE* output,int run)
+{
+    double T=0;
+    for(unsigned int i=0;i<N;i++)
+        for(unsigned int k=0;k<3;k++)
+            T+=cube[i].v[k]*cube[i].v[k];
+    T = T/(3*N);
+    fprintf(output,"%d\t%lf\n",run,T);
+}
+
 void calcCOMTemp(double* vCOM,FILE* output)
 {
     double T=0;

### 2016-11-18 src/main.cpp

@@ -51,7 +51,7 @@ int main(int argc,char** argv)
      *    q += 0.01;
      *}
      */
-    setValues(0.2,0.06,0.2,4,0.09);
+    setValues(0.2,0.06,0.2,4,0.08);
     mainLoop();
     return 0;
 }
@@ -139,9 +139,9 @@ void mainLoop() {
     FILE* COMtempout = fopen("temperature_com.dat","w");
     FILE* comData = fopen("comdata.dat","w");
     FILE* vCOMData = fopen("vCOMData.dat","w");
-    FILE* gasInData = fopen("histogram_in","w");
-    FILE* gasOutData = fopen("histogram_out","w");
-    FILE* gasRealInData = fopen("histogram_real_in","w");
+    FILE* gasInData = fopen("histogram_in.dat","w");
+    FILE* gasOutData = fopen("histogram_out.dat","w");
+    FILE* gasRealInData = fopen("histogram_real_in.dat","w");
     FILE* gasTempData = fopen("gasTempData.dat","w");
     Particle *cube = new Particle[N];
     std::list<Particle*> gas;

### 2016-11-18 test/compile_outputs.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+
+g++  outputs.cpp ../src/functions.cpp ../src/globals.cpp ../src/classes.cpp -lgsl -lgslcblas -o  outputs.o

### 2016-12-07 doc/draft1.tex
more writing, more tests, different velocities

@@
 \newcommand{\dznach}[2]{\frac{\partial^2{#1}}{\partial{#2}^2}}
-\newcommand{\vnabla}{\vec{\nabla}}
-\newcommand{\evec}[1]{\mathbf{\hat{#1}}}
+\newcommand{\vnabla}{\mathbf{\nabla}}
+\newcommand{\emathbf}[1]{\mathbf{\hat{#1}}}
 \newcommand{\bra}[1]{\langle{#1}|}
@@
 \newcommand{\fsqrt}[2]{\sqrt{\frac{#1}{#2}}}
+\newcommand{\GammaK}{{\Gamma_k}}
 
@@
 The method used here is called finite difference approach. The trajectory of the system in the phase space is cut into finite pieces of length $\Delta
-t$ and the equations of motion are solved for every segment separately (see fig. \ref{fig:finitedifference}).
+t$ and the equations of motion are solved for every segment separately (see fig. \ref{fig:finitedifference}).\\
 \begin{figure}
@@
 \end{figure}
+There are several ways to solve this kind of problem, but since we are interested in implementing it into a computer program the ideal solution
+should have some basic properties \cite{Allen1989}:
+\begin{itemize}
+    \item It should be fast and require little memory
+    \item It should permit the use of a large time step $\Delta t$
+    \item It should duplicate the classical trajectory as closely as possible
+    \item It should satisfy known conservation laws
+    \item It should be simple and easy to program
+\end{itemize}
+One algorithm that has all of the above mentioned features is the one proposed by Verlet \cite{Verlet1967}. In his paper, Verlet starts by Taylor
+expanding the coordinate mathbftor $\mathbf{r}_i$ for one particle after one time step $\Delta t$:
+\begin{equation}
+    \label{eq:verlet}
+    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + \dot{\mathbf{r}}_i \Delta t + \frac12 \ddot{\mathbf{r}}_i \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i \Delta t^3 +
+    \mathcal{O}(\Delta t^4)
+\end{equation}
+Since the first derivative of the coordinate mathbftor is the velocity, $\dot{\mathbf{r}}_i(t) = \mathbf{v}_i(t)$, and the second derivative of the coordinate 
+mathbftor is the acceleration, $\ddot{\mathbf{r}}_i(t) = \mathbf{a}_i(t)$, using Newton's law $\mathbf{F}_i(t) = m_i \mathbf{a}_i(t)$ equation \eqref{eq:verlet} 
+can be written as:
+\begin{equation}
+    \label{eq:verletnew}
+    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i(t)
+    \Delta t^3+\mathcal{O}(\Delta t^4)
+\end{equation}
+The same calculation can be carried out for one time step before $t$:
+\begin{equation}
+    \label{eq:verletnewminus}
+    \mathbf{r}_i(t-\Delta t) = \mathbf{r}_i(t) - {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 - \frac1{3!} \dddot{\mathbf{r}}_i(t) \Delta t^3+
+    \mathcal{O}(\Delta t^4)
+\end{equation}
+The sum of those two equations yields
+\begin{equation}
+    \mathbf{r}_i(t+\Delta t) + \mathbf{r}_i(t-\Delta t) = 2\mathbf{r}_i(t) + \frac1{2} {\mathbf{F}}_i(t) \Delta t^2 + \mathcal{O}(\Delta t^4)
+\end{equation}
+and from this we get the final form for the new coordinates:
+\begin{equation}
+    \label{eq:verletfinal}
+    \mathbf{r}_i(t+\Delta t) = 2\mathbf{r}_i(t)+ \mathbf{r}_i(t-\Delta t)+ \frac1{2} {\mathbf{F}}_i(t) \Delta t^2
+\end{equation}
+This means that the new coordinates can be calculated using the current coordinates, the current forces and the coordinates from the past time step.\\
+From equation \eqref{eq:verletfinal} one thing becomes clear: the velocities are not necessary to calculate the new positions and they are not
+computed in the process, but for the calculation of i.e. the kinetic energy the velocities are needed. Thus, the velocities have to be calculated with
+a combination of the Taylor expansions of the coordinate mathbftors at $t+\Delta t$ and $t-\Delta t$:
+\begin{equation}
+    \mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t) = 2\mathbf{v}_i(t)\Delta t + \mathcal{O}(\Delta t^3)
+\end{equation}
+which can be rewritten as
+\begin{equation}
+    \label{eq:velocities}
+    \mathbf{v}_i(t) = \frac{\mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t)}{2\Delta t} + \mathcal{O}(\Delta t^2)
+\end{equation}
+This approach has the advantage that it is fast, requires little memory and is reliable in the sense that there is no energy drift occuring during the
+simulation, which means that the energy is conserved. When we compare this to the list of desired properties this seems like a good algorithm.\\
+This algorithm however has two significant disatvantages. The first one is the calculation of the velocities. As can be seen in equation
+\eqref{eq:velocities}, the accuracy of the calculation is only $\mathcal{O}(\Delta t^2)$ while the positions can be calculated with an error of order
+$\Delta t^4$. The other big disatvantage is the first step of the algorithm. Since the calculation of the new positions requires the current positions
+and the ones from one time step before, which technically do not exist.\\
+The solutions to this problem is to include the stepwise calculation of the velocities \cite{swope1982}. For this we start again with the Taylor
+expansion of the coordinates mathbftor \eqref{eq:verletnew} and instead of Taylor expanding $\mathbf{r}_i(t-\Delta t)$, we write it as 
+\begin{equation}
+    \mathbf{r}_i(t) = \mathbf{r}_i(t+\Delta t) - \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
+\end{equation}
+Using \eqref{eq:verletnew} in the abvove equation we get
+\begin{equation}
+    \begin{aligned}
+        \mathbf{r}_i(t) &= \mathbf{r}_i(t) + \mathbf{v}_i(t) \Delta t + \frac1{2m}\mathbf{F}_i(t)\Delta t^2 \\
+                     &- \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
+    \end{aligned}
+\end{equation}
+and thus
+\begin{equation}
+    \label{eq:vvelocities}
+    \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \left(\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\right) \Delta t
+\end{equation}
+With the addition of this equation this algorithm is called the Velocity-Verlet algorithm, which is summarized in the following two equations:
+\begin{equation}
+    \label{eq:velocityverlet}
+    \begin{aligned}
+        \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2\\
+        \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \left(\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\right) \Delta t
+    \end{aligned}
+\end{equation}
+This algorithm is self starting, uses a small amount of memory, conserves the energy and gives a very good approximation to the originial trajectory
+in the phase space. To program this algorithm the following steps are needed: 
+\begin{enumerate}
+    \item Calculate all the forces between the particles (for the first run only)
+    \item Calculate the new positions with current velocity and forces
+    \item Calculate first half of the new velocities with current forces
+    \item Calculate all the forces for the new position
+    \item Use new forces to calculate second half of new velocities
+\end{enumerate}
+The first point only has to be carried out for the first run, because the forces have not been calculated at this point. As the forces are calculated
+for the new positions in step 4, they can be used for the next time step. For the first time step the velocities have to be chosen randomly and are 
+calculated with this algorithm from that point forward.
+
+\subsection{The Laser Beam - Energy influx}
+The glass sphere is trapped in the laser beam. While the motion of the center of mass is suppressed, which acts as a cooling mechanism, the
+individual atoms that make up the glass sphere absorb the energy from the laser which increases their velocity.\\
+To simulate this kind behaviour, thermostat algorithms \cite{Huenenberger2005} such as the Nos\'e-Hoover \cite{Nose1984,Hoover1985} or the Berendsen
+\cite{Berendsen1984} algorithms.\\
+The problem with these kind of algorithms is, that a target temperature has to be fixed which will be reached at some point. In order to simulate the
+influx of energy from the laser it would be better to have an algorithm that supplies the system with a certain amount of energy continuously.
+Fortunately, such an algorithm exists.\\
+The algorithm is called Heat Exchange Algorithm (HEX) \cite{Hafskjold1993}. Its intended purpose is the use in non-equilibrium molecular dynamics
+(NEMD) to study transport phenomena and determine transport coefficients. The algorithm works by introducing two regions in the system, one serving as
+a heat source and the other as a heat sink. A specific amount of heat is then exchanged between those two reservoirs. As it turns out however, this
+algorithm introduces an energy shift for longer simulation times. This led Wirnsberger et al \cite{Wirnsberger2015} to revisit the algorithm and 
+identify the cause of this energy drift, to create a more suitable algorithm, which they called eHEX.\\
+As in the HEX algorithm, regions are introduced to the system which act either as heat sinks or heat sources. These are labelled with $\Gamma_k$,
+where $k > 0$, and have corresponding amount of exchanged heat $\Delta Q_{\Gamma_k}$. If $\Delta Q_{\Gamma_k}$ is negative, heat is subtracted from
+the system and vice versa. Regions that neither act as heat source or heat sink are labelled with $\Gamma_0$, which are also called
+Hamiltonian regions (see fig. \ref{fig:ehex}). 
+\begin{figure}
+    \begin{center}
+        \includegraphics[scale=0.3]{images/ehex_system.png}
+        \caption{Setup for the use of the HEX and eHEX algorithms. The simulationbox $\Omega$, which is moving with a velocity of $v_\Omega$,
+        contains a heat source (red), $\Gamma_1$, which is moving with velocity $v_{\Gamma_1}$ and a heat sink (blue), $\Gamma_2$, which is moving 
+        with velocity $v_{\Gamma_2}$. The regions that are neither heat sources or heat sinks are denoted by $\Gamma_0$.}
+        \label{fig:ehex} 
+    \end{center}
+\end{figure}
+The simulation box, denoted by $\Omega$, and the regions $\Gamma_k$ are assumed to be moving with velocities $v_\Omega$ and $v_{\Gamma_k}$
+respectively.\\
+The change of the energy in a region $\Gamma_k$ is achieved by rescaling the velocities by a factor $\xi_k$ and shifted by the velocity of the
+corresponding region:
+\begin{equation}
+    \mathbf{v}_i \rightarrow \mathbf{\bar{v}}_i = \xi_k \mathbf{v}_i + (1-\xi_k)\mathbf{v}_{\Gamma_k}
+\end{equation}
+The bar over a quantity denotes the value after the exchange of heat.\\
+The factor $\xi_k$ is given by
+\begin{equation}
+    \xi_k = \sqrt{1+\frac{\Delta Q_{\Gamma_k}}{\mathcal{K}_{\Gamma_k}}}
+\end{equation}
+where $\Delta Q_{\Gamma_k}$ is the exchanged heat in the region $\Gamma_k$ and $\mathcal{K}_{\Gamma_k}$ is the non-translational kinetic energy of the
+region $\Gamma_k$ and is given by
+\begin{equation}
+    \mathcal{K}_{\Gamma_k} = \sum_{i \in \gamma_k} \frac{m_i v_i^2}{2} - \frac{m_{\Gamma_k}v_{\Gamma_k}^2}{2}
+\end{equation}
+The sum is taken over all indices in $\gamma_k$ which is the set of indices of particles in the region $\Gamma_k$.\\
+For the final version of the eHEX algorithm there are three more quantities needed. The first one is the heat flux per time step, denoted by 
+$\mathcal{F}_\GammaK$:
+\begin{equation}
+    \mathcal{F}_\GammaK = \frac{\Delta Q_\GammaK}{\Delta t}
+\end{equation}
+The second one is the thermostatting force $\mathbf{\eta}_i$, which is defined as
+\begin{equation}
+    \mathbf{\eta}_i = 
+    \begin{cases}
+        m_i \ \dfrac{{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}}{{2\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}}} \
+        \left(\mathbf{v}_i - \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)
+        & \quad \text{if } k(\mathbf{r}_i) > 0 \\
+        0 & \quad \text{otherwise}
+    \end{cases}
+    %\begin{cases}
+        %m_i \ \frac{\displaystyle{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}}{\displaystyle{2\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}}} \
+        %\left(\mathbf{v}_i - \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)
+        %& \quad \text{if } k(\mathbf{r}_i) > 0 \\
+        %0 & \quad \text{otherwise}
+    %\end{cases}
+\end{equation}
+where $k(\mathbf{r}_i)$ is the index of the region in which particle $i$ located, i.e. $k(\mathbf{r}_i) = 0$ means that the particle is in a Hamiltonian
+region and $k(\mathbf{r}_i) > 0$ denotes the heat sinks and sources.\\
+The last quantity is the one that corrects the long term energy drift of the HEX algorithm, denoted by $\mathcal{E}{r}_{i,\alpha}$.
+The analysis and derivation of this term is given in \cite{Wirnsberger2015}.
+\begin{equation}
+    \label{eq:bigeps}
+    \begin{aligned}
+        \mathcal{E}{r}_{i,\alpha} &= \dfrac{\eta_{i,\alpha}}{m_i\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}} \left[
+        \frac{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}{48} + \frac{1}{6} \sum_{j\in\gamma_{k(\mathbf{r}_i)}} \mathbf{f}_j \cdot \left(\mathbf{v}_j - 
+            \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)\right] \\
+            &- \frac{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}{12\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}} \left[
+            \frac{f_{i,\alpha}}{m_i} - \frac{1}{m_{\Gamma_{k(\mathbf{r}_i)}}} \sum_{j\in\gamma_{k(\mathbf{r}_i)}}f_{j,\alpha}\right]
+    \end{aligned}
+\end{equation}
+The $\mathbf{f}$ in the above equation denotes the force corresponding to the chosen intermolecular potential $U(\mathbf{r})$, 
+\begin{equation}
+    \mathbf{f} = -\nabla_{\mathbf{r}_i} U(\mathbf{r}_i)
+\end{equation}
+With all the necessary quantities established, the updating sequence of the eHEX algorithm can be written down:
+\begin{subequations}
+\begin{eqnarray}
+    \bar{\mathbf{v}}^n_i &=& \xi^n_{k(\mathbf{r}_i)} \mathbf{v}^n_i + \left(1-\xi^n_{k(\mathbf{r}_i)}\right) \mathbf{v}^n_{\Gamma_{k(\mathbf{r}_i)}} \\
+    \bar{\mathbf{v}}^{n+\frac12}_i &=& \bar{\mathbf{v}}^n_i + \frac{\Delta t}{2m_i} \mathbf{f}^n_i\\
+    \bar{\mathbf{r}}^{n+1}_i &=& \mathbf{r}^n_i + \Delta t \bar{\mathbf{v}}^{n+\frac12}_i\\
+    \mathbf{f}^{n+1}_i &=& \left.-\nabla_{\mathbf{r}_i} U(\mathbf{r})\right|_{\mathbf{r} = \bar{\mathbf{r}}^{n+1}}\\
+    \bar{\mathbf{v}}^{n+1}_i &=& \bar{\mathbf{v}}^{n+\frac12}_i + \frac{\Delta t}{2m_i} \mathbf{f}^{n+1}_i\\
+    \mathbf{v}^{n+1}_i &=& \bar{\xi}^{n+1}_{k(\bar{\mathbf{r}}_i)} \bar{\mathbf{v}}^{n+1}_i + \left(1-\bar{\xi}^{n+1}_{k(\bar{\mathbf{r}}_i)}\right)
+    \bar{\mathbf{v}}^{n+1}_{\Gamma_{k(\bar{\mathbf{r}}_i)}} \\
+    \mathbf{r}^{n+1}_i &=& \bar{\mathbf{r}}^{n+1} - \Delta t^3 \mathcal{E}\bar{\mathbf{r}}^{n+1}_i
+\end{eqnarray}
+\end{subequations}
+%This is done by adding for instance two regions to the simulation, one providing an influx of heat
+%and the other acting as a heat sink. The algorithm is furthermore designed to incorporate the velocities of the center of mass of the system and the
+%velocities of the regions.
 
@@
 \bibliography{references}
-\bibliographystyle{plain}
+\bibliographystyle{unsrt}
 

### 2016-12-07 include/classes.hpp

@@ -24,9 +24,9 @@ class Particle
      std::string name;
      bool surface;
 
-     std::vector<std::pair<std::string,double> >  kineticEnergy;
-     std::vector<std::pair<std::string,double> >  potentialEnergy;
-     std::vector<std::pair<std::string,double> >  totalEnergy;
+     //std::vector<std::pair<std::string,double> >  kineticEnergy;
+     //std::vector<std::pair<std::string,double> >  potentialEnergy;
+     //std::vector<std::pair<std::string,double> >  totalEnergy;
 
      Particle();
      Particle(std::string name);

### 2016-12-07 include/functions.hpp

@@ -45,6 +45,7 @@ void ComputeSoftSphere(std::list<Particle*>& gas, Particle* cube);
 void PrintAllData(Particle* cube, std::list<Particle*> gas,FILE* output);
 //void eHEX(Particle* cube,FILE* output);
 void eHEX(Particle* cube);
+void eHEXNew(Particle* cube);
 void calcPressure(Particle cube,FILE* output);
 void calcTemp(Particle* cube,FILE* output);
 void calcTemp(Particle* cube,FILE* output,int run);

### 2016-12-07 output/runs/161121_1639/temps.py

@@ -1,0 +1,20 @@
+import numpy as np
+import math
+
+temps = np.genfromtxt("gasTempData.dat",usecols=(0,1),skiprows=22)
+# for i in range(0,len(temps)):
+    # if math.isnan(temps[i][1]):
+        # print(i)
+        # # del temps[i][1]
+# print(temps[:,0])
+# print(temps[:,1])
+print("Mean inTemp: " + str(np.mean(temps[:,0])))
+print("Mean outTemp: " + str(np.mean(temps[:,1])))
+print("Mean difference " + str(np.mean(temps[:,1])-np.mean(temps[:,0])))
+
+for i in range(0,len(temps)):
+    print(temps[i][1] - temps[i][0])
+
+
+
+

### 2016-12-07 src/functions.cpp

@@ -474,8 +474,8 @@ void InitBarostatFull(std::list<Particle*>& particles)
                             ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
                             ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
                             gsl_ran_rayleigh(r,sigma),
-                            gsl_ran_gaussian(r,sigma),
-                            gsl_ran_gaussian(r,sigma),"GasIn");
+                            ((gsl_rng_uniform(r)-0.5)*2)*AmbientTemp,
+                            ((gsl_rng_uniform(r)-0.5)*2)*AmbientTemp,"GasIn");
             particles.push_back(px0);
             kinE = px0->m*(px0->v[0]*px0->v[0]+px0->v[1]*px0->v[1]+px0->v[2]*px0->v[2])*0.5;
             gsl_histogram_increment(gas_in,kinE);
@@ -489,8 +489,8 @@ void InitBarostatFull(std::list<Particle*>& particles)
                             ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
                             ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
                             gsl_ran_rayleigh(r,sigma)*(-1.),
-                            gsl_ran_gaussian(r,sigma),
-                            gsl_ran_gaussian(r,sigma),"GasIn");
+                            ((gsl_rng_uniform(r)-0.5)*2)*AmbientTemp,
+                            ((gsl_rng_uniform(r)-0.5)*2)*AmbientTemp,"GasIn");
             particles.push_back(pxL);
             kinE = pxL->m*(pxL->v[0]*pxL->v[0]+pxL->v[1]*pxL->v[1]+pxL->v[2]*pxL->v[2])*0.5;
             gsl_histogram_increment(gas_in,kinE);
@@ -503,9 +503,9 @@ void InitBarostatFull(std::list<Particle*>& particles)
             py0 = new Particle(((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
                             -L,
                             ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
-                            gsl_ran_gaussian(r,sigma),
+                            ((gsl_rng_uniform(r)-0.5)*2)*AmbientTemp,
                             gsl_ran_rayleigh(r,sigma),
-                            gsl_ran_gaussian(r,sigma),"GasIn");
[…]

### 2016-12-07 src/globals.cpp

@@ -5,9 +5,9 @@ const gsl_rng_type *T = gsl_rng_taus;
 gsl_rng *r = gsl_rng_alloc(T);
 gsl_rng *r1 = gsl_rng_alloc(T);
 
-gsl_histogram *gas_in = gsl_histogram_alloc(1000);
-gsl_histogram *gas_out = gsl_histogram_alloc(1000);
-gsl_histogram *gas_real_in = gsl_histogram_alloc(1000);
+gsl_histogram *gas_in = gsl_histogram_alloc(10000);
+gsl_histogram *gas_out = gsl_histogram_alloc(10000);
+gsl_histogram *gas_real_in = gsl_histogram_alloc(10000);
 
 int g_ID=1;
 //const unsigned int N = 864;

### 2016-12-07 src/main.cpp

@@ -51,7 +51,7 @@ int main(int argc,char** argv)
      *    q += 0.01;
      *}
      */
-    setValues(0.2,0.06,0.2,4,0.08);
+    setValues(0.1,0.05,0.2,1.1,0.1);
     mainLoop();
     return 0;
 }
@@ -116,9 +116,9 @@ void mainLoop() {
     int a;
     int run=0;
     gsl_rng_set(r,98);
-    gsl_histogram_set_ranges_uniform(gas_in,-1,5);
-    gsl_histogram_set_ranges_uniform(gas_out,-1,5);
-    gsl_histogram_set_ranges_uniform(gas_real_in,-1,5);
+    gsl_histogram_set_ranges_uniform(gas_in,-0.1,3);
+    gsl_histogram_set_ranges_uniform(gas_out,-0.1,3);
+    gsl_histogram_set_ranges_uniform(gas_real_in,-0.1,3);
     /*
      *FILE* output = fopen("output/box_test.xyz","w");
      *FILE* pressure = fopen("output/pressure.dat","w");

### 2016-12-07 test/ehexnew_test.cpp

@@ -1,0 +1,62 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <cstdlib>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../include/functions.hpp"
+#include "../include/globals.hpp"
+#include "../include/classes.hpp"
+
+
+int main(int argc, char** argv)
+{
+    setValues(0.2,0.01,0.2,4,0.1);
+
+    double* rCM = new double[3];
+    double* rCMtemp = new double[3];
+    double* rCMstart = new double[3];
+    double* vCM = new double[3];
+    double* pos = new double[3];
+    FILE* positions = fopen("ehexnew_test.xyz","w");
+    FILE* tempout = fopen("ehextemp.dat","w");
+
+    for(int i=0;i<3;i++)
+        pos[i] = 0.;
+    int run;
+
+    Particle *cube = new Particle[N];
+    std::list<Particle*> gas;
+    InitPositions(cube);
+    calcCM(cube,rCMStart,vCM);
+    calcCM(cube,rCM,vCM);
+    InitVelocities(cube);
+    ComputeAccelerations(cube);
+
+    for(run=0;run<40000;run++)
+    {
+            if(run%500==0)
+                printf("(INIT) Zeitschritt %d\n",run);
+            VelocityVerlet(cube,1,positions);
+            if(run%100 == 0 && run < 20000)
+                rescaleVelocities(cube);
+            calcTemp(cube,tempout); 
+    }
+    for(run = 0;run<20000;run++)
+    {
+        if(run%500==0)
+        {
+            printf("(MEASURE) Zeitschritt %d - Number of Gas particles: %d\n",run,gas.size());
+        }
+        eHEXNew(cube);
+        //BarostatNew(cube,gas);
+        //harmonicTrap(rCM,vCM,rCMStart,cube);
+        //if(run%400==0)
+            //GenerateOutput(cube,gas,run+10000);
+        //if(run%100==0)
+            calcTemp(cube,tempout); 
+    }
+	return 0;
+}

### 2016-12-07 test/outputs.cpp

@@ -32,7 +32,7 @@ int main(int argc, char** argv)
                 printf("(INIT) Zeitschritt %d\n",run);
             VelocityVerlet(cube,0,NULL);
             //if(run%100 == 0 && run < 50000)
-            if(run%100)
+            if(run%100 == 0)
             {
                 if(run < 50000)
                     rescaleVelocities(cube);
@@ -40,7 +40,7 @@ int main(int argc, char** argv)
                 energy = calculateEnergies(cube,gas); 
                 fprintf(eData,"%d\t%lf\n",run,energy);
             }
-            if(run%1000)
+            if(run%1000 == 0)
                 GenerateOutput(cube,gas,run);
     }
     FILE* finalpos = fopen("final_position.xyz","w");

### 2016-12-29 doc/draft1.tex
more runs

@@
 \begin{equation}
+    \label{eq:kineticK}
     \mathcal{K}_{\Gamma_k} = \sum_{i \in \gamma_k} \frac{m_i v_i^2}{2} - \frac{m_{\Gamma_k}v_{\Gamma_k}^2}{2}
@@
 \end{equation}
-The second one is the thermostatting force $\mathbf{\eta}_i$, which is defined as
+The second one is the thermostatting force $\boldsymbol{\eta}_i$, which is defined as
 \begin{equation}
-    \mathbf{\eta}_i = 
+    \boldsymbol{\eta}_i = 
     \begin{cases}
@@
 \end{subequations}
+This algorithm can be applied to the problem of the levitating nano sphere in the laser beam by adjusting the setup and some of the parameters.\\
+Firstly, the setup of the heat sinks and sources has to be changed. Since there is only energy pumped into the system from the outside, the region
+acting as a heat sink vanishes and the region acting as a heat source spans over the whole simulation box. This means that, using the 
+notation of fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither the simulation box nor the heat source are moving, i.e. 
+$\mathbf{v}_\Omega = \mathbf{v}_{\Gamma_1} = 0$. This affects the non-translational kinetic energy term $\mathcal{K}_{\Gamma_1}$, the thermostatting
+force $\boldsymbol{\eta}$ and the correction term $\mathcal{E}\mathbf{r}$. Since there is only one region acting as a heat sink, the terms $\Delta Q$,
+$\mathcal{K}$ and $\mathcal{F}$ don't need an index. The summation index in \eqref{eq:kineticK} and \eqref{eq:bigeps} can be changed to the number of
+particles in the system, $N$, since all the particles are within the heat exchanging region. The masses are set to 1, i.e. $m_i = 1$ with the chosen
+reduced units and with this the total mass of the region (in the term $1/m_{\Gamma_{k(\mathbf{r}_i)}}$ in \eqref{eq:bigeps}) is equal to the number of
+particles in the system. With these changes, the quantities can be written down as:
+\begin{eqnarray}
+    \mathcal{K} &=& \sum_{N} \frac{v_i^2}{2}\\
+    \xi &=& \sqrt{1+\frac{\Delta Q}{\mathcal{K}}}\\
+    \mathcal{F} &=& \frac{\Delta Q}{\Delta t}\\
+    \boldsymbol{\eta}_i &=& 
+    \dfrac{\mathcal{F}}{{2\mathcal{K}}} \
+        \mathbf{v}_i\\
+        \mathcal{E}{r}_{i,\alpha} &=& \dfrac{\eta_{i,\alpha}}{\mathcal{K}} \left[
+        \frac{\mathcal{F}}{48} + \frac{1}{6} \sum_{N} \mathbf{f}_j \cdot \mathbf{v}_j
+            \right] \nonumber \\
+            &-& \frac{\mathcal{F}}{12\mathcal{K}} \left[
+            {f_{i,\alpha}} - \frac{1}{N} \sum_{N}f_{j,\alpha}\right]
+\end{eqnarray}
+This means that the laser is modeled to be pumping energy into the system, which increases the velocities of the velocities of the particles in the
+system over time, while the center of mass motion is not affected by this.\\
+
+\subsection{Surrounding Gas - Barostat}
+Without any equilibrating mechanism the setup described by now would lead to the system melting, which is not desirable. In the next step we will 
+introduce the surrounding gas of the gas chamber that will absorb some of the energy in the system, leading to the final state.\\
+Generally, pressure is introduced to the system by surrounding the object of interest (in this case the glass nano sphere) with a pressure medium.
+There are two main requirements for the choice of such a pressure medium: the exerted pressure must be hydrostatic and the computation of the
+interaction between the pressure medium and the object of interest must not take up a lot of resources.\\
+The model used in this thesis was developed by Gr\"unwald and Dellago \cite{Gruenwald2006} and uses an ideal gas of non-interacting particles as
+pressure medium. The particles of this pressure medium flow into the simulation from an outside box, whose geometry is based on the form of the object
+of interest. This barostat also acts as a thermostat and thus is suitable for the application in this thesis.\\
+The gas particles interact with the object via a soft-sphere potential of the form
+\begin{equation}
+    \label{eq:softsphere}
+    U(r) = \frac{1}{r^{12}}
+\end{equation}
+in reduced units. \\
+BLABLABLA\\
+The algorithm can be performed by following these steps:
+\begin{enumerate}
+    \item Randomly draw the number of particles that are created on a single side of the minimal volume of cells, $N_\text{fac}$, from the
+        distribution 
+        \begin{equation}
+            \langle N_\text{fac}\rangle = \Delta t L^2 P \left(\frac{1}{2\pi m k_B T}\right)^\frac12
+        \end{equation}
+        where $\Delta T$ is the time step of the simulation, $L$ is the side length of the cell in which the particle is created, $P$ is the desired
+        pressure, $m$ is the mass of the gas particle, $k_B$ is the Boltzmann constant (which will be set to 1 in reduced units) and $T$ is the
+        desired temperature. This chosen number of particles is then equally distributed over the face of the cell on which they are created.
+    \item wos
+\end{enumerate}
+
 %This is done by adding for instance two regions to the simulation, one providing an influx of heat

### 2016-12-29 src/functions.cpp

@@ -38,8 +38,9 @@ unsigned int NumberOfParticles()
         //This is the calculated atmospheric pressure with LJ units
         //double P=4;
         double k=1.0;
+        double m = 0.1;
         //double T=0.9;
-        mu = dt*L*L*P*pow((1./(2*M_PI*k*AmbientTemp)),0.5);
+        mu = dt*L*L*P*pow((1./(2*m*M_PI*k*AmbientTemp)),0.5);
         //std::cout << mu << std::endl;
         return gsl_ran_poisson(r,mu);
 }
@@ -455,6 +456,8 @@ void InitBarostatFull(std::list<Particle*>& particles)
         unsigned int N;
         //double sigma=1.1;
         double sigma=AmbientTemp;
+        double lambda = 0;
+        double vel = 0;
 
         Particle *px0;
         Particle *pxL;
@@ -462,6 +465,17 @@ void InitBarostatFull(std::list<Particle*>& particles)
         Particle *pyL;
         Particle *pz0;
         Particle *pzL;
+
+    /*
+     *double vSum = 0;
+     *for(int i=0;i<N;i++)
+     *    for(int k=0;k<3;k++)
+     *        vSum += cube[i].v[k] * cube[i].v[k];
+     *double lambda = sqrt(3*(N-1)*Temp/vSum);
+     *for(int i=0;i<N;i++)
+     *    for(int k=0;k<3;k++)
+     *        cube[i].v[k] *= lambda;
+     */
 
         double kinE = 0.;
 
@@ -474,8 +488,16 @@ void InitBarostatFull(std::list<Particle*>& particles)
                             ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
                             ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
                             gsl_ran_rayleigh(r,sigma),
-                            ((gsl_rng_uniform(r)-0.5)*2)*AmbientTemp,
-                            ((gsl_rng_uniform(r)-0.5)*2)*AmbientTemp,"GasIn");
+                            ((gsl_rng_uniform(r)-0.5)*4)*AmbientTemp,
+                            ((gsl_rng_uniform(r)-0.5)*4)*AmbientTemp,"GasIn");
+            /*
+             *vel = px0->v[0]*px0->v[0]+px0->v[1]*px0->v[2]+px0->v[2]*px0->v[2];
[…]

### 2017-01-02 doc/draft1.tex
updated draft

@@
         \begin{equation}
-            \langle N_\text{fac}\rangle = \Delta t L^2 P \left(\frac{1}{2\pi m k_B T}\right)^\frac12
+            %\langle N_\text{fac}\rangle = \Delta t L^2 P \left(\frac{1}{2\pi m k_B T}\right)^\frac12
+            \langle N_\text{fac}\rangle = \Delta t L^2 P \sqrt{\frac{1}{2\pi m k_B T}}
         \end{equation}
-        where $\Delta T$ is the time step of the simulation, $L$ is the side length of the cell in which the particle is created, $P$ is the desired
+        where $\Delta t$ is the time step of the simulation, $L$ is the side length of the cell in which the particle is created, $P$ is the desired
         pressure, $m$ is the mass of the gas particle, $k_B$ is the Boltzmann constant (which will be set to 1 in reduced units) and $T$ is the
-        desired temperature. This chosen number of particles is then equally distributed over the face of the cell on which they are created.
-    \item wos
+        desired temperature. This chosen number of particles is then equally distributed over the face of the cell on which they are created. The
+        velocities of the created particles are drawn from two different random number distributions. The component of the particle perpendicular to
+        the surface is drawn from a Rayleigh distribution of the form 
+        \begin{equation}
+            p(v_i) = \frac{m}{k_B T}v_i \ e^{-\frac{mv_i^2}{2k_BT}}
+        \end{equation}
+        The other components of the particles' velocity are drawn from a Maxwell-Boltzmann distribution.
+    \item Perform the first step of the velocity Verlet algorithm to propagate the particle positions by one time step.
+    \item Check if any gas particles have left the minimal volume of cells and remove those which have.
+    \item Check if the geometry of the crystal and with it the minimal volume of cells has changed. If it has, remove all gas particles in the cells
+        that are no longer needed. Then insert new gas particles to the created cells with a number drawn from a Poisson distribution with mean value
+        \begin{equation}
+            \langle N_\text{ins}\rangle = \frac{PL^3}{k_BT}
+        \end{equation}
+        If the length of the cell $L$ is chosen to be equal to the cut-off radius $r_c$, this insertion should only be carried out with a probablity of
+        \begin{equation}
+            P_\text{ins} = e^{-\frac{U}{k_BT}} < 1
+        \end{equation}
+        (where U is the interaction energy between the gas particle and the crystal) because it is possible that the inserted particle is 
+        within the interaction range of the crystal. 
+    \item Compute all forces.
+    \item Perform the second step of the velocity Verlet and propagate the particle velocities by one time step. 
 \end{enumerate}
+This algorithm is formulated in a general fashion, so that a wide range of crystals and geometries can be used. The setup used in this thesis is a
+FCC lattice, which has a nice symmetry that is not influenced by the surrounding gas. The algorithm will therefore not be carried out using 
+cell-lists and a minimal volume of cells that is changing over the course of the simulation, but rather a fixed setup of the volume surrounding the 
+nano crystal.\\ 
+Since the crystal itself is a cube a straighforward approach is surrounding it by a bigger cube.
 

### 2017-01-02 src/main.cpp

@@ -51,7 +51,7 @@ int main(int argc,char** argv)
      *    q += 0.01;
      *}
      */
-    setValues(0.1,0.04,0.2,1.1,0.06);
+    setValues(0.1,0.04,0.2,0.9,0.06);
     mainLoop();
     return 0;
 }

### 2017-01-03 doc/draft1.tex
updated draft

@@
 vacuum chamber. The trapping of the silica nano particle (which will be referred to as \textit{glass particle}) is achieved by a gradient force of the
-laser beam acting on the particle. The experimental setup is depicted in fig. \ref{fig:setup}.\\
+laser beam acting on the particle. The experimental setup is depicted in fig.~\ref{fig:setup}.\\
 The particle fluctuates within the trap in all three spatial directions. These fluctuations can be approximated
@@
 nano crystal.\\ 
-Since the crystal itself is a cube a straighforward approach is surrounding it by a bigger cube.
+Since the crystal itself is a cube a straighforward approach is surrounding it by a bigger cube. The distance between the cube face and the nano
+crystal is chosen to be the side length of the nano crystal, so the setup will properly scale for different numbers of particles. 
+The schematic setup for this is depicted in fig.\ref{fig:gascube}.
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.8]{images/gas_volume.pdf}
+        \caption{Schematic setup for the barostat: The gas particles will be created on the face of the cube surrounding the crystal. The
+                    cube is chosen to be at a distance of one side length of the nano crystal.}
+        \label{fig:gascube}
+    \end{center}
+\end{figure}
 

### 2017-01-05 doc/draft1.tex
more drafting, more images

@@
 The model used in this thesis was developed by Gr\"unwald and Dellago \cite{Gruenwald2006} and uses an ideal gas of non-interacting particles as
-pressure medium. The particles of this pressure medium flow into the simulation from an outside box, whose geometry is based on the form of the object
-of interest. This barostat also acts as a thermostat and thus is suitable for the application in this thesis.\\
+pressure medium. The particles of this pressure medium flow into the simulation from an outside volume, whose geometry is based on the form of the
+object of interest (this will be referred to as the minimal volume of cells). This barostat also acts as a thermostat and thus is suitable for the 
+application in this thesis.\\
 The gas particles interact with the object via a soft-sphere potential of the form
@@
     \label{eq:softsphere}
-    U(r) = \frac{1}{r^{12}}
+    U(r) = \varepsilon \left(\frac{\sigma}{r}\right)^{12}
 \end{equation}
-in reduced units. \\
-BLABLABLA\\
+where $\varepsilon$ is the interaction strength and $\sigma$ is the interaction range and $r$ is the distance between the gas particle and the
+interacting particle. The gas particles do not interact with one another in this model.\\
+In order to increase the efficiency of the computing process, $\sigma$ should be chosen carefully. For larger $\sigma$, the number of interaction
+partners increases, which increases the number of force calculations which are a very time consuming part. For smaller $\sigma$ the possibility for
+gas particles reaching the inside of the nano crystal increases, which is not desirable. So $\sigma$ should be chosen small enough to keep the force
+calculations at a minimum and large enough for the particle to stay on the outside of the crystal.\\ 
 The algorithm can be performed by following these steps:
@@
         \begin{equation}
-            P_\text{ins} = e^{-\frac{U}{k_BT}} < 1
+            P_\text{ins} = e^{-\frac{U}{k_BT}}
         \end{equation}
@@
     \begin{center}
-        \includegraphics[scale=0.8]{images/gas_volume.pdf}
-        \caption{Schematic setup for the barostat: The gas particles will be created on the face of the cube surrounding the crystal. The
-                    cube is chosen to be at a distance of one side length of the nano crystal.}
+        \includegraphics[scale=0.5]{images/gas_volume_ver2.pdf}
+        \caption{Schematic setup for the barostat. The outer volume (blue) is one side length $L$ away from the nano crystal (black). 
+                    The gas particles are created along the blue lines of the outer volume. Since the geometry of the nano crystal does not change, 
+                    the geometry of the outer volume will be constant throughout the simulation.}
         \label{fig:gascube}
@@
 % results with different parameters, graphs, etc.
+Since the simulation consists of various parts that have to work together, we first need to check if every part itself works.\\
+The first element of the simulation is the setup of the FCC lattice on which the particles are placed on. To check if this is done correctly, we can
+use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano crystal can be
+seen in fig.~\ref{fig:glass}.
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.2]{images/initial_position_red.png}
+        \caption{}
+        \label{fig:gascube}
+    \end{center}
+\end{figure}
 

### 2017-01-05 src/main.cpp

@@ -51,7 +51,7 @@ int main(int argc,char** argv)
      *    q += 0.01;
      *}
      */
-    setValues(0.1,0.04,0.2,0.9,0.06);
+    setValues(0.2,0.04,0.2,0.9,0.06);
     mainLoop();
     return 0;
 }

### 2017-01-09 doc/draft1.tex
added while loop, renders

@@
 Since the simulation consists of various parts that have to work together, we first need to check if every part itself works.\\
+\subsection{The Crystal}
 The first element of the simulation is the setup of the FCC lattice on which the particles are placed on. To check if this is done correctly, we can
-use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano crystal can be
-seen in fig.~\ref{fig:glass}.
+use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano crystal 
+can be seen in fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
+where the atoms take the position on the gap between the atoms from the layer below.\\
 \begin{figure}[h]
     \begin{center}
-        \includegraphics[scale=0.2]{images/initial_position_red.png}
-        \caption{}
-        \label{fig:gascube}
+        \includegraphics[scale=0.2]{images/fcc_rendering_new.png}
+        \caption{Rendering of the model of the nano crystal before starting the simulation. 
+                    The perspective shows the 3 layers of the FCC lattice. These layers consist of atoms
+                    taking the position above the gap between two of the atoms from the layer below and match the positions of the atoms 2 layers
+                below. The black box is the outline of the particles' positions and shows the geometry of the system.}
+        \label{fig:glass}
     \end{center}

### 2017-01-10 doc/draft1.tex
added renderings, more draft

@@
 \subsection{The Velocity-Verlet Algorithm}
+\label{section:verlet}
 When we look at the system from a thermodynamical standpoint, we see that it follows some kind of path in the phase space as time progresses. Every
@@
 
+\subsection{The Laser Beam - Trapping}
+As mentioned above, the glass particle is trapped in the laser beam and the motion of the particle is suppressed in every direction. This behaviour
+has to be modeled as well.\\
+The approximation of the original paper \cite{Gieseler2014}, where the movements in the three spatial directions are decoupled will be
+used in the model as well. As to the model of the trapping force itself, the most straightforward approach is to use a harmonic oscillator potential.
+The force and the corresponding potential can be written as
+\begin{eqnarray}
+    \mathbf{F} &=& -k x\\
+    U &=& \frac12 k x^2
+\end{eqnarray}
+The trap is implemented by using the velocity Verlet algorithm with the above force to propagate the glass particle. Since the trap is acting on the
+whole system, the force is acting on the center of mass position and velocity. To calculate the center of mass positions and velocities, the positions
+and velocities of all particles have to be summed up and divided by the particle number:
+\begin{eqnarray}
+    \mathbf{r}_\text{COM} &=& \sum_{i=1}^N \frac{\mathbf{r}_i}{N}\\
+    \mathbf{v}_\text{COM} &=& \sum_{i=1}^N \frac{\mathbf{v}_i}{N}
+\end{eqnarray}
+The position of the center of mass and the center of mass velocity are then propagated by using the same procedure as introduced in
+section~\ref{section:verlet}.
+
 \subsection{Surrounding Gas - Barostat}
@@
     \begin{center}
-        \includegraphics[scale=0.2]{images/fcc_rendering_new.png}
-        \caption{Rendering of the model of the nano crystal before starting the simulation. 
-                    The perspective shows the 3 layers of the FCC lattice. These layers consist of atoms
-                    taking the position above the gap between two of the atoms from the layer below and match the positions of the atoms 2 layers
-                below. The black box is the outline of the particles' positions and shows the geometry of the system.}
+        \includegraphics[scale=0.2]{images/fcc_rendering_layer_newcoloring.png}
+        \caption{A rendered image of the FCC crystal. The chosen color-coding along the (1,1,1) Miller-index 
+                highlights the layers for this perspective. The perspective is
+                chosen to see the layering of the FCC crystal with its 3 layer structure. The black box acts as a reference point to outline the 
+                geometry of the system.}
+            %Rendering of the model of the nano crystal before starting the simulation. 
+            %The chosen perspective shows the 3 layers of the FCC crystal with a
+                    %These layers consist of atoms
+                    %taking the position above the gap between two of the atoms from the layer below and match the positions of the atoms 2 layers
+                %below. The black box is the outline of the particles' positions and shows the geometry of the system.}
         \label{fig:glass}
@@
 \end{figure}
+
+\subsection{Velocity Verlet}
+The simulation is taking place in a NVE ensemble from a thermodynamic point of view. This means that the particle number $N$, the volume $V$ and the
+total energy $E$ are constant. In contrast to the NVT ensemble used in Monte-Carlo simulations this means, that the velocities of the particle can not
+be chosen directly. The initial velocities have to be drawn from a random number generator. This means that the velocities have to be adjusted after 
+they have been drawn.\\
+The first adjustment that needs to be made is making sure that the system is staying in place and not floating away. To do this, the velocity vectors
+of all the atoms in the system have to be added to get the resulting velocity vector for the whole system that is thought to be acting on the center
+of mass. This resulting vector is then divided by the number of particles in the system and the resulting vector is then subtracted from every
+velcocity vector of every atom in the system. 
+\begin{eqnarray}
+    \mathbf{v}_\text{tot} &=& \sum_{i=1}^N \mathbf{v}_i\\
+    \mathbf{v}_\text{cut} &=& \frac{\mathbf{v}_\text{tot}}N\\
+    \mathbf{v}_i^\text{new} &=& \mathbf{v}_i - \mathbf{v}_\text{cut}
+\end{eqnarray}
+The next adjustment is a small one that makes a difference in the long run. When choosing random velocities, the system not only tends to have a
+resulting velocity vector that causes the whole system to float away, but also a resulting angular momentum that causes the system to rotate.
+The angular momentum can be defined via the moment of inertia $\mathbf{I}$
+\begin{eqnarray}
+    \mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
+    \boldsymbol{\omega} &=& \frac{\mathbf{r} \times \mathbf{v}}{r^2}
+\end{eqnarray}
+where $\boldsymbol{\omega}$ is the angular velocity of the system. 
+\\The velocities of the particle can roughly be chosen by adjusting the range of numbers the random number generator can choose from. However, the
+resulting velocities (and therefore the temperature of the system) are generally not precisely how they should be. If the system should have a
+specific starting temperature $T$, the velocities need to be scaled accordingly. This rescaling is done by summing the square of the velocities
+(this equals calculating the kinetic energy since the masses are set to 1) and rescale them by a factor $\lambda$:
+\begin{eqnarray}
+    E &=& \sum_{i=1}^N \mathbf{v}_i^2\\
+    \lambda &=& \sqrt{\frac{3(N-1)T}{E}}\\
+    \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i 
+\end{eqnarray}
+This process has to be done a couple of times, until the system equilibrates to the desired temperature. The problem with this procedure is the fact
+that it contradicts the constant energy property of the ensemble. If we look at the energy values in fig.~\ref{fig:energystart} we can see the
+irregularities before the system equilibrates to a constant value.
+
+%\begin{figure}[h]
+    %\begin{center}
+        %\begin{subfigure}[t]{0.3\textwidth}
+            %\includegraphics[scale=0.3]{images/fcc-2.pdf}
+            %\caption{Schematic figure of a face centered cubic (FCC) lattice}
+            %\label{fig:fcc}
+        %\end{subfigure} 
+        %\
+        %\begin{subfigure}[t]{0.3\textwidth}
+            %\includegraphics[scale=0.15]{images/fcc_unit.png}
+            %\caption{Schematic of the number of particles contained in one FCC unit cell. The atoms on the edges are shared by 8 cells and the ones 
+                %on the faces by 2 cells\cite{unitcell}}
+            %\label{fig:fcccut}
+        %\end{subfigure} 
+        %\
+        %%\hspace{60pt}
+        %\begin{subfigure}[t]{0.3\textwidth}
+            %\includegraphics[scale=0.3]{images/unit_cell.pdf}
+            %\caption{Setup for successively building a FCC lattice}
+            %\label{fig:unitcell}
+        %\end{subfigure}
+    %\end{center}
+%\end{figure}
 

### 2017-01-12 doc/draft1.tex
more draft, more graphs

@@
 \begin{eqnarray}
-    \mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
+    \label{eq:angularmomentum}\mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
     \boldsymbol{\omega} &=& \frac{\mathbf{r} \times \mathbf{v}}{r^2}
 \end{eqnarray}
-where $\boldsymbol{\omega}$ is the angular velocity of the system. 
+where $\boldsymbol{\omega}$ is the angular velocity of the system. Another way to define the angular momentum is given by
+\begin{equation}
+    \label{eq:angularmomentum2}
+    \mathbf{L} = \mathbf{r} \times \mathbf{p}
+\end{equation}
+where $\mathbf{p}$ is the momentum of the particle, which in this case is equal to its velocity since the mass is set to 1.The moment of inertia 
+$\mathbf{I}$ for a filled cube is given by 
+\begin{equation}
+    \mathbf{I} = 
+    \begin{pmatrix}
+        \frac{mL^2}{6} & 0 & 0\\
+        0 & \frac{mL^2}{6} & 0\\
+        0 & 0 & \frac{mL^2}{6} 
+    \end{pmatrix}
+\end{equation}
+where $m$ will be later set to 1 and $L$ is the side length of the cube. To use this equations to suppress the rotation of the cube, we need to use
+equation~\eqref{eq:angularmomentum} and rewrite it as
+\begin{equation}
+    \label{eq:omega}
+    \boldsymbol{\omega} = \mathbf{I}^{-1}\mathbf{L}
+\end{equation}
+where $\mathbf{I}^{-1}$ is the inverse matrix of $\mathbf{I}$. The calculation of the inverse matrix is fairly straightforward, as it only contains
+values in the diagonal, which means that the inverse matrix can be calculated by simply inverting these values. The angular momentum for every
+particle $\mathbf{L}_i$ in the above equation be calculated using eq.~\eqref{eq:angularmomentum2}, where the position is taken to be the distance
+between the particle and the center of mass
+\begin{equation}
+    \Delta\mathbf{r}_i = \mathbf{r}_\text{COM} - \mathbf{r}_i
+\end{equation}
+The values of the angular momentum for each paricle are added up and multiplied by the inverse of the moment of inertia, as described in
+eq.~\eqref{eq:omega}. The final step is updating the velocities of the particles:
+\begin{equation}
+    \mathbf{v}_i^\text{new} = \mathbf{v}_i - \left(\boldsymbol{\omega} \times \Delta\mathbf{r}_i\right)
+\end{equation}
 \\The velocities of the particle can roughly be chosen by adjusting the range of numbers the random number generator can choose from. However, the

### 2017-01-12 include/functions.hpp

@@ -13,6 +13,7 @@ void VelocityVerlet(Particle* particle,int WRITE,FILE* output);
 void VelocityVerlet(Particle* particle);
 
 void InitVelocities(Particle* particle);
+void InitVelocitiesTest(Particle* particle);
 double Pressure(Particle* particle);
 double* Force(Particle *particle,int i,int j);
 double Distance(Particle* particle, int i, int j);

### 2017-01-12 src/main.cpp

@@ -34,7 +34,7 @@ int main(int argc,char** argv)
 
     double t = 0.2;
     double ambient = 0.06;
-    double press = 0.5;
+    double press = 0.8;
     double q = 0.04;
     while(q < 0.1)
     {

### 2017-01-12 test/compile_inertia_test.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+sources="$(cat sources.txt)"
+g++  inertia_test.cpp $sources -lgsl -lgslcblas -o  inertia_test.o

### 2017-01-12 test/ehex_output.cpp

@@ -1,0 +1,99 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <cstdlib>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../include/globals.hpp"
+#include "../include/classes.hpp"
+#include "../include/functions.hpp"
+
+int main()
+{
+
+    setValues(0.2,0.04,0.2,0.8,0.06);
+    Particle *particles = new Particle[N];
+    InitPositions(particles);
+    FILE* tempout = fopen("ehex_temp_1.dat","w");
+    FILE* enerout = fopen("ehex_energy_1.dat","w");
+    FILE* compos = fopen("ehex_compos.dat","w");
+    FILE* comvel= fopen("ehex_comvel.dat","w");
+
+
+    InitVelocities(particles);
+    double* vCM = new double[3];
+    double* rCM = new double[3];
+    double energy = 0.;
+    //calcCM(particles,rCM,vCM);
+
+    for(int i=0;i<200000;i++)
+    {
+        if(i%100 == 0)
+        {
+            std::cout << "Step " << i << std::endl;
+            for(int k=0;k<3;k++)
+            {
+                rCM[k] = 0;
+                vCM[k] = 0;
+            }
+            for(unsigned int j=0;j<N;j++)
+                for(int k=0;k<3;k++)
+                {
+                    rCM[k] += particles[j].r[k];
+                    vCM[k] += particles[j].v[k];
+                }
+            for(int k=0;k<3;k++)
+            {
+                rCM[k] = rCM[k]/(N*1.0);
+                vCM[k] = vCM[k]/(N*1.0);
+            }
+            fprintf(compos,"%lf\t%lf\t%lf\n",rCM[0],rCM[1],rCM[2]);
+            fprintf(comvel,"%lf\t%lf\t%lf\n",vCM[0],vCM[1],vCM[2]);
+        }
+        if(i%1000 == 0)
+            if(i < 10000)
+                rescaleVelocities(particles);
+        VelocityVerlet(particles);
+        //calcTemp(particles,tempout);
+        //energy = calculateEnergies(particles);
+        //fprintf(enerout,"%lf\n",energy);
+    }
+    
+    for(int i=0;i<20000;i++)
+    {
+        if(i%100 == 0)
+        {
+            std::cout << "Step " << i << std::endl;
+            for(int k=0;k<3;k++)
+            {
[…]

### 2017-01-12 test/inertia_test.cpp

@@ -1,0 +1,42 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <cstdlib>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../include/globals.hpp"
+#include "../include/classes.hpp"
+#include "../include/functions.hpp"
+
+int main(int argc, char** argv)
+{
+    int run=0;
+    Particle *cube = new Particle[N];
+    InitPositions(cube);
+    InitVelocities(cube);
+    ComputeAccelerations(cube);
+    FILE* inertiatest = fopen("intertia_test.xyz","w");
+    fprintf(inertiatest,"%d\nCOMMENT\n",N);
+    for(int i=0;i<N;i++)
+        fprintf(inertiatest,"Ar\t%lf\t%lf\t%lf\n",cube[i].r[0],cube[i].r[1],cube[i].r[2]);
+
+    for(run=0;run<100000;run++)
+    {
+            if(run%500==0)
+                printf("(INIT) Zeitschritt %d\n",run);
+            VelocityVerlet(cube,0,NULL);
+            if(run%100 == 0)
+            {
+                //rescaleVelocities(cube);
+                fprintf(inertiatest,"%d\nCOMMENT\n",N);
+                for(int i=0;i<N;i++)
+                    fprintf(inertiatest,"Ar\t%lf\t%lf\t%lf\n",cube[i].r[0],cube[i].r[1],cube[i].r[2]);
+            }
+    }
+    fprintf(inertiatest,"%d\nCOMMENT\n",N);
+    for(int i=0;i<N;i++)
+        fprintf(inertiatest,"Ar\t%lf\t%lf\t%lf\n",cube[i].r[0],cube[i].r[1],cube[i].r[2]);
+	return 0;
+}

### 2017-01-12 test/plot/compile_verlet_output.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+sources="$(cat sources.txt)"
+g++  verlet_output.cpp  $sources -lgsl -lgslcblas -o  verlet_output.o

### 2017-01-12 test/plot/verlet_output.cpp

@@ -1,9 +1,19 @@
-#include "../include/globals.hpp"
-#include "../include/classes.hpp"
-#include "../include/functions.hpp"
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <cstdlib>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../../include/globals.hpp"
+#include "../../include/classes.hpp"
+#include "../../include/functions.hpp"
 
 int main()
 {
+
+    setValues(0.2,0.04,0.2,0.8,0.06);
     Particle *particles = new Particle[N];
     InitPositions(particles);
     FILE* tempout = fopen("verlet_temp_1.dat","w");
@@ -42,8 +52,9 @@ int main()
             fprintf(compos,"%lf\t%lf\t%lf\n",rCM[0],rCM[1],rCM[2]);
             fprintf(comvel,"%lf\t%lf\t%lf\n",vCM[0],vCM[1],vCM[2]);
         }
-        if(i%100 == 0 && i < 2000)
-            rescaleVelocities(particles);
+        if(i%1000 == 0)
+            if(i < 10000)
+                rescaleVelocities(particles);
         VelocityVerlet(particles);
         calcTemp(particles,tempout);
         energy = calculateEnergies(particles);

### 2017-01-13 doc/draft1.tex
updated

@@
 that it contradicts the constant energy property of the ensemble. If we look at the energy values in fig.~\ref{fig:energystart} we can see the
-irregularities before the system equilibrates to a constant value.
+irregularities (jumps) before the system equilibrates to a constant value. This equilibration process divides the simulation in two phases: the equilibration
+phase and the measurement phase. In the equilibration phase, the system has time to reach a configuration that is (as the name suggests) in an 
+equilibrated state. During this phase, there are usually no measurements (such as energy) performed, since the values from this phase are distorting
+the mean value of the simulation. That is why the configurations from the equilibration phase are discarded and the measurement phase takes place, 
+where the values do get calculated and are used to determine the mean values.
+
+\subsection{eHEX}
+The next checkpoint is the eHEX algorithm. Since it contains several steps and variables that have to be calculated, this algorithm is very prone
+to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\DeltaQ$. 
+The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
+The investigation of the effects on the system can be done in two ways: comparing the temperature and the energy of the system during the 
+application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\DeltaQ$.\\
+Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
+with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\
 

### 2017-01-13 test/plot/ehex_output.cpp

@@ -13,13 +13,13 @@
 int main()
 {
 
-    setValues(0.2,0.04,0.2,0.8,0.06);
+    setValues(0.2,0.09,0.2,0.8,0.06);
     Particle *particles = new Particle[N];
     InitPositions(particles);
-    FILE* tempout = fopen("ehex_temp_1.dat","w");
-    FILE* enerout = fopen("ehex_energy_1.dat","w");
-    FILE* compos = fopen("ehex_compos.dat","w");
-    FILE* comvel= fopen("ehex_comvel.dat","w");
+    FILE* tempout = fopen("ehex_temp_009_1.dat","w");
+    FILE* enerout = fopen("ehex_energy_009_1.dat","w");
+    FILE* compos = fopen("ehex_compos009.dat","w");
+    FILE* comvel= fopen("ehex_comvel009.dat","w");
 
 
     InitVelocities(particles);
@@ -55,6 +55,40 @@ int main()
         if(i%100 == 0)
             if(i < 10000)
                 rescaleVelocities(particles);
+        VelocityVerlet(particles);
+        //calcTemp(particles,tempout);
+        //energy = calculateEnergies(particles);
+        //fprintf(enerout,"%lf\n",energy);
+    }
+    
+
+    for(int i=0;i<20000;i++)
+    {
+        if(i%100 == 0)
+        {
+            std::cout << "Step " << i << std::endl;
+            for(int k=0;k<3;k++)
+            {
+                rCM[k] = 0;
+                vCM[k] = 0;
+            }
+            for(unsigned int j=0;j<N;j++)
+                for(int k=0;k<3;k++)
+                {
+                    rCM[k] += particles[j].r[k];
+                    vCM[k] += particles[j].v[k];
+                }
+            for(int k=0;k<3;k++)
+            {
+                rCM[k] = rCM[k]/(N*1.0);
+                vCM[k] = vCM[k]/(N*1.0);
+            }
+            fprintf(compos,"%lf\t%lf\t%lf\n",rCM[0],rCM[1],rCM[2]);
+            fprintf(comvel,"%lf\t%lf\t%lf\n",vCM[0],vCM[1],vCM[2]);
+        }
+        //if(i%100 == 0)
+            //if(i < 10000)
+                //rescaleVelocities(particles);
         VelocityVerlet(particles);
         calcTemp(particles,tempout);
         energy = calculateEnergies(particles);

### 2017-01-14 doc/draft1.tex

@@
 \begin{eqnarray}
-    E &=& \sum_{i=1}^N \mathbf{v}_i^2\\
-    \lambda &=& \sqrt{\frac{3(N-1)T}{E}}\\
-    \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i 
+ \label{eq:rescale1}   E &=& \sum_{i=1}^N \mathbf{v}_i^2 \\
+ \label{eq:rescale2}   \lambda &=& \sqrt{\frac{3(N-1)T}{E}} \\
+ \label{eq:rescale3}   \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i  
 \end{eqnarray}
@@
 where the values do get calculated and are used to determine the mean values.  
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.3]{images/verlet_output_te.pdf}
+        \caption{Temperature and energy in the equilibration phase of the velocity Verlet algorithm. The velocity rescaling (see eq. 
+        \eqref{eq:rescale1}-\eqref{eq:rescale3}) is taking place every 1000th step until the 9000th, which can be seen in the energy curve whre the
+    jumps are happening.}
+        \label{fig:energystart}
+    \end{center}
+\end{figure}
 
@@
 The next checkpoint is the eHEX algorithm. Since it contains several steps and variables that have to be calculated, this algorithm is very prone
-to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\DeltaQ$. 
+to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\Delta Q$. 
 The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
 The investigation of the effects on the system can be done in two ways: comparing the temperature and the energy of the system during the 
-application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\DeltaQ$.\\
-Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
-with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\
+application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\Delta Q$.\\
+%Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
+%with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\
+Since the eHEX algorithm injects heat into the system, one would expect the energy and the temperature to rise during the application of the
+algorithm. In fig.~\ref{fig:ehex_te} we can see that exactly this is happening. The graph shows a velocity Verlet phase (that follows an equilibration
+phase which is not recorded) of constant temperature and energy and an eHEX phase, where the temperature and the enrgy are constantly rising. 
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.4]{images/ehex_te.pdf}
+        \caption{Comparison of temperature (red) and energy (blue) during the application of the eHEX algorithm with $\Delta Q=0.04$ (in reduced units).
+                The first 20000 timesteps, the energy and temperature are constant during the application of the velocity Verlet algorithm and start
+            to rise, once the eHEX algortithm takes over. The values for the temperature can be seen on the ordinate on the left hand side
+            and the values for the energy on the right hand side. }
+        \label{fig:ehex_te}
+    \end{center}
+\end{figure}
+The comparison of different values of $\Delta Q$ can be seen in fig.~\ref{fig:ehex_dq}. As the correlation between the temperature and the energy has
+been shown in fig.~\ref{fig:ehex_te}, the plotting of the temperature data should suffice. 
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.4]{images/ehex_temp_dq.pdf}
+        \caption{Comparison of different values for $\Delta Q$, ranging from 0 to 0.09. As the value $\Delta Q=0$ corresponds to no heat being
+        injected into the system, it is not surprising that the temperature remains constant. The graph shows that an increase in $\Delta Q$ leads to
+    an increase in the rate of warming up, which means that for bigger $\Delta Q$, the system is heating up more quickly.}
+        \label{fig:ehex_dq}
+    \end{center}
+\end{figure}
+
+\subsection{Barostat}
+The barostat, as described in the previous section, keeps the system from overheating. The algorithm is applied at the same time as the eHEX algorithm
+which leads to the fact, that the system has time to heat up, until the particles reach the nano crystal. This can be seen in the temperature
+development in fig.~\ref{fig:baro1}.
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.4]{images/ehex_te.pdf}
+        \caption{}
+        \label{fig:baro1}
+    \end{center}
+\end{figure}
 

### 2017-01-16 doc/draft1.tex
updates

@@
 \subsection{Barostat}
-The barostat, as described in the previous section, keeps the system from overheating. The algorithm is applied at the same time as the eHEX algorithm
-which leads to the fact, that the system has time to heat up, until the particles reach the nano crystal. This can be seen in the temperature
-development in fig.~\ref{fig:baro1}.
+The barostat, as described previously, keeps the system from overheating. The barostat algorithm is applied at the same time as the eHEX algorithm,
+after the velocity Verlet has equilibrated the system. This means that the system is beginning to heat up, until the particles reach the crystal.
+Because of this, the temperature of the system rises, until it reaches a steady value again. This behaviour can be seen in fig.~\ref{fig:baro1}.
 \begin{figure}[h]

### 2017-01-16 doc/draft1.tex
updates

@@
         \eqref{eq:rescale1}-\eqref{eq:rescale3}) is taking place every 1000th step until the 9000th, which can be seen in the energy curve whre the
-    jumps are happening.}
+    jumps are happening. The desired temperature in this case is $T^* = 0.2$, which is reached after about 7000 steps.}
         \label{fig:energystart}
@@
 after the velocity Verlet has equilibrated the system. This means that the system is beginning to heat up, until the particles reach the crystal.
-Because of this, the temperature of the system rises, until it reaches a steady value again. This behaviour can be seen in fig.~\ref{fig:baro1}.
+Because of this, the temperature of the system rises, until it reaches a steady value again. This behaviour can be seen in fig.~\ref{fig:baro1}.\\
 \begin{figure}[h]
     \begin{center}
-        \includegraphics[scale=0.4]{images/ehex_te.pdf}
-        \caption{}
+        \includegraphics[scale=0.4]{images/baro_temp_start.pdf}
+        \caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
+        in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
+    an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
+reached and the temperature seems to stay at this value.}
         \label{fig:baro1}
@@
 \end{figure}
+As we are interested in the temperatures of the particles before and after the interaction with the crystal, we need some way to differentiate those
+two states. This can be done when looking at the total energy of the gas particles.\\
+The gas particles are created on the surface of the simulation box, which is further away than the interaction range between the gas particles and the
+nano crystal. Since the gas particles do not interact with each other, the total energy of the gas particles does not change until they reach the
+interaction range of the nano crystal. There they undergo a change of energy due to the interaction with the nano crystal and then make their way away
+from the nano crystal. After the interaction, when they reach a point outside of the interaction range, the enrgy stays constant again until the
+particles reach the limits of the simulation box.\\
+Therefore, the life span of a single gas particle can be partioned into 3 stages: incoming, interacting, outgoing. In a typical scene of interaction
+all three of these stages are represented by different gas particles, as can be seen in fig.~\ref{fig:barointeraction}.\\
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.4]{images/baro_170105_new_medium.png}
+        \caption{Rendering of the application of the barostat algorithm. The red particles are incoming gas particles, that have not yet interacted, 
+            pink particles are currently interacting and green particles are leaving the simulation box and
+        moving away from the nano crystal (blue). To render this image, a cut has been made to make a view into the simulation box possible.}
+        \label{fig:barointeraction}
+    \end{center}
+\end{figure}
+Since the gas particles are interacting with the nano crystal, the temperatures and energies are values of interest. The energies are recorded using
+histograms and the temperature is calculated instantaneously every 100th timestep.\\
+The calculation of the energies of the incoming particles has to be done with care. The problem are particles that enter the simulation box, but fly
+by the nano crystal at a distance that is greater than the interaction range. Since the change of the energy is the interesting part, only the energy 
+particles of that \textit{do} interact with the nano crystal shall be calculated.
 

### 2017-01-18 doc/draft1.tex
updates

--- 2017-01-14 doc/draft1.tex
@@
 \subsection{Barostat}
-The barostat, as described in the previous section, keeps the system from overheating. The algorithm is applied at the same time as the eHEX algorithm
-which leads to the fact, that the system has time to heat up, until the particles reach the nano crystal. This can be seen in the temperature
-development in fig.~\ref{fig:baro1}.
+The barostat, as described previously, keeps the system from overheating. The barostat algorithm is applied at the same time as the eHEX algorithm,
+after the velocity Verlet has equilibrated the system.\\
+While the eHEX algorithm is pumping energy into the nano crystal while the surrounding gas particles absorb excessive energy. This leads to the
+development of a new, higher steady temperature of the nano crystal, which can be seen in fig.~\ref{fig:baro1}. The first simulations were performed
+with the eHEX and the barostat algorithm starting at the same point. This lead to the question: does the new steady temperature depend on the time it
+takes the gas particles to reach the system? The answer is: no. After applying the barostat algorithm first to fill the box with gas atoms and then
+applying the eHEX algorithm it can be seen that the resulting temperature is the same, which can be seen in fig.~\ref{fig:baro2}.
+%This means that the system is beginning to heat up, until the particles reach the crystal.
+%Because of this, the temperature of the system rises, until it reaches a steady value again. This behaviour can be seen in fig.~\ref{fig:baro1}.\\
 \begin{figure}[h]
     \begin{center}
-        \includegraphics[scale=0.4]{images/ehex_te.pdf}
-        \caption{}
+        \includegraphics[scale=0.4]{images/baro_temp_start.pdf}
+        \caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
+        in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
+    an increase in temperature until a new steady temperature is reached.}
+        %\caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
+        %in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
+    %an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
+%reached and the temperature seems to stay at this value.}
         \label{fig:baro1}
@@
 \end{figure}
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.4]{images/baro_temp_compare_other.pdf}
+        \caption{Two versions of the application of the eEHX and barostat algorithm. The shorter run (red, x-axis on the bottom) represents 
+            the simultaneous start of the eHEX and the barostat algorithm. In the longer run (blue, x-axis on the top) there are 20000 steps, where
+        the barostat algorithm is applied at the same time as the velocity Verlet to give the system time to equilibrate and only after the eHEX
+    algorithm is taking the place of the velocity Verlet. The mean value of the resulting temperature is the same, indicating that the final
+temperature is based on the values of the parameters, such as pressure and ambient temperature, rather than geometry.}
+        %\caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
+        %in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
+    %an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
+%reached and the temperature seems to stay at this value.}
+        \label{fig:baro1}
+    \end{center}
+\end{figure}
+As we are interested in the temperatures of the particles before and after the interaction with the crystal, we need some way to differentiate those
+two states. This can be done when looking at the total energy of the gas particles.\\
+The gas particles are created on the surface of the simulation box, which is further away than the interaction range between the gas particles and the
+nano crystal. Since the gas particles do not interact with each other, the total energy of the gas particles does not change until they reach the
+interaction range of the nano crystal. There they undergo a change of energy due to the interaction with the nano crystal and then make their way away
+from the nano crystal. After the interaction, when they reach a point outside of the interaction range, the enrgy stays constant again until the
+particles reach the limits of the simulation box.\\
+Therefore, the life span of a single gas particle can be partioned into 3 stages: incoming, interacting, outgoing. In a typical scene of interaction
+all three of these stages are represented by different gas particles, as can be seen in fig.~\ref{fig:barointeraction}.\\
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.4]{images/baro_170105_new_medium.png}
+        \caption{Rendering of the application of the barostat algorithm. The red particles are incoming gas particles, that have not yet interacted, 
+            pink particles are currently interacting and green particles are leaving the simulation box and
+        moving away from the nano crystal (blue). To render this image, a cut has been made to make a view into the simulation box possible.}
+        \label{fig:barointeraction}
+    \end{center}
+\end{figure}
+Since the gas particles are interacting with the nano crystal, the temperatures and energies are values of interest. The energies are recorded using
+histograms and the temperature is calculated instantaneously every 100th timestep.\\
+The calculation of the energies of the incoming particles has to be done with care. The problem are particles that enter the simulation box, but fly
+by the nano crystal at a distance that is greater than the interaction range. Since the change of the energy is the interesting part, only the energy 
+particles of that \textit{do} interact with the nano crystal shall be calculated. This can be achieved by counting particles that enter the stage of
+\textit{interacing}
 

### 2017-01-19 doc/draft1.tex
new plots, new updates

@@
 particles of that \textit{do} interact with the nano crystal shall be calculated. This can be achieved by counting particles that enter the stage of
-\textit{interacing}
+\textit{interacing}. One such histogram can be seen in fig.~\ref{fig:histogram}.
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.4]{images/histogram.pdf}
+        \caption{Histogram of total energies for incoming (red) and outgoing (blue) particles. The parameters for this histogram are $\Delta Q^*=0.04$,
+        $T_\text{ambient}^*=0.08$, $P^*=0.8$. What can be seen here is that the outgoing particles have a higher count of particles with more total
+    energy.}
+        \label{fig:histogram}
+    \end{center}
+\end{figure}
 

### 2017-01-25 doc/draft2.tex
work work work

--- 2017-01-19 doc/draft1.tex
@@
 In the experiment, a silica nano particle with a radius of about 75 nm and mass of about $3 \times 10^{-18}$kg is trapped in a laser beam within a
-vacuum chamber. The trapping of the silica nano particle (which will be referred to as \textit{glass particle}) is achieved by a gradient force of the
+vacuum chamber. The trapping of the silica nano particle is achieved by a gradient force of the
 laser beam acting on the particle. The experimental setup is depicted in fig.~\ref{fig:setup}.\\
 The particle fluctuates within the trap in all three spatial directions. These fluctuations can be approximated
-such that they are decoupled, which means that they can be described by a 1-dimensional Langevin equation:
+such that they are decoupled, which means that they can be described by a one-dimensional Langevin equation:
 \begin{equation}
@@
 of the particle is modified by a parametric drive. Both the parametric drive and the feedback are turned off and -- as in the first part -- the
-motion of the particle towards an equilibrated state is observerd.}
+motion of the particle towards an equilibrated state is observerd. Figure taken from \cite{Gieseler2014}.}
         \label{fig:setup}
@@
 \end{figure}
-On the left hand side we have the friction coefficient $\Gamma_0$ and the angular frequency that describes the fluctuation along the chosen axis. On
+On the left hand side we have the position $x$ (and its derivatives $\dot{x}$, the velocity and $\ddot{x}$, the acceleration), the friction 
+coefficient $\Gamma_0$ and the angular frequency $\Omega_0$ that describes the fluctuation along the chosen axis. On
 the right hand side, there are two forces. The first one is $F_\text{fluct}$, which describes a stochastic force caused by interactions with the gas
@@
 \begin{equation}
+    \label{eq:ffluct}
     F_\text{fluct} = \sqrt{2m\Gamma_0k_BT_0} \ \xi\left(t\right)
 \end{equation}
-where $T_0$ is the temperature of the heat bath (i.e. the surrounding gas in the vacuum chamber), $k_B$ is the Boltzmann constant and $\xi(t)$ is
+where $T_0$ is the temperature of the heat bath (i.e., the surrounding gas in the vacuum chamber), $k_B$ is the Boltzmann constant and $\xi(t)$ is
 white noise, which obeys the equations $\left\langle\xi(t)\right\rangle= 0$ and $\left\langle\xi(t)\xi(t')\right\rangle= \delta(t-t')$, which means
-that it is a random force. The term $\Gamma_0$ appears in the formulat due to the fluctuation-dissipation theorem, which links the damping rate to the
-stochastic force.\\
-The external force $F_\text{ext}$ is part of the experimental setup, where the frequency of the fluctuation along an axis, $\Omega_0$, is measured and
-used to suppress the motion along said axis. This causes a decrease in the particle fluctuations and thus acts as a cooling mechanism for the particle
-in the trap. This process creates a non-equilibrium steady state $\rho_{ss}(u,\alpha)$, which is not known analytically. This state is the starting
-point of this thesis.\\
+that it is a random force. The term $\Gamma_0$ appears in the formula \eqref{eq:ffluct} 
+due to the fluctuation-dissipation theorem, which links the damping rate to the stochastic force.\\
+%The external force $F_\text{ext}$ is part of the experimental setup, where the frequency of the fluctuation along an axis, $\Omega_0$, is measured and
+%used to suppress the motion along said axis. This causes a decrease in the particle fluctuations and thus acts as a cooling mechanism for the particle
+%in the trap. This process creates a non-equilibrium steady state $\rho_{ss}(u,\alpha)$, which is not known analytically. This state is the starting
+%point of this thesis.\\
 
@@
 The problem at hand can be studied on an atomic level with the use of computer simulation. There is a variety of methods for computer simulations
-that are widely used, one of which being Molecular Dynamics (MD) simulations. The following section will give a brief overview of the concepts of this
-method, which is followed by the application to the simulation of the experiment.
+that are widely used, one of which being molecular dynamics (MD) simulations.\\ 
+The goal is to simulate the experiment depicted in Fig.~\ref{fig:setup} as accurately as possible. To achieve this, the setup has to be broken down
+into individual pieced that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, sitting on an FCC
+lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, as it acts as a trapping device as well
+as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, that acts as a thermostat in the simulation, that will be
+modeled as a box surrounding the nano particle. This box is filled with gas particles that interact with the nano particle via a soft-sphere potential
+and do not interact with one another.\\
+The following section will give a brief overview of the concepts used to simulate every part of the experiment.
 
 \subsection{Molecular Dynamics}
-Molecular Dynamics\cite{Frenkel2001} simulations is a technique for simulating, as the name suggests, the dynamics of a classical many-body system. In this case,
+Molecular dynamics~\cite{Frenkel2001} simulations is a technique for simulating, as the name suggests, the dynamics of a classical many-body system. In this case,
 classical means, that the trajectories of the individual particles are calculated using classical mechanics rather then quantum mechanics. For
 relatively big atoms/molecules this is a very good approximation, whereas for systems consisting of hydrogen or helium the effects of quantum
-mechanics cannot be neglected and other methods (such as ab-initio simulation) has to be used.\\
+mechanics cannot be neglected and other methods have to be used.\\
 The dynamics of the system are obtained by solving Newton's equations of motion for every particle.
 
-\subsection{The Glass Particle}
+\subsection{The Nano Particle}
 The glass particle from the experiment will be modeled as a system of particles interacting via a Lennard-Jones pair potential, 
@@
 \end{figure}
-Since $\varepsilon$ and $\sigma$ are crucial parameters for the simulation and do (usually) not change over time, it is practical to use them to
+Since $\varepsilon$ and $\sigma$ are crucial parameters for the simulation and do not change over time, it is practical to use them to
 define the dimensions of the system. This means that the unit of distance is $\sigma$, the unit of energy is $\varepsilon$ and the unit of mass is 
@@
 \end{equation}
-Since the reduced units will be used throughout the rest of this thesis, i will drop the asterisk henceforth.\\
+Since the reduced units will be used throughout the rest of this thesis, I will drop the asterisk henceforth.\\
 From the Lennard-Jones potential the corresponding force can be calculated by taking the derivative with respect to the direction of interest:
@@
 The initial configuration of the particles is a face centered cubic (FCC) lattice. A schematic of the FCC lattice is depicted in fig. \ref{fig:fcc}.
-With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since once FCC cell (as
-depicted) shares its atoms with its next neighbours, the number of atom per unit cell is 4, as depicted in fig. \ref{fig:wiki}. The whole system of
-atoms is then created by repeating this cell structure. One convenient way is to arrange the unit cells in a cubic system, so if there are $M$ FCC
-unit cells on one edge, the whole system consists of $M^3$ cells. Since there are 4 particles per cell, there are ideal or so called \textit{magic
-numbers} for atoms for which this setup works perfectly: $N = 4M^3 = 4,32,108,256,500,864,\ldots$.\\
+With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since once FCC cell 
+shares its atoms with its next neighbours, the number of atom per unit cell is 4 -- 1/8 of a particle on eight corners and 1/2 of a particle on six faces.
+The whole system of atoms is then created by repeating this cell structure. One convenient way is to arrange the unit cells in a cubic system, 
+so if there are $M$ FCC unit cells on one edge, the whole system consists of $M^3$ cells. Since there are 4 particles per cell, 
+there are ideal or so called \textit{magic numbers} for atoms for which this setup works perfectly: $N = 4M^3 = 4,32,108,256,500,864,\ldots$.\\
 \begin{figure}[h]
     \begin{center}
-        \begin{subfigure}[t]{0.3\textwidth}
+        \begin{subfigure}[t]{0.4\textwidth}
             \includegraphics[scale=0.3]{images/fcc-2.pdf}
@@
         \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.15]{images/fcc_unit.png}
-            \caption{Schematic of the number of particles contained in one FCC unit cell. The atoms on the edges are shared by 8 cells and the ones 
-                on the faces by 2 cells\cite{unitcell}}
-            \label{fig:fcccut}
-        \end{subfigure} 
-        \
         %\hspace{60pt}
-        \begin{subfigure}[t]{0.3\textwidth}
+        \begin{subfigure}[t]{0.4\textwidth}
             \includegraphics[scale=0.3]{images/unit_cell.pdf}
@@
     p_1 &=& \{0,0,0\}\\
-    p_2 &=& \{0.5,0.5,0\}\\
-    p_3 &=& \{0.5,0,0.5\}\\
-    p_4 &=& \{0,0.5,0.5\}
+    p_2 &=& \{0.5*a,0.5*a,0\}\\
+    p_3 &=& \{0.5*a,0,0.5*a\}\\
+    p_4 &=& \{0,0.5*a,0.5*a\}
 \end{eqnarray*}
-From the particle number $N$ and the number of FCC unit cells per edge $M$ the lattice constant $a$ can be calculated
+From the particle number $N$ and the number of FCC unit cells per edge $M$ the \textit{lattice constant} $a$ can be calculated
 \begin{equation}
@@
 \label{section:verlet}
-When we look at the system from a thermodynamical standpoint, we see that it follows some kind of path in the phase space as time progresses. Every
+When we look at the system from a microscopic standpoint, we see that it follows some kind of path in the phase space as time progresses. Every
 point in this space corresponds to a set of positions and momenta and the connection between two points corresponds to the evolution of the system
-from one state to another. As mentioned above, this evolution (the dynamics of the system) is a crucial element to Molecular Dynamics. Since the
+from one state to another. As mentioned above, this evolution (the dynamics of the system) is a crucial element to molecular dynamics. Since the
 equations of motion cannot be solved analytically in general, we need to approximate the solution.\\
@@
 One algorithm that has all of the above mentioned features is the one proposed by Verlet \cite{Verlet1967}. In his paper, Verlet starts by Taylor
-expanding the coordinate mathbftor $\mathbf{r}_i$ for one particle after one time step $\Delta t$:
+expanding the coordinate vector $\mathbf{r}_i$ for one particle after one time step $\Delta t$:
 \begin{equation}
@@
 \end{equation}
-Since the first derivative of the coordinate mathbftor is the velocity, $\dot{\mathbf{r}}_i(t) = \mathbf{v}_i(t)$, and the second derivative of the coordinate 
-mathbftor is the acceleration, $\ddot{\mathbf{r}}_i(t) = \mathbf{a}_i(t)$, using Newton's law $\mathbf{F}_i(t) = m_i \mathbf{a}_i(t)$ equation \eqref{eq:verlet} 
+Since the first derivative of the coordinate is the velocity, $\dot{\mathbf{r}}_i(t) = \mathbf{v}_i(t)$, and the second derivative of the coordinate 
+is the acceleration, $\ddot{\mathbf{r}}_i(t) = \mathbf{a}_i(t)$, using Newton's law $\mathbf{F}_i(t) = m_i \mathbf{a}_i(t)$ equation \eqref{eq:verlet} 
 can be written as:
@@
 computed in the process, but for the calculation of i.e. the kinetic energy the velocities are needed. Thus, the velocities have to be calculated with
-a combination of the Taylor expansions of the coordinate mathbftors at $t+\Delta t$ and $t-\Delta t$:
+a combination of the Taylor expansions of the coordinate vectors at $t+\Delta t$ and $t-\Delta t$:
 \begin{equation}
@@
         \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2\\
-        \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \left(\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\right) \Delta t
+        \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \Big[\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\Big] \Delta t
     \end{aligned}
 \end{equation}
-This algorithm is self starting, uses a small amount of memory, conserves the energy and gives a very good approximation to the originial trajectory
+This algorithm is self starting, uses a small amount of memory, conserves the energy (not exactly) 
+and gives a very good approximation to the originial trajectory
 in the phase space. To program this algorithm the following steps are needed: 
 \begin{enumerate}
-    \item Calculate all the forces between the particles (for the first run only)
+    \item Calculate all the forces between the particles (for the first step only)
     \item Calculate the new positions with current velocity and forces
@@
 \end{enumerate}
-The first point only has to be carried out for the first run, because the forces have not been calculated at this point. As the forces are calculated
+The first point only has to be carried out for the first step, because the forces have not been calculated at this point. As the forces are calculated
 for the new positions in step 4, they can be used for the next time step. For the first time step the velocities have to be chosen randomly and are 
@@
 
-\subsection{The Laser Beam - Energy influx}
-The glass sphere is trapped in the laser beam. While the motion of the center of mass is suppressed, which acts as a cooling mechanism, the
+\subsection{The Laser Beam - Energy Influx}
+The nano particle is trapped in the laser beam. While the motion of the center of mass is localized, the
 individual atoms that make up the glass sphere absorb the energy from the laser which increases their velocity.\\
-To simulate this kind behaviour, thermostat algorithms \cite{Huenenberger2005} such as the Nos\'e-Hoover \cite{Nose1984,Hoover1985} or the Berendsen
-\cite{Berendsen1984} algorithms.\\
+To simulate this kind of behaviour, thermostat algorithms \cite{Huenenberger2005} such as the Nos\'e-Hoover \cite{Nose1984,Hoover1985} or the Andersen
+\cite{Andersen1980} algorithms are often used in simulation to change the temperature of the system in a controllable way.\\
 The problem with these kind of algorithms is, that a target temperature has to be fixed which will be reached at some point. In order to simulate the
@@
 a heat source and the other as a heat sink. A specific amount of heat is then exchanged between those two reservoirs. As it turns out however, this
-algorithm introduces an energy shift for longer simulation times. This led Wirnsberger et al \cite{Wirnsberger2015} to revisit the algorithm and 
+algorithm introduces an energy shift for longer simulation times. This led Wirnsberger et al. \cite{Wirnsberger2015} to revisit the algorithm and 
 identify the cause of this energy drift, to create a more suitable algorithm, which they called eHEX.\\
@@
 \end{figure}
-The simulation box, denoted by $\Omega$, and the regions $\Gamma_k$ are assumed to be moving with velocities $v_\Omega$ and $v_{\Gamma_k}$
-respectively.\\
+The centers of mass of the particles in simulation box, denoted by $\Omega$, and the regions $\Gamma_k$ 
+are assumed to be moving with velocities $v_\Omega$ and $v_{\Gamma_k}$ respectively.\\
 The change of the energy in a region $\Gamma_k$ is achieved by rescaling the velocities by a factor $\xi_k$ and shifted by the velocity of the
@@
 \end{subequations}
-This algorithm can be applied to the problem of the levitating nano sphere in the laser beam by adjusting the setup and some of the parameters.\\
+This algorithm can be adapted for the problem of the levitating nano sphere in the laser beam by adjusting the setup and some of the parameters.\\
 Firstly, the setup of the heat sinks and sources has to be changed. Since there is only energy pumped into the system from the outside, the region
 acting as a heat sink vanishes and the region acting as a heat source spans over the whole simulation box. This means that, using the 
-notation of fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither the simulation box nor the heat source are moving, i.e. 
+notation of fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither center of mass of the particles the simulation box nor 
+the heat source are moving, i.e. 
 $\mathbf{v}_\Omega = \mathbf{v}_{\Gamma_1} = 0$. This affects the non-translational kinetic energy term $\mathcal{K}_{\Gamma_1}$, the thermostatting
-force $\boldsymbol{\eta}$ and the correction term $\mathcal{E}\mathbf{r}$. Since there is only one region acting as a heat sink, the terms $\Delta Q$,
+force $\boldsymbol{\eta}$ and the correction term $\mathcal{E}\mathbf{r}$. Since there is only one region acting as a heat source, the terms $\Delta Q$,
 $\mathcal{K}$ and $\mathcal{F}$ don't need an index. The summation index in \eqref{eq:kineticK} and \eqref{eq:bigeps} can be changed to the number of
@@
 \end{eqnarray}
-This means that the laser is modeled to be pumping energy into the system, which increases the velocities of the velocities of the particles in the
+This means that the laser is modeled to be pumping energy into the system, which increases the velocities of the particles in the
 system over time, while the center of mass motion is not affected by this.\\
@@
 \subsection{The Laser Beam - Trapping}
-As mentioned above, the glass particle is trapped in the laser beam and the motion of the particle is suppressed in every direction. This behaviour
+As mentioned above, the glass particle is trapped in the laser beam and the position of the particle is localized. This behaviour
 has to be modeled as well.\\
@@
 \begin{eqnarray}
-    \mathbf{F} &=& -k x\\
-    U &=& \frac12 k x^2
+    \mathbf{F} &=& -k \Big[\mathbf{x}-\mathbf{x}_0\Big]\\
+    U &=& \frac12 k \Big[\mathbf{x}-\mathbf{x}_0\Big]^2
 \end{eqnarray}
-The trap is implemented by using the velocity Verlet algorithm with the above force to propagate the glass particle. Since the trap is acting on the
-whole system, the force is acting on the center of mass position and velocity. To calculate the center of mass positions and velocities, the positions
-and velocities of all particles have to be summed up and divided by the particle number:
-\begin{eqnarray}
-    \mathbf{r}_\text{COM} &=& \sum_{i=1}^N \frac{\mathbf{r}_i}{N}\\
-    \mathbf{v}_\text{COM} &=& \sum_{i=1}^N \frac{\mathbf{v}_i}{N}
-\end{eqnarray}
-The position of the center of mass and the center of mass velocity are then propagated by using the same procedure as introduced in
-section~\ref{section:verlet}.
+where $\mathbf{x}_0$ is the position of minimal potential energy. 
+Since the trap is acting on the whole system, the force is acting on the center of mass. 
+To calculate the center of mass positions and velocities, the positions and velocities of all particles have to be summed up and 
+divided by the particle number:
+\begin{equation}
+    \mathbf{r}_\text{COM} = \frac1N\sum_{i=1}^N \mathbf{r}_i
+\end{equation}
 
-\subsection{Surrounding Gas - Barostat}
-Without any equilibrating mechanism the setup described by now would lead to the system melting, which is not desirable. In the next step we will 
-introduce the surrounding gas of the gas chamber that will absorb some of the energy in the system, leading to the final state.\\
-Generally, pressure is introduced to the system by surrounding the object of interest (in this case the glass nano sphere) with a pressure medium.
+\subsection{Surrounding Gas - Thermostat}
+Without any equilibrating mechanism the setup described by now would lead to the system heating up indefinetly, which is not desirable. Furthermore,
+the goal is to recreate the experiment as close as possible. Since the nano particle is trapped in a laser beam within a vacuum chamber, the
+surrounding gas has to be modeled as well. In the next step we will introduce the surrounding gas of the gas chamber that will absorb some of the 
+energy in the system, leading to the final state.\\
+Generally, pressure is introduced to the system by surrounding the object of interest (in this case the glass nano sphere) with a thermostatting
+pressure medium.
 There are two main requirements for the choice of such a pressure medium: the exerted pressure must be hydrostatic and the computation of the
@@
 The model used in this thesis was developed by Gr\"unwald and Dellago \cite{Gruenwald2006} and uses an ideal gas of non-interacting particles as
-pressure medium. The particles of this pressure medium flow into the simulation from an outside volume, whose geometry is based on the form of the
-object of interest (this will be referred to as the minimal volume of cells). This barostat also acts as a thermostat and thus is suitable for the 
-application in this thesis.\\
+thermostatting pressure medium. The particles of this pressure medium flow into the simulation from an outside volume, whose geometry is based on the form of the
+object of interest (this will be referred to as the minimal volume of cells) and leave the simulation box if their position reaches the boundary of
+this minimal volume of cells. This behaviour is very close to the real experiment which makes this thermostat an ideal candidate for the simulation of
+the experiment. 
+%This barostat also acts as a thermostat and thus is suitable for the 
+%application in this thesis.\\
 The gas particles interact with the object via a soft-sphere potential of the form
@@
 partners increases, which increases the number of force calculations which are a very time consuming part. For smaller $\sigma$ the possibility for
-gas particles reaching the inside of the nano crystal increases, which is not desirable. So $\sigma$ should be chosen small enough to keep the force
-calculations at a minimum and large enough for the particle to stay on the outside of the crystal.\\ 
+gas particles reaching the inside of the nano particle increases, which is not desirable. So $\sigma$ should be chosen small enough to keep the force
+calculations at a minimum and large enough for the particle to stay on the outside of the crystal. In the computer simulations performed for this
+thesis, the interaction length is chosen to be $\sigma = 1$ with a cut-off radius of $r_c = 2.5$.\\ 
 The algorithm can be performed by following these steps:
@@
 \end{enumerate}
-This algorithm is formulated in a general fashion, so that a wide range of crystals and geometries can be used. The setup used in this thesis is a
-FCC lattice, which has a nice symmetry that is not influenced by the surrounding gas. The algorithm will therefore not be carried out using 
-cell-lists and a minimal volume of cells that is changing over the course of the simulation, but rather a fixed setup of the volume surrounding the 
-nano crystal.\\ 
-Since the crystal itself is a cube a straighforward approach is surrounding it by a bigger cube. The distance between the cube face and the nano
-crystal is chosen to be the side length of the nano crystal, so the setup will properly scale for different numbers of particles. 
-The schematic setup for this is depicted in fig.\ref{fig:gascube}.
+This algorithm is formulated in a general fashion, so that a wide range of crystals and geometries can be used. 
+%The setup used in this thesis is a FCC lattice, which has a nice symmetry that is not influenced by the surrounding gas. 
+Although the shape of the nano particle changes over the course of the simulation, it is not necessary to change the geometry of the surrounding box.
+The algorithm will therefore not be carried out using cell-lists and a minimal volume of cells that is changing over the course of the simulation, 
+but rather a fixed setup of the volume surrounding the nano particle.\\ 
+Since the particle itself is modeled as a cube, a straighforward approach is surrounding it by a bigger cube. The distance between the cube face and the nano
+particle is chosen to be the side length of the nano particle, so the setup will properly scale for different numbers of particles. 
+The schematic setup for this is depicted in Fig.\ref{fig:gascube}.
 \begin{figure}[h]
     \begin{center}
-        \includegraphics[scale=0.5]{images/gas_volume_ver2.pdf}
-        \caption{Schematic setup for the barostat. The outer volume (blue) is one side length $L$ away from the nano crystal (black). 
-                    The gas particles are created along the blue lines of the outer volume. Since the geometry of the nano crystal does not change, 
+        \includegraphics[scale=0.5]{images/gas_volume_ver3.pdf}
+        \caption{Schematic setup for the barostat. The outer volume (blue) is one side length $s$ away from the nano particle (black) which means that
+            the side length of the face of the surrounding box is $L=3s$. 
+                    The gas particles are created along the blue lines of the outer volume. Since the overall geometry of the nano particle does 
+                    not change, 
                     the geometry of the outer volume will be constant throughout the simulation.}
@@
 The first element of the simulation is the setup of the FCC lattice on which the particles are placed on. To check if this is done correctly, we can
-use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano crystal 
+use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano particle 
 can be seen in fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
@@
         \includegraphics[scale=0.2]{images/fcc_rendering_layer_newcoloring.png}
-        \caption{A rendered image of the FCC crystal. The chosen color-coding along the (1,1,1) Miller-index 
+        \caption{A rendered image of the FCC lattice. The chosen color-coding along the (1,1,1) Miller-index 
                 highlights the layers for this perspective. The perspective is
-                chosen to see the layering of the FCC crystal with its 3 layer structure. The black box acts as a reference point to outline the 
+                chosen to see the layering of the FCC lattice with its 3 layer structure. The black box acts as a reference point to outline the 
                 geometry of the system.}
@@
 \subsection{Velocity Verlet}
-The simulation is taking place in a NVE ensemble from a thermodynamic point of view. This means that the particle number $N$, the volume $V$ and the
+The velocity Verlet part of the simulation is taking place in a NVE ensemble from a thermodynamic point of view. 
+This means that the particle number $N$, the volume $V$ and the
 total energy $E$ are constant. In contrast to the NVT ensemble used in Monte-Carlo simulations this means, that the velocities of the particle can not
@@
 \end{eqnarray}
-The next adjustment is a small one that makes a difference in the long run. When choosing random velocities, the system not only tends to have a
-resulting velocity vector that causes the whole system to float away, but also a resulting angular momentum that causes the system to rotate.
-The angular momentum can be defined via the moment of inertia $\mathbf{I}$
-\begin{eqnarray}
-    \label{eq:angularmomentum}\mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
-    \boldsymbol{\omega} &=& \frac{\mathbf{r} \times \mathbf{v}}{r^2}
-\end{eqnarray}
-where $\boldsymbol{\omega}$ is the angular velocity of the system. Another way to define the angular momentum is given by
-\begin{equation}
-    \label{eq:angularmomentum2}
-    \mathbf{L} = \mathbf{r} \times \mathbf{p}
-\end{equation}
-where $\mathbf{p}$ is the momentum of the particle, which in this case is equal to its velocity since the mass is set to 1.The moment of inertia 
-$\mathbf{I}$ for a filled cube is given by 
-\begin{equation}
-    \mathbf{I} = 
-    \begin{pmatrix}
-        \frac{mL^2}{6} & 0 & 0\\
-        0 & \frac{mL^2}{6} & 0\\
-        0 & 0 & \frac{mL^2}{6} 
-    \end{pmatrix}
-\end{equation}
-where $m$ will be later set to 1 and $L$ is the side length of the cube. To use this equations to suppress the rotation of the cube, we need to use
-equation~\eqref{eq:angularmomentum} and rewrite it as
-\begin{equation}
-    \label{eq:omega}
-    \boldsymbol{\omega} = \mathbf{I}^{-1}\mathbf{L}
-\end{equation}
-where $\mathbf{I}^{-1}$ is the inverse matrix of $\mathbf{I}$. The calculation of the inverse matrix is fairly straightforward, as it only contains
-values in the diagonal, which means that the inverse matrix can be calculated by simply inverting these values. The angular momentum for every
-particle $\mathbf{L}_i$ in the above equation be calculated using eq.~\eqref{eq:angularmomentum2}, where the position is taken to be the distance
-between the particle and the center of mass
-\begin{equation}
-    \Delta\mathbf{r}_i = \mathbf{r}_\text{COM} - \mathbf{r}_i
-\end{equation}
-The values of the angular momentum for each paricle are added up and multiplied by the inverse of the moment of inertia, as described in
-eq.~\eqref{eq:omega}. The final step is updating the velocities of the particles:
-\begin{equation}
-    \mathbf{v}_i^\text{new} = \mathbf{v}_i - \left(\boldsymbol{\omega} \times \Delta\mathbf{r}_i\right)
-\end{equation}
+%The next adjustment is a small one that makes a difference in the long run. When choosing random velocities, the system not only tends to have a
+%resulting velocity vector that causes the whole system to float away, but also a resulting angular momentum that causes the system to rotate.
+%The angular momentum can be defined via the moment of inertia $\mathbf{I}$
+%\begin{eqnarray}
+    %\label{eq:angularmomentum}\mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
+    %\boldsymbol{\omega} &=& \frac{\mathbf{r} \times \mathbf{v}}{r^2}
+%\end{eqnarray}
+%where $\boldsymbol{\omega}$ is the angular velocity of the system. Another way to define the angular momentum is given by
+%\begin{equation}
+    %\label{eq:angularmomentum2}
+    %\mathbf{L} = \mathbf{r} \times \mathbf{p}
+%\end{equation}
+%where $\mathbf{p}$ is the momentum of the particle, which in this case is equal to its velocity since the mass is set to 1.The moment of inertia 
+%$\mathbf{I}$ for a filled cube is given by 
+%\begin{equation}
+    %\mathbf{I} = 
+    %\begin{pmatrix}
+        %\frac{mL^2}{6} & 0 & 0\\
+        %0 & \frac{mL^2}{6} & 0\\
+        %0 & 0 & \frac{mL^2}{6} 
+    %\end{pmatrix}
+%\end{equation}
+%where $m$ will be later set to 1 and $L$ is the side length of the cube. To use this equations to suppress the rotation of the cube, we need to use
+%equation~\eqref{eq:angularmomentum} and rewrite it as
+%\begin{equation}
+    %\label{eq:omega}
+    %\boldsymbol{\omega} = \mathbf{I}^{-1}\mathbf{L}
+%\end{equation}
+%where $\mathbf{I}^{-1}$ is the inverse matrix of $\mathbf{I}$. The calculation of the inverse matrix is fairly straightforward, as it only contains
+%values in the diagonal, which means that the inverse matrix can be calculated by simply inverting these values. The angular momentum for every
+%particle $\mathbf{L}_i$ in the above equation be calculated using eq.~\eqref{eq:angularmomentum2}, where the position is taken to be the distance
+%between the particle and the center of mass
+%\begin{equation}
+    %\Delta\mathbf{r}_i = \mathbf{r}_\text{COM} - \mathbf{r}_i
+%\end{equation}
+%The values of the angular momentum for each paricle are added up and multiplied by the inverse of the moment of inertia, as described in
+%eq.~\eqref{eq:omega}. The final step is updating the velocities of the particles:
+%\begin{equation}
+    %\mathbf{v}_i^\text{new} = \mathbf{v}_i - \left(\boldsymbol{\omega} \times \Delta\mathbf{r}_i\right)
+%\end{equation}
 \\The velocities of the particle can roughly be chosen by adjusting the range of numbers the random number generator can choose from. However, the
@@
 \begin{eqnarray}
- \label{eq:rescale1}   E &=& \sum_{i=1}^N \mathbf{v}_i^2 \\
- \label{eq:rescale2}   \lambda &=& \sqrt{\frac{3(N-1)T}{E}} \\
+\label{eq:rescale1}   E &=& \sum_{i=1}^N \frac{\mathbf{v}_i^2}{2} \\
+ \label{eq:rescale2}   \lambda &=& \sqrt{\frac{3(N-1)T}{2E}} \\
  \label{eq:rescale3}   \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i  
@@
 
-\subsection{Barostat}
+\subsection{Thermostat}
 The barostat, as described previously, keeps the system from overheating. The barostat algorithm is applied at the same time as the eHEX algorithm,
 after the velocity Verlet has equilibrated the system.\\
-While the eHEX algorithm is pumping energy into the nano crystal while the surrounding gas particles absorb excessive energy. This leads to the
-development of a new, higher steady temperature of the nano crystal, which can be seen in fig.~\ref{fig:baro1}. The first simulations were performed
-with the eHEX and the barostat algorithm starting at the same point. This lead to the question: does the new steady temperature depend on the time it
-takes the gas particles to reach the system? The answer is: no. After applying the barostat algorithm first to fill the box with gas atoms and then
-applying the eHEX algorithm it can be seen that the resulting temperature is the same, which can be seen in fig.~\ref{fig:baro2}.
+While the eHEX algorithm is pumping energy into the nano particle while the surrounding gas particles absorb excessive energy. This leads to the
+development of a new, higher steady temperature of the nano particle, which can be seen in Fig.~\ref{fig:baro1}. 
+%The first simulations were performed
+%with the eHEX and the barostat algorithm starting at the same point. This lead to the question: does the new steady temperature depend on the time it
+%takes the gas particles to reach the system? The answer is: no. After applying the barostat algorithm first to fill the box with gas atoms and then
+%applying the eHEX algorithm it can be seen that the resulting temperature is the same, which can be seen in fig.~\ref{fig:baro2}.
 %This means that the system is beginning to heat up, until the particles reach the crystal.
@@
 \end{figure}
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/baro_temp_compare_other.pdf}
-        \caption{Two versions of the application of the eEHX and barostat algorithm. The shorter run (red, x-axis on the bottom) represents 
-            the simultaneous start of the eHEX and the barostat algorithm. In the longer run (blue, x-axis on the top) there are 20000 steps, where
-        the barostat algorithm is applied at the same time as the velocity Verlet to give the system time to equilibrate and only after the eHEX
-    algorithm is taking the place of the velocity Verlet. The mean value of the resulting temperature is the same, indicating that the final
-temperature is based on the values of the parameters, such as pressure and ambient temperature, rather than geometry.}
-        %\caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
-        %in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
-    %an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
-%reached and the temperature seems to stay at this value.}
-        \label{fig:baro1}
-    \end{center}
-\end{figure}
-As we are interested in the temperatures of the particles before and after the interaction with the crystal, we need some way to differentiate those
+%\begin{figure}[h]
+    %\begin{center}
+        %\includegraphics[scale=0.4]{images/baro_temp_compare_other.pdf}
+        %\caption{Two versions of the application of the eEHX and barostat algorithm. The shorter run (red, x-axis on the bottom) represents 
+            %the simultaneous start of the eHEX and the barostat algorithm. In the longer run (blue, x-axis on the top) there are 20000 steps, where
+        %the barostat algorithm is applied at the same time as the velocity Verlet to give the system time to equilibrate and only after the eHEX
+    %algorithm is taking the place of the velocity Verlet. The mean value of the resulting temperature is the same, indicating that the final
+%temperature is based on the values of the parameters, such as pressure and ambient temperature, rather than geometry.}
+        %%\caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
+        %%in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
+    %%an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
+%%reached and the temperature seems to stay at this value.}
+        %\label{fig:baro1}
+    %\end{center}
+%\end{figure}
+As we are interested in the temperatures of the particles before and after the interaction with the nano particle, we need some way to differentiate those
 two states. This can be done when looking at the total energy of the gas particles.\\
 The gas particles are created on the surface of the simulation box, which is further away than the interaction range between the gas particles and the
-nano crystal. Since the gas particles do not interact with each other, the total energy of the gas particles does not change until they reach the
-interaction range of the nano crystal. There they undergo a change of energy due to the interaction with the nano crystal and then make their way away
-from the nano crystal. After the interaction, when they reach a point outside of the interaction range, the enrgy stays constant again until the
+nano particle. Since the gas particles do not interact with each other, the total energy of the gas particles does not change until they reach the
+interaction range of the nano particle. There they undergo a change of energy due to the interaction with the nano particle and then make their way away
+from the nano particle. After the interaction, when they reach a point outside of the interaction range, the enrgy stays constant again until the
 particles reach the limits of the simulation box.\\
@@
             pink particles are currently interacting and green particles are leaving the simulation box and
-        moving away from the nano crystal (blue). To render this image, a cut has been made to make a view into the simulation box possible.}
+        moving away from the nano particle (blue). To render this image, a cut has been made to make a view into the simulation box possible.}
         \label{fig:barointeraction}
@@
 \end{figure}
-Since the gas particles are interacting with the nano crystal, the temperatures and energies are values of interest. The energies are recorded using
+Since the gas particles are interacting with the nano particle, the temperatures and energies are values of interest. The energies are recorded using
 histograms and the temperature is calculated instantaneously every 100th timestep.\\
 The calculation of the energies of the incoming particles has to be done with care. The problem are particles that enter the simulation box, but fly
-by the nano crystal at a distance that is greater than the interaction range. Since the change of the energy is the interesting part, only the energy 
-particles of that \textit{do} interact with the nano crystal shall be calculated. This can be achieved by counting particles that enter the stage of
-\textit{interacing}. One such histogram can be seen in fig.~\ref{fig:histogram}.
+by the nano particle at a distance that is greater than the interaction range. Since the change of the energy is the interesting part, only the energy 
+particles of that \textit{do} interact with the nano particle shall be calculated. This can be achieved by counting particles that enter the stage of
+\textit{interacing}. One such histogram can be seen in Fig.~\ref{fig:histogram}.
 \begin{figure}[h]
     \begin{center}
-        \includegraphics[scale=0.4]{images/histogram.pdf}
+        \includegraphics[scale=0.4]{images/histogram_2.pdf}
         \caption{Histogram of total energies for incoming (red) and outgoing (blue) particles. The parameters for this histogram are $\Delta Q^*=0.04$,

### 2017-01-25 src/functions.cpp

@@ -612,7 +612,7 @@ void InitBarostatFull(std::list<Particle*>& particles)
              */
             particles.push_back(px0);
             kinE = px0->m*(px0->v[0]*px0->v[0]+px0->v[1]*px0->v[1]+px0->v[2]*px0->v[2])*0.5;
-            //gsl_histogram_increment(gas_in,kinE);
+            gsl_histogram_increment(gas_in,kinE);
         }
 
         //side xL
@@ -635,7 +635,7 @@ void InitBarostatFull(std::list<Particle*>& particles)
              */
             particles.push_back(pxL);
             kinE = pxL->m*(pxL->v[0]*pxL->v[0]+pxL->v[1]*pxL->v[1]+pxL->v[2]*pxL->v[2])*0.5;
-            //gsl_histogram_increment(gas_in,kinE);
+            gsl_histogram_increment(gas_in,kinE);
         }   
 
         //side y0
@@ -681,7 +681,7 @@ void InitBarostatFull(std::list<Particle*>& particles)
              */
             particles.push_back(pyL);
             kinE = pyL->m*(pyL->v[0]*pyL->v[0]+pyL->v[1]*pyL->v[1]+pyL->v[2]*pyL->v[2])*0.5;
-            //gsl_histogram_increment(gas_in,kinE);
+            gsl_histogram_increment(gas_in,kinE);
         }   
         
         //side z0
@@ -705,7 +705,7 @@ void InitBarostatFull(std::list<Particle*>& particles)
              */
             particles.push_back(pz0);
             kinE = pz0->m*(pz0->v[0]*pz0->v[0]+pz0->v[1]*pz0->v[1]+pz0->v[2]*pz0->v[2])*0.5;
-            //gsl_histogram_increment(gas_in,kinE);
+            gsl_histogram_increment(gas_in,kinE);
         }   
 
         //side zL
@@ -729,7 +729,7 @@ void InitBarostatFull(std::list<Particle*>& particles)
              */
             particles.push_back(pzL);
             kinE = pzL->m*(pzL->v[0]*pzL->v[0]+pzL->v[1]*pzL->v[1]+pzL->v[2]*pzL->v[2])*0.5;
-            //gsl_histogram_increment(gas_in,kinE);
+            gsl_histogram_increment(gas_in,kinE);
         }   
 /*
  *
@@ -3456,10 +3456,10 @@ void verletBaroAccelerations(Particle* cube, std::list<Particle*> gas)
[…]

### 2017-01-25 src/globals.cpp

@@ -16,8 +16,8 @@ int g_ID=1;
 //const unsigned int N = 864;
 //const unsigned int N = 500;
 //const unsigned int N = 108;
-//const unsigned int N = 256;
-const unsigned int N = 32;
+const unsigned int N = 256;
+//const unsigned int N = 32;
 
 double* rCMStart = new double[3];
 double* center = new double[3];

### 2017-01-25 src/main.cpp

@@ -37,7 +37,7 @@ int main(int argc,char** argv)
     double t = 0.1;
     double ambient = 0.09;
     double press = 0.8;
-    double q = 0.01;
+    double q = 0.02;
     while(ambient > 0.01)
     {
         setValues(t,q,0.2,press,ambient);

### 2017-01-26 doc/draft2.tex
updates

@@
 \begin{document}
-\title{Temperature Measurements in Optical Tweezer Experiments}
+\title{Simulation of Nano Particles in a Laser Trap}
 \author{Mathias H\"old, BSc.}
@@
 %introduction of the experiment, the problem and the idea
-\section{The experiment}
 The starting point of this thesis is an experiment conducted by Gieseler et al \cite{Gieseler2014}. It is an optical tweezer experiment, where the
@@
 vacuum chamber. The trapping of the silica nano particle is achieved by a gradient force of the
-laser beam acting on the particle. The experimental setup is depicted in fig.~\ref{fig:setup}.\\
+laser beam acting on the particle. The experimental setup is depicted in Fig.~\ref{fig:setup}.\\
 The particle fluctuates within the trap in all three spatial directions. These fluctuations can be approximated
@@
 \begin{equation}
+    \label{eq:langevin}
     \ddot{x} + \Gamma_0 \dot{x} + \Omega^2_0x = \frac 1 m \left(F_\text{fluct} + F_\text{ext}\right)
@@
 %point of this thesis.\\
+\subsection{Model of the Surrounding Gas}
+J. Millen et al.~\cite{MillenJ.2014} worked on a similar experimental setup (without the feedback mechanism) and investigated the heating of the
+particle in the trap as well as the gas surrounding the particle.\\
+They start from the premise that there are 4 different temperatures one has to consider in such an experiment: the temperature of the gas particles
+before interacting with the nano particle $T_\text{imp}$ (for \textit{impinging}), the temperature of the gas particles after 
+interaction $T_\text{em}$ (for emerging), the surface temperature of the nano particle $T_\text{sur}$ and the temperature of the 
+center of mass of the nano particle $T_\text{COM}$. The experimental setup
+and the different temperatures are depicted in Fig.~\ref{fig:levitation}.\\
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[scale=0.2]{images/nnano_millen.jpg}
+        \caption{The experimental setup used by Millen et al. In \textbf{a} the experiment is schematically depicted and \textbf{b} shows a photograph
+            of the actual experimental setup. In \textbf{c} the 4 different temperatures that are present during the experiment are depicted and
+        \textbf{d} shows the position of the particle in the trap over time. The image was taken from~\cite{MillenJ.2014}}
+        \label{fig:levitation}
+    \end{center}
+\end{figure}
+Since the two temperatures of the gas, $T_\text{imp}$ and $T_\text{em}$, are not equal, the gas surrounding the nano particle is not in a thermal
+equilibrium, which is in contrast to the assumption in many optical tweezer experiments. Since the gas particles only interact with the nano particle
+in the trap and not wich each other, this situation creates two heat baths of different temperatures with the nano particle acting as a mediator
+between those two baths. This situation is depicted in
+Fig.~\ref{fig:heatbaths}.
+\begin{figure}[H]
+    \begin{center}
+        \includegraphics[scale=0.5]{images/nano_nonequilibrium_cropped.jpg}
+        \caption{Schematic picture of the two heat baths. The (hot) nano particle interacts with the (cold) incoming gas particles and they leave with
+        a higher temperature. The incoming and outgoing particles do not interact with each other and thus create two separate heat baths with
+    different temperatures. Image modified from \cite{Kroy2014}}
+        \label{fig:heatbaths}
+    \end{center}
+\end{figure}
+The one dimensional Langevin equation \eqref{eq:langevin} has to be modeled to include the different friction coefficients of the two heat baths. The
+model derived in \cite{MillenJ.2014} is described in the following equation:
+\begin{equation}
+    M\ddot{x}(t) + M\left(\Gamma^\text{imp}+\Gamma^\text{em}\right)\dot{x}(t) + M\omega^2x(t) =F^\text{imp}+F^\text{em}
+\end{equation}
+where $x$ is the position of the nano particle, $\dot{x}$ and $\ddot{x}$ are the velocity and the acceleration, respectively, $M$ is the mass,
+$\omega$ is the trap frequency, $\Gamma^\text{imp}$ and $\Gamma^\text{em}$ are the damping constants for the 
+impinging particles and the emerging particles, respectively and $F^\text{imp}$ and $F^\text{em}$ are the associated noise terms, 
+which satisfy $\langle F^\text{imp}\rangle= \langle F^\text{em} \rangle = 0$ individually.\\
+Miller et al. investigate the dependence of the emerging temperature $T_\text{em}$ on the intensity of the laser for different sized nano particles
+and different pressures.\\
+
+%These four temperatures introduced in the paper, together with the surrounding pressure and the intensity of the laser 
+%The existence of the two temperatures of the gas implies that, contrary to the assumption in many of the experiments, the gas is not in a thermal
+%equilibrium. The Langevin equation (like the one in \eqref{eq:langevin}) has to be modeled k
 
@@
 The force in the y and z direction can be calculated analogously.\\
-The initial configuration of the particles is a face centered cubic (FCC) lattice. A schematic of the FCC lattice is depicted in fig. \ref{fig:fcc}.
+The initial configuration of the particles is a face centered cubic (FCC) lattice. A schematic of the FCC lattice is depicted in Fig. \ref{fig:fcc}.
 With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since once FCC cell 
@@
 There are several ways to achieve this initial configuration and the one used in this thesis \cite{buffalo} was to create a kind of unit cell 
-consisting of four atoms, as shown in fig. \ref{fig:unitcell}, which can be described by a set of points
+consisting of four atoms, as shown in Fig. \ref{fig:unitcell}, which can be described by a set of points
 \begin{eqnarray*}
@@
 The method used here is called finite difference approach. The trajectory of the system in the phase space is cut into finite pieces of length $\Delta
-t$ and the equations of motion are solved for every segment separately (see fig. \ref{fig:finitedifference}).\\
+t$ and the equations of motion are solved for every segment separately (see Fig. \ref{fig:finitedifference}).\\
 \begin{figure}
@@
 the system and vice versa. Regions that neither act as heat source or heat sink are labelled with $\Gamma_0$, which are also called
-Hamiltonian regions (see fig. \ref{fig:ehex}). 
+Hamiltonian regions (see Fig. \ref{fig:ehex}). 
 \begin{figure}
@@
 acting as a heat sink vanishes and the region acting as a heat source spans over the whole simulation box. This means that, using the 
-notation of fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither center of mass of the particles the simulation box nor 
+notation of Fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither center of mass of the particles the simulation box nor 
 the heat source are moving, i.e. 
@@
 use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano particle 
-can be seen in fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
+can be seen in Fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
 where the atoms take the position on the gap between the atoms from the layer below.\\
@@
 This process has to be done a couple of times, until the system equilibrates to the desired temperature. The problem with this procedure is the fact
-that it contradicts the constant energy property of the ensemble. If we look at the energy values in fig.~\ref{fig:energystart} we can see the
+that it contradicts the constant energy property of the ensemble. If we look at the energy values in Fig.~\ref{fig:energystart} we can see the
 irregularities (jumps) before the system equilibrates to a constant value. This equilibration process divides the simulation in two phases: the equilibration
@@
 Since the eHEX algorithm injects heat into the system, one would expect the energy and the temperature to rise during the application of the
-algorithm. In fig.~\ref{fig:ehex_te} we can see that exactly this is happening. The graph shows a velocity Verlet phase (that follows an equilibration
+algorithm. In Fig.~\ref{fig:ehex_te} we can see that exactly this is happening. The graph shows a velocity Verlet phase (that follows an equilibration
 phase which is not recorded) of constant temperature and energy and an eHEX phase, where the temperature and the enrgy are constantly rising. 
@@
 \end{figure}
-The comparison of different values of $\Delta Q$ can be seen in fig.~\ref{fig:ehex_dq}. As the correlation between the temperature and the energy has
-been shown in fig.~\ref{fig:ehex_te}, the plotting of the temperature data should suffice. 
+The comparison of different values of $\Delta Q$ can be seen in Fig.~\ref{fig:ehex_dq}. As the correlation between the temperature and the energy has
+been shown in Fig.~\ref{fig:ehex_te}, the plotting of the temperature data should suffice. 
 \begin{figure}[h]
@@
 Therefore, the life span of a single gas particle can be partioned into 3 stages: incoming, interacting, outgoing. In a typical scene of interaction
-all three of these stages are represented by different gas particles, as can be seen in fig.~\ref{fig:barointeraction}.\\
+all three of these stages are represented by different gas particles, as can be seen in Fig.~\ref{fig:barointeraction}.\\
 \begin{figure}[h]

### 2017-01-26 src/functions.cpp

@@ -2983,7 +2983,7 @@ void trackParticle(Particle* cube, std::list<Particle*> gas, int partID, FILE* o
      {
          if((*gasIter)->ID == partID)
          {
-             fprintf(output,"%d\ntest\n",N+1);
+             //fprintf(output,"%d\ntest\n",N+1);
              /*
               *fprintf(output,"%s\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\t%lf\n",
               *        ((*gasIter)->name).c_str(),(*gasIter)->r[0],(*gasIter)->r[1],(*gasIter)->r[2],\
@@ -3019,10 +3019,12 @@ void trackParticle(Particle* cube, std::list<Particle*> gas, int partID, FILE* o
              fprintf(output,"\t%lf",(*gasIter)->potE);
              fprintf(output,"\t%lf",(*gasIter)->m*((*gasIter)->v[0]*(*gasIter)->v[0]+(*gasIter)->v[1]*(*gasIter)->v[1]+(*gasIter)->v[2]*(*gasIter)->v[2])/2.);
              fprintf(output,"\n");
-             for(unsigned int j=0;j<N;j++)
-             {
-                     fprintf(output,"%s\t\%lf\t\%lf\t%lf\n",(cube[j].name).c_str(),cube[j].r[0],cube[j].r[1],cube[j].r[2]);
-             }
+             /*
+              *for(unsigned int j=0;j<N;j++)
+              *{
+              *        fprintf(output,"%s\t\%lf\t\%lf\t%lf\n",(cube[j].name).c_str(),cube[j].r[0],cube[j].r[1],cube[j].r[2]);
+              *}
+              */
 
 
          }

### 2017-01-27 doc/draft2.tex
much changes
achtung! geschwindigkeits-initialisierung geaendert!

@@
 The velocity Verlet part of the simulation is taking place in a NVE ensemble from a thermodynamic point of view. 
-This means that the particle number $N$, the volume $V$ and the
-total energy $E$ are constant. In contrast to the NVT ensemble used in Monte-Carlo simulations this means, that the velocities of the particle can not
-be chosen directly. The initial velocities have to be drawn from a random number generator. This means that the velocities have to be adjusted after 
-they have been drawn.\\
+This means that the particle number $N$, the volume $V$ and the total energy $E$ are constant. 
+The initial velocities of the particles are chosen randomly from a Maxwell-Boltzmann distribution. To ensure that the system behaves the way it is
+intended to, small adjustments have to be made in the beginning of the simulation of the nano particle.\\
+%In contrast to the NVT ensemble used in Monte-Carlo simulations this means, that the velocities of the particle can not
+%be chosen directly. The initial velocities have to be drawn from a random number generator. This means that the velocities have to be adjusted after 
+%they have been drawn.\\
 The first adjustment that needs to be made is making sure that the system is staying in place and not floating away. To do this, the velocity vectors
-of all the atoms in the system have to be added to get the resulting velocity vector for the whole system that is thought to be acting on the center
-of mass. This resulting vector is then divided by the number of particles in the system and the resulting vector is then subtracted from every
+of all the atoms in the system have to be added to get the resulting velocity vector for the whole system. 
+This resulting vector is then divided by the number of particles in the system and the resulting vector is then subtracted from every
 velcocity vector of every atom in the system. 
@@
 \end{eqnarray}
+This adjustment only needs to be done just after randomly choosing the velocities, as there is no force present at this point that would cause the
+nano particle to move.
 %The next adjustment is a small one that makes a difference in the long run. When choosing random velocities, the system not only tends to have a
@@
 %\end{equation}
-\\The velocities of the particle can roughly be chosen by adjusting the range of numbers the random number generator can choose from. However, the
-resulting velocities (and therefore the temperature of the system) are generally not precisely how they should be. If the system should have a
+%\\The velocities of the particle can roughly be chosen by adjusting the range of numbers the random number generator can choose from. However, the
+%resulting velocities (and therefore the temperature of the system) are generally not precisely how they should be. 
+\\As mentioned above, the velocities are chosen from a Maxwell-Boltzmann distribution, which means choosing each component from a Gaussian
+distribution. However, the positions on the FCC lattice are not very natural and so the atoms equilibrate to a different temperature than desired.
+If the system should have a
 specific starting temperature $T$, the velocities need to be scaled accordingly. This rescaling is done by summing the square of the velocities
@@
 \end{eqnarray}
-This process has to be done a couple of times, until the system equilibrates to the desired temperature. The problem with this procedure is the fact
-that it contradicts the constant energy property of the ensemble. If we look at the energy values in Fig.~\ref{fig:energystart} we can see the
+This process has to be done a couple of times, until the system equilibrates to the desired temperature. 
+%The problem with this procedure is the fact
+%that it is not compatible with the requirement of constant energy of the system. 
+If we look at the energy values in Fig.~\ref{fig:energystart} we can see the
 irregularities (jumps) before the system equilibrates to a constant value. This equilibration process divides the simulation in two phases: the equilibration

### 2017-01-27 doc/expose.tex
much changes
achtung! geschwindigkeits-initialisierung geaendert!

--- 2017-01-25 doc/draft2.tex
@@
 \begin{document}
-\title{Temperature Measurements in Optical Tweezer Experiments}
+\title{Expos\'e - Simulation of Nano Particles in a Laser Trap}
 \author{Mathias H\"old, BSc.}
-\date{2016}
 \maketitle
 \thispagestyle{empty}
-\newpage
-\section{Introduction}
-%something about computer science in general, optical tweezers and the need to study the
-%systems on computers yada yada
+\section{Konzept}
+In dieser Masterarbeit wird mit Hilfe von Methoden der Computational Physics ein Experiment simuliert, in dem Nanopartikel in einer Laserfalle
+gefangen werden. Das Hauptaugenmerk dabei liegt auf der Interaktion des Nanopartikels mit dem umgebenden Gas und dem Einfluss des Lasers auf die
+Bewegung des Nanopartikels in der Falle. Die benutzte Technik in dieser Simulation nennt sich ,,molecular dynamics'' 
+und wird dazu verwendet, um die zeitliche Entwicklung des Zustandes des Nanopartikels und des Gases zu berechnen.\\
+Das Experiment besteht aus einem Nanopartikel, in diesem Fall eine Siliziumdioxid-Kugel mit einem Durchmesser von $r \approx 75 nm$, das in von einem
+Laser durch Gradientenkraft lokalisiert wird. Das Nanopartikel befindet sich in einer Vakuumkammer mit variablem Druck.\\
+Die Gasteilchen, die das Nanopartikel umgeben, interagieren nicht miteinander, wenn die mittlere freie Wegl\"ange der Gasteilchen viel gr\"o\ss er ist
+als der Radius des Nanopartikels. Das bedeutet, dass die Interaktion nur zwischen den Gas Teilchen und dem Nanopartikel in der Falle stattfindet. 
+Dadurch entstehen zwei W\"armereservoirs des umgebenden -- ein ,,kaltes'', bestehend aus Teilchen vor der Interaktion mit dem Nanopartikel und ein
+,,hei\ss es'', bestehend aus Teilchen nach der Interaktion mit dem Nanopartikel.\\
+Dem Nanopartikel selbst sind auch zwei Temperaturen zuzuweisen. Da es sich in einer Laserfalle befindet, absorbiert das Nanopartikel die Energie des
+Lasers in Form von W\"arme, die an die innere Energie der Atome, aus dem das Nanopartikel besteht, abgegeben wird. Die Temperatur, die aus dieser
+inneren Energie abzuleiten ist, wird ,,Oberfl\"achentemperatur'' genannt. Die zweite Temperatur, die dem Nanopartikel zuzuordnen ist, ist seine
+Schwerpunktstemperatur -- jene Temperatur, die aus der Geschwindigkeit der Schwerpunktsbewegung berechnet wird. Durch die Lokalisierung in der
+Laserfalle ist wird diese Bewegung minimiert. \\
+Die Simulation dieses Experiments wird durch Modellierung der Einzelnen Bestandteile durchgef\"uhrt. Das Nanopartikel wird durch einen FCC Kristall
+approximiert, in dem die Teilchen miteinander \"uber ein Lennard-Jones Potential miteinander wechselwirken. Der Laser wird durch zwei unterschiedliche
+Teile modelliert. Zum einen wird die durch den Laser abgegebene und das Nanopartikel aufgenommene Energie durch den so genannten ,,eHEX'' -- enhanced
+heat exchange algorithm -- modelliert. Die Lokalisierung des Nanopartikels durch den Laser wird durch ein harmonisches Potential modelliert, das auf
+den Schwerpunkt wirkt. Das umgebende Gas wird durch einen Barostat modelliert, der in diesem Fall eher eine Funktion als Thermostat einnimmt. Die
+Bewegung der einzelnen Atome wird durch den velocity Verlet Algorithmus berechnet.\\
+Die Fragestellung, der in dieser Arbeit nachgegangen wird ist: Hat die erh\"ohte Temperatur der inneren Freiheitsgrade des Nanopartikels einen
+Einfluss auf dessen Schwerpunktsbewegung?
 
-\newpage
-\section{Motivation}
-%introduction of the experiment, the problem and the idea
-\section{The experiment}
-The starting point of this thesis is an experiment conducted by Gieseler et al \cite{Gieseler2014}. It is an optical tweezer experiment, where the
-motion of a glass nanoparticle in a laser trap was used to investigate the fluctuation theorem\cite{Crooks1999}.
-\subsection{Experimental setup}
-In the experiment, a silica nano particle with a radius of about 75 nm and mass of about $3 \times 10^{-18}$kg is trapped in a laser beam within a
-vacuum chamber. The trapping of the silica nano particle is achieved by a gradient force of the
-laser beam acting on the particle. The experimental setup is depicted in fig.~\ref{fig:setup}.\\
-The particle fluctuates within the trap in all three spatial directions. These fluctuations can be approximated
-such that they are decoupled, which means that they can be described by a one-dimensional Langevin equation:
-\begin{equation}
-    \ddot{x} + \Gamma_0 \dot{x} + \Omega^2_0x = \frac 1 m \left(F_\text{fluct} + F_\text{ext}\right)
-\end{equation}
-%The experiment is set up as follows: a glass sphere with radius $r \approx 75$ nm and mass $m \approx 3 \times 10^{-18}$ kg is trapped in a laser beam
-%in a vacuum chamber. The initial state of the glass particle is prepared by modulation of the laser, such that the oscillation of the glass particle
-%is suppressed in every direction. This is done by measuring the oscillation of the particle in every direction (assuming that the oscillation in every
-%direction is decoupled) and using this information as feedback for the laser. This experimental setup can be seen in figure \ref{fig:setup}.
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[scale=0.3]{images/experimental_setup.jpg}
-        \caption{Experimental setup of the optical tweezer experiment. A silica nano particle is trapped in a laser beam via gradient force in a
-        vacuum. The feedback is used to cool down the particle and create a non-equilibrium steady state. In the first part of the experiment, the
-    feedback is turned off and the motion of the particle towards an equilibrated state is observed. In the second part of the experiment, the steady state
-of the particle is modified by a parametric drive. Both the parametric drive and the feedback are turned off and -- as in the first part -- the
-motion of the particle towards an equilibrated state is observerd. Figure taken from \cite{Gieseler2014}.}
-        \label{fig:setup}
-    \end{center}
-\end{figure}
-On the left hand side we have the position $x$ (and its derivatives $\dot{x}$, the velocity and $\ddot{x}$, the acceleration), the friction 
-coefficient $\Gamma_0$ and the angular frequency $\Omega_0$ that describes the fluctuation along the chosen axis. On
-the right hand side, there are two forces. The first one is $F_\text{fluct}$, which describes a stochastic force caused by interactions with the gas
-in the vacuum chamber. This force is given by
-\begin{equation}
-    \label{eq:ffluct}
-    F_\text{fluct} = \sqrt{2m\Gamma_0k_BT_0} \ \xi\left(t\right)
-\end{equation}
-where $T_0$ is the temperature of the heat bath (i.e., the surrounding gas in the vacuum chamber), $k_B$ is the Boltzmann constant and $\xi(t)$ is
-white noise, which obeys the equations $\left\langle\xi(t)\right\rangle= 0$ and $\left\langle\xi(t)\xi(t')\right\rangle= \delta(t-t')$, which means
-that it is a random force. The term $\Gamma_0$ appears in the formula \eqref{eq:ffluct} 
-due to the fluctuation-dissipation theorem, which links the damping rate to the stochastic force.\\
-%The external force $F_\text{ext}$ is part of the experimental setup, where the frequency of the fluctuation along an axis, $\Omega_0$, is measured and
-%used to suppress the motion along said axis. This causes a decrease in the particle fluctuations and thus acts as a cooling mechanism for the particle
-%in the trap. This process creates a non-equilibrium steady state $\rho_{ss}(u,\alpha)$, which is not known analytically. This state is the starting
-%point of this thesis.\\
+\section{Vorl\"aufiges Inhaltsverzeichnis}
+Die Arbeit wird in folgende Kapitel unterteilt sein:
+\begin{itemize}
+    \item{Introduction}
+    \item{Motivation}
+    \item{Simulation}
+    \item{Results}
+    \item{Conclusion}
+\end{itemize}
 
-\newpage
-\section{Simulation}
-%simulation techniques and general concept of the simulation itself, used methods and so on 
-The problem at hand can be studied on an atomic level with the use of computer simulation. There is a variety of methods for computer simulations
-that are widely used, one of which being molecular dynamics (MD) simulations.\\ 
-The goal is to simulate the experiment depicted in Fig.~\ref{fig:setup} as accurately as possible. To achieve this, the setup has to be broken down
-into individual pieced that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, sitting on an FCC
-lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, as it acts as a trapping device as well
-as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, that acts as a thermostat in the simulation, that will be
-modeled as a box surrounding the nano particle. This box is filled with gas particles that interact with the nano particle via a soft-sphere potential
-and do not interact with one another.\\
-The following section will give a brief overview of the concepts used to simulate every part of the experiment.
-
-\subsection{Molecular Dynamics}
-Molecular dynamics~\cite{Frenkel2001} simulations is a technique for simulating, as the name suggests, the dynamics of a classical many-body system. In this case,
-classical means, that the trajectories of the individual particles are calculated using classical mechanics rather then quantum mechanics. For
-relatively big atoms/molecules this is a very good approximation, whereas for systems consisting of hydrogen or helium the effects of quantum
-mechanics cannot be neglected and other methods have to be used.\\
-The dynamics of the system are obtained by solving Newton's equations of motion for every particle.
-
-\subsection{The Nano Particle}
-The glass particle from the experiment will be modeled as a system of particles interacting via a Lennard-Jones pair potential, 
-\begin{equation}
-    \label{eq:lj}
-    U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right]
-\end{equation}
-where $\varepsilon$ is the depth of the potential well (and thus its unit is energy) and $\sigma$ is the distance at which the potential is zero. 
-The form of the potential and the relation to the parameters is depicted in Fig. \ref{fig:lj}.
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.2]{images/LJ-2.pdf}
-        \caption{The Lennard-Jones 12-6 potential from \eqref{eq:lj}. The x-axis is the particle distance divided by $\sigma$ and  the y-axis is the
-        potential divided by the depth of the potential well.}
-        \label{fig:lj}
-    \end{center}
-\end{figure}
-Since $\varepsilon$ and $\sigma$ are crucial parameters for the simulation and do not change over time, it is practical to use them to
-define the dimensions of the system. This means that the unit of distance is $\sigma$, the unit of energy is $\varepsilon$ and the unit of mass is 
-the mass of the simulated particle. The so called \textit{reduced units} can be constructed from these three parameters and put into relation to the
-original units. Here are some examples:
-\begin{itemize}
-    \item {distance:} $r^* = r/\sigma$
-    \item {potential energy:} $U^* = U/\varepsilon$
-    \item {temperature:} $T^* = k_B T/\varepsilon$
-    \item {time:} $t^* = t\sqrt{\varepsilon/(m\sigma^2)}$
-    \item {pressure:} $P^* = P\sigma^3/\varepsilon$
-    \item {density:} $\rho^* = \rho \sigma^3$
-\end{itemize}
-One very popular choice for the simulated atoms is Argon because it is an inert gas and the atoms behave 
-approximately like hard spheres which attract each other with weak van der Waals forces, which justifies the use of the Lennard-Jones potential. 
-Argon has a mass of m = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$m and $\varepsilon = 1.65 \times 10^{-21}$J.\\
-With the above introduced reduced units, the Lennard-Jones potential can be written as
-\begin{equation}
-    U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right].
-\end{equation}
-Since the reduced units will be used throughout the rest of this thesis, I will drop the asterisk henceforth.\\
-From the Lennard-Jones potential the corresponding force can be calculated by taking the derivative with respect to the direction of interest:
-\begin{eqnarray}
-    F_{x} &=& -\frac{\partial}{\partial x} U(r) \nonumber\\
-                &=& -\frac{\partial}{\partial x} 4\left[{r}^{-12} - {r}^{-6}\right] \nonumber\\
-                &=& -4 \left[(-12){r}^{-13} - (-6){r}^{-7}\right] \frac{\partial r}{\partial x} \nonumber\\
-                &=& 48 \left[r^{-13} - 0.5 \ r^{-7}\right] \frac{x}{r} \nonumber\\
-    \label{eq:ljforce} &=& 48 \left[r^{-14} - 0.5 \ r^{-8}\right] x
-\end{eqnarray}
-The force in the y and z direction can be calculated analogously.\\
-The initial configuration of the particles is a face centered cubic (FCC) lattice. A schematic of the FCC lattice is depicted in fig. \ref{fig:fcc}.
-With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since once FCC cell 
-shares its atoms with its next neighbours, the number of atom per unit cell is 4 -- 1/8 of a particle on eight corners and 1/2 of a particle on six faces.
-The whole system of atoms is then created by repeating this cell structure. One convenient way is to arrange the unit cells in a cubic system, 
-so if there are $M$ FCC unit cells on one edge, the whole system consists of $M^3$ cells. Since there are 4 particles per cell, 
-there are ideal or so called \textit{magic numbers} for atoms for which this setup works perfectly: $N = 4M^3 = 4,32,108,256,500,864,\ldots$.\\
-\begin{figure}[h]
-    \begin{center}
-        \begin{subfigure}[t]{0.4\textwidth}
-            \includegraphics[scale=0.3]{images/fcc-2.pdf}
-            \caption{Schematic figure of a face centered cubic (FCC) lattice}
-            \label{fig:fcc}
-        \end{subfigure} 
-        \
-        %\hspace{60pt}
-        \begin{subfigure}[t]{0.4\textwidth}
-            \includegraphics[scale=0.3]{images/unit_cell.pdf}
-            \caption{Setup for successively building a FCC lattice}
-            \label{fig:unitcell}
-        \end{subfigure}
-    \end{center}
-\end{figure}
-%\begin{figure}
-    %\begin{center}
-        %\includegraphics[scale=0.3]{images/fcc_unit.png}
-        %\caption{Schematic of the number of particles contained in one FCC unit cell. The atoms on the edges are shared by 8 cells and the ones on the
-            %faces by 2 cells. Image taken from
-            %{https://upload.wikimedia.org/wikipedia/commons/2/27/Elementarzelle\_einer\_kubisch\_raumzentrierten\_Elementarzelle.png}}
-    %\end{center}
-%\end{figure}
-% do i need to credit the uni buffalo course? 
-% http://www.physics.buffalo.edu/phy411-506/
-There are several ways to achieve this initial configuration and the one used in this thesis \cite{buffalo} was to create a kind of unit cell 
-consisting of four atoms, as shown in fig. \ref{fig:unitcell}, which can be described by a set of points
-\begin{eqnarray*}
-    p_1 &=& \{0,0,0\}\\
-    p_2 &=& \{0.5*a,0.5*a,0\}\\
-    p_3 &=& \{0.5*a,0,0.5*a\}\\
-    p_4 &=& \{0,0.5*a,0.5*a\}
-\end{eqnarray*}
-From the particle number $N$ and the number of FCC unit cells per edge $M$ the \textit{lattice constant} $a$ can be calculated
-\begin{equation}
-    a = \frac{L}{M}
-\end{equation}
-where $L$ is the side length of the cube that is the whole system and it is calculated via the density of the system
-\begin{equation}
-    L = \sqrt[3]{\frac{N}{\rho}}
-\end{equation}
-With the lattice constant and the 4 points of the FCC cell, all the particles can be put into place.
-
-\subsection{The Velocity-Verlet Algorithm}
-\label{section:verlet}
-When we look at the system from a microscopic standpoint, we see that it follows some kind of path in the phase space as time progresses. Every
-point in this space corresponds to a set of positions and momenta and the connection between two points corresponds to the evolution of the system
-from one state to another. As mentioned above, this evolution (the dynamics of the system) is a crucial element to molecular dynamics. Since the
-equations of motion cannot be solved analytically in general, we need to approximate the solution.\\
-The method used here is called finite difference approach. The trajectory of the system in the phase space is cut into finite pieces of length $\Delta
-t$ and the equations of motion are solved for every segment separately (see fig. \ref{fig:finitedifference}).\\
-\begin{figure}
-    \begin{center}
-        \includegraphics[scale=0.5]{images/finite_approach.pdf}
-        \caption{Simplified graphical schematic of the finite difference approach. The evolution of the system from a point $(r^N(t),p^N(t))$ in the
-            phase space is approximated by slicing it up into pieces of length $\Delta t$. On every stop after the starting point ($t+\Delta t$, 
-            $t+2\Delta t$, $t+3\Delta t$,$\ldots$) the equations of motion can be solved numerically.}
-        \label{fig:finitedifference} 
-    \end{center}
-\end{figure}
-There are several ways to solve this kind of problem, but since we are interested in implementing it into a computer program the ideal solution
-should have some basic properties \cite{Allen1989}:
-\begin{itemize}
-    \item It should be fast and require little memory
-    \item It should permit the use of a large time step $\Delta t$
-    \item It should duplicate the classical trajectory as closely as possible
-    \item It should satisfy known conservation laws
-    \item It should be simple and easy to program
-\end{itemize}
-One algorithm that has all of the above mentioned features is the one proposed by Verlet \cite{Verlet1967}. In his paper, Verlet starts by Taylor
-expanding the coordinate vector $\mathbf{r}_i$ for one particle after one time step $\Delta t$:
-\begin{equation}
-    \label{eq:verlet}
-    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + \dot{\mathbf{r}}_i \Delta t + \frac12 \ddot{\mathbf{r}}_i \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i \Delta t^3 +
-    \mathcal{O}(\Delta t^4)
-\end{equation}
-Since the first derivative of the coordinate is the velocity, $\dot{\mathbf{r}}_i(t) = \mathbf{v}_i(t)$, and the second derivative of the coordinate 
-is the acceleration, $\ddot{\mathbf{r}}_i(t) = \mathbf{a}_i(t)$, using Newton's law $\mathbf{F}_i(t) = m_i \mathbf{a}_i(t)$ equation \eqref{eq:verlet} 
-can be written as:
-\begin{equation}
-    \label{eq:verletnew}
-    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i(t)
-    \Delta t^3+\mathcal{O}(\Delta t^4)
-\end{equation}
-The same calculation can be carried out for one time step before $t$:
-\begin{equation}
-    \label{eq:verletnewminus}
-    \mathbf{r}_i(t-\Delta t) = \mathbf{r}_i(t) - {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 - \frac1{3!} \dddot{\mathbf{r}}_i(t) \Delta t^3+
-    \mathcal{O}(\Delta t^4)
-\end{equation}
-The sum of those two equations yields
-\begin{equation}
-    \mathbf{r}_i(t+\Delta t) + \mathbf{r}_i(t-\Delta t) = 2\mathbf{r}_i(t) + \frac1{2} {\mathbf{F}}_i(t) \Delta t^2 + \mathcal{O}(\Delta t^4)
-\end{equation}
-and from this we get the final form for the new coordinates:
-\begin{equation}
-    \label{eq:verletfinal}
-    \mathbf{r}_i(t+\Delta t) = 2\mathbf{r}_i(t)+ \mathbf{r}_i(t-\Delta t)+ \frac1{2} {\mathbf{F}}_i(t) \Delta t^2
-\end{equation}
-This means that the new coordinates can be calculated using the current coordinates, the current forces and the coordinates from the past time step.\\
-From equation \eqref{eq:verletfinal} one thing becomes clear: the velocities are not necessary to calculate the new positions and they are not
-computed in the process, but for the calculation of i.e. the kinetic energy the velocities are needed. Thus, the velocities have to be calculated with
-a combination of the Taylor expansions of the coordinate vectors at $t+\Delta t$ and $t-\Delta t$:
-\begin{equation}
-    \mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t) = 2\mathbf{v}_i(t)\Delta t + \mathcal{O}(\Delta t^3)
-\end{equation}
-which can be rewritten as
-\begin{equation}
-    \label{eq:velocities}
-    \mathbf{v}_i(t) = \frac{\mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t)}{2\Delta t} + \mathcal{O}(\Delta t^2)
-\end{equation}
-This approach has the advantage that it is fast, requires little memory and is reliable in the sense that there is no energy drift occuring during the
-simulation, which means that the energy is conserved. When we compare this to the list of desired properties this seems like a good algorithm.\\
-This algorithm however has two significant disatvantages. The first one is the calculation of the velocities. As can be seen in equation
-\eqref{eq:velocities}, the accuracy of the calculation is only $\mathcal{O}(\Delta t^2)$ while the positions can be calculated with an error of order
-$\Delta t^4$. The other big disatvantage is the first step of the algorithm. Since the calculation of the new positions requires the current positions
-and the ones from one time step before, which technically do not exist.\\
-The solutions to this problem is to include the stepwise calculation of the velocities \cite{swope1982}. For this we start again with the Taylor
-expansion of the coordinates mathbftor \eqref{eq:verletnew} and instead of Taylor expanding $\mathbf{r}_i(t-\Delta t)$, we write it as 
-\begin{equation}
-    \mathbf{r}_i(t) = \mathbf{r}_i(t+\Delta t) - \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
-\end{equation}
-Using \eqref{eq:verletnew} in the abvove equation we get
-\begin{equation}
-    \begin{aligned}
-        \mathbf{r}_i(t) &= \mathbf{r}_i(t) + \mathbf{v}_i(t) \Delta t + \frac1{2m}\mathbf{F}_i(t)\Delta t^2 \\
-                     &- \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
-    \end{aligned}
-\end{equation}
-and thus
-\begin{equation}
-    \label{eq:vvelocities}
-    \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \left(\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\right) \Delta t
-\end{equation}
-With the addition of this equation this algorithm is called the Velocity-Verlet algorithm, which is summarized in the following two equations:
-\begin{equation}
-    \label{eq:velocityverlet}
-    \begin{aligned}
-        \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2\\
-        \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \Big[\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\Big] \Delta t
-    \end{aligned}
-\end{equation}
-This algorithm is self starting, uses a small amount of memory, conserves the energy (not exactly) 
-and gives a very good approximation to the originial trajectory
-in the phase space. To program this algorithm the following steps are needed: 
-\begin{enumerate}
-    \item Calculate all the forces between the particles (for the first step only)
-    \item Calculate the new positions with current velocity and forces
-    \item Calculate first half of the new velocities with current forces
-    \item Calculate all the forces for the new position
-    \item Use new forces to calculate second half of new velocities
-\end{enumerate}
-The first point only has to be carried out for the first step, because the forces have not been calculated at this point. As the forces are calculated
-for the new positions in step 4, they can be used for the next time step. For the first time step the velocities have to be chosen randomly and are 
-calculated with this algorithm from that point forward.
-
-\subsection{The Laser Beam - Energy Influx}
-The nano particle is trapped in the laser beam. While the motion of the center of mass is localized, the
-individual atoms that make up the glass sphere absorb the energy from the laser which increases their velocity.\\
-To simulate this kind of behaviour, thermostat algorithms \cite{Huenenberger2005} such as the Nos\'e-Hoover \cite{Nose1984,Hoover1985} or the Andersen
-\cite{Andersen1980} algorithms are often used in simulation to change the temperature of the system in a controllable way.\\
-The problem with these kind of algorithms is, that a target temperature has to be fixed which will be reached at some point. In order to simulate the
-influx of energy from the laser it would be better to have an algorithm that supplies the system with a certain amount of energy continuously.
-Fortunately, such an algorithm exists.\\
-The algorithm is called Heat Exchange Algorithm (HEX) \cite{Hafskjold1993}. Its intended purpose is the use in non-equilibrium molecular dynamics
-(NEMD) to study transport phenomena and determine transport coefficients. The algorithm works by introducing two regions in the system, one serving as
-a heat source and the other as a heat sink. A specific amount of heat is then exchanged between those two reservoirs. As it turns out however, this
-algorithm introduces an energy shift for longer simulation times. This led Wirnsberger et al. \cite{Wirnsberger2015} to revisit the algorithm and 
-identify the cause of this energy drift, to create a more suitable algorithm, which they called eHEX.\\
-As in the HEX algorithm, regions are introduced to the system which act either as heat sinks or heat sources. These are labelled with $\Gamma_k$,
-where $k > 0$, and have corresponding amount of exchanged heat $\Delta Q_{\Gamma_k}$. If $\Delta Q_{\Gamma_k}$ is negative, heat is subtracted from
-the system and vice versa. Regions that neither act as heat source or heat sink are labelled with $\Gamma_0$, which are also called
-Hamiltonian regions (see fig. \ref{fig:ehex}). 
-\begin{figure}
-    \begin{center}
-        \includegraphics[scale=0.3]{images/ehex_system.png}
-        \caption{Setup for the use of the HEX and eHEX algorithms. The simulationbox $\Omega$, which is moving with a velocity of $v_\Omega$,
-        contains a heat source (red), $\Gamma_1$, which is moving with velocity $v_{\Gamma_1}$ and a heat sink (blue), $\Gamma_2$, which is moving 
-        with velocity $v_{\Gamma_2}$. The regions that are neither heat sources or heat sinks are denoted by $\Gamma_0$.}
-        \label{fig:ehex} 
-    \end{center}
-\end{figure}
-The centers of mass of the particles in simulation box, denoted by $\Omega$, and the regions $\Gamma_k$ 
-are assumed to be moving with velocities $v_\Omega$ and $v_{\Gamma_k}$ respectively.\\
-The change of the energy in a region $\Gamma_k$ is achieved by rescaling the velocities by a factor $\xi_k$ and shifted by the velocity of the
-corresponding region:
-\begin{equation}
-    \mathbf{v}_i \rightarrow \mathbf{\bar{v}}_i = \xi_k \mathbf{v}_i + (1-\xi_k)\mathbf{v}_{\Gamma_k}
-\end{equation}
-The bar over a quantity denotes the value after the exchange of heat.\\
-The factor $\xi_k$ is given by
-\begin{equation}
-    \xi_k = \sqrt{1+\frac{\Delta Q_{\Gamma_k}}{\mathcal{K}_{\Gamma_k}}}
-\end{equation}
-where $\Delta Q_{\Gamma_k}$ is the exchanged heat in the region $\Gamma_k$ and $\mathcal{K}_{\Gamma_k}$ is the non-translational kinetic energy of the
-region $\Gamma_k$ and is given by
-\begin{equation}
-    \label{eq:kineticK}
-    \mathcal{K}_{\Gamma_k} = \sum_{i \in \gamma_k} \frac{m_i v_i^2}{2} - \frac{m_{\Gamma_k}v_{\Gamma_k}^2}{2}
-\end{equation}
-The sum is taken over all indices in $\gamma_k$ which is the set of indices of particles in the region $\Gamma_k$.\\
-For the final version of the eHEX algorithm there are three more quantities needed. The first one is the heat flux per time step, denoted by 
-$\mathcal{F}_\GammaK$:
-\begin{equation}
-    \mathcal{F}_\GammaK = \frac{\Delta Q_\GammaK}{\Delta t}
-\end{equation}
-The second one is the thermostatting force $\boldsymbol{\eta}_i$, which is defined as
-\begin{equation}
-    \boldsymbol{\eta}_i = 
-    \begin{cases}
-        m_i \ \dfrac{{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}}{{2\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}}} \
-        \left(\mathbf{v}_i - \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)
-        & \quad \text{if } k(\mathbf{r}_i) > 0 \\
-        0 & \quad \text{otherwise}
-    \end{cases}
-    %\begin{cases}
-        %m_i \ \frac{\displaystyle{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}}{\displaystyle{2\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}}} \
-        %\left(\mathbf{v}_i - \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)
-        %& \quad \text{if } k(\mathbf{r}_i) > 0 \\
-        %0 & \quad \text{otherwise}
-    %\end{cases}
-\end{equation}
-where $k(\mathbf{r}_i)$ is the index of the region in which particle $i$ located, i.e. $k(\mathbf{r}_i) = 0$ means that the particle is in a Hamiltonian
-region and $k(\mathbf{r}_i) > 0$ denotes the heat sinks and sources.\\
-The last quantity is the one that corrects the long term energy drift of the HEX algorithm, denoted by $\mathcal{E}{r}_{i,\alpha}$.
-The analysis and derivation of this term is given in \cite{Wirnsberger2015}.
-\begin{equation}
-    \label{eq:bigeps}
-    \begin{aligned}
-        \mathcal{E}{r}_{i,\alpha} &= \dfrac{\eta_{i,\alpha}}{m_i\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}} \left[
-        \frac{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}{48} + \frac{1}{6} \sum_{j\in\gamma_{k(\mathbf{r}_i)}} \mathbf{f}_j \cdot \left(\mathbf{v}_j - 
-            \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)\right] \\
-            &- \frac{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}{12\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}} \left[
-            \frac{f_{i,\alpha}}{m_i} - \frac{1}{m_{\Gamma_{k(\mathbf{r}_i)}}} \sum_{j\in\gamma_{k(\mathbf{r}_i)}}f_{j,\alpha}\right]
-    \end{aligned}
-\end{equation}
-The $\mathbf{f}$ in the above equation denotes the force corresponding to the chosen intermolecular potential $U(\mathbf{r})$, 
-\begin{equation}
-    \mathbf{f} = -\nabla_{\mathbf{r}_i} U(\mathbf{r}_i)
-\end{equation}
-With all the necessary quantities established, the updating sequence of the eHEX algorithm can be written down:
-\begin{subequations}
-\begin{eqnarray}
-    \bar{\mathbf{v}}^n_i &=& \xi^n_{k(\mathbf{r}_i)} \mathbf{v}^n_i + \left(1-\xi^n_{k(\mathbf{r}_i)}\right) \mathbf{v}^n_{\Gamma_{k(\mathbf{r}_i)}} \\
-    \bar{\mathbf{v}}^{n+\frac12}_i &=& \bar{\mathbf{v}}^n_i + \frac{\Delta t}{2m_i} \mathbf{f}^n_i\\
-    \bar{\mathbf{r}}^{n+1}_i &=& \mathbf{r}^n_i + \Delta t \bar{\mathbf{v}}^{n+\frac12}_i\\
-    \mathbf{f}^{n+1}_i &=& \left.-\nabla_{\mathbf{r}_i} U(\mathbf{r})\right|_{\mathbf{r} = \bar{\mathbf{r}}^{n+1}}\\
-    \bar{\mathbf{v}}^{n+1}_i &=& \bar{\mathbf{v}}^{n+\frac12}_i + \frac{\Delta t}{2m_i} \mathbf{f}^{n+1}_i\\
-    \mathbf{v}^{n+1}_i &=& \bar{\xi}^{n+1}_{k(\bar{\mathbf{r}}_i)} \bar{\mathbf{v}}^{n+1}_i + \left(1-\bar{\xi}^{n+1}_{k(\bar{\mathbf{r}}_i)}\right)
-    \bar{\mathbf{v}}^{n+1}_{\Gamma_{k(\bar{\mathbf{r}}_i)}} \\
-    \mathbf{r}^{n+1}_i &=& \bar{\mathbf{r}}^{n+1} - \Delta t^3 \mathcal{E}\bar{\mathbf{r}}^{n+1}_i
-\end{eqnarray}
-\end{subequations}
-This algorithm can be adapted for the problem of the levitating nano sphere in the laser beam by adjusting the setup and some of the parameters.\\
-Firstly, the setup of the heat sinks and sources has to be changed. Since there is only energy pumped into the system from the outside, the region
-acting as a heat sink vanishes and the region acting as a heat source spans over the whole simulation box. This means that, using the 
-notation of fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither center of mass of the particles the simulation box nor 
-the heat source are moving, i.e. 
-$\mathbf{v}_\Omega = \mathbf{v}_{\Gamma_1} = 0$. This affects the non-translational kinetic energy term $\mathcal{K}_{\Gamma_1}$, the thermostatting
-force $\boldsymbol{\eta}$ and the correction term $\mathcal{E}\mathbf{r}$. Since there is only one region acting as a heat source, the terms $\Delta Q$,
-$\mathcal{K}$ and $\mathcal{F}$ don't need an index. The summation index in \eqref{eq:kineticK} and \eqref{eq:bigeps} can be changed to the number of
-particles in the system, $N$, since all the particles are within the heat exchanging region. The masses are set to 1, i.e. $m_i = 1$ with the chosen
-reduced units and with this the total mass of the region (in the term $1/m_{\Gamma_{k(\mathbf{r}_i)}}$ in \eqref{eq:bigeps}) is equal to the number of
-particles in the system. With these changes, the quantities can be written down as:
-\begin{eqnarray}
-    \mathcal{K} &=& \sum_{N} \frac{v_i^2}{2}\\
-    \xi &=& \sqrt{1+\frac{\Delta Q}{\mathcal{K}}}\\
-    \mathcal{F} &=& \frac{\Delta Q}{\Delta t}\\
-    \boldsymbol{\eta}_i &=& 
-    \dfrac{\mathcal{F}}{{2\mathcal{K}}} \
-        \mathbf{v}_i\\
-        \mathcal{E}{r}_{i,\alpha} &=& \dfrac{\eta_{i,\alpha}}{\mathcal{K}} \left[
-        \frac{\mathcal{F}}{48} + \frac{1}{6} \sum_{N} \mathbf{f}_j \cdot \mathbf{v}_j
-            \right] \nonumber \\
-            &-& \frac{\mathcal{F}}{12\mathcal{K}} \left[
-            {f_{i,\alpha}} - \frac{1}{N} \sum_{N}f_{j,\alpha}\right]
-\end{eqnarray}
-This means that the laser is modeled to be pumping energy into the system, which increases the velocities of the particles in the
-system over time, while the center of mass motion is not affected by this.\\
-
-\subsection{The Laser Beam - Trapping}
-As mentioned above, the glass particle is trapped in the laser beam and the position of the particle is localized. This behaviour
-has to be modeled as well.\\
-The approximation of the original paper \cite{Gieseler2014}, where the movements in the three spatial directions are decoupled will be
-used in the model as well. As to the model of the trapping force itself, the most straightforward approach is to use a harmonic oscillator potential.
-The force and the corresponding potential can be written as
-\begin{eqnarray}
-    \mathbf{F} &=& -k \Big[\mathbf{x}-\mathbf{x}_0\Big]\\
-    U &=& \frac12 k \Big[\mathbf{x}-\mathbf{x}_0\Big]^2
-\end{eqnarray}
-where $\mathbf{x}_0$ is the position of minimal potential energy. 
-Since the trap is acting on the whole system, the force is acting on the center of mass. 
-To calculate the center of mass positions and velocities, the positions and velocities of all particles have to be summed up and 
-divided by the particle number:
-\begin{equation}
-    \mathbf{r}_\text{COM} = \frac1N\sum_{i=1}^N \mathbf{r}_i
-\end{equation}
-
-\subsection{Surrounding Gas - Thermostat}
-Without any equilibrating mechanism the setup described by now would lead to the system heating up indefinetly, which is not desirable. Furthermore,
-the goal is to recreate the experiment as close as possible. Since the nano particle is trapped in a laser beam within a vacuum chamber, the
-surrounding gas has to be modeled as well. In the next step we will introduce the surrounding gas of the gas chamber that will absorb some of the 
-energy in the system, leading to the final state.\\
-Generally, pressure is introduced to the system by surrounding the object of interest (in this case the glass nano sphere) with a thermostatting
-pressure medium.
-There are two main requirements for the choice of such a pressure medium: the exerted pressure must be hydrostatic and the computation of the
-interaction between the pressure medium and the object of interest must not take up a lot of resources.\\
-The model used in this thesis was developed by Gr\"unwald and Dellago \cite{Gruenwald2006} and uses an ideal gas of non-interacting particles as
-thermostatting pressure medium. The particles of this pressure medium flow into the simulation from an outside volume, whose geometry is based on the form of the
-object of interest (this will be referred to as the minimal volume of cells) and leave the simulation box if their position reaches the boundary of
-this minimal volume of cells. This behaviour is very close to the real experiment which makes this thermostat an ideal candidate for the simulation of
-the experiment. 
-%This barostat also acts as a thermostat and thus is suitable for the 
-%application in this thesis.\\
-The gas particles interact with the object via a soft-sphere potential of the form
-\begin{equation}
-    \label{eq:softsphere}
-    U(r) = \varepsilon \left(\frac{\sigma}{r}\right)^{12}
-\end{equation}
-where $\varepsilon$ is the interaction strength and $\sigma$ is the interaction range and $r$ is the distance between the gas particle and the
-interacting particle. The gas particles do not interact with one another in this model.\\
-In order to increase the efficiency of the computing process, $\sigma$ should be chosen carefully. For larger $\sigma$, the number of interaction
-partners increases, which increases the number of force calculations which are a very time consuming part. For smaller $\sigma$ the possibility for
-gas particles reaching the inside of the nano particle increases, which is not desirable. So $\sigma$ should be chosen small enough to keep the force
-calculations at a minimum and large enough for the particle to stay on the outside of the crystal. In the computer simulations performed for this
-thesis, the interaction length is chosen to be $\sigma = 1$ with a cut-off radius of $r_c = 2.5$.\\ 
-The algorithm can be performed by following these steps:
-\begin{enumerate}
-    \item Randomly draw the number of particles that are created on a single side of the minimal volume of cells, $N_\text{fac}$, from the
-        distribution 
-        \begin{equation}
-            %\langle N_\text{fac}\rangle = \Delta t L^2 P \left(\frac{1}{2\pi m k_B T}\right)^\frac12
-            \langle N_\text{fac}\rangle = \Delta t L^2 P \sqrt{\frac{1}{2\pi m k_B T}}
-        \end{equation}
-        where $\Delta t$ is the time step of the simulation, $L$ is the side length of the cell in which the particle is created, $P$ is the desired
-        pressure, $m$ is the mass of the gas particle, $k_B$ is the Boltzmann constant (which will be set to 1 in reduced units) and $T$ is the
-        desired temperature. This chosen number of particles is then equally distributed over the face of the cell on which they are created. The
-        velocities of the created particles are drawn from two different random number distributions. The component of the particle perpendicular to
-        the surface is drawn from a Rayleigh distribution of the form 
-        \begin{equation}
-            p(v_i) = \frac{m}{k_B T}v_i \ e^{-\frac{mv_i^2}{2k_BT}}
-        \end{equation}
-        The other components of the particles' velocity are drawn from a Maxwell-Boltzmann distribution.
-    \item Perform the first step of the velocity Verlet algorithm to propagate the particle positions by one time step.
-    \item Check if any gas particles have left the minimal volume of cells and remove those which have.
-    \item Check if the geometry of the crystal and with it the minimal volume of cells has changed. If it has, remove all gas particles in the cells
-        that are no longer needed. Then insert new gas particles to the created cells with a number drawn from a Poisson distribution with mean value
-        \begin{equation}
-            \langle N_\text{ins}\rangle = \frac{PL^3}{k_BT}
-        \end{equation}
-        If the length of the cell $L$ is chosen to be equal to the cut-off radius $r_c$, this insertion should only be carried out with a probablity of
-        \begin{equation}
-            P_\text{ins} = e^{-\frac{U}{k_BT}}
-        \end{equation}
-        (where U is the interaction energy between the gas particle and the crystal) because it is possible that the inserted particle is 
-        within the interaction range of the crystal. 
-    \item Compute all forces.
-    \item Perform the second step of the velocity Verlet and propagate the particle velocities by one time step. 
-\end{enumerate}
-This algorithm is formulated in a general fashion, so that a wide range of crystals and geometries can be used. 
-%The setup used in this thesis is a FCC lattice, which has a nice symmetry that is not influenced by the surrounding gas. 
-Although the shape of the nano particle changes over the course of the simulation, it is not necessary to change the geometry of the surrounding box.
-The algorithm will therefore not be carried out using cell-lists and a minimal volume of cells that is changing over the course of the simulation, 
-but rather a fixed setup of the volume surrounding the nano particle.\\ 
-Since the particle itself is modeled as a cube, a straighforward approach is surrounding it by a bigger cube. The distance between the cube face and the nano
-particle is chosen to be the side length of the nano particle, so the setup will properly scale for different numbers of particles. 
-The schematic setup for this is depicted in Fig.\ref{fig:gascube}.
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.5]{images/gas_volume_ver3.pdf}
-        \caption{Schematic setup for the barostat. The outer volume (blue) is one side length $s$ away from the nano particle (black) which means that
-            the side length of the face of the surrounding box is $L=3s$. 
-                    The gas particles are created along the blue lines of the outer volume. Since the overall geometry of the nano particle does 
-                    not change, 
-                    the geometry of the outer volume will be constant throughout the simulation.}
-        \label{fig:gascube}
-    \end{center}
-\end{figure}
-
-%This is done by adding for instance two regions to the simulation, one providing an influx of heat
-%and the other acting as a heat sink. The algorithm is furthermore designed to incorporate the velocities of the center of mass of the system and the
-%velocities of the regions.
-
-%\subsection{The Glass Nanoparticle}
-%The glass particle from the experiment will be represented by a system of 864 particles, aligned regularly on a FCC (face centered cubic) lattice.
-%This alignment is achieved by defining a minimum cell, containing points:
-%\begin{eqnarray*}
-    %p_1 &=& \{0,0,0\}\\
-    %p_2 &=& \{0.5,0.5,0\}\\
-    %p_3 &=& \{0.5,0,0.5\}\\
-    %p_4 &=& \{0,0.5,0.5\}
-%\end{eqnarray*}
-%The whole system is then  created by copying this unit cell.\\
-%The interaction between the atoms is modeled by the Lennard-Jones potential. It has the form
-%\begin{equation}
-    %U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right].
-%\end{equation}
-%To simplify the equation and computation of the potential and other quantities, like the forces or pressure, it is useful to introduce reduced units.
-%In general, reduced units BLABLABLA THINK OF GOOD TEXT HEREJKA\\
-%The basic units in a system with Lennard-Jones interaction are length ($\sigma$), energy ($\varepsilon$ or $\varepsilon / k_B$) and mass ($m$). Every 
-%quantity can now be written in terms of this units, so they become reduced quantities, denotet by an asterisk. The most important ones are:
-%\begin{eqnarray*}
-    %r^* &=& r/\sigma\\
-    %T^* &=& k_B/\varepsilon \ T\\
-    %U^* &=& U/\varepsilon \\
-    %P^* &=& P \sigma^3/\varepsilon 
-%\end{eqnarray*}
-%With the above introduced reduced units, the Lennard-Jones potential can be written as
-%\begin{equation}
-    %U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right].
-%\end{equation}
-%Since this is the most practical form, it will be used without the asterisk from here on.\\
-%The Lennard-Jones potential is an additive pair-potential, so the total energy of the system can be calculated by summing over all pairs of atoms:
-%\begin{equation}
-    %U_{\text{tot}} = \sum_{i=1}^{N-1}\sum_{j=i+1}^N4\left[{r_{ij}}^{-12} - {r_{ij}}^{-6}\right]
-%\end{equation}
-%where $r_{ij}$ denotes the distance between atom $i$ and $j$. \\
-%The Velocity-Verlet algorithm uses the forces to calculate the positions and velocities for the next timestep, we need to calculate the derivative of
-%the potential energy:
-%\begin{eqnarray}
-    %F_{x} &=& -\frac{\partial}{\partial x} U(r) \nonumber\\
-                %&=& -\frac{\partial}{\partial x} 4\left[{r}^{-12} - {r}^{-6}\right] \nonumber\\
-                %&=& -4 \left[(-12){r}^{-13} - (-6){r}^{-7}\right] \frac{\partial r}{\partial x} \nonumber\\
-                %&=& 48 \left[r^{-13} - 0.5 \ r^{-7}\right] \frac{x}{r} \nonumber\\
-    %\label{eq:ljforce} &=& 48 \left[r^{-14} - 0.5 \ r^{-8}\right] x
-%\end{eqnarray}
-%This is the component of the force in x-direction -- the other components are calculated analogously.
-
-\newpage
-\section{Results}
-% results with different parameters, graphs, etc.
-Since the simulation consists of various parts that have to work together, we first need to check if every part itself works.\\
-\subsection{The Crystal}
-The first element of the simulation is the setup of the FCC lattice on which the particles are placed on. To check if this is done correctly, we can
-use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano particle 
-can be seen in fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
-where the atoms take the position on the gap between the atoms from the layer below.\\
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.2]{images/fcc_rendering_layer_newcoloring.png}
-        \caption{A rendered image of the FCC lattice. The chosen color-coding along the (1,1,1) Miller-index 
-                highlights the layers for this perspective. The perspective is
-                chosen to see the layering of the FCC lattice with its 3 layer structure. The black box acts as a reference point to outline the 
-                geometry of the system.}
-            %Rendering of the model of the nano crystal before starting the simulation. 
-            %The chosen perspective shows the 3 layers of the FCC crystal with a
-                    %These layers consist of atoms
-                    %taking the position above the gap between two of the atoms from the layer below and match the positions of the atoms 2 layers
-                %below. The black box is the outline of the particles' positions and shows the geometry of the system.}
-        \label{fig:glass}
-    \end{center}
-\end{figure}
-
-\subsection{Velocity Verlet}
-The velocity Verlet part of the simulation is taking place in a NVE ensemble from a thermodynamic point of view. 
-This means that the particle number $N$, the volume $V$ and the
-total energy $E$ are constant. In contrast to the NVT ensemble used in Monte-Carlo simulations this means, that the velocities of the particle can not
-be chosen directly. The initial velocities have to be drawn from a random number generator. This means that the velocities have to be adjusted after 
-they have been drawn.\\
-The first adjustment that needs to be made is making sure that the system is staying in place and not floating away. To do this, the velocity vectors
-of all the atoms in the system have to be added to get the resulting velocity vector for the whole system that is thought to be acting on the center
-of mass. This resulting vector is then divided by the number of particles in the system and the resulting vector is then subtracted from every
-velcocity vector of every atom in the system. 
-\begin{eqnarray}
-    \mathbf{v}_\text{tot} &=& \sum_{i=1}^N \mathbf{v}_i\\
-    \mathbf{v}_\text{cut} &=& \frac{\mathbf{v}_\text{tot}}N\\
-    \mathbf{v}_i^\text{new} &=& \mathbf{v}_i - \mathbf{v}_\text{cut}
-\end{eqnarray}
-%The next adjustment is a small one that makes a difference in the long run. When choosing random velocities, the system not only tends to have a
-%resulting velocity vector that causes the whole system to float away, but also a resulting angular momentum that causes the system to rotate.
-%The angular momentum can be defined via the moment of inertia $\mathbf{I}$
-%\begin{eqnarray}
-    %\label{eq:angularmomentum}\mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
-    %\boldsymbol{\omega} &=& \frac{\mathbf{r} \times \mathbf{v}}{r^2}
-%\end{eqnarray}
-%where $\boldsymbol{\omega}$ is the angular velocity of the system. Another way to define the angular momentum is given by
-%\begin{equation}
-    %\label{eq:angularmomentum2}
-    %\mathbf{L} = \mathbf{r} \times \mathbf{p}
-%\end{equation}
-%where $\mathbf{p}$ is the momentum of the particle, which in this case is equal to its velocity since the mass is set to 1.The moment of inertia 
-%$\mathbf{I}$ for a filled cube is given by 
-%\begin{equation}
-    %\mathbf{I} = 
-    %\begin{pmatrix}
-        %\frac{mL^2}{6} & 0 & 0\\
-        %0 & \frac{mL^2}{6} & 0\\
-        %0 & 0 & \frac{mL^2}{6} 
-    %\end{pmatrix}
-%\end{equation}
-%where $m$ will be later set to 1 and $L$ is the side length of the cube. To use this equations to suppress the rotation of the cube, we need to use
-%equation~\eqref{eq:angularmomentum} and rewrite it as
-%\begin{equation}
-    %\label{eq:omega}
-    %\boldsymbol{\omega} = \mathbf{I}^{-1}\mathbf{L}
-%\end{equation}
-%where $\mathbf{I}^{-1}$ is the inverse matrix of $\mathbf{I}$. The calculation of the inverse matrix is fairly straightforward, as it only contains
-%values in the diagonal, which means that the inverse matrix can be calculated by simply inverting these values. The angular momentum for every
-%particle $\mathbf{L}_i$ in the above equation be calculated using eq.~\eqref{eq:angularmomentum2}, where the position is taken to be the distance
-%between the particle and the center of mass
-%\begin{equation}
-    %\Delta\mathbf{r}_i = \mathbf{r}_\text{COM} - \mathbf{r}_i
-%\end{equation}
-%The values of the angular momentum for each paricle are added up and multiplied by the inverse of the moment of inertia, as described in
-%eq.~\eqref{eq:omega}. The final step is updating the velocities of the particles:
-%\begin{equation}
-    %\mathbf{v}_i^\text{new} = \mathbf{v}_i - \left(\boldsymbol{\omega} \times \Delta\mathbf{r}_i\right)
-%\end{equation}
-\\The velocities of the particle can roughly be chosen by adjusting the range of numbers the random number generator can choose from. However, the
-resulting velocities (and therefore the temperature of the system) are generally not precisely how they should be. If the system should have a
-specific starting temperature $T$, the velocities need to be scaled accordingly. This rescaling is done by summing the square of the velocities
-(this equals calculating the kinetic energy since the masses are set to 1) and rescale them by a factor $\lambda$:
-\begin{eqnarray}
-\label{eq:rescale1}   E &=& \sum_{i=1}^N \frac{\mathbf{v}_i^2}{2} \\
- \label{eq:rescale2}   \lambda &=& \sqrt{\frac{3(N-1)T}{2E}} \\
- \label{eq:rescale3}   \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i  
-\end{eqnarray}
-This process has to be done a couple of times, until the system equilibrates to the desired temperature. The problem with this procedure is the fact
-that it contradicts the constant energy property of the ensemble. If we look at the energy values in fig.~\ref{fig:energystart} we can see the
-irregularities (jumps) before the system equilibrates to a constant value. This equilibration process divides the simulation in two phases: the equilibration
-phase and the measurement phase. In the equilibration phase, the system has time to reach a configuration that is (as the name suggests) in an 
-equilibrated state. During this phase, there are usually no measurements (such as energy) performed, since the values from this phase are distorting
-the mean value of the simulation. That is why the configurations from the equilibration phase are discarded and the measurement phase takes place, 
-where the values do get calculated and are used to determine the mean values.  
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.3]{images/verlet_output_te.pdf}
-        \caption{Temperature and energy in the equilibration phase of the velocity Verlet algorithm. The velocity rescaling (see eq. 
-        \eqref{eq:rescale1}-\eqref{eq:rescale3}) is taking place every 1000th step until the 9000th, which can be seen in the energy curve whre the
-    jumps are happening. The desired temperature in this case is $T^* = 0.2$, which is reached after about 7000 steps.}
-        \label{fig:energystart}
-    \end{center}
-\end{figure}
-
-\subsection{eHEX}
-The next checkpoint is the eHEX algorithm. Since it contains several steps and variables that have to be calculated, this algorithm is very prone
-to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\Delta Q$. 
-The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
-The investigation of the effects on the system can be done in two ways: comparing the temperature and the energy of the system during the 
-application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\Delta Q$.\\
-%Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
-%with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\
-Since the eHEX algorithm injects heat into the system, one would expect the energy and the temperature to rise during the application of the
-algorithm. In fig.~\ref{fig:ehex_te} we can see that exactly this is happening. The graph shows a velocity Verlet phase (that follows an equilibration
-phase which is not recorded) of constant temperature and energy and an eHEX phase, where the temperature and the enrgy are constantly rising. 
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/ehex_te.pdf}
-        \caption{Comparison of temperature (red) and energy (blue) during the application of the eHEX algorithm with $\Delta Q=0.04$ (in reduced units).
-                The first 20000 timesteps, the energy and temperature are constant during the application of the velocity Verlet algorithm and start
-            to rise, once the eHEX algortithm takes over. The values for the temperature can be seen on the ordinate on the left hand side
-            and the values for the energy on the right hand side. }
-        \label{fig:ehex_te}
-    \end{center}
-\end{figure}
-The comparison of different values of $\Delta Q$ can be seen in fig.~\ref{fig:ehex_dq}. As the correlation between the temperature and the energy has
-been shown in fig.~\ref{fig:ehex_te}, the plotting of the temperature data should suffice. 
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/ehex_temp_dq.pdf}
-        \caption{Comparison of different values for $\Delta Q$, ranging from 0 to 0.09. As the value $\Delta Q=0$ corresponds to no heat being
-        injected into the system, it is not surprising that the temperature remains constant. The graph shows that an increase in $\Delta Q$ leads to
-    an increase in the rate of warming up, which means that for bigger $\Delta Q$, the system is heating up more quickly.}
-        \label{fig:ehex_dq}
-    \end{center}
-\end{figure}
-
-\subsection{Thermostat}
-The barostat, as described previously, keeps the system from overheating. The barostat algorithm is applied at the same time as the eHEX algorithm,
-after the velocity Verlet has equilibrated the system.\\
-While the eHEX algorithm is pumping energy into the nano particle while the surrounding gas particles absorb excessive energy. This leads to the
-development of a new, higher steady temperature of the nano particle, which can be seen in Fig.~\ref{fig:baro1}. 
-%The first simulations were performed
-%with the eHEX and the barostat algorithm starting at the same point. This lead to the question: does the new steady temperature depend on the time it
-%takes the gas particles to reach the system? The answer is: no. After applying the barostat algorithm first to fill the box with gas atoms and then
-%applying the eHEX algorithm it can be seen that the resulting temperature is the same, which can be seen in fig.~\ref{fig:baro2}.
-%This means that the system is beginning to heat up, until the particles reach the crystal.
-%Because of this, the temperature of the system rises, until it reaches a steady value again. This behaviour can be seen in fig.~\ref{fig:baro1}.\\
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/baro_temp_start.pdf}
-        \caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
-        in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
-    an increase in temperature until a new steady temperature is reached.}
-        %\caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
-        %in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
-    %an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
-%reached and the temperature seems to stay at this value.}
-        \label{fig:baro1}
-    \end{center}
-\end{figure}
-%\begin{figure}[h]
-    %\begin{center}
-        %\includegraphics[scale=0.4]{images/baro_temp_compare_other.pdf}
-        %\caption{Two versions of the application of the eEHX and barostat algorithm. The shorter run (red, x-axis on the bottom) represents 
-            %the simultaneous start of the eHEX and the barostat algorithm. In the longer run (blue, x-axis on the top) there are 20000 steps, where
-        %the barostat algorithm is applied at the same time as the velocity Verlet to give the system time to equilibrate and only after the eHEX
-    %algorithm is taking the place of the velocity Verlet. The mean value of the resulting temperature is the same, indicating that the final
-%temperature is based on the values of the parameters, such as pressure and ambient temperature, rather than geometry.}
-        %%\caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
-        %%in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
-    %%an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
-%%reached and the temperature seems to stay at this value.}
-        %\label{fig:baro1}
-    %\end{center}
-%\end{figure}
-As we are interested in the temperatures of the particles before and after the interaction with the nano particle, we need some way to differentiate those
-two states. This can be done when looking at the total energy of the gas particles.\\
-The gas particles are created on the surface of the simulation box, which is further away than the interaction range between the gas particles and the
-nano particle. Since the gas particles do not interact with each other, the total energy of the gas particles does not change until they reach the
-interaction range of the nano particle. There they undergo a change of energy due to the interaction with the nano particle and then make their way away
-from the nano particle. After the interaction, when they reach a point outside of the interaction range, the enrgy stays constant again until the
-particles reach the limits of the simulation box.\\
-Therefore, the life span of a single gas particle can be partioned into 3 stages: incoming, interacting, outgoing. In a typical scene of interaction
-all three of these stages are represented by different gas particles, as can be seen in fig.~\ref{fig:barointeraction}.\\
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/baro_170105_new_medium.png}
-        \caption{Rendering of the application of the barostat algorithm. The red particles are incoming gas particles, that have not yet interacted, 
-            pink particles are currently interacting and green particles are leaving the simulation box and
-        moving away from the nano particle (blue). To render this image, a cut has been made to make a view into the simulation box possible.}
-        \label{fig:barointeraction}
-    \end{center}
-\end{figure}
-Since the gas particles are interacting with the nano particle, the temperatures and energies are values of interest. The energies are recorded using
-histograms and the temperature is calculated instantaneously every 100th timestep.\\
-The calculation of the energies of the incoming particles has to be done with care. The problem are particles that enter the simulation box, but fly
-by the nano particle at a distance that is greater than the interaction range. Since the change of the energy is the interesting part, only the energy 
-particles of that \textit{do} interact with the nano particle shall be calculated. This can be achieved by counting particles that enter the stage of
-\textit{interacing}. One such histogram can be seen in Fig.~\ref{fig:histogram}.
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/histogram_2.pdf}
-        \caption{Histogram of total energies for incoming (red) and outgoing (blue) particles. The parameters for this histogram are $\Delta Q^*=0.04$,
-        $T_\text{ambient}^*=0.08$, $P^*=0.8$. What can be seen here is that the outgoing particles have a higher count of particles with more total
-    energy.}
-        \label{fig:histogram}
-    \end{center}
-\end{figure}
-
-%\begin{figure}[h]
-    %\begin{center}
-        %\begin{subfigure}[t]{0.3\textwidth}
-            %\includegraphics[scale=0.3]{images/fcc-2.pdf}
-            %\caption{Schematic figure of a face centered cubic (FCC) lattice}
-            %\label{fig:fcc}
-        %\end{subfigure} 
-        %\
-        %\begin{subfigure}[t]{0.3\textwidth}
-            %\includegraphics[scale=0.15]{images/fcc_unit.png}
-            %\caption{Schematic of the number of particles contained in one FCC unit cell. The atoms on the edges are shared by 8 cells and the ones 
-                %on the faces by 2 cells\cite{unitcell}}
-            %\label{fig:fcccut}
-        %\end{subfigure} 
-        %\
-        %%\hspace{60pt}
-        %\begin{subfigure}[t]{0.3\textwidth}
-            %\includegraphics[scale=0.3]{images/unit_cell.pdf}
-            %\caption{Setup for successively building a FCC lattice}
-            %\label{fig:unitcell}
-        %\end{subfigure}
-    %\end{center}
-%\end{figure}
-
-\newpage
-\section{Conclusion}
-% what does this all tell us? do we need to study more of this?
-
-\newpage
+\section{Auswahlbibliographie}
+\nocite{*} 
 \bibliography{references}

### 2017-01-27 src/functions.cpp

@@ -562,7 +562,7 @@ void InitBarostat(std::list<Particle*>& particles)
          */
 }
 
-void InitBarostatFull(std::list<Particle*>& particles)
+void InitBarostatFullNewTemp(std::list<Particle*>& particles)
 {
         unsigned int i;
         unsigned int Num;
@@ -841,6 +841,117 @@ void InitBarostatFull(std::list<Particle*>& particles)
         //delete pyL; NO segfault
         //delete pz0; segfault
         //delete pzL; NO segfault
+}
+void InitBarostatFull(std::list<Particle*>& particles)
+{
+        unsigned int i;
+        unsigned int Num;
+        //double sigma=1.1;
+        double sigma=AmbientTemp;
+        double lambda = 0;
+        double vel = 0;
+
+        Particle *px0;
+        Particle *pxL;
+        Particle *py0;
+        Particle *pyL;
+        Particle *pz0;
+        Particle *pzL;
+
+        double kinE = 0.;
+
+        //side: x0
+        Num = NumberOfParticles();
+        for(i=0;i<Num;i++)
+        {
+            
+            px0 = new Particle(-L,
+                            ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
+                            ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
+                            gsl_ran_rayleigh(r,sigma),
+                            gsl_ran_gaussian(r,0.001)+AmbientTemp,
+                            gsl_ran_gaussian(r,0.001)+AmbientTemp,"GasIn");
+            particles.push_back(px0);
+            kinE = px0->m*(px0->v[0]*px0->v[0]+px0->v[1]*px0->v[1]+px0->v[2]*px0->v[2])*0.5;
+            gsl_histogram_increment(gas_in,kinE);
+        }
+
+        //side xL
+        Num = NumberOfParticles();
+        for(i=0;i<Num;i++)
+        {
+            pxL = new Particle(2.*L,
+                            ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
+                            ((gsl_rng_uniform(r)-0.5)*3*L)+0.5*L,
+                            gsl_ran_rayleigh(r,sigma)*(-1.),
+                            gsl_ran_gaussian(r,0.001)+AmbientTemp,
[…]

### 2017-01-27 src/main.cpp

@@ -35,10 +35,10 @@ int main(int argc,char** argv)
      */
 
     double t = 0.1;
-    double ambient = 0.09;
+    double ambient = 0.05;
     double press = 0.8;
-    double q = 0.02;
-    while(ambient > 0.01)
+    double q = 0.01;
+    while(q < 0.1)
     {
         setValues(t,q,0.2,press,ambient);
         std::cout << "==================================================" << std::endl;
@@ -48,7 +48,7 @@ int main(int argc,char** argv)
         std::cout << "at: " << ambient << std::endl;
         std::cout << "==================================================" << std::endl;
         mainLoop();
-        ambient -= 0.01;
+        q += 0.01;
     }
     //setValues(0.2,0.04,0.2,4,0.8);
     mainLoop();

### 2017-02-02 doc/draft2.tex
added test for 1 gas particle

@@
 and different pressures.\\
+With these four temperatures introduced, some questions arise about the relationship between these temperatures. For this theses, I want to focus on
+the question: how does the laser intensity influence the motion of the nano particle in the laser trap and its center of mass velocity?\\
+In the next section I will introduce the methods that will be used to simulate the system, followed by the results of the simulations and the
+conclusion of this thesis.
 

### 2017-02-02 scripts/data/create_gnuplot.sh

@@ -6,7 +6,9 @@ echo -e 'Tout=system("./get_number.sh GasOut '$datetime'_statistics.dat")+0' >> 
 echo -e 'Tcm=system("./get_number.sh T_COM '$datetime'_statistics.dat")+0' >> $datetime"_plot.gnuplot"
 echo -e 'plot "'$datetime'_histogram_vx.dat" with boxes, \
     sqrt(1./(2.*pi*Tcm))*exp(-x**2/(2*Tcm)), \
-    sqrt(1./(2.*pi*Tin))*exp(-x**2/(2*Tin)), \
-    sqrt(1./(2.*pi*Tout))*exp(-x**2/(2*Tout))' >> $datetime"_plot.gnuplot"
+    sqrt(0.1/(2.*pi*Tin))*exp(-x**2*0.1/(2*Tin)), \
+    sqrt(0.1/(2.*pi*Tout))*exp(-x**2*0.1/(2*Tout))' >> $datetime"_plot.gnuplot"
+    #sqrt(1./(2.*pi*Tin))*exp(-x**2/(2*Tin)), \
+    #sqrt(1./(2.*pi*Tout))*exp(-x**2/(2*Tout))' >> $datetime"_plot.gnuplot"
 
 

### 2017-02-02 scripts/data/vCOMhistogram.py

@@ -31,8 +31,8 @@ inTemp = np.mean(gasTempDataIn)
 outTemp = np.mean(gasTempDataOut)
 
 statistics = open(datetime+"statistics.dat","w")
-statistics.write("GasIn: " + str(inTemp*0.5)+"\n")
-statistics.write("GasOut: " + str(outTemp*0.5)+"\n")
+statistics.write("GasIn: " + str(inTemp)+"\n")
+statistics.write("GasOut: " + str(outTemp)+"\n")
 statistics.write("T_COM: " + str(2./3. * vSqdMean)+"\n")
 
 statistics.write("Mu_x " + str(np.mean(vCOMData_x))+"\n")

### 2017-02-02 scripts/data2/copy_data.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+rm [0-9]*
+
+cd ~/codes/OpticalTweezer/output/runs
+folders="$(ls -d */ | tail -n 8)"
+
+for i in $folders
+do
+    #cp $i"vCOMData.dat" ../../scripts/data/"vCOMData_"${i%/}".dat";
+    #cp $i"parameters.txt" ../../scripts/data/"parameters_"${i%/}".txt";
+    #cp $i"gasTempData.dat" ../../scripts/data/"gasTempData"${i%/}".dat";
+    cp $i"vCOMData.dat" ../../scripts/data2/${i%/}"_vCOMData.dat";
+    cp $i"parameters.txt" ../../scripts/data2/${i%/}"_parameters.txt";
+    cp $i"gasTempData.dat" ../../scripts/data2/${i%/}"_gasTempData.dat";
+done

### 2017-02-02 scripts/data2/create_stats.sh

@@ -1,0 +1,10 @@
+#!/bin/bash
+
+datetime="$(ls | grep -P -o "[0-9]{6}_[0-9]{4}" | sort -u)"
+for i in $datetime
+do
+    python vCOMhistogram.py $i
+    ./create_gnuplot.sh $i
+done
+
+

### 2017-02-02 src/functions.cpp

@@ -3676,6 +3676,7 @@ void verletBaroAccelerations(Particle* cube, std::list<Particle*> gas)
     
 
 }
+
 void verletBaro(Particle* cube, std::list<Particle*>& gas)
 {
     InitBarostatFull(gas);
@@ -3809,6 +3810,7 @@ void eHEXBaroNewTemp(Particle* cube, std::list<Particle*>& gas)
         for(unsigned int k=0;k<3;k++)
             (*gasIter)->v[k] += 0.5 * (*gasIter)->a[k] * dt / (*gasIter)->m;
 }
+
 void eHEXBaro(Particle* cube, std::list<Particle*>& gas)
 {
 	double eps[N][3];

### 2017-02-02 test/new_temp_test/compile_new_temp_test.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+rm combined/*
+g++  new_temp_test.cpp ../../src/functions.cpp ../../src/globals.cpp ../../src/classes.cpp -lgsl -lgslcblas -o  new_temp_test.o

### 2017-02-02 test/new_temp_test/new_temp_test.cpp

@@ -1,0 +1,327 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <cstdlib>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../../include/functions.hpp"
+#include "../../include/globals.hpp"
+#include "../../include/classes.hpp"
+
+void out(Particle* cube, int Run)
+{
+        std::string filename = "combined/combined_";
+        filename += numberToString(Run);
+        filename += ".xyz";
+        /*
+         *if(input != NULL)
+         *{
+         *    filename = "output/combined_";
+         *    std::string prefix(input);
+         *    filename += prefix;
+         *    filename += "/combined_";
+         *    filename += numberToString(Run);
+         *    filename += ".xyz";
+         *}
+         *else
+         *{
+         *    std::string filename = "output/combined/combined_";
+         *    filename += numberToString(Run);
+         *    filename += ".xyz";
+         *}
+         */
+        
+        FILE* combinedOut = fopen(filename.c_str(),"w");
+        //int printed = 0;
+        fprintf(combinedOut,"%ld\nFrame 0\n",N);
+        unsigned int t=0;
+        for(unsigned int j=0;j<N;j++)
+        {
+                fprintf(combinedOut,"%s\t\%lf\t\%lf\t%lf\t%d\n",(cube[j].name).c_str(),cube[j].r[0],cube[j].r[1],cube[j].r[2],cube[j].ID);
+                //printed++;
+        }
+
+        fclose(combinedOut);
+}
+void out(Particle* cube, Particle* gas, int Run)
+{
+        std::string filename = "combined/combined_";
+        filename += numberToString(Run);
+        filename += ".xyz";
+        /*
+         *if(input != NULL)
+         *{
+         *    filename = "output/combined_";
+         *    std::string prefix(input);
+         *    filename += prefix;
+         *    filename += "/combined_";
+         *    filename += numberToString(Run);
+         *    filename += ".xyz";
+         *}
+         *else
+         *{
[…]

### 2017-02-09 doc/draft2.tex

@@
 
+\subsection{Measurement}
+Now that the functionality of every individual part has been established, the main question can be tackled: how does the laser intensity influence the
+center of mass motion of the nano particle? And which influence does the surrounding gas exert?\\
+In order to investigate the origin of the motion of the center of mass, we first need to check whether the eHEX algorithm itself has any influence on
+the center of mass motion. This is done by applying the algorithm without the thermostat and calculating the center of mass velocity. This turns out
+to be the case, i.e. the center of mass motion stands still.\\
+To check the integrity of the algortihm, a simple case needs to be tested. Such a simple case is $\Delta Q = 0$, i.e. no energy begin pumped into the
+system. The expected outcome of this case is a thermal equilibration between the incoming gas particles, the outgoing gas particles and the center of
+mass motion of the nano particle with the incoming gas temperature being the reference value. 
+The measurement yields this result and backs up the claim. Furthermore, the internal temperature of the nano particle also equilibrates to the 
+temperature of the surrounding gas. This was checked on a system with pressure $P=0.8$, temperature of the incoming particles $T_\text{imp} = 0.05$ 
+and internal temperature of the nano particle (set with the velocity verlet) $T_\text{intern} = 0.2$. The resulting temperatures are:
+\begin{eqnarray*}
+    T_\text{em} = 0.0494 \pm 0.0009\\
+    T_\text{COM} = 0.051 \pm 0.108\\
+    T_\text{int} = 0.048 \pm 0.007\\
+\end{eqnarray*}
+
 %\begin{figure}[h]

### 2017-02-09 doc/draft2.tex

--- 2017-02-02 doc/draft2.tex
@@
             %\langle N_\text{fac}\rangle = \Delta t L^2 P \left(\frac{1}{2\pi m k_B T}\right)^\frac12
+            \label{eq:numberofparticles}
             \langle N_\text{fac}\rangle = \Delta t L^2 P \sqrt{\frac{1}{2\pi m k_B T}}
@@
 %\end{figure}
-As we are interested in the temperatures of the particles before and after the interaction with the nano particle, we need some way to differentiate those
-two states. This can be done when looking at the total energy of the gas particles.\\
-The gas particles are created on the surface of the simulation box, which is further away than the interaction range between the gas particles and the
-nano particle. Since the gas particles do not interact with each other, the total energy of the gas particles does not change until they reach the
-interaction range of the nano particle. There they undergo a change of energy due to the interaction with the nano particle and then make their way away
-from the nano particle. After the interaction, when they reach a point outside of the interaction range, the enrgy stays constant again until the
-particles reach the limits of the simulation box.\\
-Therefore, the life span of a single gas particle can be partioned into 3 stages: incoming, interacting, outgoing. In a typical scene of interaction
-all three of these stages are represented by different gas particles, as can be seen in Fig.~\ref{fig:barointeraction}.\\
+%As we are interested in the temperatures of the particles before and after the interaction with the nano particle, we need some way to differentiate those
+%two states. This can be done when looking at the total energy of the gas particles.\\
+%The gas particles are created on the surface of the simulation box, which is further away than the interaction range between the gas particles and the
+%nano particle. Since the gas particles do not interact with each other, the total energy of the gas particles does not change until they reach the
+%interaction range of the nano particle. There they undergo a change of energy due to the interaction with the nano particle and then make their way away
+%from the nano particle. After the interaction, when they reach a point outside of the interaction range, the enrgy stays constant again until the
+%particles reach the limits of the simulation box.\\
+%Therefore, the life span of a single gas particle can be partioned into 3 stages: incoming, interacting, outgoing. In a typical scene of interaction
+%all three of these stages are represented by different gas particles, as can be seen in Fig.~\ref{fig:barointeraction}.\\
+%\begin{figure}[h]
+    %\begin{center}
+        %\includegraphics[scale=0.4]{images/baro_170105_new_medium.png}
+        %\caption{Rendering of the application of the barostat algorithm. The red particles are incoming gas particles, that have not yet interacted, 
+            %pink particles are currently interacting and green particles are leaving the simulation box and
+        %moving away from the nano particle (blue). To render this image, a cut has been made to make a view into the simulation box possible.}
+        %\label{fig:barointeraction}
+    %\end{center}
+%\end{figure}
+%Since the gas particles are interacting with the nano particle, the temperatures and energies are values of interest. The energies are recorded using
+%histograms and the temperature is calculated instantaneously every 100th timestep.\\
+%The calculation of the energies of the incoming particles has to be done with care. The problem are particles that enter the simulation box, but fly
+%by the nano particle at a distance that is greater than the interaction range. Since the change of the energy is the interesting part, only the energy 
+%particles of that \textit{do} interact with the nano particle shall be calculated. This can be achieved by counting particles that enter the stage of
+%\textit{interacing}. One such histogram can be seen in Fig.~\ref{fig:histogram}.
+%\begin{figure}[h]
+    %\begin{center}
+        %\includegraphics[scale=0.4]{images/histogram_2.pdf}
+        %\caption{Histogram of total energies for incoming (red) and outgoing (blue) particles. The parameters for this histogram are $\Delta Q^*=0.04$,
+        %$T_\text{ambient}^*=0.08$, $P^*=0.8$. What can be seen here is that the outgoing particles have a higher count of particles with more total
+    %energy.}
+        %\label{fig:histogram}
+    %\end{center}
+%\end{figure}
+
+\subsection{Measurement}
+Now that the functionality of every individual part has been established, the main question can be tackled: how does the laser intensity influence the
+center of mass motion of the nano particle? And which influence does the surrounding gas exert?\\
+In order to investigate the origin of the motion of the center of mass, we first need to check whether the eHEX algorithm itself has any influence on
+the center of mass motion. This is done by applying the algorithm without the thermostat and calculating the center of mass velocity. This turns out
+to be the case, i.e. the center of mass motion stands still.\\
+To check the integrity of the algortihm, a simple case needs to be tested. Such a simple case is $\Delta Q = 0$, i.e. no energy begin pumped into the
+system. The expected outcome of this case is a thermal equilibration between the incoming gas particles, the outgoing gas particles and the center of
+mass motion of the nano particle with the incoming gas temperature being the reference value. 
+The measurement yields this result and backs up the claim. Furthermore, the internal temperature of the nano particle also equilibrates to the 
+temperature of the surrounding gas. This was checked on a system with pressure $P=0.8$, temperature of the incoming particles $T_\text{imp} = 0.05$, 
+internal temperature of the nano particle (set with the velocity verlet) $T_\text{intern} = 0.2$ and a system size of $N=32$ atoms in the nano
+particle. The resulting temperatures are:
+\begin{eqnarray*}
+    T_\text{em} = 0.0494 \pm 0.0009\\
+    T_\text{COM} = 0.051 \pm 0.108\\
+    T_\text{int} = 0.048 \pm 0.007\\
+\end{eqnarray*}
+The calculation of the temperature is done by using the following equations:
+\begin{eqnarray}
+    \label{eq:ekin}E_\text{Kin} &=& \frac{m}{2 M} \sum_i \vec{v}_i^{\, 2} \\
+    \label{eq:temp}E &=& \frac{f}{2} k_B T
+\end{eqnarray}
+In the first line, $m$ denotes the mass, which is $m=0.1$ for the gas particles, $m=1$ for the atoms in the nano particle and $m=32$ for the center
+of mass. $M$ denotes the number of particles that are being summed over, which is $M=1$ for the center of mass, $M=32$ for the internal temperature
+and fluctuates for the outgoing gas particles. The number of the outgoing particles is not constant and depends heavily on the pressure and the
+temperature of the impinging gas particles, as can be seen in \eqref{eq:numberofparticles}. In the second line, $f$ denotes the number of degrees of
+freedom of the system, which is $f=3$ for all three systems. Putting these two equations together, we get an equation for the temperature:
+\begin{equation}
+    T = \frac{m}{3 M} \sum_i \vec{v}_i^{\, 2} \\
+\end{equation}
+Since this describes an instantaneous value, the mean value of the velocities over the course of the simulation has to be used to get the mean value
+of the temperature of each subsystem. The simulation results compared with the mean values can be seen in
+Figs.~\ref{fig:outgoingequib},~\ref{fig:vcomsqdequib} and \ref{fig:internalequib}.
 \begin{figure}[h]
     \begin{center}
-        \includegraphics[scale=0.4]{images/baro_170105_new_medium.png}
-        \caption{Rendering of the application of the barostat algorithm. The red particles are incoming gas particles, that have not yet interacted, 
-            pink particles are currently interacting and green particles are leaving the simulation box and
-        moving away from the nano particle (blue). To render this image, a cut has been made to make a view into the simulation box possible.}
-        \label{fig:barointeraction}
+        \includegraphics[scale=0.4]{images/gastemp.pdf}
+        \caption{Instantaneous and mean temperature of the outgoing gas particles of the surrounding gas}
+        \label{fig:outgoingequib}
     \end{center}
 \end{figure}
-Since the gas particles are interacting with the nano particle, the temperatures and energies are values of interest. The energies are recorded using
-histograms and the temperature is calculated instantaneously every 100th timestep.\\
-The calculation of the energies of the incoming particles has to be done with care. The problem are particles that enter the simulation box, but fly
-by the nano particle at a distance that is greater than the interaction range. Since the change of the energy is the interesting part, only the energy 
-particles of that \textit{do} interact with the nano particle shall be calculated. This can be achieved by counting particles that enter the stage of
-\textit{interacing}. One such histogram can be seen in Fig.~\ref{fig:histogram}.
 \begin{figure}[h]
     \begin{center}
-        \includegraphics[scale=0.4]{images/histogram_2.pdf}
-        \caption{Histogram of total energies for incoming (red) and outgoing (blue) particles. The parameters for this histogram are $\Delta Q^*=0.04$,
-        $T_\text{ambient}^*=0.08$, $P^*=0.8$. What can be seen here is that the outgoing particles have a higher count of particles with more total
-    energy.}
-        \label{fig:histogram}
+        \includegraphics[scale=0.4]{images/vcomsqd.pdf}
+        \caption{Instantaneous and mean temperature of the center of mass of the nano particle in the laser trap}
+        \label{fig:vcomsqdequib}
     \end{center}
 \end{figure}
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.4]{images/t_int.pdf}
+        \caption{Instantaneous and mean internal temperature of the nano particle in the laser trap}
+        \label{fig:internalequib}
+    \end{center}
+\end{figure}
+\\With the calculated mean temperature of the center of mass, we can take a look at the distribution of the components of the center of mass velocity.
+The expected distribution is a gaussian with $\sigma=\sqrt{\frac{k_B T}{m}}$, i.e.
+\begin{equation}
+    \label{eq:gaussianvel}
+    p(v_i) = \frac{m}{2\pi k_B T} e^{-\frac{x^2 m}{2 k_B T}}
+\end{equation}
+where $i$ denotes the chosen component of the center of mass velocity vector. A histogram of the x-component of the center of mass velocity  
+with a curve representing \eqref{eq:gaussianvel} can be seen in Fig.~\ref{fig:comvel}. Since the y and z axis yield very similar histograms, the
+x-component has been chosen to represent the velocity components. The graph in Fig.~\ref{fig:comvel} clearly shows, that the velocities indeed follow
+the gaussian distribution.
+\begin{figure}[h]
+    \begin{center}
+        \includegraphics[scale=0.4]{images/vcomtemp.pdf}
+        \caption{A histogram of the x-component of the center of mass velocities with the expected gaussian distribution}
+        \label{fig:comvel}
+    \end{center}
+\end{figure}
+\\The case of $\Delta Q=0$ is a good way to start and to check the integrity of the algorithms, but the more interesting cases are the ones where
+$\Delta Q > 0$. Besides $\Delta Q$, there are two values that determine the dynamics of the system: the pressure $P$ and the temperature of the
+impinging gas particles $T_\text{imp}$.\\
+Since the surrounding gas acts as a thermostat, the pressure of the impinging gas at a given gas temperature determines the maximum amount of heat that
+can be pumped into the system without particles of the nano particle leaving the simulation box -- the higher the pressure, the more energy can be
+pumped into the system. The algorithm is applied at two surrounding gas temperatures of $T_\text{imp} = 0.05$ and $T_\text{imp} = 0.1$, 
+which are one quarter and one half the temperature of the nano particle before applying the eEHX and the thermostat algorithm. The pressure varied
+from $P=0.1$ to $P=1.2$ in steps of $\Delta P=0.1$ and the heat is generally varied from $\Delta Q=0$ to $\Delta Q=0.5$ in steps of 0.05. The term
+''generally'' is used because there are cases where the stepsize has to be decreased or one value for $\Delta Q$ is skipped over. The stepsize has to
+be decresed for low pressures, such as $P=0.1$ or $P=0.2$ because there would otherwise not be enough data points. Some values of $\Delta Q$ need to
+be skipped because it can happen that atoms of the nano particle leave the system due to statistical fluctuations of velocities, which leads to
+unusable results.\\
 

### 2017-02-10 doc/draft2.tex
updates

@@
 unusable results.\\
+To see the comparison between the different temperatures of the impinging gas temperatures, three values for the pressure are chosen: $P=0.1$, $P=0.5$
+and $P=1.0$. The simulation results for the center of mass temperature, the internal temperature and the temperature of the outgoing gas particles are
+plotted for each of the pressure values and both temperatures.
 

### 2017-02-10 scripts/p0.1_at0.05/vCOMhistogramMass.py

@@ -15,7 +15,10 @@ gasTempDataOut = np.genfromtxt(datetime+"gasTempData.dat",usecols=1,skip_header=
 vCOMData_x = np.genfromtxt(datetime+"vCOMData.dat",usecols=0,skip_header=100)
 vCOMData_y = np.genfromtxt(datetime+"vCOMData.dat",usecols=1,skip_header=100)
 vCOMData_z = np.genfromtxt(datetime+"vCOMData.dat",usecols=2,skip_header=100)
+internalTempData = np.genfromtxt(datetime+"temperature_internal.dat",skip_header=200)
 N = 32
+
+internalTemp = np.mean(internalTempData)
 
 vSqd = []
 for i in range(0,len(vCOMData_x)):
@@ -34,9 +37,10 @@ inTemp = np.mean(gasTempDataIn)
 outTemp = np.mean(gasTempDataOut)
 
 statistics = open(datetime+"statistics_mass.dat","w")
-statistics.write("GasIn: " + str(inTemp)+"\n")
-statistics.write("GasOut: " + str(outTemp)+"\n")
-statistics.write("T_COM: " + str(2./3. * vSqdMean)+"\n")
+statistics.write("GasIn: " + str(inTemp)+" +- " + str(np.std(gasTempDataIn)) + "\n")
+statistics.write("GasOut: " + str(outTemp)+str(np.std(gasTempDataOut)) + "\n")
+statistics.write("T_COM: " + str(2./3. * vSqdMean)+str(np.std(vSqd)) + "\n")
+statistics.write("T_INT: " + str(internalTemp)+str(np.std(internalTempData)) + "\n")
 
 statistics.write("Mu_x " + str(np.mean(vCOMData_x))+"\n")
 statistics.write("Sigma_x: " + str(np.std(vCOMData_x))+"\n")

### 2017-02-10 scripts/p0.8_at0.05/create_plot.sh

@@ -13,10 +13,14 @@ for i in $datetimes
 do
     q="$(cat "$i"_parameters.txt | grep "dQ" | cut -f3 -d " ")"
     tcom="$(cat "$i"_statistics_mass.dat | grep "COM" | cut -f2 -d " ")"
+    terr="$(cat "$i"_statistics_mass.dat | grep "COM" | cut -f4 -d " ")"
     ti="$(cat "$i"_statistics_mass.dat | grep "INT" | cut -f2 -d " ")"
+    interr="$(cat "$i"_statistics_mass.dat | grep "INT" | cut -f4 -d " ")"
     gin="$(cat "$i"_statistics_mass.dat | grep "In" | cut -f2 -d " ")"
+    ginerr="$(cat "$i"_statistics_mass.dat | grep "In" | cut -f4 -d " ")"
     gout="$(cat "$i"_statistics_mass.dat | grep "Out" | cut -f2 -d " ")"
-    echo -e $q"\t"$tcom"\t"$ti"\t"$gin"\t"$gout >> plotfile.dat
+    gouterr="$(cat "$i"_statistics_mass.dat | grep "Out" | cut -f4 -d " ")"
+    echo -e $q"\t"$tcom"\t"$terr"\t"$ti"\t"$interr"\t"$gin"\t"$ginerr"\t"$gout"\t"$gouterr>> plotfile.dat
 done
 
 cp plotfile.dat plotfile_clean.dat

### 2017-02-10 scripts/spring15_p0.5_at0.1/init.sh

@@ -1,0 +1,32 @@
+#!/bin/bash
+dir="$(pwd)"
+target="$(basename $dir)"
+#target=$1
+folders="$(ls -d ../../test/output/runs/$target*)"
+
+for i in $folders
+do
+    basename $i >> folderlist.txt
+done
+
+names="$(cat folderlist.txt)"
+
+for i in $names
+do
+    ./copy_data_date.sh $i
+done
+
+for i in $names
+do
+    python vCOMhistogramMass.py $i
+done
+
+./create_png.sh
+
+gnuplot -e "\
+    files=system('ls *.gnuplot');
+    do for [ i in files ] {
+        load i
+    }"
+
+./create_plot.sh

### 2017-02-10 scripts/vCOMhistogramMass.py

@@ -37,10 +37,10 @@ inTemp = np.mean(gasTempDataIn)
 outTemp = np.mean(gasTempDataOut)
 
 statistics = open(datetime+"statistics_mass.dat","w")
-statistics.write("GasIn: " + str(inTemp)+"\n")
-statistics.write("GasOut: " + str(outTemp)+"\n")
-statistics.write("T_COM: " + str(2./3. * vSqdMean)+"\n")
-statistics.write("T_INT: " + str(internalTemp)+"\n")
+statistics.write("GasIn: " + str(inTemp)+" +- " + str(np.std(gasTempDataIn)) + "\n")
+statistics.write("GasOut: " + str(outTemp)+str(np.std(gasTempDataOut)) + "\n")
+statistics.write("T_COM: " + str(2./3. * vSqdMean)+str(np.std(vSqd)) + "\n")
+statistics.write("T_INT: " + str(internalTemp)+str(np.std(internalTempData)) + "\n")
 
 statistics.write("Mu_x " + str(np.mean(vCOMData_x))+"\n")
 statistics.write("Sigma_x: " + str(np.std(vCOMData_x))+"\n")

### 2017-02-10 test/plot/baro_output.cpp

@@ -17,10 +17,10 @@ int main()
     Particle *particles = new Particle[N];
     InitPositions(particles);
     std::list<Particle*> gas;
-    FILE* tempout = fopen("baro_temp_004_trackdata_1.dat","w");
-    FILE* enerout = fopen("baro_energy_004_trackdata_1.dat","w");
-    FILE* compos = fopen("baro_compos004_trackdata.dat","w");
-    FILE* comvel= fopen("baro_comvel004_trackdata.dat","w");
+    FILE* tempout = fopen("baro_temp_004_trackdata_1_new.dat","w");
+    FILE* enerout = fopen("baro_energy_004_trackdata_1_new.dat","w");
+    FILE* compos = fopen("baro_compos004_trackdata_new.dat","w");
+    FILE* comvel= fopen("baro_comvel004_trackdata_new.dat","w");
 
     FILE* trackdata[1000];
     for(int i=0;i<1000;i++)

### 2017-02-13 doc/draft2.tex
updates

@@
 %\usepackage[font=footnotesize,labelfont=bf]{caption}
-\usepackage{subcaption}
+\usepackage[font=scriptsize]{subcaption}
 \usepackage{fancyhdr}
@@
 \newpage
+\tableofcontents
+\newpage
 \section{Introduction}
@@
 %systems on computers yada yada
+In the year 1953, Metropolis et al.~\cite{Metropolis1953} used the MANIAC (Mathematical Analyzer Numerical Integrator And Computer) 
+at the Los Alamos National Laboratories in the United States to
+simulate particles in a liquid state to calculate the equation of state. This was the first publication in which scientists used a computer to
+calculate numerical results based on a physical theory in a process that is to this day called \textit{Metropolis Monte Carlo}. To say that a lot has
+changed in the field of computer simulation (or computers in general) would be a vast understatement.\\
+Not only the processing power of computers has incresed exponentially (Moore's law), which allows the simulation of a greater number of particles for
+a longer period of time, but also new techniques have been developed, such as molecular dynamics. These techniques are not only used by physicists,
+but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. This thesis is an
+attempt to apply methods of simulation to a field with a lot of recent research happening in it: optical trapping.
+Optical trapping is a technique that has been used and developed for over 40 years~\cite{Marago2013}. The idea for this goes back to the beginning of
+the 20th century, where Lebedev~\cite{Lebedew1901}, Nichols and Hull~\cite{Nichols1901} demonstrated the existence of a force that light exerts on 
+matter. Due to the technical limitations at that time, the idea was discarded until 1970, where Ashkin~\cite{Ashkin2000} used lasers to manipulate the
+motion of neutral atoms and micrometer sized objects and thus brought the idea of optical trapping back to life.\\
+With this advancement it was possible to develop methods for trapping and manipulating objects of different sizes ranging from the micrometer scale
+down to the subnano scale. At first, this technique was used to either trap objects in the micrometer range (like cells) or in the subnano scale,
+where lasers were used for the cooling of atoms. The range between those two, with objects at the scale of one to a hundred nanometer, was not
+explored as much because neither downscaling the methods for micrometer sized objects nor upscaling the methods for atoms seemed feasible. In recent
+years however, new developments in the field opened the possibilities to work with a variety of objects in the desired nanometer scale, such as metal
+nanoparticles, nano tubes, quantum dots and much more. This non-invasive method of trapping objects allows very precise measurements of various
+properties and also allows the cooling of nanoparticles towards their quantum mechanical ground state.\\
+The small size of the objects used in the optical tweezer experiments make them ideal candidates for computer simulations. In this thesis, I will
+model an optical tweezer setup, based on a real experiment, to investigate different temperatures in a non-equilibrium state.\\
+The first chapter will be the introduction of the experiment and the introduction of the problem statement. The second chapter contains basic
+information on the techniques used to model the experiment and simulate the dynamic properties. In the third chapter, I will present my findings which
+are followed by the conclusions and further remarks.
+
+%Due to the properties of light it was possible
+%to create methods of trapping and manipulating objects of different sizes from the micrometer to the subnanometer-scale.
+
+%At first, the micrometer
+%ranges (objects like cells) and subnanometer-ranges (lasercooling of atoms) were explored, but the methods used in these ranges were not fit to be
+%applied to objects in the range of one to a hundred nanometers.
 
@@
 plotted for each of the pressure values and both temperatures.
+\begin{figure}[h]
+    \begin{center}
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.13]{images/p01_com.pdf}
+        \end{subfigure} 
+        \
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.13]{images/p05_com.pdf}
+        \end{subfigure} 
+        \
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.13]{images/p1_com.pdf}
+        \end{subfigure} 
+        \\
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.13]{images/p01_int.pdf}
+        \end{subfigure} 
+        \
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.13]{images/p05_int.pdf}
+        \end{subfigure} 
+        \
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.13]{images/p1_int.pdf}
+        \end{subfigure} 
+        \\
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.13]{images/p01_out.pdf}
+        \end{subfigure} 
+        \
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.13]{images/p05_out.pdf}
+        \end{subfigure} 
+        \
+        \begin{subfigure}[t]{0.3\textwidth}
+            \includegraphics[scale=0.13]{images/p1_out.pdf}
+        \end{subfigure} 
+    \end{center}
+\end{figure}
 

### 2017-02-13 scripts/data3/vCOMhistogramMass.py

@@ -15,11 +15,14 @@ gasTempDataOut = np.genfromtxt(datetime+"gasTempData.dat",usecols=1,skip_header=
 vCOMData_x = np.genfromtxt(datetime+"vCOMData.dat",usecols=0,skip_header=100)
 vCOMData_y = np.genfromtxt(datetime+"vCOMData.dat",usecols=1,skip_header=100)
 vCOMData_z = np.genfromtxt(datetime+"vCOMData.dat",usecols=2,skip_header=100)
+internalTempData = np.genfromtxt(datetime+"temperature_internal.dat",skip_header=200)
 N = 32
+
+internalTemp = np.mean(internalTempData)
 
 vSqd = []
 for i in range(0,len(vCOMData_x)):
-    vSqd.append(32*(vCOMData_x[i]*vCOMData_x[i]+vCOMData_y[i]*vCOMData_y[i]+vCOMData_z[i]*vCOMData_z[i])*0.5)
+    vSqd.append(32*(vCOMData_x[i]*vCOMData_x[i]+vCOMData_x[i]*vCOMData_x[i]+vCOMData_x[i]*vCOMData_x[i])*0.5)
 
 vSqdMean = np.mean(vSqd)
 
@@ -34,9 +37,10 @@ inTemp = np.mean(gasTempDataIn)
 outTemp = np.mean(gasTempDataOut)
 
 statistics = open(datetime+"statistics_mass.dat","w")
-statistics.write("GasIn: " + str(inTemp)+"\n")
-statistics.write("GasOut: " + str(outTemp)+"\n")
-statistics.write("T_COM: " + str(2./3. * vSqdMean)+"\n")
+statistics.write("GasIn: " + str(inTemp)+" +- " + str(np.std(gasTempDataIn)) + "\n")
+statistics.write("GasOut: " + str(outTemp)+" +- " +str(np.std(gasTempDataOut)) + "\n")
+statistics.write("T_COM: " + str(2./3. * vSqdMean)+" +- " +str(np.std(vSqd)) + "\n")
+statistics.write("T_INT: " + str(internalTemp)+" +- " +str(np.std(internalTempData)) + "\n")
 
 statistics.write("Mu_x " + str(np.mean(vCOMData_x))+"\n")
 statistics.write("Sigma_x: " + str(np.std(vCOMData_x))+"\n")

### 2017-02-13 scripts/p0.3_at0.05/vCOMhistogramMass.py

@@ -15,7 +15,10 @@ gasTempDataOut = np.genfromtxt(datetime+"gasTempData.dat",usecols=1,skip_header=
 vCOMData_x = np.genfromtxt(datetime+"vCOMData.dat",usecols=0,skip_header=100)
 vCOMData_y = np.genfromtxt(datetime+"vCOMData.dat",usecols=1,skip_header=100)
 vCOMData_z = np.genfromtxt(datetime+"vCOMData.dat",usecols=2,skip_header=100)
+internalTempData = np.genfromtxt(datetime+"temperature_internal.dat",skip_header=200)
 N = 32
+
+internalTemp = np.mean(internalTempData)
 
 vSqd = []
 for i in range(0,len(vCOMData_x)):
@@ -34,9 +37,10 @@ inTemp = np.mean(gasTempDataIn)
 outTemp = np.mean(gasTempDataOut)
 
 statistics = open(datetime+"statistics_mass.dat","w")
-statistics.write("GasIn: " + str(inTemp)+"\n")
-statistics.write("GasOut: " + str(outTemp)+"\n")
-statistics.write("T_COM: " + str(2./3. * vSqdMean)+"\n")
+statistics.write("GasIn: " + str(inTemp)+" +- " + str(np.std(gasTempDataIn)) + "\n")
+statistics.write("GasOut: " + str(outTemp)+" +- " +str(np.std(gasTempDataOut)) + "\n")
+statistics.write("T_COM: " + str(2./3. * vSqdMean)+" +- " +str(np.std(vSqd)) + "\n")
+statistics.write("T_INT: " + str(internalTemp)+" +- " +str(np.std(internalTempData)) + "\n")
 
 statistics.write("Mu_x " + str(np.mean(vCOMData_x))+"\n")
 statistics.write("Sigma_x: " + str(np.std(vCOMData_x))+"\n")

### 2017-02-13 scripts/p0.6_at0.05/create_plot.sh

@@ -13,10 +13,14 @@ for i in $datetimes
 do
     q="$(cat "$i"_parameters.txt | grep "dQ" | cut -f3 -d " ")"
     tcom="$(cat "$i"_statistics_mass.dat | grep "COM" | cut -f2 -d " ")"
+    terr="$(cat "$i"_statistics_mass.dat | grep "COM" | cut -f4 -d " ")"
     ti="$(cat "$i"_statistics_mass.dat | grep "INT" | cut -f2 -d " ")"
+    interr="$(cat "$i"_statistics_mass.dat | grep "INT" | cut -f4 -d " ")"
     gin="$(cat "$i"_statistics_mass.dat | grep "In" | cut -f2 -d " ")"
+    ginerr="$(cat "$i"_statistics_mass.dat | grep "In" | cut -f4 -d " ")"
     gout="$(cat "$i"_statistics_mass.dat | grep "Out" | cut -f2 -d " ")"
-    echo -e $q"\t"$tcom"\t"$ti"\t"$gin"\t"$gout >> plotfile.dat
+    gouterr="$(cat "$i"_statistics_mass.dat | grep "Out" | cut -f4 -d " ")"
+    echo -e $q"\t\t"$tcom"\t\t"$terr"\t\t"$ti"\t\t"$interr"\t\t"$gin"\t\t"$ginerr"\t\t"$gout"\t\t"$gouterr>> plotfile.dat
 done
 
 cp plotfile.dat plotfile_clean.dat

### 2017-02-13 scripts/p1_at0.05/vCOMhistogramMass.py

@@ -37,10 +37,10 @@ inTemp = np.mean(gasTempDataIn)
 outTemp = np.mean(gasTempDataOut)
 
 statistics = open(datetime+"statistics_mass.dat","w")
-statistics.write("GasIn: " + str(inTemp)+"\n")
-statistics.write("GasOut: " + str(outTemp)+"\n")
-statistics.write("T_COM: " + str(2./3. * vSqdMean)+"\n")
-statistics.write("T_INT: " + str(internalTemp)+"\n")
+statistics.write("GasIn: " + str(inTemp)+" +- " + str(np.std(gasTempDataIn)) + "\n")
+statistics.write("GasOut: " + str(outTemp)+" +- " +str(np.std(gasTempDataOut)) + "\n")
+statistics.write("T_COM: " + str(2./3. * vSqdMean)+" +- " +str(np.std(vSqd)) + "\n")
+statistics.write("T_INT: " + str(internalTemp)+" +- " +str(np.std(internalTempData)) + "\n")
 
 statistics.write("Mu_x " + str(np.mean(vCOMData_x))+"\n")
 statistics.write("Sigma_x: " + str(np.std(vCOMData_x))+"\n")

### 2017-02-13 test/no_ehex_test.cpp

@@ -34,27 +34,25 @@ int main(int argc,char** argv)
      *    }
      */
 
-    /*
-     *double t = 0.2;
-     *double ambient = 0.05;
-     *double press = 0.5;
-     *double q = 0.005;
-     *while(q < 0.5)
-     *{
-     *    setValues(t,q,0.2,press,ambient);
-     *    std::cout << "==================================================" << std::endl;
-     *    std::cout << "t: " << t << std::endl;
-     *    std::cout << "p: " << press << std::endl;
-     *    std::cout << "q: " << q << std::endl;
-     *    std::cout << "at: " << ambient << std::endl;
-     *    std::cout << "==================================================" << std::endl;
-     *    mainLoop();
-     *    q += 0.05;
-     *    //ambient -= 0.01;
-     *}
-     */
-    setValues(0.2,0.1,0.2,0.5,0.1);
-    mainLoop();
+    double t = 0.2;
+    double ambient = 0.1;
+    double press = 0.5;
+    double q = 0.1;
+    while(q < 0.4)
+    {
+        setValues(t,q,0.2,press,ambient);
+        std::cout << "==================================================" << std::endl;
+        std::cout << "t: " << t << std::endl;
+        std::cout << "p: " << press << std::endl;
+        std::cout << "q: " << q << std::endl;
+        std::cout << "at: " << ambient << std::endl;
+        std::cout << "==================================================" << std::endl;
+        mainLoop();
+        q += 0.05;
+        //ambient -= 0.01;
+    }
+    //setValues(0.2,0.5,0.2,1.2,0.1);
+    //mainLoop();
     return 0;
 }
 

### 2017-02-14 doc/draft2.tex
updates

@@
 The force in the y and z direction can be calculated analogously.\\
+As can be seen in Fig.~\ref{fig:lj}, the potential is very close to 0 at about $r/\sigma=2.5$. This means that any contribution of the potential
+energy after this point is vanishingly small. To save computing time it is common practice to cut off the potential at a certain point and don't
+calculate the contribution for distances beyond that. The cut-off distance for this thesis was set to $r = 2.5\sigma$.\\
 The initial configuration of the particles is a face centered cubic (FCC) lattice. A schematic of the FCC lattice is depicted in Fig. \ref{fig:fcc}.
@@
     \begin{center}
-        \includegraphics[scale=0.4]{images/baro_temp_start.pdf}
+        \includegraphics[scale=0.4]{images/baro_temp_long_new.pdf}
         \caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
@@
 and $P=1.0$. The simulation results for the center of mass temperature, the internal temperature and the temperature of the outgoing gas particles are
-plotted for each of the pressure values and both temperatures.
+plotted for each of the pressure values and both temperatures, which can be seen in~\ref{fig:comintout}.
 \begin{figure}[h]
@@
         \end{subfigure} 
-    \end{center}
-\end{figure}
+        \caption{The temperature of the center of mass, the internal temperature and the outgoing temperature of the surrounding gas as functions of
+            heat being pumped into the system, for pressures
+        $P=0.1$, $P=0.5$ and $P=1$ at impinging gas temperatures of $T_\text{imp}=0.05$ (red) and $T_\text{imp}=0.1$. The three columns each show 
+        a fixed pressure value and the rows show fixed measured quantities. As the pressure relates to the maximum amount of heat that can be injected
+    into the system, the ranges on the abscissa in one row is different for every plot.}
+        \label{fig:comintout}
+    \end{center}
+\end{figure}
+The three columns of the plot correspond to fixed pressure values and different measured quantities, whereas every row corresponds to a fixed measured
+quantity at different pressures. As previously mentioned, the amount of pressure relates to the number of surrounding particles and thus limits the
+amount of heat that can be pumped into the system. Therefore, the range of values in the abscissa differs for different pressures -- at $P=0.1$ the
+values of $\Delta Q$ range from $0$ to $0.02$ and at $P=1.0$ from $0$ to $0.5$.\\
+The first thing that is noticable, is the increase of the temperature of the center of mass, the internal temperature and the outgoing gas temperature
+with increasing $\Delta Q$. This makes sense, as the amount of heat that is put into the system is also transfered in the interaction between the nano
+particle and the surrounding gas.\\
+One observation that is not so obvious due to the different ranges of $\Delta Q$ values: all temperature values increase faster for low pressures and
+increase slower for higher pressures. Take the center of mass temperature for $P=0.1$ and $P=1$ at $T=0.05$ (red line) as an example: while it reaches
+a value of a $T_\text{com} = 0.1$ at about $\Delta Q = 0.015$ at $P=0.1$, it takes a $\Delta Q$ of a little under $0.1$ to achieve the same value.
+This can be observed for every one of the three quantities. This is to be expected, as the surrounding gas acts as a thermostat. \\
+The development of the different temperatures seem to be depending of the temperature of the surrounding gas as well. The blue curves (which
+correspont do a surrounding gas temperature of $T=0.1$) are growing slightly faster than the red curves ($T=0.05$).\\
+In their paper~\cite{MillenJ.2014} Millen et al. give a formula for the center of mass temperature:
+\begin{equation}
+    T_\text{COM} = \frac{T_\text{imp}^{3/2}+\tfrac\pi 8 \ T_\text{em}^{3/2}}{T_\text{imp}^{1/2}+\tfrac\pi8 T_\text{em}^{1/2}}
+\end{equation}
+This formula together with the measured values of $T_\text{imp}$ and $T_\text{em}$ can be used to compare the results. The comparison between the
+calculated and the measured value can be seen for a representative choice of parameters in Fig.~\ref{fig:measurevscalc}.
 
@@
 \newpage
-\section{Conclusion}
+\section{Conclusion and Further Remarks}
 % what does this all tell us? do we need to study more of this? 
+In this thesis I have modeled an optical tweezer experiment with methods from the field of computational physics. The experiment was broken down into
+different sections, that have been modeled separately. With the use of molecular dynamics I simulated the behaviour of a nano particle in a laser trap
+and investigated the effects of the laser power and temperature of the surrounding gas on the center of mass movement of the nano particle.\\
+The results show, that the center of mass temperature (and thus its movement) does indeed depend on the energy of the laser, as well as on the
+pressure and temperature of the surrounding gas.\\
+The development of the center of mass temperature shows a curve that seems to have a convex form, whereas the internal temperature shows a more
+concave behaviour. So while the internal temperature initially grows faster, it could be possible for the center of mass temperature to pass the
+internal temperature.\\
+There are several ways to explore this subject further. The first idea would be to decrease the pressure even more and investiate the case of very low
+energies, since this is not easily feasible in an experimental setup. Another idea would be to increase the number of atoms in the nano particle to
+see if the surface of the particle plays a role in the development of the temperatures. Furthermore, the interaction between the gas particles and the
+nano particle in the trap could be varied to include an attractive part in the potential. Generally, these simulations could be expanded by running
+them on faster computers to investigate long term behaviour of the systems and expand the ranges of the different values like pressure or laser
+energy.
 

### 2017-02-14 scripts/p1_at0.1/vCOMhistogramMass.py

@@ -62,7 +62,8 @@ for i in range(0,len(histogram_x)):
 # plt.hist(vCOMData_x,bins=100)
 # plt.figure(2)
 # plt.hist(vCOMData_y,bins=100)
-# plt.figure(3)
+#(histogram)):
+outfile.write(str(bins[i]) + "\t" + histogram[i] + "\n")plt.figure(3)
 # plt.hist(vCOMData_z,bins=100)
 # plt.show()
 # plt.figure(1)

### 2017-02-14 test/no_ehex_test.cpp

@@ -53,7 +53,7 @@ int main(int argc,char** argv)
      *    //ambient -= 0.01;
      *}
      */
-    setValues(0.2,0.02,0.2,0.5,0.05);
+    setValues(0.2,0.1,0.2,0.5,0.05);
     mainLoop();
     return 0;
 }
@@ -71,7 +71,7 @@ void mainLoop() {
      *system(makedir.c_str());
      *system("rm output/combined/ *");
      */
-    std::string folderName = "output/runs/p";
+    std::string folderName = "output/runs/test_p";
     folderName += numberToString(P);
     folderName += "_at";
     folderName += numberToString(AmbientTemp);
@@ -304,6 +304,7 @@ void mainLoop() {
         }
         if(run%100==0)
         {
+            gasTempLog(gas,run);
             temp = calcTempNew(cube,tempout); 
             if(!temp)
                 break;

### 2017-02-14 test/plot/baro_output.cpp

@@ -104,7 +104,7 @@ int main()
         //VelocityVerlet(particles);
         //BarostatNew(particles,gas);
         //harmonicTrap(rCM,vCM,rCMStart,particles);
-        eHEXBaro(particles,gas);
+        eHEXBaroNewTemp(particles,gas);
         calcTemp(particles,tempout);
         energy = calculateEnergies(particles);
         fprintf(enerout,"%lf\n",energy);

### 2017-02-14 test/plot/compile_baro_output_new.sh

@@ -1,0 +1,3 @@
+#!/bin/bash
+sources="$(cat sources.txt)"
+g++  baro_output_new.cpp  $sources -lgsl -lgslcblas -o  baro_output_new.o

### 2017-02-16 doc/draft2.tex
updates

@@
 This formula together with the measured values of $T_\text{imp}$ and $T_\text{em}$ can be used to compare the results. The comparison between the
-calculated and the measured value can be seen for a representative choice of parameters in Fig.~\ref{fig:measurevscalc}.
+calculated and the measured value can be seen for a representative choice of parameters in Fig.~\ref{fig:measurevscalc}. There it becomes obvious,
+that the measured center of mass temperature rises quicker than the calculated value. This is also the case for different pressures and impinging gas
+temperatures. The reason for this may be in the difference of the accomodation factor $\alpha$. In the paper, Millen et al. use an accomodation factor
+for a sphere with a value of $\alpha = 0.777$, whereas the accomodation factor for the setup in the simulation for the chosen values is around
+$\alpha = 0.016$. This means that the real and the modeled interaction between the gas particles and the nano particle differ from one another.
+\begin{figure}
+    \begin{center}
+        \includegraphics[scale=0.3]{images/measurevscalc.pdf}
+        \caption{Comparison between the measured (red) and calculated (blue) center of mass temperature according to \cite{MillenJ.2014}.}
+        \label{fig:measurevscalc}
+    \end{center}
+\end{figure}
 

### 2017-02-17 doc/abstract_de.tex
updates

--- 2017-02-13 doc/draft2.tex
@@
 \begin{document}
-\title{Simulation of Nano Particles in a Laser Trap}
-\author{Mathias H\"old, BSc.}
-\date{2016}
-\maketitle
-\thispagestyle{empty}
-\newpage
-\tableofcontents
-\newpage
-\section{Introduction}
-%something about computer science in general, optical tweezers and the need to study the
-%systems on computers yada yada
-In the year 1953, Metropolis et al.~\cite{Metropolis1953} used the MANIAC (Mathematical Analyzer Numerical Integrator And Computer) 
-at the Los Alamos National Laboratories in the United States to
-simulate particles in a liquid state to calculate the equation of state. This was the first publication in which scientists used a computer to
-calculate numerical results based on a physical theory in a process that is to this day called \textit{Metropolis Monte Carlo}. To say that a lot has
-changed in the field of computer simulation (or computers in general) would be a vast understatement.\\
-Not only the processing power of computers has incresed exponentially (Moore's law), which allows the simulation of a greater number of particles for
-a longer period of time, but also new techniques have been developed, such as molecular dynamics. These techniques are not only used by physicists,
-but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. This thesis is an
-attempt to apply methods of simulation to a field with a lot of recent research happening in it: optical trapping.
-Optical trapping is a technique that has been used and developed for over 40 years~\cite{Marago2013}. The idea for this goes back to the beginning of
-the 20th century, where Lebedev~\cite{Lebedew1901}, Nichols and Hull~\cite{Nichols1901} demonstrated the existence of a force that light exerts on 
-matter. Due to the technical limitations at that time, the idea was discarded until 1970, where Ashkin~\cite{Ashkin2000} used lasers to manipulate the
-motion of neutral atoms and micrometer sized objects and thus brought the idea of optical trapping back to life.\\
-With this advancement it was possible to develop methods for trapping and manipulating objects of different sizes ranging from the micrometer scale
-down to the subnano scale. At first, this technique was used to either trap objects in the micrometer range (like cells) or in the subnano scale,
-where lasers were used for the cooling of atoms. The range between those two, with objects at the scale of one to a hundred nanometer, was not
-explored as much because neither downscaling the methods for micrometer sized objects nor upscaling the methods for atoms seemed feasible. In recent
-years however, new developments in the field opened the possibilities to work with a variety of objects in the desired nanometer scale, such as metal
-nanoparticles, nano tubes, quantum dots and much more. This non-invasive method of trapping objects allows very precise measurements of various
-properties and also allows the cooling of nanoparticles towards their quantum mechanical ground state.\\
-The small size of the objects used in the optical tweezer experiments make them ideal candidates for computer simulations. In this thesis, I will
-model an optical tweezer setup, based on a real experiment, to investigate different temperatures in a non-equilibrium state.\\
-The first chapter will be the introduction of the experiment and the introduction of the problem statement. The second chapter contains basic
-information on the techniques used to model the experiment and simulate the dynamic properties. In the third chapter, I will present my findings which
-are followed by the conclusions and further remarks.
-
-%Due to the properties of light it was possible
-%to create methods of trapping and manipulating objects of different sizes from the micrometer to the subnanometer-scale.
-
-%At first, the micrometer
-%ranges (objects like cells) and subnanometer-ranges (lasercooling of atoms) were explored, but the methods used in these ranges were not fit to be
-%applied to objects in the range of one to a hundred nanometers.
-
-\newpage
-\section{Motivation}
-%introduction of the experiment, the problem and the idea
-The starting point of this thesis is an experiment conducted by Gieseler et al \cite{Gieseler2014}. It is an optical tweezer experiment, where the
-motion of a glass nanoparticle in a laser trap was used to investigate the fluctuation theorem\cite{Crooks1999}.
-\subsection{Experimental setup}
-In the experiment, a silica nano particle with a radius of about 75 nm and mass of about $3 \times 10^{-18}$kg is trapped in a laser beam within a
-vacuum chamber. The trapping of the silica nano particle is achieved by a gradient force of the
-laser beam acting on the particle. The experimental setup is depicted in Fig.~\ref{fig:setup}.\\
-The particle fluctuates within the trap in all three spatial directions. These fluctuations can be approximated
-such that they are decoupled, which means that they can be described by a one-dimensional Langevin equation:
-\begin{equation}
-    \label{eq:langevin}
-    \ddot{x} + \Gamma_0 \dot{x} + \Omega^2_0x = \frac 1 m \left(F_\text{fluct} + F_\text{ext}\right)
-\end{equation}
-%The experiment is set up as follows: a glass sphere with radius $r \approx 75$ nm and mass $m \approx 3 \times 10^{-18}$ kg is trapped in a laser beam
-%in a vacuum chamber. The initial state of the glass particle is prepared by modulation of the laser, such that the oscillation of the glass particle
-%is suppressed in every direction. This is done by measuring the oscillation of the particle in every direction (assuming that the oscillation in every
-%direction is decoupled) and using this information as feedback for the laser. This experimental setup can be seen in figure \ref{fig:setup}.
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[scale=0.3]{images/experimental_setup.jpg}
-        \caption{Experimental setup of the optical tweezer experiment. A silica nano particle is trapped in a laser beam via gradient force in a
-        vacuum. The feedback is used to cool down the particle and create a non-equilibrium steady state. In the first part of the experiment, the
-    feedback is turned off and the motion of the particle towards an equilibrated state is observed. In the second part of the experiment, the steady state
-of the particle is modified by a parametric drive. Both the parametric drive and the feedback are turned off and -- as in the first part -- the
-motion of the particle towards an equilibrated state is observerd. Figure taken from \cite{Gieseler2014}.}
-        \label{fig:setup}
-    \end{center}
-\end{figure}
-On the left hand side we have the position $x$ (and its derivatives $\dot{x}$, the velocity and $\ddot{x}$, the acceleration), the friction 
-coefficient $\Gamma_0$ and the angular frequency $\Omega_0$ that describes the fluctuation along the chosen axis. On
-the right hand side, there are two forces. The first one is $F_\text{fluct}$, which describes a stochastic force caused by interactions with the gas
-in the vacuum chamber. This force is given by
-\begin{equation}
-    \label{eq:ffluct}
-    F_\text{fluct} = \sqrt{2m\Gamma_0k_BT_0} \ \xi\left(t\right)
-\end{equation}
-where $T_0$ is the temperature of the heat bath (i.e., the surrounding gas in the vacuum chamber), $k_B$ is the Boltzmann constant and $\xi(t)$ is
-white noise, which obeys the equations $\left\langle\xi(t)\right\rangle= 0$ and $\left\langle\xi(t)\xi(t')\right\rangle= \delta(t-t')$, which means
-that it is a random force. The term $\Gamma_0$ appears in the formula \eqref{eq:ffluct} 
-due to the fluctuation-dissipation theorem, which links the damping rate to the stochastic force.\\
-%The external force $F_\text{ext}$ is part of the experimental setup, where the frequency of the fluctuation along an axis, $\Omega_0$, is measured and
-%used to suppress the motion along said axis. This causes a decrease in the particle fluctuations and thus acts as a cooling mechanism for the particle
-%in the trap. This process creates a non-equilibrium steady state $\rho_{ss}(u,\alpha)$, which is not known analytically. This state is the starting
-%point of this thesis.\\
-\subsection{Model of the Surrounding Gas}
-J. Millen et al.~\cite{MillenJ.2014} worked on a similar experimental setup (without the feedback mechanism) and investigated the heating of the
-particle in the trap as well as the gas surrounding the particle.\\
-They start from the premise that there are 4 different temperatures one has to consider in such an experiment: the temperature of the gas particles
-before interacting with the nano particle $T_\text{imp}$ (for \textit{impinging}), the temperature of the gas particles after 
-interaction $T_\text{em}$ (for emerging), the surface temperature of the nano particle $T_\text{sur}$ and the temperature of the 
-center of mass of the nano particle $T_\text{COM}$. The experimental setup
-and the different temperatures are depicted in Fig.~\ref{fig:levitation}.\\
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[scale=0.2]{images/nnano_millen.jpg}
-        \caption{The experimental setup used by Millen et al. In \textbf{a} the experiment is schematically depicted and \textbf{b} shows a photograph
-            of the actual experimental setup. In \textbf{c} the 4 different temperatures that are present during the experiment are depicted and
-        \textbf{d} shows the position of the particle in the trap over time. The image was taken from~\cite{MillenJ.2014}}
-        \label{fig:levitation}
-    \end{center}
-\end{figure}
-Since the two temperatures of the gas, $T_\text{imp}$ and $T_\text{em}$, are not equal, the gas surrounding the nano particle is not in a thermal
-equilibrium, which is in contrast to the assumption in many optical tweezer experiments. Since the gas particles only interact with the nano particle
-in the trap and not wich each other, this situation creates two heat baths of different temperatures with the nano particle acting as a mediator
-between those two baths. This situation is depicted in
-Fig.~\ref{fig:heatbaths}.
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[scale=0.5]{images/nano_nonequilibrium_cropped.jpg}
-        \caption{Schematic picture of the two heat baths. The (hot) nano particle interacts with the (cold) incoming gas particles and they leave with
-        a higher temperature. The incoming and outgoing particles do not interact with each other and thus create two separate heat baths with
-    different temperatures. Image modified from \cite{Kroy2014}}
-        \label{fig:heatbaths}
-    \end{center}
-\end{figure}
-The one dimensional Langevin equation \eqref{eq:langevin} has to be modeled to include the different friction coefficients of the two heat baths. The
-model derived in \cite{MillenJ.2014} is described in the following equation:
-\begin{equation}
-    M\ddot{x}(t) + M\left(\Gamma^\text{imp}+\Gamma^\text{em}\right)\dot{x}(t) + M\omega^2x(t) =F^\text{imp}+F^\text{em}
-\end{equation}
-where $x$ is the position of the nano particle, $\dot{x}$ and $\ddot{x}$ are the velocity and the acceleration, respectively, $M$ is the mass,
-$\omega$ is the trap frequency, $\Gamma^\text{imp}$ and $\Gamma^\text{em}$ are the damping constants for the 
-impinging particles and the emerging particles, respectively and $F^\text{imp}$ and $F^\text{em}$ are the associated noise terms, 
-which satisfy $\langle F^\text{imp}\rangle= \langle F^\text{em} \rangle = 0$ individually.\\
-Miller et al. investigate the dependence of the emerging temperature $T_\text{em}$ on the intensity of the laser for different sized nano particles
-and different pressures.\\
-With these four temperatures introduced, some questions arise about the relationship between these temperatures. For this theses, I want to focus on
-the question: how does the laser intensity influence the motion of the nano particle in the laser trap and its center of mass velocity?\\
-In the next section I will introduce the methods that will be used to simulate the system, followed by the results of the simulations and the
-conclusion of this thesis.
-
-%These four temperatures introduced in the paper, together with the surrounding pressure and the intensity of the laser 
-%The existence of the two temperatures of the gas implies that, contrary to the assumption in many of the experiments, the gas is not in a thermal
-%equilibrium. The Langevin equation (like the one in \eqref{eq:langevin}) has to be modeled k
-
-\newpage
-\section{Simulation}
-%simulation techniques and general concept of the simulation itself, used methods and so on 
-The problem at hand can be studied on an atomic level with the use of computer simulation. There is a variety of methods for computer simulations
-that are widely used, one of which being molecular dynamics (MD) simulations.\\ 
-The goal is to simulate the experiment depicted in Fig.~\ref{fig:setup} as accurately as possible. To achieve this, the setup has to be broken down
-into individual pieced that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, sitting on an FCC
-lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, as it acts as a trapping device as well
-as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, that acts as a thermostat in the simulation, that will be
-modeled as a box surrounding the nano particle. This box is filled with gas particles that interact with the nano particle via a soft-sphere potential
-and do not interact with one another.\\
-The following section will give a brief overview of the concepts used to simulate every part of the experiment.
-
-\subsection{Molecular Dynamics}
-Molecular dynamics~\cite{Frenkel2001} simulations is a technique for simulating, as the name suggests, the dynamics of a classical many-body system. In this case,
-classical means, that the trajectories of the individual particles are calculated using classical mechanics rather then quantum mechanics. For
-relatively big atoms/molecules this is a very good approximation, whereas for systems consisting of hydrogen or helium the effects of quantum
-mechanics cannot be neglected and other methods have to be used.\\
-The dynamics of the system are obtained by solving Newton's equations of motion for every particle.
-
-\subsection{The Nano Particle}
-The glass particle from the experiment will be modeled as a system of particles interacting via a Lennard-Jones pair potential, 
-\begin{equation}
-    \label{eq:lj}
-    U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right]
-\end{equation}
-where $\varepsilon$ is the depth of the potential well (and thus its unit is energy) and $\sigma$ is the distance at which the potential is zero. 
-The form of the potential and the relation to the parameters is depicted in Fig. \ref{fig:lj}.
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.2]{images/LJ-2.pdf}
-        \caption{The Lennard-Jones 12-6 potential from \eqref{eq:lj}. The x-axis is the particle distance divided by $\sigma$ and  the y-axis is the
-        potential divided by the depth of the potential well.}
-        \label{fig:lj}
-    \end{center}
-\end{figure}
-Since $\varepsilon$ and $\sigma$ are crucial parameters for the simulation and do not change over time, it is practical to use them to
-define the dimensions of the system. This means that the unit of distance is $\sigma$, the unit of energy is $\varepsilon$ and the unit of mass is 
-the mass of the simulated particle. The so called \textit{reduced units} can be constructed from these three parameters and put into relation to the
-original units. Here are some examples:
-\begin{itemize}
-    \item {distance:} $r^* = r/\sigma$
-    \item {potential energy:} $U^* = U/\varepsilon$
-    \item {temperature:} $T^* = k_B T/\varepsilon$
-    \item {time:} $t^* = t\sqrt{\varepsilon/(m\sigma^2)}$
-    \item {pressure:} $P^* = P\sigma^3/\varepsilon$
-    \item {density:} $\rho^* = \rho \sigma^3$
-\end{itemize}
-One very popular choice for the simulated atoms is Argon because it is an inert gas and the atoms behave 
-approximately like hard spheres which attract each other with weak van der Waals forces, which justifies the use of the Lennard-Jones potential. 
-Argon has a mass of m = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$m and $\varepsilon = 1.65 \times 10^{-21}$J.\\
-With the above introduced reduced units, the Lennard-Jones potential can be written as
-\begin{equation}
-    U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right].
-\end{equation}
-Since the reduced units will be used throughout the rest of this thesis, I will drop the asterisk henceforth.\\
-From the Lennard-Jones potential the corresponding force can be calculated by taking the derivative with respect to the direction of interest:
-\begin{eqnarray}
-    F_{x} &=& -\frac{\partial}{\partial x} U(r) \nonumber\\
-                &=& -\frac{\partial}{\partial x} 4\left[{r}^{-12} - {r}^{-6}\right] \nonumber\\
-                &=& -4 \left[(-12){r}^{-13} - (-6){r}^{-7}\right] \frac{\partial r}{\partial x} \nonumber\\
-                &=& 48 \left[r^{-13} - 0.5 \ r^{-7}\right] \frac{x}{r} \nonumber\\
-    \label{eq:ljforce} &=& 48 \left[r^{-14} - 0.5 \ r^{-8}\right] x
-\end{eqnarray}
-The force in the y and z direction can be calculated analogously.\\
-The initial configuration of the particles is a face centered cubic (FCC) lattice. A schematic of the FCC lattice is depicted in Fig. \ref{fig:fcc}.
-With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since once FCC cell 
-shares its atoms with its next neighbours, the number of atom per unit cell is 4 -- 1/8 of a particle on eight corners and 1/2 of a particle on six faces.
-The whole system of atoms is then created by repeating this cell structure. One convenient way is to arrange the unit cells in a cubic system, 
-so if there are $M$ FCC unit cells on one edge, the whole system consists of $M^3$ cells. Since there are 4 particles per cell, 
-there are ideal or so called \textit{magic numbers} for atoms for which this setup works perfectly: $N = 4M^3 = 4,32,108,256,500,864,\ldots$.\\
-\begin{figure}[h]
-    \begin{center}
-        \begin{subfigure}[t]{0.4\textwidth}
-            \includegraphics[scale=0.3]{images/fcc-2.pdf}
-            \caption{Schematic figure of a face centered cubic (FCC) lattice}
-            \label{fig:fcc}
-        \end{subfigure} 
-        \
-        %\hspace{60pt}
-        \begin{subfigure}[t]{0.4\textwidth}
-            \includegraphics[scale=0.3]{images/unit_cell.pdf}
-            \caption{Setup for successively building a FCC lattice}
-            \label{fig:unitcell}
-        \end{subfigure}
-    \end{center}
-\end{figure}
-%\begin{figure}
-    %\begin{center}
-        %\includegraphics[scale=0.3]{images/fcc_unit.png}
-        %\caption{Schematic of the number of particles contained in one FCC unit cell. The atoms on the edges are shared by 8 cells and the ones on the
-            %faces by 2 cells. Image taken from
-            %{https://upload.wikimedia.org/wikipedia/commons/2/27/Elementarzelle\_einer\_kubisch\_raumzentrierten\_Elementarzelle.png}}
-    %\end{center}
-%\end{figure}
-% do i need to credit the uni buffalo course? 
-% http://www.physics.buffalo.edu/phy411-506/
-There are several ways to achieve this initial configuration and the one used in this thesis \cite{buffalo} was to create a kind of unit cell 
-consisting of four atoms, as shown in Fig. \ref{fig:unitcell}, which can be described by a set of points
-\begin{eqnarray*}
-    p_1 &=& \{0,0,0\}\\
-    p_2 &=& \{0.5*a,0.5*a,0\}\\
-    p_3 &=& \{0.5*a,0,0.5*a\}\\
-    p_4 &=& \{0,0.5*a,0.5*a\}
-\end{eqnarray*}
-From the particle number $N$ and the number of FCC unit cells per edge $M$ the \textit{lattice constant} $a$ can be calculated
-\begin{equation}
-    a = \frac{L}{M}
-\end{equation}
-where $L$ is the side length of the cube that is the whole system and it is calculated via the density of the system
-\begin{equation}
-    L = \sqrt[3]{\frac{N}{\rho}}
-\end{equation}
-With the lattice constant and the 4 points of the FCC cell, all the particles can be put into place.
-
-\subsection{The Velocity-Verlet Algorithm}
-\label{section:verlet}
-When we look at the system from a microscopic standpoint, we see that it follows some kind of path in the phase space as time progresses. Every
-point in this space corresponds to a set of positions and momenta and the connection between two points corresponds to the evolution of the system
-from one state to another. As mentioned above, this evolution (the dynamics of the system) is a crucial element to molecular dynamics. Since the
-equations of motion cannot be solved analytically in general, we need to approximate the solution.\\
-The method used here is called finite difference approach. The trajectory of the system in the phase space is cut into finite pieces of length $\Delta
-t$ and the equations of motion are solved for every segment separately (see Fig. \ref{fig:finitedifference}).\\
-\begin{figure}
-    \begin{center}
-        \includegraphics[scale=0.5]{images/finite_approach.pdf}
-        \caption{Simplified graphical schematic of the finite difference approach. The evolution of the system from a point $(r^N(t),p^N(t))$ in the
-            phase space is approximated by slicing it up into pieces of length $\Delta t$. On every stop after the starting point ($t+\Delta t$, 
-            $t+2\Delta t$, $t+3\Delta t$,$\ldots$) the equations of motion can be solved numerically.}
-        \label{fig:finitedifference} 
-    \end{center}
-\end{figure}
-There are several ways to solve this kind of problem, but since we are interested in implementing it into a computer program the ideal solution
-should have some basic properties \cite{Allen1989}:
-\begin{itemize}
-    \item It should be fast and require little memory
-    \item It should permit the use of a large time step $\Delta t$
-    \item It should duplicate the classical trajectory as closely as possible
-    \item It should satisfy known conservation laws
-    \item It should be simple and easy to program
-\end{itemize}
-One algorithm that has all of the above mentioned features is the one proposed by Verlet \cite{Verlet1967}. In his paper, Verlet starts by Taylor
-expanding the coordinate vector $\mathbf{r}_i$ for one particle after one time step $\Delta t$:
-\begin{equation}
-    \label{eq:verlet}
-    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + \dot{\mathbf{r}}_i \Delta t + \frac12 \ddot{\mathbf{r}}_i \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i \Delta t^3 +
-    \mathcal{O}(\Delta t^4)
-\end{equation}
-Since the first derivative of the coordinate is the velocity, $\dot{\mathbf{r}}_i(t) = \mathbf{v}_i(t)$, and the second derivative of the coordinate 
-is the acceleration, $\ddot{\mathbf{r}}_i(t) = \mathbf{a}_i(t)$, using Newton's law $\mathbf{F}_i(t) = m_i \mathbf{a}_i(t)$ equation \eqref{eq:verlet} 
-can be written as:
-\begin{equation}
-    \label{eq:verletnew}
-    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i(t)
-    \Delta t^3+\mathcal{O}(\Delta t^4)
-\end{equation}
-The same calculation can be carried out for one time step before $t$:
-\begin{equation}
-    \label{eq:verletnewminus}
-    \mathbf{r}_i(t-\Delta t) = \mathbf{r}_i(t) - {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 - \frac1{3!} \dddot{\mathbf{r}}_i(t) \Delta t^3+
-    \mathcal{O}(\Delta t^4)
-\end{equation}
-The sum of those two equations yields
-\begin{equation}
-    \mathbf{r}_i(t+\Delta t) + \mathbf{r}_i(t-\Delta t) = 2\mathbf{r}_i(t) + \frac1{2} {\mathbf{F}}_i(t) \Delta t^2 + \mathcal{O}(\Delta t^4)
-\end{equation}
-and from this we get the final form for the new coordinates:
-\begin{equation}
-    \label{eq:verletfinal}
-    \mathbf{r}_i(t+\Delta t) = 2\mathbf{r}_i(t)+ \mathbf{r}_i(t-\Delta t)+ \frac1{2} {\mathbf{F}}_i(t) \Delta t^2
-\end{equation}
-This means that the new coordinates can be calculated using the current coordinates, the current forces and the coordinates from the past time step.\\
-From equation \eqref{eq:verletfinal} one thing becomes clear: the velocities are not necessary to calculate the new positions and they are not
-computed in the process, but for the calculation of i.e. the kinetic energy the velocities are needed. Thus, the velocities have to be calculated with
-a combination of the Taylor expansions of the coordinate vectors at $t+\Delta t$ and $t-\Delta t$:
-\begin{equation}
-    \mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t) = 2\mathbf{v}_i(t)\Delta t + \mathcal{O}(\Delta t^3)
-\end{equation}
-which can be rewritten as
-\begin{equation}
-    \label{eq:velocities}
-    \mathbf{v}_i(t) = \frac{\mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t)}{2\Delta t} + \mathcal{O}(\Delta t^2)
-\end{equation}
-This approach has the advantage that it is fast, requires little memory and is reliable in the sense that there is no energy drift occuring during the
-simulation, which means that the energy is conserved. When we compare this to the list of desired properties this seems like a good algorithm.\\
-This algorithm however has two significant disatvantages. The first one is the calculation of the velocities. As can be seen in equation
-\eqref{eq:velocities}, the accuracy of the calculation is only $\mathcal{O}(\Delta t^2)$ while the positions can be calculated with an error of order
-$\Delta t^4$. The other big disatvantage is the first step of the algorithm. Since the calculation of the new positions requires the current positions
-and the ones from one time step before, which technically do not exist.\\
-The solutions to this problem is to include the stepwise calculation of the velocities \cite{swope1982}. For this we start again with the Taylor
-expansion of the coordinates mathbftor \eqref{eq:verletnew} and instead of Taylor expanding $\mathbf{r}_i(t-\Delta t)$, we write it as 
-\begin{equation}
-    \mathbf{r}_i(t) = \mathbf{r}_i(t+\Delta t) - \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
-\end{equation}
-Using \eqref{eq:verletnew} in the abvove equation we get
-\begin{equation}
-    \begin{aligned}
-        \mathbf{r}_i(t) &= \mathbf{r}_i(t) + \mathbf{v}_i(t) \Delta t + \frac1{2m}\mathbf{F}_i(t)\Delta t^2 \\
-                     &- \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
-    \end{aligned}
-\end{equation}
-and thus
-\begin{equation}
-    \label{eq:vvelocities}
-    \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \left(\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\right) \Delta t
-\end{equation}
-With the addition of this equation this algorithm is called the Velocity-Verlet algorithm, which is summarized in the following two equations:
-\begin{equation}
-    \label{eq:velocityverlet}
-    \begin{aligned}
-        \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2\\
-        \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \Big[\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\Big] \Delta t
-    \end{aligned}
-\end{equation}
-This algorithm is self starting, uses a small amount of memory, conserves the energy (not exactly) 
-and gives a very good approximation to the originial trajectory
-in the phase space. To program this algorithm the following steps are needed: 
-\begin{enumerate}
-    \item Calculate all the forces between the particles (for the first step only)
-    \item Calculate the new positions with current velocity and forces
-    \item Calculate first half of the new velocities with current forces
-    \item Calculate all the forces for the new position
-    \item Use new forces to calculate second half of new velocities
-\end{enumerate}
-The first point only has to be carried out for the first step, because the forces have not been calculated at this point. As the forces are calculated
-for the new positions in step 4, they can be used for the next time step. For the first time step the velocities have to be chosen randomly and are 
-calculated with this algorithm from that point forward.
-
-\subsection{The Laser Beam - Energy Influx}
-The nano particle is trapped in the laser beam. While the motion of the center of mass is localized, the
-individual atoms that make up the glass sphere absorb the energy from the laser which increases their velocity.\\
-To simulate this kind of behaviour, thermostat algorithms \cite{Huenenberger2005} such as the Nos\'e-Hoover \cite{Nose1984,Hoover1985} or the Andersen
-\cite{Andersen1980} algorithms are often used in simulation to change the temperature of the system in a controllable way.\\
-The problem with these kind of algorithms is, that a target temperature has to be fixed which will be reached at some point. In order to simulate the
-influx of energy from the laser it would be better to have an algorithm that supplies the system with a certain amount of energy continuously.
-Fortunately, such an algorithm exists.\\
-The algorithm is called Heat Exchange Algorithm (HEX) \cite{Hafskjold1993}. Its intended purpose is the use in non-equilibrium molecular dynamics
-(NEMD) to study transport phenomena and determine transport coefficients. The algorithm works by introducing two regions in the system, one serving as
-a heat source and the other as a heat sink. A specific amount of heat is then exchanged between those two reservoirs. As it turns out however, this
-algorithm introduces an energy shift for longer simulation times. This led Wirnsberger et al. \cite{Wirnsberger2015} to revisit the algorithm and 
-identify the cause of this energy drift, to create a more suitable algorithm, which they called eHEX.\\
-As in the HEX algorithm, regions are introduced to the system which act either as heat sinks or heat sources. These are labelled with $\Gamma_k$,
-where $k > 0$, and have corresponding amount of exchanged heat $\Delta Q_{\Gamma_k}$. If $\Delta Q_{\Gamma_k}$ is negative, heat is subtracted from
-the system and vice versa. Regions that neither act as heat source or heat sink are labelled with $\Gamma_0$, which are also called
-Hamiltonian regions (see Fig. \ref{fig:ehex}). 
-\begin{figure}
-    \begin{center}
-        \includegraphics[scale=0.3]{images/ehex_system.png}
-        \caption{Setup for the use of the HEX and eHEX algorithms. The simulationbox $\Omega$, which is moving with a velocity of $v_\Omega$,
-        contains a heat source (red), $\Gamma_1$, which is moving with velocity $v_{\Gamma_1}$ and a heat sink (blue), $\Gamma_2$, which is moving 
-        with velocity $v_{\Gamma_2}$. The regions that are neither heat sources or heat sinks are denoted by $\Gamma_0$.}
-        \label{fig:ehex} 
-    \end{center}
-\end{figure}
-The centers of mass of the particles in simulation box, denoted by $\Omega$, and the regions $\Gamma_k$ 
-are assumed to be moving with velocities $v_\Omega$ and $v_{\Gamma_k}$ respectively.\\
-The change of the energy in a region $\Gamma_k$ is achieved by rescaling the velocities by a factor $\xi_k$ and shifted by the velocity of the
-corresponding region:
-\begin{equation}
-    \mathbf{v}_i \rightarrow \mathbf{\bar{v}}_i = \xi_k \mathbf{v}_i + (1-\xi_k)\mathbf{v}_{\Gamma_k}
-\end{equation}
-The bar over a quantity denotes the value after the exchange of heat.\\
-The factor $\xi_k$ is given by
-\begin{equation}
-    \xi_k = \sqrt{1+\frac{\Delta Q_{\Gamma_k}}{\mathcal{K}_{\Gamma_k}}}
-\end{equation}
-where $\Delta Q_{\Gamma_k}$ is the exchanged heat in the region $\Gamma_k$ and $\mathcal{K}_{\Gamma_k}$ is the non-translational kinetic energy of the
-region $\Gamma_k$ and is given by
-\begin{equation}
-    \label{eq:kineticK}
-    \mathcal{K}_{\Gamma_k} = \sum_{i \in \gamma_k} \frac{m_i v_i^2}{2} - \frac{m_{\Gamma_k}v_{\Gamma_k}^2}{2}
-\end{equation}
-The sum is taken over all indices in $\gamma_k$ which is the set of indices of particles in the region $\Gamma_k$.\\
-For the final version of the eHEX algorithm there are three more quantities needed. The first one is the heat flux per time step, denoted by 
-$\mathcal{F}_\GammaK$:
-\begin{equation}
-    \mathcal{F}_\GammaK = \frac{\Delta Q_\GammaK}{\Delta t}
-\end{equation}
-The second one is the thermostatting force $\boldsymbol{\eta}_i$, which is defined as
-\begin{equation}
-    \boldsymbol{\eta}_i = 
-    \begin{cases}
-        m_i \ \dfrac{{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}}{{2\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}}} \
-        \left(\mathbf{v}_i - \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)
-        & \quad \text{if } k(\mathbf{r}_i) > 0 \\
-        0 & \quad \text{otherwise}
-    \end{cases}
-    %\begin{cases}
-        %m_i \ \frac{\displaystyle{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}}{\displaystyle{2\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}}} \
-        %\left(\mathbf{v}_i - \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)
-        %& \quad \text{if } k(\mathbf{r}_i) > 0 \\
-        %0 & \quad \text{otherwise}
-    %\end{cases}
-\end{equation}
-where $k(\mathbf{r}_i)$ is the index of the region in which particle $i$ located, i.e. $k(\mathbf{r}_i) = 0$ means that the particle is in a Hamiltonian
-region and $k(\mathbf{r}_i) > 0$ denotes the heat sinks and sources.\\
-The last quantity is the one that corrects the long term energy drift of the HEX algorithm, denoted by $\mathcal{E}{r}_{i,\alpha}$.
-The analysis and derivation of this term is given in \cite{Wirnsberger2015}.
-\begin{equation}
-    \label{eq:bigeps}
-    \begin{aligned}
-        \mathcal{E}{r}_{i,\alpha} &= \dfrac{\eta_{i,\alpha}}{m_i\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}} \left[
-        \frac{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}{48} + \frac{1}{6} \sum_{j\in\gamma_{k(\mathbf{r}_i)}} \mathbf{f}_j \cdot \left(\mathbf{v}_j - 
-            \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)\right] \\
-            &- \frac{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}{12\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}} \left[
-            \frac{f_{i,\alpha}}{m_i} - \frac{1}{m_{\Gamma_{k(\mathbf{r}_i)}}} \sum_{j\in\gamma_{k(\mathbf{r}_i)}}f_{j,\alpha}\right]
-    \end{aligned}
-\end{equation}
-The $\mathbf{f}$ in the above equation denotes the force corresponding to the chosen intermolecular potential $U(\mathbf{r})$, 
-\begin{equation}
-    \mathbf{f} = -\nabla_{\mathbf{r}_i} U(\mathbf{r}_i)
-\end{equation}
-With all the necessary quantities established, the updating sequence of the eHEX algorithm can be written down:
-\begin{subequations}
-\begin{eqnarray}
-    \bar{\mathbf{v}}^n_i &=& \xi^n_{k(\mathbf{r}_i)} \mathbf{v}^n_i + \left(1-\xi^n_{k(\mathbf{r}_i)}\right) \mathbf{v}^n_{\Gamma_{k(\mathbf{r}_i)}} \\
-    \bar{\mathbf{v}}^{n+\frac12}_i &=& \bar{\mathbf{v}}^n_i + \frac{\Delta t}{2m_i} \mathbf{f}^n_i\\
-    \bar{\mathbf{r}}^{n+1}_i &=& \mathbf{r}^n_i + \Delta t \bar{\mathbf{v}}^{n+\frac12}_i\\
-    \mathbf{f}^{n+1}_i &=& \left.-\nabla_{\mathbf{r}_i} U(\mathbf{r})\right|_{\mathbf{r} = \bar{\mathbf{r}}^{n+1}}\\
-    \bar{\mathbf{v}}^{n+1}_i &=& \bar{\mathbf{v}}^{n+\frac12}_i + \frac{\Delta t}{2m_i} \mathbf{f}^{n+1}_i\\
-    \mathbf{v}^{n+1}_i &=& \bar{\xi}^{n+1}_{k(\bar{\mathbf{r}}_i)} \bar{\mathbf{v}}^{n+1}_i + \left(1-\bar{\xi}^{n+1}_{k(\bar{\mathbf{r}}_i)}\right)
-    \bar{\mathbf{v}}^{n+1}_{\Gamma_{k(\bar{\mathbf{r}}_i)}} \\
-    \mathbf{r}^{n+1}_i &=& \bar{\mathbf{r}}^{n+1} - \Delta t^3 \mathcal{E}\bar{\mathbf{r}}^{n+1}_i
-\end{eqnarray}
-\end{subequations}
-This algorithm can be adapted for the problem of the levitating nano sphere in the laser beam by adjusting the setup and some of the parameters.\\
-Firstly, the setup of the heat sinks and sources has to be changed. Since there is only energy pumped into the system from the outside, the region
-acting as a heat sink vanishes and the region acting as a heat source spans over the whole simulation box. This means that, using the 
-notation of Fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither center of mass of the particles the simulation box nor 
-the heat source are moving, i.e. 
-$\mathbf{v}_\Omega = \mathbf{v}_{\Gamma_1} = 0$. This affects the non-translational kinetic energy term $\mathcal{K}_{\Gamma_1}$, the thermostatting
-force $\boldsymbol{\eta}$ and the correction term $\mathcal{E}\mathbf{r}$. Since there is only one region acting as a heat source, the terms $\Delta Q$,
-$\mathcal{K}$ and $\mathcal{F}$ don't need an index. The summation index in \eqref{eq:kineticK} and \eqref{eq:bigeps} can be changed to the number of
-particles in the system, $N$, since all the particles are within the heat exchanging region. The masses are set to 1, i.e. $m_i = 1$ with the chosen
-reduced units and with this the total mass of the region (in the term $1/m_{\Gamma_{k(\mathbf{r}_i)}}$ in \eqref{eq:bigeps}) is equal to the number of
-particles in the system. With these changes, the quantities can be written down as:
-\begin{eqnarray}
-    \mathcal{K} &=& \sum_{N} \frac{v_i^2}{2}\\
-    \xi &=& \sqrt{1+\frac{\Delta Q}{\mathcal{K}}}\\
-    \mathcal{F} &=& \frac{\Delta Q}{\Delta t}\\
-    \boldsymbol{\eta}_i &=& 
-    \dfrac{\mathcal{F}}{{2\mathcal{K}}} \
-        \mathbf{v}_i\\
-        \mathcal{E}{r}_{i,\alpha} &=& \dfrac{\eta_{i,\alpha}}{\mathcal{K}} \left[
-        \frac{\mathcal{F}}{48} + \frac{1}{6} \sum_{N} \mathbf{f}_j \cdot \mathbf{v}_j
-            \right] \nonumber \\
-            &-& \frac{\mathcal{F}}{12\mathcal{K}} \left[
-            {f_{i,\alpha}} - \frac{1}{N} \sum_{N}f_{j,\alpha}\right]
-\end{eqnarray}
-This means that the laser is modeled to be pumping energy into the system, which increases the velocities of the particles in the
-system over time, while the center of mass motion is not affected by this.\\
-
-\subsection{The Laser Beam - Trapping}
-As mentioned above, the glass particle is trapped in the laser beam and the position of the particle is localized. This behaviour
-has to be modeled as well.\\
-The approximation of the original paper \cite{Gieseler2014}, where the movements in the three spatial directions are decoupled will be
-used in the model as well. As to the model of the trapping force itself, the most straightforward approach is to use a harmonic oscillator potential.
-The force and the corresponding potential can be written as
-\begin{eqnarray}
-    \mathbf{F} &=& -k \Big[\mathbf{x}-\mathbf{x}_0\Big]\\
-    U &=& \frac12 k \Big[\mathbf{x}-\mathbf{x}_0\Big]^2
-\end{eqnarray}
-where $\mathbf{x}_0$ is the position of minimal potential energy. 
-Since the trap is acting on the whole system, the force is acting on the center of mass. 
-To calculate the center of mass positions and velocities, the positions and velocities of all particles have to be summed up and 
-divided by the particle number:
-\begin{equation}
-    \mathbf{r}_\text{COM} = \frac1N\sum_{i=1}^N \mathbf{r}_i
-\end{equation}
-
-\subsection{Surrounding Gas - Thermostat}
-Without any equilibrating mechanism the setup described by now would lead to the system heating up indefinetly, which is not desirable. Furthermore,
-the goal is to recreate the experiment as close as possible. Since the nano particle is trapped in a laser beam within a vacuum chamber, the
-surrounding gas has to be modeled as well. In the next step we will introduce the surrounding gas of the gas chamber that will absorb some of the 
-energy in the system, leading to the final state.\\
-Generally, pressure is introduced to the system by surrounding the object of interest (in this case the glass nano sphere) with a thermostatting
-pressure medium.
-There are two main requirements for the choice of such a pressure medium: the exerted pressure must be hydrostatic and the computation of the
-interaction between the pressure medium and the object of interest must not take up a lot of resources.\\
-The model used in this thesis was developed by Gr\"unwald and Dellago \cite{Gruenwald2006} and uses an ideal gas of non-interacting particles as
-thermostatting pressure medium. The particles of this pressure medium flow into the simulation from an outside volume, whose geometry is based on the form of the
-object of interest (this will be referred to as the minimal volume of cells) and leave the simulation box if their position reaches the boundary of
-this minimal volume of cells. This behaviour is very close to the real experiment which makes this thermostat an ideal candidate for the simulation of
-the experiment. 
-%This barostat also acts as a thermostat and thus is suitable for the 
-%application in this thesis.\\
-The gas particles interact with the object via a soft-sphere potential of the form
-\begin{equation}
-    \label{eq:softsphere}
-    U(r) = \varepsilon \left(\frac{\sigma}{r}\right)^{12}
-\end{equation}
-where $\varepsilon$ is the interaction strength and $\sigma$ is the interaction range and $r$ is the distance between the gas particle and the
-interacting particle. The gas particles do not interact with one another in this model.\\
-In order to increase the efficiency of the computing process, $\sigma$ should be chosen carefully. For larger $\sigma$, the number of interaction
-partners increases, which increases the number of force calculations which are a very time consuming part. For smaller $\sigma$ the possibility for
-gas particles reaching the inside of the nano particle increases, which is not desirable. So $\sigma$ should be chosen small enough to keep the force
-calculations at a minimum and large enough for the particle to stay on the outside of the crystal. In the computer simulations performed for this
-thesis, the interaction length is chosen to be $\sigma = 1$ with a cut-off radius of $r_c = 2.5$.\\ 
-The algorithm can be performed by following these steps:
-\begin{enumerate}
-    \item Randomly draw the number of particles that are created on a single side of the minimal volume of cells, $N_\text{fac}$, from the
-        distribution 
-        \begin{equation}
-            %\langle N_\text{fac}\rangle = \Delta t L^2 P \left(\frac{1}{2\pi m k_B T}\right)^\frac12
-            \label{eq:numberofparticles}
-            \langle N_\text{fac}\rangle = \Delta t L^2 P \sqrt{\frac{1}{2\pi m k_B T}}
-        \end{equation}
-        where $\Delta t$ is the time step of the simulation, $L$ is the side length of the cell in which the particle is created, $P$ is the desired
-        pressure, $m$ is the mass of the gas particle, $k_B$ is the Boltzmann constant (which will be set to 1 in reduced units) and $T$ is the
-        desired temperature. This chosen number of particles is then equally distributed over the face of the cell on which they are created. The
-        velocities of the created particles are drawn from two different random number distributions. The component of the particle perpendicular to
-        the surface is drawn from a Rayleigh distribution of the form 
-        \begin{equation}
-            p(v_i) = \frac{m}{k_B T}v_i \ e^{-\frac{mv_i^2}{2k_BT}}
-        \end{equation}
-        The other components of the particles' velocity are drawn from a Maxwell-Boltzmann distribution.
-    \item Perform the first step of the velocity Verlet algorithm to propagate the particle positions by one time step.
-    \item Check if any gas particles have left the minimal volume of cells and remove those which have.
-    \item Check if the geometry of the crystal and with it the minimal volume of cells has changed. If it has, remove all gas particles in the cells
-        that are no longer needed. Then insert new gas particles to the created cells with a number drawn from a Poisson distribution with mean value
-        \begin{equation}
-            \langle N_\text{ins}\rangle = \frac{PL^3}{k_BT}
-        \end{equation}
-        If the length of the cell $L$ is chosen to be equal to the cut-off radius $r_c$, this insertion should only be carried out with a probablity of
-        \begin{equation}
-            P_\text{ins} = e^{-\frac{U}{k_BT}}
-        \end{equation}
-        (where U is the interaction energy between the gas particle and the crystal) because it is possible that the inserted particle is 
-        within the interaction range of the crystal. 
-    \item Compute all forces.
-    \item Perform the second step of the velocity Verlet and propagate the particle velocities by one time step. 
-\end{enumerate}
-This algorithm is formulated in a general fashion, so that a wide range of crystals and geometries can be used. 
-%The setup used in this thesis is a FCC lattice, which has a nice symmetry that is not influenced by the surrounding gas. 
-Although the shape of the nano particle changes over the course of the simulation, it is not necessary to change the geometry of the surrounding box.
-The algorithm will therefore not be carried out using cell-lists and a minimal volume of cells that is changing over the course of the simulation, 
-but rather a fixed setup of the volume surrounding the nano particle.\\ 
-Since the particle itself is modeled as a cube, a straighforward approach is surrounding it by a bigger cube. The distance between the cube face and the nano
-particle is chosen to be the side length of the nano particle, so the setup will properly scale for different numbers of particles. 
-The schematic setup for this is depicted in Fig.\ref{fig:gascube}.
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.5]{images/gas_volume_ver3.pdf}
-        \caption{Schematic setup for the barostat. The outer volume (blue) is one side length $s$ away from the nano particle (black) which means that
-            the side length of the face of the surrounding box is $L=3s$. 
-                    The gas particles are created along the blue lines of the outer volume. Since the overall geometry of the nano particle does 
-                    not change, 
-                    the geometry of the outer volume will be constant throughout the simulation.}
-        \label{fig:gascube}
-    \end{center}
-\end{figure}
-
-%This is done by adding for instance two regions to the simulation, one providing an influx of heat
-%and the other acting as a heat sink. The algorithm is furthermore designed to incorporate the velocities of the center of mass of the system and the
-%velocities of the regions.
-
-%\subsection{The Glass Nanoparticle}
-%The glass particle from the experiment will be represented by a system of 864 particles, aligned regularly on a FCC (face centered cubic) lattice.
-%This alignment is achieved by defining a minimum cell, containing points:
-%\begin{eqnarray*}
-    %p_1 &=& \{0,0,0\}\\
-    %p_2 &=& \{0.5,0.5,0\}\\
-    %p_3 &=& \{0.5,0,0.5\}\\
-    %p_4 &=& \{0,0.5,0.5\}
-%\end{eqnarray*}
-%The whole system is then  created by copying this unit cell.\\
-%The interaction between the atoms is modeled by the Lennard-Jones potential. It has the form
-%\begin{equation}
-    %U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right].
-%\end{equation}
-%To simplify the equation and computation of the potential and other quantities, like the forces or pressure, it is useful to introduce reduced units.
-%In general, reduced units BLABLABLA THINK OF GOOD TEXT HEREJKA\\
-%The basic units in a system with Lennard-Jones interaction are length ($\sigma$), energy ($\varepsilon$ or $\varepsilon / k_B$) and mass ($m$). Every 
-%quantity can now be written in terms of this units, so they become reduced quantities, denotet by an asterisk. The most important ones are:
-%\begin{eqnarray*}
-    %r^* &=& r/\sigma\\
-    %T^* &=& k_B/\varepsilon \ T\\
-    %U^* &=& U/\varepsilon \\
-    %P^* &=& P \sigma^3/\varepsilon 
-%\end{eqnarray*}
-%With the above introduced reduced units, the Lennard-Jones potential can be written as
-%\begin{equation}
-    %U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right].
-%\end{equation}
-%Since this is the most practical form, it will be used without the asterisk from here on.\\
-%The Lennard-Jones potential is an additive pair-potential, so the total energy of the system can be calculated by summing over all pairs of atoms:
-%\begin{equation}
-    %U_{\text{tot}} = \sum_{i=1}^{N-1}\sum_{j=i+1}^N4\left[{r_{ij}}^{-12} - {r_{ij}}^{-6}\right]
-%\end{equation}
-%where $r_{ij}$ denotes the distance between atom $i$ and $j$. \\
-%The Velocity-Verlet algorithm uses the forces to calculate the positions and velocities for the next timestep, we need to calculate the derivative of
-%the potential energy:
-%\begin{eqnarray}
-    %F_{x} &=& -\frac{\partial}{\partial x} U(r) \nonumber\\
-                %&=& -\frac{\partial}{\partial x} 4\left[{r}^{-12} - {r}^{-6}\right] \nonumber\\
-                %&=& -4 \left[(-12){r}^{-13} - (-6){r}^{-7}\right] \frac{\partial r}{\partial x} \nonumber\\
-                %&=& 48 \left[r^{-13} - 0.5 \ r^{-7}\right] \frac{x}{r} \nonumber\\
-    %\label{eq:ljforce} &=& 48 \left[r^{-14} - 0.5 \ r^{-8}\right] x
-%\end{eqnarray}
-%This is the component of the force in x-direction -- the other components are calculated analogously.
-
-\newpage
-\section{Results}
-% results with different parameters, graphs, etc.
-Since the simulation consists of various parts that have to work together, we first need to check if every part itself works.\\
-\subsection{The Crystal}
-The first element of the simulation is the setup of the FCC lattice on which the particles are placed on. To check if this is done correctly, we can
-use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano particle 
-can be seen in Fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
-where the atoms take the position on the gap between the atoms from the layer below.\\
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.2]{images/fcc_rendering_layer_newcoloring.png}
-        \caption{A rendered image of the FCC lattice. The chosen color-coding along the (1,1,1) Miller-index 
-                highlights the layers for this perspective. The perspective is
-                chosen to see the layering of the FCC lattice with its 3 layer structure. The black box acts as a reference point to outline the 
-                geometry of the system.}
-            %Rendering of the model of the nano crystal before starting the simulation. 
-            %The chosen perspective shows the 3 layers of the FCC crystal with a
-                    %These layers consist of atoms
-                    %taking the position above the gap between two of the atoms from the layer below and match the positions of the atoms 2 layers
-                %below. The black box is the outline of the particles' positions and shows the geometry of the system.}
-        \label{fig:glass}
-    \end{center}
-\end{figure}
-
-\subsection{Velocity Verlet}
-The velocity Verlet part of the simulation is taking place in a NVE ensemble from a thermodynamic point of view. 
-This means that the particle number $N$, the volume $V$ and the total energy $E$ are constant. 
-The initial velocities of the particles are chosen randomly from a Maxwell-Boltzmann distribution. To ensure that the system behaves the way it is
-intended to, small adjustments have to be made in the beginning of the simulation of the nano particle.\\
-%In contrast to the NVT ensemble used in Monte-Carlo simulations this means, that the velocities of the particle can not
-%be chosen directly. The initial velocities have to be drawn from a random number generator. This means that the velocities have to be adjusted after 
-%they have been drawn.\\
-The first adjustment that needs to be made is making sure that the system is staying in place and not floating away. To do this, the velocity vectors
-of all the atoms in the system have to be added to get the resulting velocity vector for the whole system. 
-This resulting vector is then divided by the number of particles in the system and the resulting vector is then subtracted from every
-velcocity vector of every atom in the system. 
-\begin{eqnarray}
-    \mathbf{v}_\text{tot} &=& \sum_{i=1}^N \mathbf{v}_i\\
-    \mathbf{v}_\text{cut} &=& \frac{\mathbf{v}_\text{tot}}N\\
-    \mathbf{v}_i^\text{new} &=& \mathbf{v}_i - \mathbf{v}_\text{cut}
-\end{eqnarray}
-This adjustment only needs to be done just after randomly choosing the velocities, as there is no force present at this point that would cause the
-nano particle to move.
-%The next adjustment is a small one that makes a difference in the long run. When choosing random velocities, the system not only tends to have a
-%resulting velocity vector that causes the whole system to float away, but also a resulting angular momentum that causes the system to rotate.
-%The angular momentum can be defined via the moment of inertia $\mathbf{I}$
-%\begin{eqnarray}
-    %\label{eq:angularmomentum}\mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
-    %\boldsymbol{\omega} &=& \frac{\mathbf{r} \times \mathbf{v}}{r^2}
-%\end{eqnarray}
-%where $\boldsymbol{\omega}$ is the angular velocity of the system. Another way to define the angular momentum is given by
-%\begin{equation}
-    %\label{eq:angularmomentum2}
-    %\mathbf{L} = \mathbf{r} \times \mathbf{p}
-%\end{equation}
-%where $\mathbf{p}$ is the momentum of the particle, which in this case is equal to its velocity since the mass is set to 1.The moment of inertia 
-%$\mathbf{I}$ for a filled cube is given by 
-%\begin{equation}
-    %\mathbf{I} = 
-    %\begin{pmatrix}
-        %\frac{mL^2}{6} & 0 & 0\\
-        %0 & \frac{mL^2}{6} & 0\\
-        %0 & 0 & \frac{mL^2}{6} 
-    %\end{pmatrix}
-%\end{equation}
-%where $m$ will be later set to 1 and $L$ is the side length of the cube. To use this equations to suppress the rotation of the cube, we need to use
-%equation~\eqref{eq:angularmomentum} and rewrite it as
-%\begin{equation}
-    %\label{eq:omega}
-    %\boldsymbol{\omega} = \mathbf{I}^{-1}\mathbf{L}
-%\end{equation}
-%where $\mathbf{I}^{-1}$ is the inverse matrix of $\mathbf{I}$. The calculation of the inverse matrix is fairly straightforward, as it only contains
-%values in the diagonal, which means that the inverse matrix can be calculated by simply inverting these values. The angular momentum for every
-%particle $\mathbf{L}_i$ in the above equation be calculated using eq.~\eqref{eq:angularmomentum2}, where the position is taken to be the distance
-%between the particle and the center of mass
-%\begin{equation}
-    %\Delta\mathbf{r}_i = \mathbf{r}_\text{COM} - \mathbf{r}_i
-%\end{equation}
-%The values of the angular momentum for each paricle are added up and multiplied by the inverse of the moment of inertia, as described in
-%eq.~\eqref{eq:omega}. The final step is updating the velocities of the particles:
-%\begin{equation}
-    %\mathbf{v}_i^\text{new} = \mathbf{v}_i - \left(\boldsymbol{\omega} \times \Delta\mathbf{r}_i\right)
-%\end{equation}
-%\\The velocities of the particle can roughly be chosen by adjusting the range of numbers the random number generator can choose from. However, the
-%resulting velocities (and therefore the temperature of the system) are generally not precisely how they should be. 
-\\As mentioned above, the velocities are chosen from a Maxwell-Boltzmann distribution, which means choosing each component from a Gaussian
-distribution. However, the positions on the FCC lattice are not very natural and so the atoms equilibrate to a different temperature than desired.
-If the system should have a
-specific starting temperature $T$, the velocities need to be scaled accordingly. This rescaling is done by summing the square of the velocities
-(this equals calculating the kinetic energy since the masses are set to 1) and rescale them by a factor $\lambda$:
-\begin{eqnarray}
-\label{eq:rescale1}   E &=& \sum_{i=1}^N \frac{\mathbf{v}_i^2}{2} \\
- \label{eq:rescale2}   \lambda &=& \sqrt{\frac{3(N-1)T}{2E}} \\
- \label{eq:rescale3}   \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i  
-\end{eqnarray}
-This process has to be done a couple of times, until the system equilibrates to the desired temperature. 
-%The problem with this procedure is the fact
-%that it is not compatible with the requirement of constant energy of the system. 
-If we look at the energy values in Fig.~\ref{fig:energystart} we can see the
-irregularities (jumps) before the system equilibrates to a constant value. This equilibration process divides the simulation in two phases: the equilibration
-phase and the measurement phase. In the equilibration phase, the system has time to reach a configuration that is (as the name suggests) in an 
-equilibrated state. During this phase, there are usually no measurements (such as energy) performed, since the values from this phase are distorting
-the mean value of the simulation. That is why the configurations from the equilibration phase are discarded and the measurement phase takes place, 
-where the values do get calculated and are used to determine the mean values.  
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.3]{images/verlet_output_te.pdf}
-        \caption{Temperature and energy in the equilibration phase of the velocity Verlet algorithm. The velocity rescaling (see eq. 
-        \eqref{eq:rescale1}-\eqref{eq:rescale3}) is taking place every 1000th step until the 9000th, which can be seen in the energy curve whre the
-    jumps are happening. The desired temperature in this case is $T^* = 0.2$, which is reached after about 7000 steps.}
-        \label{fig:energystart}
-    \end{center}
-\end{figure}
-
-\subsection{eHEX}
-The next checkpoint is the eHEX algorithm. Since it contains several steps and variables that have to be calculated, this algorithm is very prone
-to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\Delta Q$. 
-The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
-The investigation of the effects on the system can be done in two ways: comparing the temperature and the energy of the system during the 
-application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\Delta Q$.\\
-%Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
-%with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\
-Since the eHEX algorithm injects heat into the system, one would expect the energy and the temperature to rise during the application of the
-algorithm. In Fig.~\ref{fig:ehex_te} we can see that exactly this is happening. The graph shows a velocity Verlet phase (that follows an equilibration
-phase which is not recorded) of constant temperature and energy and an eHEX phase, where the temperature and the enrgy are constantly rising. 
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/ehex_te.pdf}
-        \caption{Comparison of temperature (red) and energy (blue) during the application of the eHEX algorithm with $\Delta Q=0.04$ (in reduced units).
-                The first 20000 timesteps, the energy and temperature are constant during the application of the velocity Verlet algorithm and start
-            to rise, once the eHEX algortithm takes over. The values for the temperature can be seen on the ordinate on the left hand side
-            and the values for the energy on the right hand side. }
-        \label{fig:ehex_te}
-    \end{center}
-\end{figure}
-The comparison of different values of $\Delta Q$ can be seen in Fig.~\ref{fig:ehex_dq}. As the correlation between the temperature and the energy has
-been shown in Fig.~\ref{fig:ehex_te}, the plotting of the temperature data should suffice. 
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/ehex_temp_dq.pdf}
-        \caption{Comparison of different values for $\Delta Q$, ranging from 0 to 0.09. As the value $\Delta Q=0$ corresponds to no heat being
-        injected into the system, it is not surprising that the temperature remains constant. The graph shows that an increase in $\Delta Q$ leads to
-    an increase in the rate of warming up, which means that for bigger $\Delta Q$, the system is heating up more quickly.}
-        \label{fig:ehex_dq}
-    \end{center}
-\end{figure}
-
-\subsection{Thermostat}
-The barostat, as described previously, keeps the system from overheating. The barostat algorithm is applied at the same time as the eHEX algorithm,
-after the velocity Verlet has equilibrated the system.\\
-While the eHEX algorithm is pumping energy into the nano particle while the surrounding gas particles absorb excessive energy. This leads to the
-development of a new, higher steady temperature of the nano particle, which can be seen in Fig.~\ref{fig:baro1}. 
-%The first simulations were performed
-%with the eHEX and the barostat algorithm starting at the same point. This lead to the question: does the new steady temperature depend on the time it
-%takes the gas particles to reach the system? The answer is: no. After applying the barostat algorithm first to fill the box with gas atoms and then
-%applying the eHEX algorithm it can be seen that the resulting temperature is the same, which can be seen in fig.~\ref{fig:baro2}.
-%This means that the system is beginning to heat up, until the particles reach the crystal.
-%Because of this, the temperature of the system rises, until it reaches a steady value again. This behaviour can be seen in fig.~\ref{fig:baro1}.\\
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/baro_temp_start.pdf}
-        \caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
-        in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
-    an increase in temperature until a new steady temperature is reached.}
-        %\caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
-        %in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
-    %an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
-%reached and the temperature seems to stay at this value.}
-        \label{fig:baro1}
-    \end{center}
-\end{figure}
-%\begin{figure}[h]
-    %\begin{center}
-        %\includegraphics[scale=0.4]{images/baro_temp_compare_other.pdf}
-        %\caption{Two versions of the application of the eEHX and barostat algorithm. The shorter run (red, x-axis on the bottom) represents 
-            %the simultaneous start of the eHEX and the barostat algorithm. In the longer run (blue, x-axis on the top) there are 20000 steps, where
-        %the barostat algorithm is applied at the same time as the velocity Verlet to give the system time to equilibrate and only after the eHEX
-    %algorithm is taking the place of the velocity Verlet. The mean value of the resulting temperature is the same, indicating that the final
-%temperature is based on the values of the parameters, such as pressure and ambient temperature, rather than geometry.}
-        %%\caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
-        %%in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
-    %%an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
-%%reached and the temperature seems to stay at this value.}
-        %\label{fig:baro1}
-    %\end{center}
-%\end{figure}
-%As we are interested in the temperatures of the particles before and after the interaction with the nano particle, we need some way to differentiate those
-%two states. This can be done when looking at the total energy of the gas particles.\\
-%The gas particles are created on the surface of the simulation box, which is further away than the interaction range between the gas particles and the
-%nano particle. Since the gas particles do not interact with each other, the total energy of the gas particles does not change until they reach the
-%interaction range of the nano particle. There they undergo a change of energy due to the interaction with the nano particle and then make their way away
-%from the nano particle. After the interaction, when they reach a point outside of the interaction range, the enrgy stays constant again until the
-%particles reach the limits of the simulation box.\\
-%Therefore, the life span of a single gas particle can be partioned into 3 stages: incoming, interacting, outgoing. In a typical scene of interaction
-%all three of these stages are represented by different gas particles, as can be seen in Fig.~\ref{fig:barointeraction}.\\
-%\begin{figure}[h]
-    %\begin{center}
-        %\includegraphics[scale=0.4]{images/baro_170105_new_medium.png}
-        %\caption{Rendering of the application of the barostat algorithm. The red particles are incoming gas particles, that have not yet interacted, 
-            %pink particles are currently interacting and green particles are leaving the simulation box and
-        %moving away from the nano particle (blue). To render this image, a cut has been made to make a view into the simulation box possible.}
-        %\label{fig:barointeraction}
-    %\end{center}
-%\end{figure}
-%Since the gas particles are interacting with the nano particle, the temperatures and energies are values of interest. The energies are recorded using
-%histograms and the temperature is calculated instantaneously every 100th timestep.\\
-%The calculation of the energies of the incoming particles has to be done with care. The problem are particles that enter the simulation box, but fly
-%by the nano particle at a distance that is greater than the interaction range. Since the change of the energy is the interesting part, only the energy 
-%particles of that \textit{do} interact with the nano particle shall be calculated. This can be achieved by counting particles that enter the stage of
-%\textit{interacing}. One such histogram can be seen in Fig.~\ref{fig:histogram}.
-%\begin{figure}[h]
-    %\begin{center}
-        %\includegraphics[scale=0.4]{images/histogram_2.pdf}
-        %\caption{Histogram of total energies for incoming (red) and outgoing (blue) particles. The parameters for this histogram are $\Delta Q^*=0.04$,
-        %$T_\text{ambient}^*=0.08$, $P^*=0.8$. What can be seen here is that the outgoing particles have a higher count of particles with more total
-    %energy.}
-        %\label{fig:histogram}
-    %\end{center}
-%\end{figure}
-
-\subsection{Measurement}
-Now that the functionality of every individual part has been established, the main question can be tackled: how does the laser intensity influence the
-center of mass motion of the nano particle? And which influence does the surrounding gas exert?\\
-In order to investigate the origin of the motion of the center of mass, we first need to check whether the eHEX algorithm itself has any influence on
-the center of mass motion. This is done by applying the algorithm without the thermostat and calculating the center of mass velocity. This turns out
-to be the case, i.e. the center of mass motion stands still.\\
-To check the integrity of the algortihm, a simple case needs to be tested. Such a simple case is $\Delta Q = 0$, i.e. no energy begin pumped into the
-system. The expected outcome of this case is a thermal equilibration between the incoming gas particles, the outgoing gas particles and the center of
-mass motion of the nano particle with the incoming gas temperature being the reference value. 
-The measurement yields this result and backs up the claim. Furthermore, the internal temperature of the nano particle also equilibrates to the 
-temperature of the surrounding gas. This was checked on a system with pressure $P=0.8$, temperature of the incoming particles $T_\text{imp} = 0.05$, 
-internal temperature of the nano particle (set with the velocity verlet) $T_\text{intern} = 0.2$ and a system size of $N=32$ atoms in the nano
-particle. The resulting temperatures are:
-\begin{eqnarray*}
-    T_\text{em} = 0.0494 \pm 0.0009\\
-    T_\text{COM} = 0.051 \pm 0.108\\
-    T_\text{int} = 0.048 \pm 0.007\\
-\end{eqnarray*}
-The calculation of the temperature is done by using the following equations:
-\begin{eqnarray}
-    \label{eq:ekin}E_\text{Kin} &=& \frac{m}{2 M} \sum_i \vec{v}_i^{\, 2} \\
-    \label{eq:temp}E &=& \frac{f}{2} k_B T
-\end{eqnarray}
-In the first line, $m$ denotes the mass, which is $m=0.1$ for the gas particles, $m=1$ for the atoms in the nano particle and $m=32$ for the center
-of mass. $M$ denotes the number of particles that are being summed over, which is $M=1$ for the center of mass, $M=32$ for the internal temperature
-and fluctuates for the outgoing gas particles. The number of the outgoing particles is not constant and depends heavily on the pressure and the
-temperature of the impinging gas particles, as can be seen in \eqref{eq:numberofparticles}. In the second line, $f$ denotes the number of degrees of
-freedom of the system, which is $f=3$ for all three systems. Putting these two equations together, we get an equation for the temperature:
-\begin{equation}
-    T = \frac{m}{3 M} \sum_i \vec{v}_i^{\, 2} \\
-\end{equation}
-Since this describes an instantaneous value, the mean value of the velocities over the course of the simulation has to be used to get the mean value
-of the temperature of each subsystem. The simulation results compared with the mean values can be seen in
-Figs.~\ref{fig:outgoingequib},~\ref{fig:vcomsqdequib} and \ref{fig:internalequib}.
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/gastemp.pdf}
-        \caption{Instantaneous and mean temperature of the outgoing gas particles of the surrounding gas}
-        \label{fig:outgoingequib}
-    \end{center}
-\end{figure}
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/vcomsqd.pdf}
-        \caption{Instantaneous and mean temperature of the center of mass of the nano particle in the laser trap}
-        \label{fig:vcomsqdequib}
-    \end{center}
-\end{figure}
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/t_int.pdf}
-        \caption{Instantaneous and mean internal temperature of the nano particle in the laser trap}
-        \label{fig:internalequib}
-    \end{center}
-\end{figure}
-\\With the calculated mean temperature of the center of mass, we can take a look at the distribution of the components of the center of mass velocity.
-The expected distribution is a gaussian with $\sigma=\sqrt{\frac{k_B T}{m}}$, i.e.
-\begin{equation}
-    \label{eq:gaussianvel}
-    p(v_i) = \frac{m}{2\pi k_B T} e^{-\frac{x^2 m}{2 k_B T}}
-\end{equation}
-where $i$ denotes the chosen component of the center of mass velocity vector. A histogram of the x-component of the center of mass velocity  
-with a curve representing \eqref{eq:gaussianvel} can be seen in Fig.~\ref{fig:comvel}. Since the y and z axis yield very similar histograms, the
-x-component has been chosen to represent the velocity components. The graph in Fig.~\ref{fig:comvel} clearly shows, that the velocities indeed follow
-the gaussian distribution.
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/vcomtemp.pdf}
-        \caption{A histogram of the x-component of the center of mass velocities with the expected gaussian distribution}
-        \label{fig:comvel}
-    \end{center}
-\end{figure}
-\\The case of $\Delta Q=0$ is a good way to start and to check the integrity of the algorithms, but the more interesting cases are the ones where
-$\Delta Q > 0$. Besides $\Delta Q$, there are two values that determine the dynamics of the system: the pressure $P$ and the temperature of the
-impinging gas particles $T_\text{imp}$.\\
-Since the surrounding gas acts as a thermostat, the pressure of the impinging gas at a given gas temperature determines the maximum amount of heat that
-can be pumped into the system without particles of the nano particle leaving the simulation box -- the higher the pressure, the more energy can be
-pumped into the system. The algorithm is applied at two surrounding gas temperatures of $T_\text{imp} = 0.05$ and $T_\text{imp} = 0.1$, 
-which are one quarter and one half the temperature of the nano particle before applying the eEHX and the thermostat algorithm. The pressure varied
-from $P=0.1$ to $P=1.2$ in steps of $\Delta P=0.1$ and the heat is generally varied from $\Delta Q=0$ to $\Delta Q=0.5$ in steps of 0.05. The term
-''generally'' is used because there are cases where the stepsize has to be decreased or one value for $\Delta Q$ is skipped over. The stepsize has to
-be decresed for low pressures, such as $P=0.1$ or $P=0.2$ because there would otherwise not be enough data points. Some values of $\Delta Q$ need to
-be skipped because it can happen that atoms of the nano particle leave the system due to statistical fluctuations of velocities, which leads to
-unusable results.\\
-To see the comparison between the different temperatures of the impinging gas temperatures, three values for the pressure are chosen: $P=0.1$, $P=0.5$
-and $P=1.0$. The simulation results for the center of mass temperature, the internal temperature and the temperature of the outgoing gas particles are
-plotted for each of the pressure values and both temperatures.
-\begin{figure}[h]
-    \begin{center}
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p01_com.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p05_com.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p1_com.pdf}
-        \end{subfigure} 
-        \\
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p01_int.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p05_int.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p1_int.pdf}
-        \end{subfigure} 
-        \\
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p01_out.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p05_out.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p1_out.pdf}
-        \end{subfigure} 
-    \end{center}
-\end{figure}
-
-%\begin{figure}[h]
-    %\begin{center}
-        %\begin{subfigure}[t]{0.3\textwidth}
-            %\includegraphics[scale=0.3]{images/fcc-2.pdf}
-            %\caption{Schematic figure of a face centered cubic (FCC) lattice}
-            %\label{fig:fcc}
-        %\end{subfigure} 
-        %\
-        %\begin{subfigure}[t]{0.3\textwidth}
-            %\includegraphics[scale=0.15]{images/fcc_unit.png}
-            %\caption{Schematic of the number of particles contained in one FCC unit cell. The atoms on the edges are shared by 8 cells and the ones 
-                %on the faces by 2 cells\cite{unitcell}}
-            %\label{fig:fcccut}
-        %\end{subfigure} 
-        %\
-        %%\hspace{60pt}
-        %\begin{subfigure}[t]{0.3\textwidth}
-            %\includegraphics[scale=0.3]{images/unit_cell.pdf}
-            %\caption{Setup for successively building a FCC lattice}
-            %\label{fig:unitcell}
-        %\end{subfigure}
-    %\end{center}
-%\end{figure}
-
-\newpage
-\section{Conclusion}
-% what does this all tell us? do we need to study more of this?
-
-\newpage
-\bibliography{references}
-\bibliographystyle{unsrt}
-
 \end{document}

### 2017-02-17 doc/draft3.tex
updates

--- 2017-02-16 doc/draft2.tex
@@
 changed in the field of computer simulation (or computers in general) would be a vast understatement.\\
-Not only the processing power of computers has incresed exponentially (Moore's law), which allows the simulation of a greater number of particles for
+Not only has the processing power of computers increased exponentially (Moore's law), which allows the simulation of a greater number of particles for
 a longer period of time, but also new techniques have been developed, such as molecular dynamics. These techniques are not only used by physicists,
 but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. This thesis is an
-attempt to apply methods of simulation to a field with a lot of recent research happening in it: optical trapping.
+attempt to apply methods of simulation to a field of active research: optical trapping.
 Optical trapping is a technique that has been used and developed for over 40 years~\cite{Marago2013}. The idea for this goes back to the beginning of
@@
 \subsection{Experimental setup}
-In the experiment, a silica nano particle with a radius of about 75 nm and mass of about $3 \times 10^{-18}$kg is trapped in a laser beam within a
+In the experiment, a silica nano particle with a radius of about 75 nm and mass of about $3 \times 10^{-18}$ kg is trapped in a laser beam within a
 vacuum chamber. The trapping of the silica nano particle is achieved by a gradient force of the
@@
 center of mass of the nano particle $T_\text{COM}$. The experimental setup
-and the different temperatures are depicted in Fig.~\ref{fig:levitation}.\\
+and the different temperatures are shown in Fig.~\ref{fig:levitation}.\\
 \begin{figure}[H]
@@
         \caption{The experimental setup used by Millen et al. In \textbf{a} the experiment is schematically depicted and \textbf{b} shows a photograph
-            of the actual experimental setup. In \textbf{c} the 4 different temperatures that are present during the experiment are depicted and
-        \textbf{d} shows the position of the particle in the trap over time. The image was taken from~\cite{MillenJ.2014}}
+            of the actual experimental setup. In \textbf{c} the 4 different temperatures that are present during the experiment are illustrated and
+        \textbf{d} shows the position of the particle in the trap over time. The image was taken from~\cite{MillenJ.2014}.}
         \label{fig:levitation}
@@
 Since the two temperatures of the gas, $T_\text{imp}$ and $T_\text{em}$, are not equal, the gas surrounding the nano particle is not in a thermal
-equilibrium, which is in contrast to the assumption in many optical tweezer experiments. Since the gas particles only interact with the nano particle
-in the trap and not wich each other, this situation creates two heat baths of different temperatures with the nano particle acting as a mediator
+equilibrium, which is in contrast to the assumptions made in many optical tweezer experiments. Since the gas particles only interact with the nano particle
+in the trap and not which each other, this situation creates two heat baths of different temperatures with the nano particle acting as a mediator
 between those two baths. This situation is depicted in
@@
         a higher temperature. The incoming and outgoing particles do not interact with each other and thus create two separate heat baths with
-    different temperatures. Image modified from \cite{Kroy2014}}
+    different temperatures. Image modified from \cite{Kroy2014}.}
         \label{fig:heatbaths}
@@
 The goal is to simulate the experiment depicted in Fig.~\ref{fig:setup} as accurately as possible. To achieve this, the setup has to be broken down
-into individual pieced that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, sitting on an FCC
-lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, as it acts as a trapping device as well
-as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, that acts as a thermostat in the simulation, that will be
-modeled as a box surrounding the nano particle. This box is filled with gas particles that interact with the nano particle via a soft-sphere potential
-and do not interact with one another.\\
+into individual pieces that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, sitting on a FCC
+(face-centered cubic) lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, 
+as it acts as a trapping device as well as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, 
+that acts as a thermostat in the simulation, which will be modeled as a cubic box surrounding the nano particle. 
+This box is filled with gas particles that interact with the nano particle via a soft-sphere potential and do not interact with one another.\\
 The following section will give a brief overview of the concepts used to simulate every part of the experiment.
@@
 approximately like hard spheres which attract each other with weak van der Waals forces, which justifies the use of the Lennard-Jones potential. 
-Argon has a mass of m = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$m and $\varepsilon = 1.65 \times 10^{-21}$J.\\
+Argon has a mass of m = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$ m and $\varepsilon = 1.65 \times 10^{-21}$ J.\\
 With the above introduced reduced units, the Lennard-Jones potential can be written as
@@
 The initial configuration of the particles is a face centered cubic (FCC) lattice. A schematic of the FCC lattice is depicted in Fig. \ref{fig:fcc}.
-With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since once FCC cell 
+With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since one FCC cell 
 shares its atoms with its next neighbours, the number of atom per unit cell is 4 -- 1/8 of a particle on eight corners and 1/2 of a particle on six faces.
@@
             \includegraphics[scale=0.3]{images/fcc-2.pdf}
-            \caption{Schematic figure of a face centered cubic (FCC) lattice}
+            \caption{Schematic figure of a face-centered cubic (FCC) lattice.}
             \label{fig:fcc}
@@
             \includegraphics[scale=0.3]{images/unit_cell.pdf}
-            \caption{Setup for successively building a FCC lattice}
+            \caption{Setup for successively building a FCC lattice.}
             \label{fig:unitcell}
@@
 \end{equation}
-where $L$ is the side length of the cube that is the whole system and it is calculated via the density of the system
+where $L$ is the side length of the cube that represents the nano particle and it is calculated via the density of the system
 \begin{equation}
@@
 The solutions to this problem is to include the stepwise calculation of the velocities \cite{swope1982}. For this we start again with the Taylor
-expansion of the coordinates mathbftor \eqref{eq:verletnew} and instead of Taylor expanding $\mathbf{r}_i(t-\Delta t)$, we write it as 
+expansion of the coordinates vector \eqref{eq:verletnew} and instead of Taylor expanding $\mathbf{r}_i(t-\Delta t)$, we write it as 
 \begin{equation}
@@
 acting as a heat sink vanishes and the region acting as a heat source spans over the whole simulation box. This means that, using the 
-notation of Fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither center of mass of the particles the simulation box nor 
+notation of Fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither the center of mass of the particles the simulation box nor 
 the heat source are moving, i.e. 
@@
 \end{equation}
-where $\varepsilon$ is the interaction strength and $\sigma$ is the interaction range and $r$ is the distance between the gas particle and the
+where $\varepsilon$ is the interaction strength, $\sigma$ is the interaction range and $r$ is the distance between the gas particle and the
 interacting particle. The gas particles do not interact with one another in this model.\\
@@
         \end{equation}
-        If the length of the cell $L$ is chosen to be equal to the cut-off radius $r_c$, this insertion should only be carried out with a probablity of
+        If the length of the cell $L$ is chosen to be equal to the cut-off radius $r_c$, this insertion should only be carried out with a probability of
         \begin{equation}
@@
         \includegraphics[scale=0.4]{images/gastemp.pdf}
-        \caption{Instantaneous and mean temperature of the outgoing gas particles of the surrounding gas}
+        \caption{Instantaneous and mean temperature of the outgoing gas particles of the surrounding gas.}
         \label{fig:outgoingequib}
@@
         \includegraphics[scale=0.4]{images/vcomsqd.pdf}
-        \caption{Instantaneous and mean temperature of the center of mass of the nano particle in the laser trap}
+        \caption{Instantaneous and mean temperature of the center of mass of the nano particle in the laser trap.}
         \label{fig:vcomsqdequib}
@@
         \includegraphics[scale=0.4]{images/t_int.pdf}
-        \caption{Instantaneous and mean internal temperature of the nano particle in the laser trap}
+        \caption{Instantaneous and mean internal temperature of the nano particle in the laser trap.}
         \label{fig:internalequib}
@@
         \includegraphics[scale=0.4]{images/vcomtemp.pdf}
-        \caption{A histogram of the x-component of the center of mass velocities with the expected gaussian distribution}
+        \caption{A histogram of the x-component of the center of mass velocities with the expected gaussian distribution.}
         \label{fig:comvel}
@@
 values of $\Delta Q$ range from $0$ to $0.02$ and at $P=1.0$ from $0$ to $0.5$.\\
-The first thing that is noticable, is the increase of the temperature of the center of mass, the internal temperature and the outgoing gas temperature
+The first thing that is noticeable, is the increase of the temperature of the center of mass, the internal temperature and the outgoing gas temperature
 with increasing $\Delta Q$. This makes sense, as the amount of heat that is put into the system is also transfered in the interaction between the nano
@@
 internal temperature.\\
-There are several ways to explore this subject further. The first idea would be to decrease the pressure even more and investiate the case of very low
+There are several ways to explore this subject further. The first idea would be to decrease the pressure even more and investigate the case of very low
 energies, since this is not easily feasible in an experimental setup. Another idea would be to increase the number of atoms in the nano particle to

### 2017-02-22 doc/draft4.tex
updates

--- 2017-02-17 doc/draft3.tex
@@
 
-\lhead[\ ]{\ }
+\lhead[\ ]{\leftmark }
 \chead[\ ]{\ }
-\rhead[\ ]{\ }
+\rhead[\ ]{\rightmark }
 
 \lfoot[\ ]{\ }
-\cfoot[\ ]{\ }
+\cfoot[\ ]{\thepage }
 \rfoot[\ ]{\ }
@@
 simulate particles in a liquid state to calculate the equation of state. This was the first publication in which scientists used a computer to
-calculate numerical results based on a physical theory in a process that is to this day called \textit{Metropolis Monte Carlo}. To say that a lot has
+calculate numerical results based on a physical theory in a process that is to this day called \textit{Metropolis Monte Carlo simulation}. 
+To say that since then a lot has
 changed in the field of computer simulation (or computers in general) would be a vast understatement.\\
@@
 a longer period of time, but also new techniques have been developed, such as molecular dynamics. These techniques are not only used by physicists,
-but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. This thesis is an
-attempt to apply methods of simulation to a field of active research: optical trapping.
+but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. In thhis thesis we 
+apply methods of simulation to a field of active research: optical trapping.
 Optical trapping is a technique that has been used and developed for over 40 years~\cite{Marago2013}. The idea for this goes back to the beginning of
@@
 properties and also allows the cooling of nanoparticles towards their quantum mechanical ground state.\\
-The small size of the objects used in the optical tweezer experiments make them ideal candidates for computer simulations. In this thesis, I will
+The small size of the objects used in the optical tweezer experiments make them amenable candidates for computer simulations. In this thesis, I will
 model an optical tweezer setup, based on a real experiment, to investigate different temperatures in a non-equilibrium state.\\
@@
 information on the techniques used to model the experiment and simulate the dynamic properties. In the third chapter, I will present my findings which
-are followed by the conclusions and further remarks.
+are followed by conclusions and further remarks.
 
@@
 laser beam acting on the particle. The experimental setup is depicted in Fig.~\ref{fig:setup}.\\
-The particle fluctuates within the trap in all three spatial directions. These fluctuations can be approximated
-such that they are decoupled, which means that they can be described by a one-dimensional Langevin equation:
+The particle fluctuates within the trap in all three spatial directions. These fluctuations are approximatedly
+decoupled, which means that they can be described by a one-dimensional Langevin equation:
 \begin{equation}
@@
 \end{figure}
-The one dimensional Langevin equation \eqref{eq:langevin} has to be modeled to include the different friction coefficients of the two heat baths. The
+The one dimensional Langevin equation \eqref{eq:langevin} has to be modified to include the different friction coefficients of the two heat baths. The
 model derived in \cite{MillenJ.2014} is described in the following equation:
@@
 The goal is to simulate the experiment depicted in Fig.~\ref{fig:setup} as accurately as possible. To achieve this, the setup has to be broken down
-into individual pieces that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, sitting on a FCC
+into individual pieces that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, located on the
+sites of a FCC
 (face-centered cubic) lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, 
@@
 approximately like hard spheres which attract each other with weak van der Waals forces, which justifies the use of the Lennard-Jones potential. 
-Argon has a mass of m = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$ m and $\varepsilon = 1.65 \times 10^{-21}$ J.\\
+Argon has a mass of $m$ = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$ m and $\varepsilon = 1.65 \times 10^{-21}$ J.\\
 With the above introduced reduced units, the Lennard-Jones potential can be written as
@@
 \end{eqnarray}
-The force in the y and z direction can be calculated analogously.\\
+The force in the $y$ and $z$ direction can be calculated analogously.\\
 As can be seen in Fig.~\ref{fig:lj}, the potential is very close to 0 at about $r/\sigma=2.5$. This means that any contribution of the potential
-energy after this point is vanishingly small. To save computing time it is common practice to cut off the potential at a certain point and don't
+energy after this point is vanishingly small. To save computing time it is common practice to cut off the potential at a certain point and not to
 calculate the contribution for distances beyond that. The cut-off distance for this thesis was set to $r = 2.5\sigma$.\\
@@
     p_1 &=& \{0,0,0\}\\
-    p_2 &=& \{0.5*a,0.5*a,0\}\\
-    p_3 &=& \{0.5*a,0,0.5*a\}\\
-    p_4 &=& \{0,0.5*a,0.5*a\}
+    p_2 &=& \{0.5 \ a,0.5 \ a,0\}\\
+    p_3 &=& \{0.5 \ a,0,0.5 \ a\}\\
+    p_4 &=& \{0,0.5 \ a,0.5 \ a\}
 \end{eqnarray*}
@@
 \end{equation}
-The sum of those two equations yields
+The sum of these two equations yields
 \begin{equation}
@@
 \cite{Andersen1980} algorithms are often used in simulation to change the temperature of the system in a controllable way.\\
-The problem with these kind of algorithms is, that a target temperature has to be fixed which will be reached at some point. In order to simulate the
+The problem with these kind of algorithms is that a target temperature has to be fixed which will be reached at some point. In order to simulate the
 influx of energy from the laser it would be better to have an algorithm that supplies the system with a certain amount of energy continuously.
@@
 \subsection{Surrounding Gas - Thermostat}
-Without any equilibrating mechanism the setup described by now would lead to the system heating up indefinetly, which is not desirable. Furthermore,
+Without any equilibrating mechanism the setup described by now would lead to the system heating up indefinitely, which is not desirable. Furthermore,
 the goal is to recreate the experiment as close as possible. Since the nano particle is trapped in a laser beam within a vacuum chamber, the
@@
 \begin{enumerate}
-    \item Randomly draw the number of particles that are created on a single side of the minimal volume of cells, $N_\text{fac}$, from the
-        distribution 
+    \item Randomly draw the number of particles that are created on a single side of the minimal volume of cells, $N_\text{fac}$, from a Poisson
+        distribution with mean value
         \begin{equation}
@@
         pressure, $m$ is the mass of the gas particle, $k_B$ is the Boltzmann constant (which will be set to 1 in reduced units) and $T$ is the
-        desired temperature. This chosen number of particles is then equally distributed over the face of the cell on which they are created. The
-        velocities of the created particles are drawn from two different random number distributions. The component of the particle perpendicular to
+        desired temperature. This chosen number of particles is then uniformly distributed over the face of the cell on which they are created. The
+        velocities of the created particles are drawn from two different random number distributions. The component of the particle velocity perpendicular to
         the surface is drawn from a Rayleigh distribution of the form 
@@
         \end{equation}
-        (where U is the interaction energy between the gas particle and the crystal) because it is possible that the inserted particle is 
+        (where $U$ is the interaction energy between the gas particle and the crystal) because it is possible that the inserted particle is 
         within the interaction range of the crystal. 
@@
 \begin{eqnarray}
-    \mathbf{v}_\text{tot} &=& \sum_{i=1}^N \mathbf{v}_i\\
-    \mathbf{v}_\text{cut} &=& \frac{\mathbf{v}_\text{tot}}N\\
-    \mathbf{v}_i^\text{new} &=& \mathbf{v}_i - \mathbf{v}_\text{cut}
+    \mathbf{v}_\text{COM} &=& \frac1N\sum_{i=1}^N \mathbf{v}_i\\
+    %\mathbf{v}_\text{cut} &=& \frac{\mathbf{v}_\text{tot}}N\\
+    \mathbf{v}_i^\text{new} &=& \mathbf{v}_i - \mathbf{v}_\text{COM}
 \end{eqnarray}
@@
 \subsection{Thermostat}
-The barostat, as described previously, keeps the system from overheating. The barostat algorithm is applied at the same time as the eHEX algorithm,
+The ideal gas thermostat/barostat, as described previously, 
+keeps the system from overheating. The barostat algorithm is applied at the same time as the eHEX algorithm,
 after the velocity Verlet has equilibrated the system.\\
-While the eHEX algorithm is pumping energy into the nano particle while the surrounding gas particles absorb excessive energy. This leads to the
+The eHEX algorithm is pumping energy into the nano particle while the surrounding gas particles absorb excessive energy. This leads to the
 development of a new, higher steady temperature of the nano particle, which can be seen in Fig.~\ref{fig:baro1}. 
@@
 to be the case, i.e. the center of mass motion stands still.\\
-To check the integrity of the algortihm, a simple case needs to be tested. Such a simple case is $\Delta Q = 0$, i.e. no energy begin pumped into the
+To check the integrity of the algortihm, a simple case needs to be tested. Such a simple case is $\Delta Q = 0$, i.e. no energy being pumped into the
 system. The expected outcome of this case is a thermal equilibration between the incoming gas particles, the outgoing gas particles and the center of
@@
 \begin{eqnarray*}
-    T_\text{em} = 0.0494 \pm 0.0009\\
-    T_\text{COM} = 0.051 \pm 0.108\\
-    T_\text{int} = 0.048 \pm 0.007\\
+    T_\text{em} &=& 0.0494 \pm 0.0009\\
+    T_\text{COM} &=& 0.051 \pm 0.108\\
+    T_\text{int} &=& 0.048 \pm 0.007\\
 \end{eqnarray*}
@@
 \begin{eqnarray}
-    \label{eq:ekin}E_\text{Kin} &=& \frac{m}{2 M} \sum_i \vec{v}_i^{\, 2} \\
-    \label{eq:temp}E &=& \frac{f}{2} k_B T
+    \label{eq:ekin}E &=& \frac{m}{2 M} \sum_i \vec{v}_i^{\, 2} \\
+    \label{eq:temp}E_\text{Kin} &=& \frac{f}{2} k_B T
 \end{eqnarray}
@@
 \begin{equation}
-    T = \frac{m}{3 M} \sum_i \vec{v}_i^{\, 2} \\
-\end{equation}
-Since this describes an instantaneous value, the mean value of the velocities over the course of the simulation has to be used to get the mean value
-of the temperature of each subsystem. The simulation results compared with the mean values can be seen in
+    k_B T = \frac{m}{3 M} \sum_i \vec{v}_i^{\, 2} \\
+\end{equation}
+Since this describes an instantaneous value, the mean value of the kinetic energy over the course of the simulation has to be used to get the mean value
+of the temperature of each subsystem. The instantaneous values compared with the mean values can be seen in
 Figs.~\ref{fig:outgoingequib},~\ref{fig:vcomsqdequib} and \ref{fig:internalequib}.
@@
 \\With the calculated mean temperature of the center of mass, we can take a look at the distribution of the components of the center of mass velocity.
-The expected distribution is a gaussian with $\sigma=\sqrt{\frac{k_B T}{m}}$, i.e.
+The expected distribution is a Gaussian with variance $\sigma^2=\frac{k_B T}{m}$, i.e.
 \begin{equation}
     \label{eq:gaussianvel}
-    p(v_i) = \frac{m}{2\pi k_B T} e^{-\frac{x^2 m}{2 k_B T}}
-\end{equation}
-where $i$ denotes the chosen component of the center of mass velocity vector. A histogram of the x-component of the center of mass velocity  
-with a curve representing \eqref{eq:gaussianvel} can be seen in Fig.~\ref{fig:comvel}. Since the y and z axis yield very similar histograms, the
-x-component has been chosen to represent the velocity components. The graph in Fig.~\ref{fig:comvel} clearly shows, that the velocities indeed follow
-the gaussian distribution.
+    p(v_i) = \sqrt{\frac{m}{2\pi k_B T}} e^{-\frac{x^2 m}{2 k_B T}}
+\end{equation}
+where $i$ denotes the chosen component of the center of mass velocity vector. A histogram of the $x$-component of the center of mass velocity  
+with a curve representing eq.~\eqref{eq:gaussianvel} can be seen in Fig.~\ref{fig:comvel}. Since the $y$ and $z$ axis yield very similar histograms, the
+$x$-component has been chosen to represent the velocity components. The graph in Fig.~\ref{fig:comvel} clearly shows, that the velocities indeed follow
+the Gaussian distribution.
 \begin{figure}[h]
@@
             heat being pumped into the system, for pressures
-        $P=0.1$, $P=0.5$ and $P=1$ at impinging gas temperatures of $T_\text{imp}=0.05$ (red) and $T_\text{imp}=0.1$. The three columns each show 
+            $P=0.1$, $P=0.5$ and $P=1$ at impinging gas temperatures of $T_\text{imp}=0.05$ (red) and $T_\text{imp}=0.1$ (blue). The three columns each show 
         a fixed pressure value and the rows show fixed measured quantities. As the pressure relates to the maximum amount of heat that can be injected
@@
 increase slower for higher pressures. Take the center of mass temperature for $P=0.1$ and $P=1$ at $T=0.05$ (red line) as an example: while it reaches
-a value of a $T_\text{com} = 0.1$ at about $\Delta Q = 0.015$ at $P=0.1$, it takes a $\Delta Q$ of a little under $0.1$ to achieve the same value.
+a value of a $T_\text{com} = 0.1$ at about $\Delta Q = 0.015$ at $P=0.1$, it takes a $\Delta Q$ of a little under $0.1$ to achieve the same value
+for $P = 1$.
 This can be observed for every one of the three quantities. This is to be expected, as the surrounding gas acts as a thermostat. \\
 The development of the different temperatures seem to be depending of the temperature of the surrounding gas as well. The blue curves (which
-correspont do a surrounding gas temperature of $T=0.1$) are growing slightly faster than the red curves ($T=0.05$).\\
+correspond to a surrounding gas temperature of $T=0.1$) are growing slightly faster than the red curves ($T=0.05$).\\
 In their paper~\cite{MillenJ.2014} Millen et al. give a formula for the center of mass temperature:

### 2017-02-28 doc/draft4.tex
more writing

@@
 Not only has the processing power of computers increased exponentially (Moore's law), which allows the simulation of a greater number of particles for
-a longer period of time, but also new techniques have been developed, such as molecular dynamics. These techniques are not only used by physicists,
-but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. In thhis thesis we 
+a longer period of time, but also new techniques have been developed, such as molecular dynamics, transition path sampling, finite element method etc. 
+These techniques are not only used by physicists,
+but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. In this thesis we 
 apply methods of simulation to a field of active research: optical trapping. 
+Optical trapping refers to the experimental method of localising small objects with a laser beam. The laser is highly focused on one point, where 
+the object is trapped due to the gradient force of the laser. The size of the trapped objects ranges from a few micrometers down to the subatomic 
+scale.\\
+%Optical trapping refers to experiments where a focused laser beam is 
+%used to keep small objects with a size in the range of a few micrometers down to a subatomic scahe in place.\\
 Optical trapping is a technique that has been used and developed for over 40 years~\cite{Marago2013}. The idea for this goes back to the beginning of
-the 20th century, where Lebedev~\cite{Lebedew1901}, Nichols and Hull~\cite{Nichols1901} demonstrated the existence of a force that light exerts on 
+the 20th century, when Lebedev~\cite{Lebedew1901}, Nichols and Hull~\cite{Nichols1901} demonstrated the existence of a force that light exerts on 
 matter. Due to the technical limitations at that time, the idea was discarded until 1970, where Ashkin~\cite{Ashkin2000} used lasers to manipulate the
@@
 years however, new developments in the field opened the possibilities to work with a variety of objects in the desired nanometer scale, such as metal
-nanoparticles, nano tubes, quantum dots and much more. This non-invasive method of trapping objects allows very precise measurements of various
+nanoparticles~\cite{Svoboda1994}, nano tubes~\cite{Tan2004}, quantum dots~\cite{Jauffred2008} and much more. 
+This non-invasive method of trapping objects allows very precise measurements of various
 properties and also allows the cooling of nanoparticles towards their quantum mechanical ground state.\\
 The small size of the objects used in the optical tweezer experiments make them amenable candidates for computer simulations. In this thesis, I will
-model an optical tweezer setup, based on a real experiment, to investigate different temperatures in a non-equilibrium state.\\
-The first chapter will be the introduction of the experiment and the introduction of the problem statement. The second chapter contains basic
-information on the techniques used to model the experiment and simulate the dynamic properties. In the third chapter, I will present my findings which
+model an optical tweezer setup, based on a real experiment, to investigate different temperatures in a non-equilibrium state. 
+The experiment features a silica nano sphere trapped in a laser beam within a vacuum chamber. To simulate this experiment, each of these individual parts 
+of the experiment need to be modeled. The silica nano sphere is approximated by a system of argon atoms that interact with each other via a 
+Lennard-Jones potential. These atoms are absorbing energy from the laser and are thus heated up using the enhanced heat exchange algorithm (eHEX) 
+and a harmonic trap is keeping them in place. The gas particles in the gas chamber are acting as a thermostat/barostat and are modeled by a ideal
+gas pressure bath.\\
+In the first chapter, I will introduce two optical tweezer experiments: one conducted by Gieseler et al. to investigate the relaxation of a nano particle
+from a non-equilibrium steady state and another conducted by Millen et al. to examine different temperature regimes in an optical tweezer experiment. 
+This is followed by the problem statement. The second chapter contains basic
+information on the techniques used to model the individual parts of the experiment and 
+simulate the dynamic properties. In the third chapter, I will present my findings which
 are followed by conclusions and further remarks.
@@
 \end{figure}
-On the left hand side we have the position $x$ (and its derivatives $\dot{x}$, the velocity and $\ddot{x}$, the acceleration), the friction 
-coefficient $\Gamma_0$ and the angular frequency $\Omega_0$ that describes the fluctuation along the chosen axis. On
+On the left hand side of this equation we have the position $x$ of the nano particle (and its derivatives $\dot{x}$, the velocity and $\ddot{x}$, 
+the acceleration), the friction coefficient $\Gamma_0$ and the angular frequency $\Omega_0$ that describes the fluctuation along the chosen axis. On
 the right hand side, there are two forces. The first one is $F_\text{fluct}$, which describes a stochastic force caused by interactions with the gas
@@
 due to the fluctuation-dissipation theorem, which links the damping rate to the stochastic force.\\
+The other force, $F_\text{ext}$ is a time-dependent external force that is applied to the nano particle to create a non-equilibrium steady state. At
+time $t = t_\text{off}$ the force is deactivated and the particle relaxes to a thermally equilibrated state in the laser trap. The evolution of the nano 
+particle from the non equilibrium steady state to the equilibrated state is examined to investigate the fluctuation theorem, which can be written 
+down as 
+\begin{equation}
+    \label{eq:fluctuationtheorem}
+    \frac{p(-\Delta S)}{\Delta S} = e^{-\Delta S}
+\end{equation}
+where $\Delta S$ is the relative entropy change given by
+\begin{equation}
+    \Delta S = \beta_0 Q + \Delta \phi
+\end{equation}
+In the above equation, $Q$ is the heat absorbed by the bath, $\beta_0$ is the reciprocal temperature at which the heat is absorbed and 
+$\Delta \phi$ is the difference of the trajectory dependent entropy. The fluctuation theorem~\cite{Searles1999}, as expressed in 
+\eqref{eq:fluctuationtheorem}, states that there is a measurable probablity for a finite system to violate the second law of thermodynamics
+for finite times. This does not stand in contradiction to the second law of thermodynamics itself, as it applies only to short times and does not 
+happen for large systems or long time scales.\\
+The investigation of the fluctuation theorem is performed for two different external forces. The first one is a feedback force and the second 
+one is the feedback force combined with a parametric drive. The results show that the relaxation process for both initial non-equilibrium steady 
+states generated by the external forces demonstrates the validity of the fluctuation theorem.\\
+
 %The external force $F_\text{ext}$ is part of the experimental setup, where the frequency of the fluctuation along an axis, $\Omega_0$, is measured and
@@
 before interacting with the nano particle $T_\text{imp}$ (for \textit{impinging}), the temperature of the gas particles after 
-interaction $T_\text{em}$ (for emerging), the surface temperature of the nano particle $T_\text{sur}$ and the temperature of the 
-center of mass of the nano particle $T_\text{COM}$. The experimental setup
+interaction $T_\text{em}$ (for \textit{emerging}), the \textit{surface} temperature of the nano particle $T_\text{sur}$ and the temperature of the 
+\textit{center of mass} of the nano particle $T_\text{COM}$. The experimental setup
 and the different temperatures are shown in Fig.~\ref{fig:levitation}.\\
@@
 \end{figure}
-Since the two temperatures of the gas, $T_\text{imp}$ and $T_\text{em}$, are not equal, the gas surrounding the nano particle is not in a thermal
-equilibrium, which is in contrast to the assumptions made in many optical tweezer experiments. Since the gas particles only interact with the nano particle
+The nano particle in the trap is absorbing heat from the laser which causes the surface temperature of the nano particle to be higher than the 
+temperature of the surrounding gas. The impinging gas particles interact with the sphere, absorb some of the energy and emerge with a different
+temperature, which is in general not equal to the surface temperature. This means that
+the two temperatures of the gas, $T_\text{imp}$ and $T_\text{em}$, are not equal, the gas surrounding the nano particle is not in a thermal
+equilibrium, which is in contrast to the assumptions made in many optical tweezer experiments. 
+Since the gas particles only interact with the nano particle
 in the trap and not which each other, this situation creates two heat baths of different temperatures with the nano particle acting as a mediator
@@
 sites of a FCC
-(face-centered cubic) lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, 
-as it acts as a trapping device as well as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, 
+(face-centered cubic) lattice that interact via a Lennard-Jones potential. The laser serves two purposes: heating up and trapping the particle. The 
+nano particle will be heated up using the eHEX algorithm and trapped using a harmonic potential.
+%Since the laser serves a dual purpose -- heating and trapping --
+%both of these are modeled separately.
+%The laser will be broken down into its two main purposes, 
+%as it acts as a trapping device as well as a heat source for the nano particle. 
+Surrounding the nano particle will be a gas chamber, 
 that acts as a thermostat in the simulation, which will be modeled as a cubic box surrounding the nano particle. 
@@
 
+\subsection{The Velocity-Verlet Algorithm}
+\label{section:verlet}
+When we look at the system from a microscopic standpoint, we see that it follows some kind of path in the phase space as time progresses. Every
+point in this space corresponds to a set of positions and momenta and the connection between two points corresponds to the evolution of the system
+from one state to another. As mentioned above, this evolution (the dynamics of the system) is a crucial element to molecular dynamics. Since the
+equations of motion cannot be solved analytically in general, we need to approximate the solution.\\
+The method used here is called finite difference approach. The trajectory of the system in the phase space is cut into finite pieces of length $\Delta
+t$ and the equations of motion are solved for every segment separately (see Fig. \ref{fig:finitedifference}).\\
+\begin{figure}
+    \begin{center}
+        \includegraphics[scale=0.5]{images/finite_approach.pdf}
+        \caption{Simplified graphical schematic of the finite difference approach. The evolution of the system from a point $(r^N(t),p^N(t))$ in the
+            phase space is approximated by slicing it up into pieces of length $\Delta t$. On every stop after the starting point ($t+\Delta t$, 
+            $t+2\Delta t$, $t+3\Delta t$,$\ldots$) the equations of motion can be solved numerically.}
+        \label{fig:finitedifference} 
+    \end{center}
+\end{figure}
+There are several ways to solve this kind of problem, but since we are interested in implementing it into a computer program the ideal solution
+should have some basic properties \cite{Allen1989}:
+\begin{itemize}
+    \item It should be fast and require little memory
+    \item It should permit the use of a large time step $\Delta t$
+    \item It should duplicate the classical trajectory as closely as possible
+    \item It should satisfy known conservation laws
+    \item It should be simple and easy to program
+\end{itemize}
+One algorithm that has all of the above mentioned features is the one proposed by Verlet \cite{Verlet1967}. In his paper, Verlet starts by Taylor
+expanding the coordinate vector $\mathbf{r}_i$ for one particle after one time step $\Delta t$:
+\begin{equation}
+    \label{eq:verlet}
+    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + \dot{\mathbf{r}}_i \Delta t + \frac12 \ddot{\mathbf{r}}_i \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i \Delta t^3 +
+    \mathcal{O}(\Delta t^4)
+\end{equation}
+Since the first derivative of the coordinate is the velocity, $\dot{\mathbf{r}}_i(t) = \mathbf{v}_i(t)$, and the second derivative of the coordinate 
+is the acceleration, $\ddot{\mathbf{r}}_i(t) = \mathbf{a}_i(t)$, using Newton's law $\mathbf{F}_i(t) = m_i \mathbf{a}_i(t)$ equation \eqref{eq:verlet} 
+can be written as:
+\begin{equation}
+    \label{eq:verletnew}
+    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i(t)
+    \Delta t^3+\mathcal{O}(\Delta t^4)
+\end{equation}
+The same calculation can be carried out for one time step before $t$:
+\begin{equation}
+    \label{eq:verletnewminus}
+    \mathbf{r}_i(t-\Delta t) = \mathbf{r}_i(t) - {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 - \frac1{3!} \dddot{\mathbf{r}}_i(t) \Delta t^3+
+    \mathcal{O}(\Delta t^4)
+\end{equation}
+The sum of these two equations yields
+\begin{equation}
+    \mathbf{r}_i(t+\Delta t) + \mathbf{r}_i(t-\Delta t) = 2\mathbf{r}_i(t) + \frac1{2} {\mathbf{F}}_i(t) \Delta t^2 + \mathcal{O}(\Delta t^4)
+\end{equation}
+and from this we get the final form for the new coordinates:
+\begin{equation}
+    \label{eq:verletfinal}
+    \mathbf{r}_i(t+\Delta t) = 2\mathbf{r}_i(t)+ \mathbf{r}_i(t-\Delta t)+ \frac1{2} {\mathbf{F}}_i(t) \Delta t^2
+\end{equation}
+This means that the new coordinates can be calculated using the current coordinates, the current forces and the coordinates from the past time step.\\
+From equation \eqref{eq:verletfinal} one thing becomes clear: the velocities are not necessary to calculate the new positions and they are not
+computed in the process, but for the calculation of i.e. the kinetic energy the velocities are needed. Thus, the velocities have to be calculated with
+a combination of the Taylor expansions of the coordinate vectors at $t+\Delta t$ and $t-\Delta t$:
+\begin{equation}
+    \mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t) = 2\mathbf{v}_i(t)\Delta t + \mathcal{O}(\Delta t^3)
+\end{equation}
+which can be rewritten as
+\begin{equation}
+    \label{eq:velocities}
+    \mathbf{v}_i(t) = \frac{\mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t)}{2\Delta t} + \mathcal{O}(\Delta t^2)
+\end{equation}
+This approach has the advantage that it is fast, requires little memory and is reliable in the sense that there is no energy drift occuring during the
+simulation, which means that the energy is conserved. When we compare this to the list of desired properties this seems like a good algorithm.\\
+This algorithm however has two significant disatvantages. The first one is the calculation of the velocities. As can be seen in equation
+\eqref{eq:velocities}, the accuracy of the calculation is only $\mathcal{O}(\Delta t^2)$ while the positions can be calculated with an error of order
+$\Delta t^4$. The other big disatvantage is the first step of the algorithm. Since the calculation of the new positions requires the current positions
+and the ones from one time step before, which technically do not exist.\\
+The solutions to this problem is to include the stepwise calculation of the velocities \cite{swope1982}. For this we start again with the Taylor
+expansion of the coordinates vector \eqref{eq:verletnew} and instead of Taylor expanding $\mathbf{r}_i(t-\Delta t)$, we write it as 
+\begin{equation}
+    \mathbf{r}_i(t) = \mathbf{r}_i(t+\Delta t) - \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
+\end{equation}
+Using \eqref{eq:verletnew} in the abvove equation we get
+\begin{equation}
+    \begin{aligned}
+        \mathbf{r}_i(t) &= \mathbf{r}_i(t) + \mathbf{v}_i(t) \Delta t + \frac1{2m}\mathbf{F}_i(t)\Delta t^2 \\
+                     &- \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
+    \end{aligned}
+\end{equation}
+and thus
+\begin{equation}
+    \label{eq:vvelocities}
+    \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \left(\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\right) \Delta t
+\end{equation}
+With the addition of this equation this algorithm is called the Velocity-Verlet algorithm, which is summarized in the following two equations:
+\begin{equation}
+    \label{eq:velocityverlet}
+    \begin{aligned}
+        \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2\\
+        \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \Big[\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\Big] \Delta t
+    \end{aligned}
+\end{equation}
+This algorithm is self starting, uses a small amount of memory, conserves the energy (not exactly) 
+and gives a very good approximation to the originial trajectory
+in the phase space. To program this algorithm the following steps are needed: 
+\begin{enumerate}
+    \item Calculate all the forces between the particles (for the first step only)
+    \item Calculate the new positions with current velocity and forces
+    \item Calculate first half of the new velocities with current forces
+    \item Calculate all the forces for the new position
+    \item Use new forces to calculate second half of new velocities
+\end{enumerate}
+The first point only has to be carried out for the first step, because the forces have not been calculated at this point. As the forces are calculated
+for the new positions in step 4, they can be used for the next time step. For the first time step the velocities have to be chosen randomly and are 
+calculated with this algorithm from that point forward.
 \subsection{The Nano Particle}
@@
 With the lattice constant and the 4 points of the FCC cell, all the particles can be put into place.
-
-\subsection{The Velocity-Verlet Algorithm}
-\label{section:verlet}
-When we look at the system from a microscopic standpoint, we see that it follows some kind of path in the phase space as time progresses. Every
-point in this space corresponds to a set of positions and momenta and the connection between two points corresponds to the evolution of the system
-from one state to another. As mentioned above, this evolution (the dynamics of the system) is a crucial element to molecular dynamics. Since the
-equations of motion cannot be solved analytically in general, we need to approximate the solution.\\
-The method used here is called finite difference approach. The trajectory of the system in the phase space is cut into finite pieces of length $\Delta
-t$ and the equations of motion are solved for every segment separately (see Fig. \ref{fig:finitedifference}).\\
-\begin{figure}
-    \begin{center}
-        \includegraphics[scale=0.5]{images/finite_approach.pdf}
-        \caption{Simplified graphical schematic of the finite difference approach. The evolution of the system from a point $(r^N(t),p^N(t))$ in the
-            phase space is approximated by slicing it up into pieces of length $\Delta t$. On every stop after the starting point ($t+\Delta t$, 
-            $t+2\Delta t$, $t+3\Delta t$,$\ldots$) the equations of motion can be solved numerically.}
-        \label{fig:finitedifference} 
-    \end{center}
-\end{figure}
-There are several ways to solve this kind of problem, but since we are interested in implementing it into a computer program the ideal solution
-should have some basic properties \cite{Allen1989}:
-\begin{itemize}
-    \item It should be fast and require little memory
-    \item It should permit the use of a large time step $\Delta t$
-    \item It should duplicate the classical trajectory as closely as possible
-    \item It should satisfy known conservation laws
-    \item It should be simple and easy to program
-\end{itemize}
-One algorithm that has all of the above mentioned features is the one proposed by Verlet \cite{Verlet1967}. In his paper, Verlet starts by Taylor
-expanding the coordinate vector $\mathbf{r}_i$ for one particle after one time step $\Delta t$:
-\begin{equation}
-    \label{eq:verlet}
-    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + \dot{\mathbf{r}}_i \Delta t + \frac12 \ddot{\mathbf{r}}_i \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i \Delta t^3 +
-    \mathcal{O}(\Delta t^4)
-\end{equation}
-Since the first derivative of the coordinate is the velocity, $\dot{\mathbf{r}}_i(t) = \mathbf{v}_i(t)$, and the second derivative of the coordinate 
-is the acceleration, $\ddot{\mathbf{r}}_i(t) = \mathbf{a}_i(t)$, using Newton's law $\mathbf{F}_i(t) = m_i \mathbf{a}_i(t)$ equation \eqref{eq:verlet} 
-can be written as:
-\begin{equation}
-    \label{eq:verletnew}
-    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i(t)
-    \Delta t^3+\mathcal{O}(\Delta t^4)
-\end{equation}
-The same calculation can be carried out for one time step before $t$:
-\begin{equation}
-    \label{eq:verletnewminus}
-    \mathbf{r}_i(t-\Delta t) = \mathbf{r}_i(t) - {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 - \frac1{3!} \dddot{\mathbf{r}}_i(t) \Delta t^3+
-    \mathcal{O}(\Delta t^4)
-\end{equation}
-The sum of these two equations yields
-\begin{equation}
-    \mathbf{r}_i(t+\Delta t) + \mathbf{r}_i(t-\Delta t) = 2\mathbf{r}_i(t) + \frac1{2} {\mathbf{F}}_i(t) \Delta t^2 + \mathcal{O}(\Delta t^4)
-\end{equation}
-and from this we get the final form for the new coordinates:
-\begin{equation}
-    \label{eq:verletfinal}
-    \mathbf{r}_i(t+\Delta t) = 2\mathbf{r}_i(t)+ \mathbf{r}_i(t-\Delta t)+ \frac1{2} {\mathbf{F}}_i(t) \Delta t^2
-\end{equation}
-This means that the new coordinates can be calculated using the current coordinates, the current forces and the coordinates from the past time step.\\
-From equation \eqref{eq:verletfinal} one thing becomes clear: the velocities are not necessary to calculate the new positions and they are not
-computed in the process, but for the calculation of i.e. the kinetic energy the velocities are needed. Thus, the velocities have to be calculated with
-a combination of the Taylor expansions of the coordinate vectors at $t+\Delta t$ and $t-\Delta t$:
-\begin{equation}
-    \mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t) = 2\mathbf{v}_i(t)\Delta t + \mathcal{O}(\Delta t^3)
-\end{equation}
-which can be rewritten as
-\begin{equation}
-    \label{eq:velocities}
-    \mathbf{v}_i(t) = \frac{\mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t)}{2\Delta t} + \mathcal{O}(\Delta t^2)
-\end{equation}
-This approach has the advantage that it is fast, requires little memory and is reliable in the sense that there is no energy drift occuring during the
-simulation, which means that the energy is conserved. When we compare this to the list of desired properties this seems like a good algorithm.\\
-This algorithm however has two significant disatvantages. The first one is the calculation of the velocities. As can be seen in equation
-\eqref{eq:velocities}, the accuracy of the calculation is only $\mathcal{O}(\Delta t^2)$ while the positions can be calculated with an error of order
-$\Delta t^4$. The other big disatvantage is the first step of the algorithm. Since the calculation of the new positions requires the current positions
-and the ones from one time step before, which technically do not exist.\\
-The solutions to this problem is to include the stepwise calculation of the velocities \cite{swope1982}. For this we start again with the Taylor
-expansion of the coordinates vector \eqref{eq:verletnew} and instead of Taylor expanding $\mathbf{r}_i(t-\Delta t)$, we write it as 
-\begin{equation}
-    \mathbf{r}_i(t) = \mathbf{r}_i(t+\Delta t) - \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
-\end{equation}
-Using \eqref{eq:verletnew} in the abvove equation we get
-\begin{equation}
-    \begin{aligned}
-        \mathbf{r}_i(t) &= \mathbf{r}_i(t) + \mathbf{v}_i(t) \Delta t + \frac1{2m}\mathbf{F}_i(t)\Delta t^2 \\
-                     &- \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
-    \end{aligned}
-\end{equation}
-and thus
-\begin{equation}
-    \label{eq:vvelocities}
-    \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \left(\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\right) \Delta t
-\end{equation}
-With the addition of this equation this algorithm is called the Velocity-Verlet algorithm, which is summarized in the following two equations:
-\begin{equation}
-    \label{eq:velocityverlet}
-    \begin{aligned}
-        \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2\\
-        \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \Big[\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\Big] \Delta t
-    \end{aligned}
-\end{equation}
-This algorithm is self starting, uses a small amount of memory, conserves the energy (not exactly) 
-and gives a very good approximation to the originial trajectory
-in the phase space. To program this algorithm the following steps are needed: 
-\begin{enumerate}
-    \item Calculate all the forces between the particles (for the first step only)
-    \item Calculate the new positions with current velocity and forces
-    \item Calculate first half of the new velocities with current forces
-    \item Calculate all the forces for the new position
-    \item Use new forces to calculate second half of new velocities
-\end{enumerate}
-The first point only has to be carried out for the first step, because the forces have not been calculated at this point. As the forces are calculated
-for the new positions in step 4, they can be used for the next time step. For the first time step the velocities have to be chosen randomly and are 
-calculated with this algorithm from that point forward.
 
@@
 \begin{eqnarray}
-    \mathbf{F} &=& -k \Big[\mathbf{x}-\mathbf{x}_0\Big]\\
-    U &=& \frac12 k \Big[\mathbf{x}-\mathbf{x}_0\Big]^2
+    \mathbf{F} &=& -k \Big[\mathbf{r}_\text{COM}-\mathbf{x}_0\Big]\\
+    U &=& \frac12 k \Big[\mathbf{r}_\text{COM}-\mathbf{x}_0\Big]^2
 \end{eqnarray}
-where $\mathbf{x}_0$ is the position of minimal potential energy. 
-Since the trap is acting on the whole system, the force is acting on the center of mass. 
+where $\mathbf{x}_0$ is the position of minimal potential energy, $\mathbf{r}_\text{COM}$ 
+is the position of the center of mass and $k$ is the spring constant.
+%Since the trap is acting on the whole system, the force is acting on the center of mass. 
 To calculate the center of mass positions and velocities, the positions and velocities of all particles have to be summed up and 
@@
 \end{equation}
+To use this in the simulation, the distance of the center of mass to the desired point of minimal potential energy and then the corresponding 
+force is calculated and added to the interaction force of each atom.
 
@@
 object of interest (this will be referred to as the minimal volume of cells) and leave the simulation box if their position reaches the boundary of
-this minimal volume of cells. This behaviour is very close to the real experiment which makes this thermostat an ideal candidate for the simulation of
-the experiment. 
+this minimal volume of cells. The minimal volume of cells is based on cell lists used in the simulation. Cell lists are a method of saving 
+computing time by dividing the simulation box in cells with a side length equal to or greater than the cutoff radius $r_C$. The particles are then
+sorted into these cells and the interaction between the particles is only calculated for other particles within the same and neighbouring cells. 
+For the minimum volume of cells used in the thermostat algorithm all cells filled with particles and one additional cell for the interaction partners
+is used.\\
+This algorithm is a very good approximation of the situation in the experiment.  There the gas particles only interact with the nano 
+particle in the laser trap and not with eatch other. After this interaction the
+gas particles emerge with a slightly higher temperature and leave the vicinity of the nano particle. Moreover, the algorithm uses an ideal gas
+as a pressure medium, which allows us to simulate the dynamics of the interaction that is happening in the experiment.\\
+
+%This behaviour is very close to the real experiment which makes this thermostat an ideal candidate for the simulation of
+%the experiment. There the gas particles only interact with the nano particle in the laser trap and not with eatch other. After this interaction the
+%gas particles emerge with a slightly higher temperature and leave the vicinity of the nano particle. 
 %This barostat also acts as a thermostat and thus is suitable for the 
@@
 use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano particle 
-can be seen in Fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
-where the atoms take the position on the gap between the atoms from the layer below.\\
+can be seen in Fig.~\ref{fig:glass}. To see that the particles form a FCC lattice a perspective is chosen, where the three layers of the FCC lattice
+are visible. \\
+These layers are generated by stacking the atoms onto the gap between two atoms. In a FCC lattice there are three distinct layers, which means that
+the atoms of the fourth layer are on the same position as the atoms in the first layer.
+%The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
+%where the atoms take the position on the gap between the atoms from the layer below.\\
 \begin{figure}[h]
@@
 \subsection{Velocity Verlet}
-The velocity Verlet part of the simulation is taking place in a NVE ensemble from a thermodynamic point of view. 
+The simulation of the nano particle without the gas is taking place in a NVE ensemble from a thermodynamic point of view. 
 This means that the particle number $N$, the volume $V$ and the total energy $E$ are constant. 

### 2017-02-28 test/plot/verlet_test.cpp

@@ -1,0 +1,48 @@
+#include <iostream>
+#include <cstdio>
+#include <cmath>
+#include <cstdlib>
+#include <fstream>
+#include <iomanip>
+#include <string>
+#include <cstring>
+#include "../../include/globals.hpp"
+#include "../../include/classes.hpp"
+#include "../../include/functions.hpp"
+
+int main()
+{
+    dt = 0.005;
+    setValues(0.2,0.04,0.2,0.8,0.06);
+    Particle *particles = new Particle[N];
+    InitPositions(particles);
+    FILE* tempout = fopen("verlet_test_temp.dat","w");
+    FILE* enerout = fopen("verlet_test_ener.dat","w");
+    FILE* compos = fopen("verlet_test_compos.dat","w");
+    FILE* comvel= fopen("verlet_test_comvel.dat","w");
+
+
+    InitVelocities(particles);
+    double* vCM = new double[3];
+    double* rCM = new double[3];
+    double energy = 0.;
+    //calcCM(particles,rCM,vCM);
+
+    for(int i=0;i<200000;i++)
+    {
+        if(i%100 == 0)
+            std::cout << "Step " << i << std::endl;
+        if(i%1000 == 0)
+            if(i < 20000)
+                rescaleVelocities(particles);
+        VelocityVerlet(particles);
+        calcTemp(particles,tempout);
+        energy = calculateEnergies(particles);
+        fprintf(enerout,"%lf\n",energy);
+    }
+
+    fclose(tempout);
+    fclose(enerout);
+    fclose(compos);
+    fclose(comvel);
+}

### 2017-03-01 doc/draft4.tex
updates

@@
 which satisfy $\langle F^\text{imp}\rangle= \langle F^\text{em} \rangle = 0$ individually.\\
-Miller et al. investigate the dependence of the emerging temperature $T_\text{em}$ on the intensity of the laser for different sized nano particles
-and different pressures.\\
-With these four temperatures introduced, some questions arise about the relationship between these temperatures. For this theses, I want to focus on
-the question: how does the laser intensity influence the motion of the nano particle in the laser trap and its center of mass velocity?\\
+As described above, the impining and emerging gas particles form two independent heat baths with two different temperatures 
+with $T_\text{em} > T_\text{imp}$ in general. This leads to two different friction coefficients and noise terms that have to be included in the 
+Langevin equation.\\
+Millen et al. investigate the dependence of the emerging temperature $T_\text{em}$ on the intensity of the laser for different sized nano particles
+and different pressures. They show that the emerging gas temperature increases with increasing laser power and depends on the size of the nano
+particle, as the emerging gas temperature increases slower for bigger particles.\\
+Where the experiment conducted by Gieseler et al. uses only the temperature of the heat bath for their model, Millen et al. establish four different
+temperatures that are present in the experiment: the temperature of the impinging and emerging gas particles, the
+surface temperature and the center of mass temperature of the nano particle in the trap. Since the center of mass motion of the nano particles
+is the main objective in the experiment of Gieseler et al., the dependence of the center of mass temperature on 
+laser intensity and the temperature of the impinging gas should be investigated, which will be done in this thesis. The questions I would like to
+answer are: How does the laser intensity influence the motion of the nano particle in the laser trap and its center of mass velocity? What impact
+does the impinging gas temperature have on the temperature of the center of mass?\\
+%With these four temperatures introduced, some questions arise about the relationship between these temperatures. For this theses, I want to focus on
+%the question: how does the laser intensity influence the motion of the nano particle in the laser trap and its center of mass velocity?\\
 In the next section I will introduce the methods that will be used to simulate the system, followed by the results of the simulations and the
@@
 \\As mentioned above, the velocities are chosen from a Maxwell-Boltzmann distribution, which means choosing each component from a Gaussian
-distribution. However, the positions on the FCC lattice are not very natural and so the atoms equilibrate to a different temperature than desired.
+distribution. As it is very unlikely for the atoms to be sitting exactly on a FCC lattice, the first few time steps are needed for the equilibriation
+of the system. Due to this process the temperature of the system can change slightly and vary from the desired temperature. 
+%However, the positions on the FCC lattice are not very natural and so the atoms equilibrate to a different temperature than desired.
 If the system should have a
@@
 If we look at the energy values in Fig.~\ref{fig:energystart} we can see the
-irregularities (jumps) before the system equilibrates to a constant value. This equilibration process divides the simulation in two phases: the equilibration
+jumps before the system equilibrates to a constant value. This equilibration process divides the simulation in two phases: the equilibration
 phase and the measurement phase. In the equilibration phase, the system has time to reach a configuration that is (as the name suggests) in an 
@@
         \includegraphics[scale=0.3]{images/verlet_output_te.pdf}
-        \caption{Temperature and energy in the equilibration phase of the velocity Verlet algorithm. The velocity rescaling (see eq. 
+        \caption{Temperature and energy in the equilibration phase of the nano particle without the gas with a time step of
+            $\Delta t = 0.02$. The velocity rescaling (see eq. 
         \eqref{eq:rescale1}-\eqref{eq:rescale3}) is taking place every 1000th step until the 9000th, which can be seen in the energy curve whre the
-    jumps are happening. The desired temperature in this case is $T^* = 0.2$, which is reached after about 7000 steps.}
+    jumps are happening. The desired temperature in this case is $T^* = 0.2$, which is reached after about 7000 steps. The fluctuation of the energy
+    is due to the introduction of the cutoff radius $r_C$ which is set at $r_C = 2.5$.}
         \label{fig:energystart}
@@
 The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
-The investigation of the effects on the system can be done in two ways: comparing the temperature and the energy of the system during the 
+The investigation of the effects on the system can be done in two ways: examining the development of the temperature and the energy of the system 
+during the 
 application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\Delta Q$.\\
@@
         \includegraphics[scale=0.4]{images/ehex_temp_dq.pdf}
-        \caption{Comparison of different values for $\Delta Q$, ranging from 0 to 0.09. As the value $\Delta Q=0$ corresponds to no heat being
+        \caption{Comparison of different values for $\Delta Q$, ranging from 0 to 0.09. Up to time $t = 200$ only the velocity Verlet algorithm is
+            applied which is followed by the application of the eHEX algorithm at different values for $\Delta Q$.
+            As the value $\Delta Q=0$ corresponds to no heat being
         injected into the system, it is not surprising that the temperature remains constant. The graph shows that an increase in $\Delta Q$ leads to
@@
 
-\subsection{Measurement}
+\subsection{Simulation of the Experiment}
 Now that the functionality of every individual part has been established, the main question can be tackled: how does the laser intensity influence the
@@
 temperature of the surrounding gas. This was checked on a system with pressure $P=0.8$, temperature of the incoming particles $T_\text{imp} = 0.05$, 
-internal temperature of the nano particle (set with the velocity verlet) $T_\text{intern} = 0.2$ and a system size of $N=32$ atoms in the nano
+internal temperature of the nano particle $T_\text{intern} = 0.2$ and a system size of $N=32$ atoms in the nano
 particle. The resulting temperatures are:
@@
 \begin{equation}
+    \label{eq:compaper}
     T_\text{COM} = \frac{T_\text{imp}^{3/2}+\tfrac\pi 8 \ T_\text{em}^{3/2}}{T_\text{imp}^{1/2}+\tfrac\pi8 T_\text{em}^{1/2}}
@@
 that the measured center of mass temperature rises quicker than the calculated value. This is also the case for different pressures and impinging gas
-temperatures. The reason for this may be in the difference of the accomodation factor $\alpha$. In the paper, Millen et al. use an accomodation factor
+temperatures. The reason for this may be in the difference of the accomodation factor $\alpha$, which is given by~\cite{Goodman1980}
+\begin{equation}
+    \alpha = \frac{T_\text{em} - T_\text{imp}}{T_\text{sur} - T_\text{imp}}
+\end{equation}
+In the paper, Millen et al. use an accomodation factor
 for a sphere with a value of $\alpha = 0.777$, whereas the accomodation factor for the setup in the simulation for the chosen values is around
 $\alpha = 0.016$. This means that the real and the modeled interaction between the gas particles and the nano particle differ from one another.
+The thermal accomodation factor quantifies the behaviour of the interaction between the gas particles and the nano particle. Although Eq.
+\eqref{eq:compaper} does not contain an explicit dependence on the thermal accomodation factor, the difference between the value in the experiment and
+in the simulation suggests a difference in the energy transfer during interaction between the gas and the nano particle.
+
 \begin{figure}
@@
 nano particle in the trap could be varied to include an attractive part in the potential. Generally, these simulations could be expanded by running
-them on faster computers to investigate long term behaviour of the systems and expand the ranges of the different values like pressure or laser
-energy.
+them on faster computers to investigate larger systems and expand the ranges of the involved values like pressure, laser
+energy, and internal temperature.
 

### 2017-03-01 test/plot/baro_output.cpp

@@ -13,7 +13,7 @@
 int main()
 {
 
-    setValues(0.2,0.1,0.2,0.8,0.06);
+    setValues(0.2,0.05,0.2,0.8,0.06);
     Particle *particles = new Particle[N];
     InitPositions(particles);
     std::list<Particle*> gas;

### 2017-03-03 doc/draft4.tex
final version of the thesis

@@
 $\alpha = 0.016$. This means that the real and the modeled interaction between the gas particles and the nano particle differ from one another.
-The thermal accomodation factor quantifies the behaviour of the interaction between the gas particles and the nano particle. Although Eq.
+The thermal accomodation factor is a measure of the energy transfer during the interaction between the gas particles and the nano particle. Although Eq.
 \eqref{eq:compaper} does not contain an explicit dependence on the thermal accomodation factor, the difference between the value in the experiment and
@@
 \newpage
-\section{Conclusion and Further Remarks}
+\section{Conclusion}
 % what does this all tell us? do we need to study more of this? 
@@
 see if the surface of the particle plays a role in the development of the temperatures. Furthermore, the interaction between the gas particles and the
-nano particle in the trap could be varied to include an attractive part in the potential. Generally, these simulations could be expanded by running
+nano particle in the trap could be varied to include an attractive part in the potential
+to vary the thermal accomodation coefficient. Generally, these simulations could be expanded by running
 them on faster computers to investigate larger systems and expand the ranges of the involved values like pressure, laser

### 2017-03-06 doc/abstract_de.tex
added abstracts

@@
 \begin{document}
+Optical trapping bezeichnet eine experimentelle Methode, bei der Objekte mit einer Größe, die von einer subatomaren bis zur Mikrometer-Skala
+reicht, durch die Gradientenkraft eines Lasers lokalisiert werden. In dieser Arbeit benutzen wir Methoden aus dem Feld der Computational Physics, 
+um das Verhalten eines Nanoteilchens in so einer Laserfalle und seine Interaktionen mit dem Laser und dem umgebenden Gas zu simulieren.\\
+Gieseler et al. führten ein sogenanntes ,,Optical Tweezer''-Experiment durch, um das Fluktuationstheorem zu untersuchen. In diesem Experiment
+verwenden sie die Lagevin-Gleichung, um das Verhalten einer Nanokugel aus Siliciumdioxid in einer Laserfalle zu modellieren, wobei die 
+Temperatur des Umgebungsgases als Wärmereservoir in dem Modell verwendet wird, um die stochastische Kraft auf das Nanoteilchen zu beschreiben. 
+Millen et al. führen ein ähnliches Experiment durch und zeigten auf, dass zwischen vier Temperaturen unterschieden werden muss: die Temperatur 
+des eingehenden Gases, des ausgehenden Gases, der Schwerpunktsbewegung und der Oberfläche des Nanoteilchens. Mit Augenmerk auf diese verschiedenen
+Temperaturen wird das Experiment mit Molecular Dynamics Methoden simuliert, um den Einfluss der Stärke des Lasers und des Umgebungsgases auf 
+die Schwerpunktstemperatur des Nanoteilchens zu untersuchen.\\
+Um das Experiment zu simulieren, müssen die einzelnen Bestandteile modelliert werden. Das Nanoteilchen wird durch ein System aus Atomen 
+dargestellt, die über ein Lennard-Jones Potential miteinander wechselwirken. Die Gradientenkraft des Lasers, die das Nanoteilchen lokalisiert, 
+wird durch ein harmonisches Potential approximiert. Da die Energie des Lasers vom Nanoteilchen absorbiert wird, erhöht sich die interne 
+Temperatur des Nanoteilchens, was durch Anwendung des eHEX Algorithmus erreicht wird. Das umgebende Gas hat die Funktion eines Barostaten und die 
+eines Thermostaten und wird in der Simulation durch einen Algorithmus repräsentiert, der ein ideales Gas als Druckmedium verwendet.\\
+Die Simulation zeigt, dass die Schwerpunktstemperatur von der Lasertemperatur und der Temperatur des einströmenden Gases abhängt.
+
 \end{document}

### 2017-03-06 doc/danksagung.tex
added abstracts

--- 2017-02-14 doc/draft2.tex
@@
 \begin{document}
-\title{Simulation of Nano Particles in a Laser Trap}
-\author{Mathias H\"old, BSc.}
-\date{2016}
-\maketitle
-\thispagestyle{empty}
-\newpage
-\tableofcontents
-\newpage
-\section{Introduction}
-%something about computer science in general, optical tweezers and the need to study the
-%systems on computers yada yada
-In the year 1953, Metropolis et al.~\cite{Metropolis1953} used the MANIAC (Mathematical Analyzer Numerical Integrator And Computer) 
-at the Los Alamos National Laboratories in the United States to
-simulate particles in a liquid state to calculate the equation of state. This was the first publication in which scientists used a computer to
-calculate numerical results based on a physical theory in a process that is to this day called \textit{Metropolis Monte Carlo}. To say that a lot has
-changed in the field of computer simulation (or computers in general) would be a vast understatement.\\
-Not only the processing power of computers has incresed exponentially (Moore's law), which allows the simulation of a greater number of particles for
-a longer period of time, but also new techniques have been developed, such as molecular dynamics. These techniques are not only used by physicists,
-but also find applications in the fields of chemistry, biology, etc., making it possible to apply them to virtually anything. This thesis is an
-attempt to apply methods of simulation to a field with a lot of recent research happening in it: optical trapping.
-Optical trapping is a technique that has been used and developed for over 40 years~\cite{Marago2013}. The idea for this goes back to the beginning of
-the 20th century, where Lebedev~\cite{Lebedew1901}, Nichols and Hull~\cite{Nichols1901} demonstrated the existence of a force that light exerts on 
-matter. Due to the technical limitations at that time, the idea was discarded until 1970, where Ashkin~\cite{Ashkin2000} used lasers to manipulate the
-motion of neutral atoms and micrometer sized objects and thus brought the idea of optical trapping back to life.\\
-With this advancement it was possible to develop methods for trapping and manipulating objects of different sizes ranging from the micrometer scale
-down to the subnano scale. At first, this technique was used to either trap objects in the micrometer range (like cells) or in the subnano scale,
-where lasers were used for the cooling of atoms. The range between those two, with objects at the scale of one to a hundred nanometer, was not
-explored as much because neither downscaling the methods for micrometer sized objects nor upscaling the methods for atoms seemed feasible. In recent
-years however, new developments in the field opened the possibilities to work with a variety of objects in the desired nanometer scale, such as metal
-nanoparticles, nano tubes, quantum dots and much more. This non-invasive method of trapping objects allows very precise measurements of various
-properties and also allows the cooling of nanoparticles towards their quantum mechanical ground state.\\
-The small size of the objects used in the optical tweezer experiments make them ideal candidates for computer simulations. In this thesis, I will
-model an optical tweezer setup, based on a real experiment, to investigate different temperatures in a non-equilibrium state.\\
-The first chapter will be the introduction of the experiment and the introduction of the problem statement. The second chapter contains basic
-information on the techniques used to model the experiment and simulate the dynamic properties. In the third chapter, I will present my findings which
-are followed by the conclusions and further remarks.
-
-%Due to the properties of light it was possible
-%to create methods of trapping and manipulating objects of different sizes from the micrometer to the subnanometer-scale.
-
-%At first, the micrometer
-%ranges (objects like cells) and subnanometer-ranges (lasercooling of atoms) were explored, but the methods used in these ranges were not fit to be
-%applied to objects in the range of one to a hundred nanometers.
-
-\newpage
-\section{Motivation}
-%introduction of the experiment, the problem and the idea
-The starting point of this thesis is an experiment conducted by Gieseler et al \cite{Gieseler2014}. It is an optical tweezer experiment, where the
-motion of a glass nanoparticle in a laser trap was used to investigate the fluctuation theorem\cite{Crooks1999}.
-\subsection{Experimental setup}
-In the experiment, a silica nano particle with a radius of about 75 nm and mass of about $3 \times 10^{-18}$kg is trapped in a laser beam within a
-vacuum chamber. The trapping of the silica nano particle is achieved by a gradient force of the
-laser beam acting on the particle. The experimental setup is depicted in Fig.~\ref{fig:setup}.\\
-The particle fluctuates within the trap in all three spatial directions. These fluctuations can be approximated
-such that they are decoupled, which means that they can be described by a one-dimensional Langevin equation:
-\begin{equation}
-    \label{eq:langevin}
-    \ddot{x} + \Gamma_0 \dot{x} + \Omega^2_0x = \frac 1 m \left(F_\text{fluct} + F_\text{ext}\right)
-\end{equation}
-%The experiment is set up as follows: a glass sphere with radius $r \approx 75$ nm and mass $m \approx 3 \times 10^{-18}$ kg is trapped in a laser beam
-%in a vacuum chamber. The initial state of the glass particle is prepared by modulation of the laser, such that the oscillation of the glass particle
-%is suppressed in every direction. This is done by measuring the oscillation of the particle in every direction (assuming that the oscillation in every
-%direction is decoupled) and using this information as feedback for the laser. This experimental setup can be seen in figure \ref{fig:setup}.
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[scale=0.3]{images/experimental_setup.jpg}
-        \caption{Experimental setup of the optical tweezer experiment. A silica nano particle is trapped in a laser beam via gradient force in a
-        vacuum. The feedback is used to cool down the particle and create a non-equilibrium steady state. In the first part of the experiment, the
-    feedback is turned off and the motion of the particle towards an equilibrated state is observed. In the second part of the experiment, the steady state
-of the particle is modified by a parametric drive. Both the parametric drive and the feedback are turned off and -- as in the first part -- the
-motion of the particle towards an equilibrated state is observerd. Figure taken from \cite{Gieseler2014}.}
-        \label{fig:setup}
-    \end{center}
-\end{figure}
-On the left hand side we have the position $x$ (and its derivatives $\dot{x}$, the velocity and $\ddot{x}$, the acceleration), the friction 
-coefficient $\Gamma_0$ and the angular frequency $\Omega_0$ that describes the fluctuation along the chosen axis. On
-the right hand side, there are two forces. The first one is $F_\text{fluct}$, which describes a stochastic force caused by interactions with the gas
-in the vacuum chamber. This force is given by
-\begin{equation}
-    \label{eq:ffluct}
-    F_\text{fluct} = \sqrt{2m\Gamma_0k_BT_0} \ \xi\left(t\right)
-\end{equation}
-where $T_0$ is the temperature of the heat bath (i.e., the surrounding gas in the vacuum chamber), $k_B$ is the Boltzmann constant and $\xi(t)$ is
-white noise, which obeys the equations $\left\langle\xi(t)\right\rangle= 0$ and $\left\langle\xi(t)\xi(t')\right\rangle= \delta(t-t')$, which means
-that it is a random force. The term $\Gamma_0$ appears in the formula \eqref{eq:ffluct} 
-due to the fluctuation-dissipation theorem, which links the damping rate to the stochastic force.\\
-%The external force $F_\text{ext}$ is part of the experimental setup, where the frequency of the fluctuation along an axis, $\Omega_0$, is measured and
-%used to suppress the motion along said axis. This causes a decrease in the particle fluctuations and thus acts as a cooling mechanism for the particle
-%in the trap. This process creates a non-equilibrium steady state $\rho_{ss}(u,\alpha)$, which is not known analytically. This state is the starting
-%point of this thesis.\\
-\subsection{Model of the Surrounding Gas}
-J. Millen et al.~\cite{MillenJ.2014} worked on a similar experimental setup (without the feedback mechanism) and investigated the heating of the
-particle in the trap as well as the gas surrounding the particle.\\
-They start from the premise that there are 4 different temperatures one has to consider in such an experiment: the temperature of the gas particles
-before interacting with the nano particle $T_\text{imp}$ (for \textit{impinging}), the temperature of the gas particles after 
-interaction $T_\text{em}$ (for emerging), the surface temperature of the nano particle $T_\text{sur}$ and the temperature of the 
-center of mass of the nano particle $T_\text{COM}$. The experimental setup
-and the different temperatures are depicted in Fig.~\ref{fig:levitation}.\\
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[scale=0.2]{images/nnano_millen.jpg}
-        \caption{The experimental setup used by Millen et al. In \textbf{a} the experiment is schematically depicted and \textbf{b} shows a photograph
-            of the actual experimental setup. In \textbf{c} the 4 different temperatures that are present during the experiment are depicted and
-        \textbf{d} shows the position of the particle in the trap over time. The image was taken from~\cite{MillenJ.2014}}
-        \label{fig:levitation}
-    \end{center}
-\end{figure}
-Since the two temperatures of the gas, $T_\text{imp}$ and $T_\text{em}$, are not equal, the gas surrounding the nano particle is not in a thermal
-equilibrium, which is in contrast to the assumption in many optical tweezer experiments. Since the gas particles only interact with the nano particle
-in the trap and not wich each other, this situation creates two heat baths of different temperatures with the nano particle acting as a mediator
-between those two baths. This situation is depicted in
-Fig.~\ref{fig:heatbaths}.
-\begin{figure}[H]
-    \begin{center}
-        \includegraphics[scale=0.5]{images/nano_nonequilibrium_cropped.jpg}
-        \caption{Schematic picture of the two heat baths. The (hot) nano particle interacts with the (cold) incoming gas particles and they leave with
-        a higher temperature. The incoming and outgoing particles do not interact with each other and thus create two separate heat baths with
-    different temperatures. Image modified from \cite{Kroy2014}}
-        \label{fig:heatbaths}
-    \end{center}
-\end{figure}
-The one dimensional Langevin equation \eqref{eq:langevin} has to be modeled to include the different friction coefficients of the two heat baths. The
-model derived in \cite{MillenJ.2014} is described in the following equation:
-\begin{equation}
-    M\ddot{x}(t) + M\left(\Gamma^\text{imp}+\Gamma^\text{em}\right)\dot{x}(t) + M\omega^2x(t) =F^\text{imp}+F^\text{em}
-\end{equation}
-where $x$ is the position of the nano particle, $\dot{x}$ and $\ddot{x}$ are the velocity and the acceleration, respectively, $M$ is the mass,
-$\omega$ is the trap frequency, $\Gamma^\text{imp}$ and $\Gamma^\text{em}$ are the damping constants for the 
-impinging particles and the emerging particles, respectively and $F^\text{imp}$ and $F^\text{em}$ are the associated noise terms, 
-which satisfy $\langle F^\text{imp}\rangle= \langle F^\text{em} \rangle = 0$ individually.\\
-Miller et al. investigate the dependence of the emerging temperature $T_\text{em}$ on the intensity of the laser for different sized nano particles
-and different pressures.\\
-With these four temperatures introduced, some questions arise about the relationship between these temperatures. For this theses, I want to focus on
-the question: how does the laser intensity influence the motion of the nano particle in the laser trap and its center of mass velocity?\\
-In the next section I will introduce the methods that will be used to simulate the system, followed by the results of the simulations and the
-conclusion of this thesis.
-
-%These four temperatures introduced in the paper, together with the surrounding pressure and the intensity of the laser 
-%The existence of the two temperatures of the gas implies that, contrary to the assumption in many of the experiments, the gas is not in a thermal
-%equilibrium. The Langevin equation (like the one in \eqref{eq:langevin}) has to be modeled k
-
-\newpage
-\section{Simulation}
-%simulation techniques and general concept of the simulation itself, used methods and so on 
-The problem at hand can be studied on an atomic level with the use of computer simulation. There is a variety of methods for computer simulations
-that are widely used, one of which being molecular dynamics (MD) simulations.\\ 
-The goal is to simulate the experiment depicted in Fig.~\ref{fig:setup} as accurately as possible. To achieve this, the setup has to be broken down
-into individual pieced that can be modeled by using existing methods. The nano particle will be approximated by a cube of particles, sitting on an FCC
-lattice that interact via a Lennard-Jones potential. The laser will be broken down into its two main purposes, as it acts as a trapping device as well
-as a heat source for the nano particle. Surrounding the nano particle will be a gas chamber, that acts as a thermostat in the simulation, that will be
-modeled as a box surrounding the nano particle. This box is filled with gas particles that interact with the nano particle via a soft-sphere potential
-and do not interact with one another.\\
-The following section will give a brief overview of the concepts used to simulate every part of the experiment.
-
-\subsection{Molecular Dynamics}
-Molecular dynamics~\cite{Frenkel2001} simulations is a technique for simulating, as the name suggests, the dynamics of a classical many-body system. In this case,
-classical means, that the trajectories of the individual particles are calculated using classical mechanics rather then quantum mechanics. For
-relatively big atoms/molecules this is a very good approximation, whereas for systems consisting of hydrogen or helium the effects of quantum
-mechanics cannot be neglected and other methods have to be used.\\
-The dynamics of the system are obtained by solving Newton's equations of motion for every particle.
-
-\subsection{The Nano Particle}
-The glass particle from the experiment will be modeled as a system of particles interacting via a Lennard-Jones pair potential, 
-\begin{equation}
-    \label{eq:lj}
-    U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right]
-\end{equation}
-where $\varepsilon$ is the depth of the potential well (and thus its unit is energy) and $\sigma$ is the distance at which the potential is zero. 
-The form of the potential and the relation to the parameters is depicted in Fig. \ref{fig:lj}.
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.2]{images/LJ-2.pdf}
-        \caption{The Lennard-Jones 12-6 potential from \eqref{eq:lj}. The x-axis is the particle distance divided by $\sigma$ and  the y-axis is the
-        potential divided by the depth of the potential well.}
-        \label{fig:lj}
-    \end{center}
-\end{figure}
-Since $\varepsilon$ and $\sigma$ are crucial parameters for the simulation and do not change over time, it is practical to use them to
-define the dimensions of the system. This means that the unit of distance is $\sigma$, the unit of energy is $\varepsilon$ and the unit of mass is 
-the mass of the simulated particle. The so called \textit{reduced units} can be constructed from these three parameters and put into relation to the
-original units. Here are some examples:
-\begin{itemize}
-    \item {distance:} $r^* = r/\sigma$
-    \item {potential energy:} $U^* = U/\varepsilon$
-    \item {temperature:} $T^* = k_B T/\varepsilon$
-    \item {time:} $t^* = t\sqrt{\varepsilon/(m\sigma^2)}$
-    \item {pressure:} $P^* = P\sigma^3/\varepsilon$
-    \item {density:} $\rho^* = \rho \sigma^3$
-\end{itemize}
-One very popular choice for the simulated atoms is Argon because it is an inert gas and the atoms behave 
-approximately like hard spheres which attract each other with weak van der Waals forces, which justifies the use of the Lennard-Jones potential. 
-Argon has a mass of m = $6.69 \times 10^{-26}$ kg, $\sigma = 3.4 \times 10^{-10}$m and $\varepsilon = 1.65 \times 10^{-21}$J.\\
-With the above introduced reduced units, the Lennard-Jones potential can be written as
-\begin{equation}
-    U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right].
-\end{equation}
-Since the reduced units will be used throughout the rest of this thesis, I will drop the asterisk henceforth.\\
-From the Lennard-Jones potential the corresponding force can be calculated by taking the derivative with respect to the direction of interest:
-\begin{eqnarray}
-    F_{x} &=& -\frac{\partial}{\partial x} U(r) \nonumber\\
-                &=& -\frac{\partial}{\partial x} 4\left[{r}^{-12} - {r}^{-6}\right] \nonumber\\
-                &=& -4 \left[(-12){r}^{-13} - (-6){r}^{-7}\right] \frac{\partial r}{\partial x} \nonumber\\
-                &=& 48 \left[r^{-13} - 0.5 \ r^{-7}\right] \frac{x}{r} \nonumber\\
-    \label{eq:ljforce} &=& 48 \left[r^{-14} - 0.5 \ r^{-8}\right] x
-\end{eqnarray}
-The force in the y and z direction can be calculated analogously.\\
-As can be seen in Fig.~\ref{fig:lj}, the potential is very close to 0 at about $r/\sigma=2.5$. This means that any contribution of the potential
-energy after this point is vanishingly small. To save computing time it is common practice to cut off the potential at a certain point and don't
-calculate the contribution for distances beyond that. The cut-off distance for this thesis was set to $r = 2.5\sigma$.\\
-The initial configuration of the particles is a face centered cubic (FCC) lattice. A schematic of the FCC lattice is depicted in Fig. \ref{fig:fcc}.
-With the choice of FCC as initial configuration, there are optimal numbers for the numbers of the particles in the system. Since once FCC cell 
-shares its atoms with its next neighbours, the number of atom per unit cell is 4 -- 1/8 of a particle on eight corners and 1/2 of a particle on six faces.
-The whole system of atoms is then created by repeating this cell structure. One convenient way is to arrange the unit cells in a cubic system, 
-so if there are $M$ FCC unit cells on one edge, the whole system consists of $M^3$ cells. Since there are 4 particles per cell, 
-there are ideal or so called \textit{magic numbers} for atoms for which this setup works perfectly: $N = 4M^3 = 4,32,108,256,500,864,\ldots$.\\
-\begin{figure}[h]
-    \begin{center}
-        \begin{subfigure}[t]{0.4\textwidth}
-            \includegraphics[scale=0.3]{images/fcc-2.pdf}
-            \caption{Schematic figure of a face centered cubic (FCC) lattice}
-            \label{fig:fcc}
-        \end{subfigure} 
-        \
-        %\hspace{60pt}
-        \begin{subfigure}[t]{0.4\textwidth}
-            \includegraphics[scale=0.3]{images/unit_cell.pdf}
-            \caption{Setup for successively building a FCC lattice}
-            \label{fig:unitcell}
-        \end{subfigure}
-    \end{center}
-\end{figure}
-%\begin{figure}
-    %\begin{center}
-        %\includegraphics[scale=0.3]{images/fcc_unit.png}
-        %\caption{Schematic of the number of particles contained in one FCC unit cell. The atoms on the edges are shared by 8 cells and the ones on the
-            %faces by 2 cells. Image taken from
-            %{https://upload.wikimedia.org/wikipedia/commons/2/27/Elementarzelle\_einer\_kubisch\_raumzentrierten\_Elementarzelle.png}}
-    %\end{center}
-%\end{figure}
-% do i need to credit the uni buffalo course? 
-% http://www.physics.buffalo.edu/phy411-506/
-There are several ways to achieve this initial configuration and the one used in this thesis \cite{buffalo} was to create a kind of unit cell 
-consisting of four atoms, as shown in Fig. \ref{fig:unitcell}, which can be described by a set of points
-\begin{eqnarray*}
-    p_1 &=& \{0,0,0\}\\
-    p_2 &=& \{0.5*a,0.5*a,0\}\\
-    p_3 &=& \{0.5*a,0,0.5*a\}\\
-    p_4 &=& \{0,0.5*a,0.5*a\}
-\end{eqnarray*}
-From the particle number $N$ and the number of FCC unit cells per edge $M$ the \textit{lattice constant} $a$ can be calculated
-\begin{equation}
-    a = \frac{L}{M}
-\end{equation}
-where $L$ is the side length of the cube that is the whole system and it is calculated via the density of the system
-\begin{equation}
-    L = \sqrt[3]{\frac{N}{\rho}}
-\end{equation}
-With the lattice constant and the 4 points of the FCC cell, all the particles can be put into place.
-
-\subsection{The Velocity-Verlet Algorithm}
-\label{section:verlet}
-When we look at the system from a microscopic standpoint, we see that it follows some kind of path in the phase space as time progresses. Every
-point in this space corresponds to a set of positions and momenta and the connection between two points corresponds to the evolution of the system
-from one state to another. As mentioned above, this evolution (the dynamics of the system) is a crucial element to molecular dynamics. Since the
-equations of motion cannot be solved analytically in general, we need to approximate the solution.\\
-The method used here is called finite difference approach. The trajectory of the system in the phase space is cut into finite pieces of length $\Delta
-t$ and the equations of motion are solved for every segment separately (see Fig. \ref{fig:finitedifference}).\\
-\begin{figure}
-    \begin{center}
-        \includegraphics[scale=0.5]{images/finite_approach.pdf}
-        \caption{Simplified graphical schematic of the finite difference approach. The evolution of the system from a point $(r^N(t),p^N(t))$ in the
-            phase space is approximated by slicing it up into pieces of length $\Delta t$. On every stop after the starting point ($t+\Delta t$, 
-            $t+2\Delta t$, $t+3\Delta t$,$\ldots$) the equations of motion can be solved numerically.}
-        \label{fig:finitedifference} 
-    \end{center}
-\end{figure}
-There are several ways to solve this kind of problem, but since we are interested in implementing it into a computer program the ideal solution
-should have some basic properties \cite{Allen1989}:
-\begin{itemize}
-    \item It should be fast and require little memory
-    \item It should permit the use of a large time step $\Delta t$
-    \item It should duplicate the classical trajectory as closely as possible
-    \item It should satisfy known conservation laws
-    \item It should be simple and easy to program
-\end{itemize}
-One algorithm that has all of the above mentioned features is the one proposed by Verlet \cite{Verlet1967}. In his paper, Verlet starts by Taylor
-expanding the coordinate vector $\mathbf{r}_i$ for one particle after one time step $\Delta t$:
-\begin{equation}
-    \label{eq:verlet}
-    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + \dot{\mathbf{r}}_i \Delta t + \frac12 \ddot{\mathbf{r}}_i \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i \Delta t^3 +
-    \mathcal{O}(\Delta t^4)
-\end{equation}
-Since the first derivative of the coordinate is the velocity, $\dot{\mathbf{r}}_i(t) = \mathbf{v}_i(t)$, and the second derivative of the coordinate 
-is the acceleration, $\ddot{\mathbf{r}}_i(t) = \mathbf{a}_i(t)$, using Newton's law $\mathbf{F}_i(t) = m_i \mathbf{a}_i(t)$ equation \eqref{eq:verlet} 
-can be written as:
-\begin{equation}
-    \label{eq:verletnew}
-    \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 + \frac1{3!} \dddot{\mathbf{r}}_i(t)
-    \Delta t^3+\mathcal{O}(\Delta t^4)
-\end{equation}
-The same calculation can be carried out for one time step before $t$:
-\begin{equation}
-    \label{eq:verletnewminus}
-    \mathbf{r}_i(t-\Delta t) = \mathbf{r}_i(t) - {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2 - \frac1{3!} \dddot{\mathbf{r}}_i(t) \Delta t^3+
-    \mathcal{O}(\Delta t^4)
-\end{equation}
-The sum of those two equations yields
-\begin{equation}
-    \mathbf{r}_i(t+\Delta t) + \mathbf{r}_i(t-\Delta t) = 2\mathbf{r}_i(t) + \frac1{2} {\mathbf{F}}_i(t) \Delta t^2 + \mathcal{O}(\Delta t^4)
-\end{equation}
-and from this we get the final form for the new coordinates:
-\begin{equation}
-    \label{eq:verletfinal}
-    \mathbf{r}_i(t+\Delta t) = 2\mathbf{r}_i(t)+ \mathbf{r}_i(t-\Delta t)+ \frac1{2} {\mathbf{F}}_i(t) \Delta t^2
-\end{equation}
-This means that the new coordinates can be calculated using the current coordinates, the current forces and the coordinates from the past time step.\\
-From equation \eqref{eq:verletfinal} one thing becomes clear: the velocities are not necessary to calculate the new positions and they are not
-computed in the process, but for the calculation of i.e. the kinetic energy the velocities are needed. Thus, the velocities have to be calculated with
-a combination of the Taylor expansions of the coordinate vectors at $t+\Delta t$ and $t-\Delta t$:
-\begin{equation}
-    \mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t) = 2\mathbf{v}_i(t)\Delta t + \mathcal{O}(\Delta t^3)
-\end{equation}
-which can be rewritten as
-\begin{equation}
-    \label{eq:velocities}
-    \mathbf{v}_i(t) = \frac{\mathbf{r}_i(t+\Delta t) - \mathbf{r}_i(t-\Delta t)}{2\Delta t} + \mathcal{O}(\Delta t^2)
-\end{equation}
-This approach has the advantage that it is fast, requires little memory and is reliable in the sense that there is no energy drift occuring during the
-simulation, which means that the energy is conserved. When we compare this to the list of desired properties this seems like a good algorithm.\\
-This algorithm however has two significant disatvantages. The first one is the calculation of the velocities. As can be seen in equation
-\eqref{eq:velocities}, the accuracy of the calculation is only $\mathcal{O}(\Delta t^2)$ while the positions can be calculated with an error of order
-$\Delta t^4$. The other big disatvantage is the first step of the algorithm. Since the calculation of the new positions requires the current positions
-and the ones from one time step before, which technically do not exist.\\
-The solutions to this problem is to include the stepwise calculation of the velocities \cite{swope1982}. For this we start again with the Taylor
-expansion of the coordinates mathbftor \eqref{eq:verletnew} and instead of Taylor expanding $\mathbf{r}_i(t-\Delta t)$, we write it as 
-\begin{equation}
-    \mathbf{r}_i(t) = \mathbf{r}_i(t+\Delta t) - \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
-\end{equation}
-Using \eqref{eq:verletnew} in the abvove equation we get
-\begin{equation}
-    \begin{aligned}
-        \mathbf{r}_i(t) &= \mathbf{r}_i(t) + \mathbf{v}_i(t) \Delta t + \frac1{2m}\mathbf{F}_i(t)\Delta t^2 \\
-                     &- \mathbf{v}_i(t+\Delta t) \Delta t + \frac1{2m} \mathbf{F}_i(t+\Delta t) \Delta t^2
-    \end{aligned}
-\end{equation}
-and thus
-\begin{equation}
-    \label{eq:vvelocities}
-    \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \left(\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\right) \Delta t
-\end{equation}
-With the addition of this equation this algorithm is called the Velocity-Verlet algorithm, which is summarized in the following two equations:
-\begin{equation}
-    \label{eq:velocityverlet}
-    \begin{aligned}
-        \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2\\
-        \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \Big[\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\Big] \Delta t
-    \end{aligned}
-\end{equation}
-This algorithm is self starting, uses a small amount of memory, conserves the energy (not exactly) 
-and gives a very good approximation to the originial trajectory
-in the phase space. To program this algorithm the following steps are needed: 
-\begin{enumerate}
-    \item Calculate all the forces between the particles (for the first step only)
-    \item Calculate the new positions with current velocity and forces
-    \item Calculate first half of the new velocities with current forces
-    \item Calculate all the forces for the new position
-    \item Use new forces to calculate second half of new velocities
-\end{enumerate}
-The first point only has to be carried out for the first step, because the forces have not been calculated at this point. As the forces are calculated
-for the new positions in step 4, they can be used for the next time step. For the first time step the velocities have to be chosen randomly and are 
-calculated with this algorithm from that point forward.
-
-\subsection{The Laser Beam - Energy Influx}
-The nano particle is trapped in the laser beam. While the motion of the center of mass is localized, the
-individual atoms that make up the glass sphere absorb the energy from the laser which increases their velocity.\\
-To simulate this kind of behaviour, thermostat algorithms \cite{Huenenberger2005} such as the Nos\'e-Hoover \cite{Nose1984,Hoover1985} or the Andersen
-\cite{Andersen1980} algorithms are often used in simulation to change the temperature of the system in a controllable way.\\
-The problem with these kind of algorithms is, that a target temperature has to be fixed which will be reached at some point. In order to simulate the
-influx of energy from the laser it would be better to have an algorithm that supplies the system with a certain amount of energy continuously.
-Fortunately, such an algorithm exists.\\
-The algorithm is called Heat Exchange Algorithm (HEX) \cite{Hafskjold1993}. Its intended purpose is the use in non-equilibrium molecular dynamics
-(NEMD) to study transport phenomena and determine transport coefficients. The algorithm works by introducing two regions in the system, one serving as
-a heat source and the other as a heat sink. A specific amount of heat is then exchanged between those two reservoirs. As it turns out however, this
-algorithm introduces an energy shift for longer simulation times. This led Wirnsberger et al. \cite{Wirnsberger2015} to revisit the algorithm and 
-identify the cause of this energy drift, to create a more suitable algorithm, which they called eHEX.\\
-As in the HEX algorithm, regions are introduced to the system which act either as heat sinks or heat sources. These are labelled with $\Gamma_k$,
-where $k > 0$, and have corresponding amount of exchanged heat $\Delta Q_{\Gamma_k}$. If $\Delta Q_{\Gamma_k}$ is negative, heat is subtracted from
-the system and vice versa. Regions that neither act as heat source or heat sink are labelled with $\Gamma_0$, which are also called
-Hamiltonian regions (see Fig. \ref{fig:ehex}). 
-\begin{figure}
-    \begin{center}
-        \includegraphics[scale=0.3]{images/ehex_system.png}
-        \caption{Setup for the use of the HEX and eHEX algorithms. The simulationbox $\Omega$, which is moving with a velocity of $v_\Omega$,
-        contains a heat source (red), $\Gamma_1$, which is moving with velocity $v_{\Gamma_1}$ and a heat sink (blue), $\Gamma_2$, which is moving 
-        with velocity $v_{\Gamma_2}$. The regions that are neither heat sources or heat sinks are denoted by $\Gamma_0$.}
-        \label{fig:ehex} 
-    \end{center}
-\end{figure}
-The centers of mass of the particles in simulation box, denoted by $\Omega$, and the regions $\Gamma_k$ 
-are assumed to be moving with velocities $v_\Omega$ and $v_{\Gamma_k}$ respectively.\\
-The change of the energy in a region $\Gamma_k$ is achieved by rescaling the velocities by a factor $\xi_k$ and shifted by the velocity of the
-corresponding region:
-\begin{equation}
-    \mathbf{v}_i \rightarrow \mathbf{\bar{v}}_i = \xi_k \mathbf{v}_i + (1-\xi_k)\mathbf{v}_{\Gamma_k}
-\end{equation}
-The bar over a quantity denotes the value after the exchange of heat.\\
-The factor $\xi_k$ is given by
-\begin{equation}
-    \xi_k = \sqrt{1+\frac{\Delta Q_{\Gamma_k}}{\mathcal{K}_{\Gamma_k}}}
-\end{equation}
-where $\Delta Q_{\Gamma_k}$ is the exchanged heat in the region $\Gamma_k$ and $\mathcal{K}_{\Gamma_k}$ is the non-translational kinetic energy of the
-region $\Gamma_k$ and is given by
-\begin{equation}
-    \label{eq:kineticK}
-    \mathcal{K}_{\Gamma_k} = \sum_{i \in \gamma_k} \frac{m_i v_i^2}{2} - \frac{m_{\Gamma_k}v_{\Gamma_k}^2}{2}
-\end{equation}
-The sum is taken over all indices in $\gamma_k$ which is the set of indices of particles in the region $\Gamma_k$.\\
-For the final version of the eHEX algorithm there are three more quantities needed. The first one is the heat flux per time step, denoted by 
-$\mathcal{F}_\GammaK$:
-\begin{equation}
-    \mathcal{F}_\GammaK = \frac{\Delta Q_\GammaK}{\Delta t}
-\end{equation}
-The second one is the thermostatting force $\boldsymbol{\eta}_i$, which is defined as
-\begin{equation}
-    \boldsymbol{\eta}_i = 
-    \begin{cases}
-        m_i \ \dfrac{{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}}{{2\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}}} \
-        \left(\mathbf{v}_i - \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)
-        & \quad \text{if } k(\mathbf{r}_i) > 0 \\
-        0 & \quad \text{otherwise}
-    \end{cases}
-    %\begin{cases}
-        %m_i \ \frac{\displaystyle{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}}{\displaystyle{2\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}}} \
-        %\left(\mathbf{v}_i - \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)
-        %& \quad \text{if } k(\mathbf{r}_i) > 0 \\
-        %0 & \quad \text{otherwise}
-    %\end{cases}
-\end{equation}
-where $k(\mathbf{r}_i)$ is the index of the region in which particle $i$ located, i.e. $k(\mathbf{r}_i) = 0$ means that the particle is in a Hamiltonian
-region and $k(\mathbf{r}_i) > 0$ denotes the heat sinks and sources.\\
-The last quantity is the one that corrects the long term energy drift of the HEX algorithm, denoted by $\mathcal{E}{r}_{i,\alpha}$.
-The analysis and derivation of this term is given in \cite{Wirnsberger2015}.
-\begin{equation}
-    \label{eq:bigeps}
-    \begin{aligned}
-        \mathcal{E}{r}_{i,\alpha} &= \dfrac{\eta_{i,\alpha}}{m_i\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}} \left[
-        \frac{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}{48} + \frac{1}{6} \sum_{j\in\gamma_{k(\mathbf{r}_i)}} \mathbf{f}_j \cdot \left(\mathbf{v}_j - 
-            \mathbf{v}_{\Gamma_{k(\mathbf{r}_i)}}\right)\right] \\
-            &- \frac{\mathcal{F}_{\Gamma_{k(\mathbf{r}_i)}}}{12\mathcal{K}_{\Gamma_{k(\mathbf{r}_i)}}} \left[
-            \frac{f_{i,\alpha}}{m_i} - \frac{1}{m_{\Gamma_{k(\mathbf{r}_i)}}} \sum_{j\in\gamma_{k(\mathbf{r}_i)}}f_{j,\alpha}\right]
-    \end{aligned}
-\end{equation}
-The $\mathbf{f}$ in the above equation denotes the force corresponding to the chosen intermolecular potential $U(\mathbf{r})$, 
-\begin{equation}
-    \mathbf{f} = -\nabla_{\mathbf{r}_i} U(\mathbf{r}_i)
-\end{equation}
-With all the necessary quantities established, the updating sequence of the eHEX algorithm can be written down:
-\begin{subequations}
-\begin{eqnarray}
-    \bar{\mathbf{v}}^n_i &=& \xi^n_{k(\mathbf{r}_i)} \mathbf{v}^n_i + \left(1-\xi^n_{k(\mathbf{r}_i)}\right) \mathbf{v}^n_{\Gamma_{k(\mathbf{r}_i)}} \\
-    \bar{\mathbf{v}}^{n+\frac12}_i &=& \bar{\mathbf{v}}^n_i + \frac{\Delta t}{2m_i} \mathbf{f}^n_i\\
-    \bar{\mathbf{r}}^{n+1}_i &=& \mathbf{r}^n_i + \Delta t \bar{\mathbf{v}}^{n+\frac12}_i\\
-    \mathbf{f}^{n+1}_i &=& \left.-\nabla_{\mathbf{r}_i} U(\mathbf{r})\right|_{\mathbf{r} = \bar{\mathbf{r}}^{n+1}}\\
-    \bar{\mathbf{v}}^{n+1}_i &=& \bar{\mathbf{v}}^{n+\frac12}_i + \frac{\Delta t}{2m_i} \mathbf{f}^{n+1}_i\\
-    \mathbf{v}^{n+1}_i &=& \bar{\xi}^{n+1}_{k(\bar{\mathbf{r}}_i)} \bar{\mathbf{v}}^{n+1}_i + \left(1-\bar{\xi}^{n+1}_{k(\bar{\mathbf{r}}_i)}\right)
-    \bar{\mathbf{v}}^{n+1}_{\Gamma_{k(\bar{\mathbf{r}}_i)}} \\
-    \mathbf{r}^{n+1}_i &=& \bar{\mathbf{r}}^{n+1} - \Delta t^3 \mathcal{E}\bar{\mathbf{r}}^{n+1}_i
-\end{eqnarray}
-\end{subequations}
-This algorithm can be adapted for the problem of the levitating nano sphere in the laser beam by adjusting the setup and some of the parameters.\\
-Firstly, the setup of the heat sinks and sources has to be changed. Since there is only energy pumped into the system from the outside, the region
-acting as a heat sink vanishes and the region acting as a heat source spans over the whole simulation box. This means that, using the 
-notation of Fig. \ref{fig:ehex}, $\Omega = \Gamma_1$. Furthermore, neither center of mass of the particles the simulation box nor 
-the heat source are moving, i.e. 
-$\mathbf{v}_\Omega = \mathbf{v}_{\Gamma_1} = 0$. This affects the non-translational kinetic energy term $\mathcal{K}_{\Gamma_1}$, the thermostatting
-force $\boldsymbol{\eta}$ and the correction term $\mathcal{E}\mathbf{r}$. Since there is only one region acting as a heat source, the terms $\Delta Q$,
-$\mathcal{K}$ and $\mathcal{F}$ don't need an index. The summation index in \eqref{eq:kineticK} and \eqref{eq:bigeps} can be changed to the number of
-particles in the system, $N$, since all the particles are within the heat exchanging region. The masses are set to 1, i.e. $m_i = 1$ with the chosen
-reduced units and with this the total mass of the region (in the term $1/m_{\Gamma_{k(\mathbf{r}_i)}}$ in \eqref{eq:bigeps}) is equal to the number of
-particles in the system. With these changes, the quantities can be written down as:
-\begin{eqnarray}
-    \mathcal{K} &=& \sum_{N} \frac{v_i^2}{2}\\
-    \xi &=& \sqrt{1+\frac{\Delta Q}{\mathcal{K}}}\\
-    \mathcal{F} &=& \frac{\Delta Q}{\Delta t}\\
-    \boldsymbol{\eta}_i &=& 
-    \dfrac{\mathcal{F}}{{2\mathcal{K}}} \
-        \mathbf{v}_i\\
-        \mathcal{E}{r}_{i,\alpha} &=& \dfrac{\eta_{i,\alpha}}{\mathcal{K}} \left[
-        \frac{\mathcal{F}}{48} + \frac{1}{6} \sum_{N} \mathbf{f}_j \cdot \mathbf{v}_j
-            \right] \nonumber \\
-            &-& \frac{\mathcal{F}}{12\mathcal{K}} \left[
-            {f_{i,\alpha}} - \frac{1}{N} \sum_{N}f_{j,\alpha}\right]
-\end{eqnarray}
-This means that the laser is modeled to be pumping energy into the system, which increases the velocities of the particles in the
-system over time, while the center of mass motion is not affected by this.\\
-
-\subsection{The Laser Beam - Trapping}
-As mentioned above, the glass particle is trapped in the laser beam and the position of the particle is localized. This behaviour
-has to be modeled as well.\\
-The approximation of the original paper \cite{Gieseler2014}, where the movements in the three spatial directions are decoupled will be
-used in the model as well. As to the model of the trapping force itself, the most straightforward approach is to use a harmonic oscillator potential.
-The force and the corresponding potential can be written as
-\begin{eqnarray}
-    \mathbf{F} &=& -k \Big[\mathbf{x}-\mathbf{x}_0\Big]\\
-    U &=& \frac12 k \Big[\mathbf{x}-\mathbf{x}_0\Big]^2
-\end{eqnarray}
-where $\mathbf{x}_0$ is the position of minimal potential energy. 
-Since the trap is acting on the whole system, the force is acting on the center of mass. 
-To calculate the center of mass positions and velocities, the positions and velocities of all particles have to be summed up and 
-divided by the particle number:
-\begin{equation}
-    \mathbf{r}_\text{COM} = \frac1N\sum_{i=1}^N \mathbf{r}_i
-\end{equation}
-
-\subsection{Surrounding Gas - Thermostat}
-Without any equilibrating mechanism the setup described by now would lead to the system heating up indefinetly, which is not desirable. Furthermore,
-the goal is to recreate the experiment as close as possible. Since the nano particle is trapped in a laser beam within a vacuum chamber, the
-surrounding gas has to be modeled as well. In the next step we will introduce the surrounding gas of the gas chamber that will absorb some of the 
-energy in the system, leading to the final state.\\
-Generally, pressure is introduced to the system by surrounding the object of interest (in this case the glass nano sphere) with a thermostatting
-pressure medium.
-There are two main requirements for the choice of such a pressure medium: the exerted pressure must be hydrostatic and the computation of the
-interaction between the pressure medium and the object of interest must not take up a lot of resources.\\
-The model used in this thesis was developed by Gr\"unwald and Dellago \cite{Gruenwald2006} and uses an ideal gas of non-interacting particles as
-thermostatting pressure medium. The particles of this pressure medium flow into the simulation from an outside volume, whose geometry is based on the form of the
-object of interest (this will be referred to as the minimal volume of cells) and leave the simulation box if their position reaches the boundary of
-this minimal volume of cells. This behaviour is very close to the real experiment which makes this thermostat an ideal candidate for the simulation of
-the experiment. 
-%This barostat also acts as a thermostat and thus is suitable for the 
-%application in this thesis.\\
-The gas particles interact with the object via a soft-sphere potential of the form
-\begin{equation}
-    \label{eq:softsphere}
-    U(r) = \varepsilon \left(\frac{\sigma}{r}\right)^{12}
-\end{equation}
-where $\varepsilon$ is the interaction strength and $\sigma$ is the interaction range and $r$ is the distance between the gas particle and the
-interacting particle. The gas particles do not interact with one another in this model.\\
-In order to increase the efficiency of the computing process, $\sigma$ should be chosen carefully. For larger $\sigma$, the number of interaction
-partners increases, which increases the number of force calculations which are a very time consuming part. For smaller $\sigma$ the possibility for
-gas particles reaching the inside of the nano particle increases, which is not desirable. So $\sigma$ should be chosen small enough to keep the force
-calculations at a minimum and large enough for the particle to stay on the outside of the crystal. In the computer simulations performed for this
-thesis, the interaction length is chosen to be $\sigma = 1$ with a cut-off radius of $r_c = 2.5$.\\ 
-The algorithm can be performed by following these steps:
-\begin{enumerate}
-    \item Randomly draw the number of particles that are created on a single side of the minimal volume of cells, $N_\text{fac}$, from the
-        distribution 
-        \begin{equation}
-            %\langle N_\text{fac}\rangle = \Delta t L^2 P \left(\frac{1}{2\pi m k_B T}\right)^\frac12
-            \label{eq:numberofparticles}
-            \langle N_\text{fac}\rangle = \Delta t L^2 P \sqrt{\frac{1}{2\pi m k_B T}}
-        \end{equation}
-        where $\Delta t$ is the time step of the simulation, $L$ is the side length of the cell in which the particle is created, $P$ is the desired
-        pressure, $m$ is the mass of the gas particle, $k_B$ is the Boltzmann constant (which will be set to 1 in reduced units) and $T$ is the
-        desired temperature. This chosen number of particles is then equally distributed over the face of the cell on which they are created. The
-        velocities of the created particles are drawn from two different random number distributions. The component of the particle perpendicular to
-        the surface is drawn from a Rayleigh distribution of the form 
-        \begin{equation}
-            p(v_i) = \frac{m}{k_B T}v_i \ e^{-\frac{mv_i^2}{2k_BT}}
-        \end{equation}
-        The other components of the particles' velocity are drawn from a Maxwell-Boltzmann distribution.
-    \item Perform the first step of the velocity Verlet algorithm to propagate the particle positions by one time step.
-    \item Check if any gas particles have left the minimal volume of cells and remove those which have.
-    \item Check if the geometry of the crystal and with it the minimal volume of cells has changed. If it has, remove all gas particles in the cells
-        that are no longer needed. Then insert new gas particles to the created cells with a number drawn from a Poisson distribution with mean value
-        \begin{equation}
-            \langle N_\text{ins}\rangle = \frac{PL^3}{k_BT}
-        \end{equation}
-        If the length of the cell $L$ is chosen to be equal to the cut-off radius $r_c$, this insertion should only be carried out with a probablity of
-        \begin{equation}
-            P_\text{ins} = e^{-\frac{U}{k_BT}}
-        \end{equation}
-        (where U is the interaction energy between the gas particle and the crystal) because it is possible that the inserted particle is 
-        within the interaction range of the crystal. 
-    \item Compute all forces.
-    \item Perform the second step of the velocity Verlet and propagate the particle velocities by one time step. 
-\end{enumerate}
-This algorithm is formulated in a general fashion, so that a wide range of crystals and geometries can be used. 
-%The setup used in this thesis is a FCC lattice, which has a nice symmetry that is not influenced by the surrounding gas. 
-Although the shape of the nano particle changes over the course of the simulation, it is not necessary to change the geometry of the surrounding box.
-The algorithm will therefore not be carried out using cell-lists and a minimal volume of cells that is changing over the course of the simulation, 
-but rather a fixed setup of the volume surrounding the nano particle.\\ 
-Since the particle itself is modeled as a cube, a straighforward approach is surrounding it by a bigger cube. The distance between the cube face and the nano
-particle is chosen to be the side length of the nano particle, so the setup will properly scale for different numbers of particles. 
-The schematic setup for this is depicted in Fig.\ref{fig:gascube}.
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.5]{images/gas_volume_ver3.pdf}
-        \caption{Schematic setup for the barostat. The outer volume (blue) is one side length $s$ away from the nano particle (black) which means that
-            the side length of the face of the surrounding box is $L=3s$. 
-                    The gas particles are created along the blue lines of the outer volume. Since the overall geometry of the nano particle does 
-                    not change, 
-                    the geometry of the outer volume will be constant throughout the simulation.}
-        \label{fig:gascube}
-    \end{center}
-\end{figure}
-
-%This is done by adding for instance two regions to the simulation, one providing an influx of heat
-%and the other acting as a heat sink. The algorithm is furthermore designed to incorporate the velocities of the center of mass of the system and the
-%velocities of the regions.
-
-%\subsection{The Glass Nanoparticle}
-%The glass particle from the experiment will be represented by a system of 864 particles, aligned regularly on a FCC (face centered cubic) lattice.
-%This alignment is achieved by defining a minimum cell, containing points:
-%\begin{eqnarray*}
-    %p_1 &=& \{0,0,0\}\\
-    %p_2 &=& \{0.5,0.5,0\}\\
-    %p_3 &=& \{0.5,0,0.5\}\\
-    %p_4 &=& \{0,0.5,0.5\}
-%\end{eqnarray*}
-%The whole system is then  created by copying this unit cell.\\
-%The interaction between the atoms is modeled by the Lennard-Jones potential. It has the form
-%\begin{equation}
-    %U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right].
-%\end{equation}
-%To simplify the equation and computation of the potential and other quantities, like the forces or pressure, it is useful to introduce reduced units.
-%In general, reduced units BLABLABLA THINK OF GOOD TEXT HEREJKA\\
-%The basic units in a system with Lennard-Jones interaction are length ($\sigma$), energy ($\varepsilon$ or $\varepsilon / k_B$) and mass ($m$). Every 
-%quantity can now be written in terms of this units, so they become reduced quantities, denotet by an asterisk. The most important ones are:
-%\begin{eqnarray*}
-    %r^* &=& r/\sigma\\
-    %T^* &=& k_B/\varepsilon \ T\\
-    %U^* &=& U/\varepsilon \\
-    %P^* &=& P \sigma^3/\varepsilon 
-%\end{eqnarray*}
-%With the above introduced reduced units, the Lennard-Jones potential can be written as
-%\begin{equation}
-    %U(r^*) = 4\left[{r^*}^{-12} - {r^*}^{-6}\right].
-%\end{equation}
-%Since this is the most practical form, it will be used without the asterisk from here on.\\
-%The Lennard-Jones potential is an additive pair-potential, so the total energy of the system can be calculated by summing over all pairs of atoms:
-%\begin{equation}
-    %U_{\text{tot}} = \sum_{i=1}^{N-1}\sum_{j=i+1}^N4\left[{r_{ij}}^{-12} - {r_{ij}}^{-6}\right]
-%\end{equation}
-%where $r_{ij}$ denotes the distance between atom $i$ and $j$. \\
-%The Velocity-Verlet algorithm uses the forces to calculate the positions and velocities for the next timestep, we need to calculate the derivative of
-%the potential energy:
-%\begin{eqnarray}
-    %F_{x} &=& -\frac{\partial}{\partial x} U(r) \nonumber\\
-                %&=& -\frac{\partial}{\partial x} 4\left[{r}^{-12} - {r}^{-6}\right] \nonumber\\
-                %&=& -4 \left[(-12){r}^{-13} - (-6){r}^{-7}\right] \frac{\partial r}{\partial x} \nonumber\\
-                %&=& 48 \left[r^{-13} - 0.5 \ r^{-7}\right] \frac{x}{r} \nonumber\\
-    %\label{eq:ljforce} &=& 48 \left[r^{-14} - 0.5 \ r^{-8}\right] x
-%\end{eqnarray}
-%This is the component of the force in x-direction -- the other components are calculated analogously.
-
-\newpage
-\section{Results}
-% results with different parameters, graphs, etc.
-Since the simulation consists of various parts that have to work together, we first need to check if every part itself works.\\
-\subsection{The Crystal}
-The first element of the simulation is the setup of the FCC lattice on which the particles are placed on. To check if this is done correctly, we can
-use a visualisation and rendering software such as VMD\cite{Humphrey1996} or OVITO\cite{Stukowski2010}. The rendered image of the nano particle 
-can be seen in Fig.~\ref{fig:glass}. The chosen perspective shows the structure of the FCC lattice, where 3 layers are forming 
-where the atoms take the position on the gap between the atoms from the layer below.\\
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.2]{images/fcc_rendering_layer_newcoloring.png}
-        \caption{A rendered image of the FCC lattice. The chosen color-coding along the (1,1,1) Miller-index 
-                highlights the layers for this perspective. The perspective is
-                chosen to see the layering of the FCC lattice with its 3 layer structure. The black box acts as a reference point to outline the 
-                geometry of the system.}
-            %Rendering of the model of the nano crystal before starting the simulation. 
-            %The chosen perspective shows the 3 layers of the FCC crystal with a
-                    %These layers consist of atoms
-                    %taking the position above the gap between two of the atoms from the layer below and match the positions of the atoms 2 layers
-                %below. The black box is the outline of the particles' positions and shows the geometry of the system.}
-        \label{fig:glass}
-    \end{center}
-\end{figure}
-
-\subsection{Velocity Verlet}
-The velocity Verlet part of the simulation is taking place in a NVE ensemble from a thermodynamic point of view. 
-This means that the particle number $N$, the volume $V$ and the total energy $E$ are constant. 
-The initial velocities of the particles are chosen randomly from a Maxwell-Boltzmann distribution. To ensure that the system behaves the way it is
-intended to, small adjustments have to be made in the beginning of the simulation of the nano particle.\\
-%In contrast to the NVT ensemble used in Monte-Carlo simulations this means, that the velocities of the particle can not
-%be chosen directly. The initial velocities have to be drawn from a random number generator. This means that the velocities have to be adjusted after 
-%they have been drawn.\\
-The first adjustment that needs to be made is making sure that the system is staying in place and not floating away. To do this, the velocity vectors
-of all the atoms in the system have to be added to get the resulting velocity vector for the whole system. 
-This resulting vector is then divided by the number of particles in the system and the resulting vector is then subtracted from every
-velcocity vector of every atom in the system. 
-\begin{eqnarray}
-    \mathbf{v}_\text{tot} &=& \sum_{i=1}^N \mathbf{v}_i\\
-    \mathbf{v}_\text{cut} &=& \frac{\mathbf{v}_\text{tot}}N\\
-    \mathbf{v}_i^\text{new} &=& \mathbf{v}_i - \mathbf{v}_\text{cut}
-\end{eqnarray}
-This adjustment only needs to be done just after randomly choosing the velocities, as there is no force present at this point that would cause the
-nano particle to move.
-%The next adjustment is a small one that makes a difference in the long run. When choosing random velocities, the system not only tends to have a
-%resulting velocity vector that causes the whole system to float away, but also a resulting angular momentum that causes the system to rotate.
-%The angular momentum can be defined via the moment of inertia $\mathbf{I}$
-%\begin{eqnarray}
-    %\label{eq:angularmomentum}\mathbf{L} &=& \mathbf{I}\boldsymbol{\omega}\\
-    %\boldsymbol{\omega} &=& \frac{\mathbf{r} \times \mathbf{v}}{r^2}
-%\end{eqnarray}
-%where $\boldsymbol{\omega}$ is the angular velocity of the system. Another way to define the angular momentum is given by
-%\begin{equation}
-    %\label{eq:angularmomentum2}
-    %\mathbf{L} = \mathbf{r} \times \mathbf{p}
-%\end{equation}
-%where $\mathbf{p}$ is the momentum of the particle, which in this case is equal to its velocity since the mass is set to 1.The moment of inertia 
-%$\mathbf{I}$ for a filled cube is given by 
-%\begin{equation}
-    %\mathbf{I} = 
-    %\begin{pmatrix}
-        %\frac{mL^2}{6} & 0 & 0\\
-        %0 & \frac{mL^2}{6} & 0\\
-        %0 & 0 & \frac{mL^2}{6} 
-    %\end{pmatrix}
-%\end{equation}
-%where $m$ will be later set to 1 and $L$ is the side length of the cube. To use this equations to suppress the rotation of the cube, we need to use
-%equation~\eqref{eq:angularmomentum} and rewrite it as
-%\begin{equation}
-    %\label{eq:omega}
-    %\boldsymbol{\omega} = \mathbf{I}^{-1}\mathbf{L}
-%\end{equation}
-%where $\mathbf{I}^{-1}$ is the inverse matrix of $\mathbf{I}$. The calculation of the inverse matrix is fairly straightforward, as it only contains
-%values in the diagonal, which means that the inverse matrix can be calculated by simply inverting these values. The angular momentum for every
-%particle $\mathbf{L}_i$ in the above equation be calculated using eq.~\eqref{eq:angularmomentum2}, where the position is taken to be the distance
-%between the particle and the center of mass
-%\begin{equation}
-    %\Delta\mathbf{r}_i = \mathbf{r}_\text{COM} - \mathbf{r}_i
-%\end{equation}
-%The values of the angular momentum for each paricle are added up and multiplied by the inverse of the moment of inertia, as described in
-%eq.~\eqref{eq:omega}. The final step is updating the velocities of the particles:
-%\begin{equation}
-    %\mathbf{v}_i^\text{new} = \mathbf{v}_i - \left(\boldsymbol{\omega} \times \Delta\mathbf{r}_i\right)
-%\end{equation}
-%\\The velocities of the particle can roughly be chosen by adjusting the range of numbers the random number generator can choose from. However, the
-%resulting velocities (and therefore the temperature of the system) are generally not precisely how they should be. 
-\\As mentioned above, the velocities are chosen from a Maxwell-Boltzmann distribution, which means choosing each component from a Gaussian
-distribution. However, the positions on the FCC lattice are not very natural and so the atoms equilibrate to a different temperature than desired.
-If the system should have a
-specific starting temperature $T$, the velocities need to be scaled accordingly. This rescaling is done by summing the square of the velocities
-(this equals calculating the kinetic energy since the masses are set to 1) and rescale them by a factor $\lambda$:
-\begin{eqnarray}
-\label{eq:rescale1}   E &=& \sum_{i=1}^N \frac{\mathbf{v}_i^2}{2} \\
- \label{eq:rescale2}   \lambda &=& \sqrt{\frac{3(N-1)T}{2E}} \\
- \label{eq:rescale3}   \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i  
-\end{eqnarray}
-This process has to be done a couple of times, until the system equilibrates to the desired temperature. 
-%The problem with this procedure is the fact
-%that it is not compatible with the requirement of constant energy of the system. 
-If we look at the energy values in Fig.~\ref{fig:energystart} we can see the
-irregularities (jumps) before the system equilibrates to a constant value. This equilibration process divides the simulation in two phases: the equilibration
-phase and the measurement phase. In the equilibration phase, the system has time to reach a configuration that is (as the name suggests) in an 
-equilibrated state. During this phase, there are usually no measurements (such as energy) performed, since the values from this phase are distorting
-the mean value of the simulation. That is why the configurations from the equilibration phase are discarded and the measurement phase takes place, 
-where the values do get calculated and are used to determine the mean values.  
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.3]{images/verlet_output_te.pdf}
-        \caption{Temperature and energy in the equilibration phase of the velocity Verlet algorithm. The velocity rescaling (see eq. 
-        \eqref{eq:rescale1}-\eqref{eq:rescale3}) is taking place every 1000th step until the 9000th, which can be seen in the energy curve whre the
-    jumps are happening. The desired temperature in this case is $T^* = 0.2$, which is reached after about 7000 steps.}
-        \label{fig:energystart}
-    \end{center}
-\end{figure}
-
-\subsection{eHEX}
-The next checkpoint is the eHEX algorithm. Since it contains several steps and variables that have to be calculated, this algorithm is very prone
-to errors. Perhaps the most crucial of those variables in the algorithm is the amount of heat injected into the system, $\Delta Q$. 
-The effect of the value of this variable on the system has therefore to be checked to make sure the system is not overheating.\\
-The investigation of the effects on the system can be done in two ways: comparing the temperature and the energy of the system during the 
-application of the algorithm (similar to the way it's been done with the velocity Verlet) and comparing the effect of different values of $\Delta Q$.\\
-%Like mentioned above, there is an equilibration phase, which takes place before the recording of the values for this comparison. The graph starts 
-%with a velocity Verlet phase after which the eHEX algorithm starts, to emphasize the effect of the change in temperature and energy.\\
-Since the eHEX algorithm injects heat into the system, one would expect the energy and the temperature to rise during the application of the
-algorithm. In Fig.~\ref{fig:ehex_te} we can see that exactly this is happening. The graph shows a velocity Verlet phase (that follows an equilibration
-phase which is not recorded) of constant temperature and energy and an eHEX phase, where the temperature and the enrgy are constantly rising. 
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/ehex_te.pdf}
-        \caption{Comparison of temperature (red) and energy (blue) during the application of the eHEX algorithm with $\Delta Q=0.04$ (in reduced units).
-                The first 20000 timesteps, the energy and temperature are constant during the application of the velocity Verlet algorithm and start
-            to rise, once the eHEX algortithm takes over. The values for the temperature can be seen on the ordinate on the left hand side
-            and the values for the energy on the right hand side. }
-        \label{fig:ehex_te}
-    \end{center}
-\end{figure}
-The comparison of different values of $\Delta Q$ can be seen in Fig.~\ref{fig:ehex_dq}. As the correlation between the temperature and the energy has
-been shown in Fig.~\ref{fig:ehex_te}, the plotting of the temperature data should suffice. 
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/ehex_temp_dq.pdf}
-        \caption{Comparison of different values for $\Delta Q$, ranging from 0 to 0.09. As the value $\Delta Q=0$ corresponds to no heat being
-        injected into the system, it is not surprising that the temperature remains constant. The graph shows that an increase in $\Delta Q$ leads to
-    an increase in the rate of warming up, which means that for bigger $\Delta Q$, the system is heating up more quickly.}
-        \label{fig:ehex_dq}
-    \end{center}
-\end{figure}
-
-\subsection{Thermostat}
-The barostat, as described previously, keeps the system from overheating. The barostat algorithm is applied at the same time as the eHEX algorithm,
-after the velocity Verlet has equilibrated the system.\\
-While the eHEX algorithm is pumping energy into the nano particle while the surrounding gas particles absorb excessive energy. This leads to the
-development of a new, higher steady temperature of the nano particle, which can be seen in Fig.~\ref{fig:baro1}. 
-%The first simulations were performed
-%with the eHEX and the barostat algorithm starting at the same point. This lead to the question: does the new steady temperature depend on the time it
-%takes the gas particles to reach the system? The answer is: no. After applying the barostat algorithm first to fill the box with gas atoms and then
-%applying the eHEX algorithm it can be seen that the resulting temperature is the same, which can be seen in fig.~\ref{fig:baro2}.
-%This means that the system is beginning to heat up, until the particles reach the crystal.
-%Because of this, the temperature of the system rises, until it reaches a steady value again. This behaviour can be seen in fig.~\ref{fig:baro1}.\\
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/baro_temp_long_new.pdf}
-        \caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
-        in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
-    an increase in temperature until a new steady temperature is reached.}
-        %\caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
-        %in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
-    %an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
-%reached and the temperature seems to stay at this value.}
-        \label{fig:baro1}
-    \end{center}
-\end{figure}
-%\begin{figure}[h]
-    %\begin{center}
-        %\includegraphics[scale=0.4]{images/baro_temp_compare_other.pdf}
-        %\caption{Two versions of the application of the eEHX and barostat algorithm. The shorter run (red, x-axis on the bottom) represents 
-            %the simultaneous start of the eHEX and the barostat algorithm. In the longer run (blue, x-axis on the top) there are 20000 steps, where
-        %the barostat algorithm is applied at the same time as the velocity Verlet to give the system time to equilibrate and only after the eHEX
-    %algorithm is taking the place of the velocity Verlet. The mean value of the resulting temperature is the same, indicating that the final
-%temperature is based on the values of the parameters, such as pressure and ambient temperature, rather than geometry.}
-        %%\caption{Temperature of the system during the application of the eHEX and the barostat algorithm. During the first 20000 steps the system is
-        %%in equilibrium while the velocity Verlet algorithm is applied. After that, the eHEX and the barostat algorithms are turned on. This leads to
-    %%an increase in temperature until the gas particles reach the nano crystal to slow down the heating of the system. After some time, a plateau is
-%%reached and the temperature seems to stay at this value.}
-        %\label{fig:baro1}
-    %\end{center}
-%\end{figure}
-%As we are interested in the temperatures of the particles before and after the interaction with the nano particle, we need some way to differentiate those
-%two states. This can be done when looking at the total energy of the gas particles.\\
-%The gas particles are created on the surface of the simulation box, which is further away than the interaction range between the gas particles and the
-%nano particle. Since the gas particles do not interact with each other, the total energy of the gas particles does not change until they reach the
-%interaction range of the nano particle. There they undergo a change of energy due to the interaction with the nano particle and then make their way away
-%from the nano particle. After the interaction, when they reach a point outside of the interaction range, the enrgy stays constant again until the
-%particles reach the limits of the simulation box.\\
-%Therefore, the life span of a single gas particle can be partioned into 3 stages: incoming, interacting, outgoing. In a typical scene of interaction
-%all three of these stages are represented by different gas particles, as can be seen in Fig.~\ref{fig:barointeraction}.\\
-%\begin{figure}[h]
-    %\begin{center}
-        %\includegraphics[scale=0.4]{images/baro_170105_new_medium.png}
-        %\caption{Rendering of the application of the barostat algorithm. The red particles are incoming gas particles, that have not yet interacted, 
-            %pink particles are currently interacting and green particles are leaving the simulation box and
-        %moving away from the nano particle (blue). To render this image, a cut has been made to make a view into the simulation box possible.}
-        %\label{fig:barointeraction}
-    %\end{center}
-%\end{figure}
-%Since the gas particles are interacting with the nano particle, the temperatures and energies are values of interest. The energies are recorded using
-%histograms and the temperature is calculated instantaneously every 100th timestep.\\
-%The calculation of the energies of the incoming particles has to be done with care. The problem are particles that enter the simulation box, but fly
-%by the nano particle at a distance that is greater than the interaction range. Since the change of the energy is the interesting part, only the energy 
-%particles of that \textit{do} interact with the nano particle shall be calculated. This can be achieved by counting particles that enter the stage of
-%\textit{interacing}. One such histogram can be seen in Fig.~\ref{fig:histogram}.
-%\begin{figure}[h]
-    %\begin{center}
-        %\includegraphics[scale=0.4]{images/histogram_2.pdf}
-        %\caption{Histogram of total energies for incoming (red) and outgoing (blue) particles. The parameters for this histogram are $\Delta Q^*=0.04$,
-        %$T_\text{ambient}^*=0.08$, $P^*=0.8$. What can be seen here is that the outgoing particles have a higher count of particles with more total
-    %energy.}
-        %\label{fig:histogram}
-    %\end{center}
-%\end{figure}
-
-\subsection{Measurement}
-Now that the functionality of every individual part has been established, the main question can be tackled: how does the laser intensity influence the
-center of mass motion of the nano particle? And which influence does the surrounding gas exert?\\
-In order to investigate the origin of the motion of the center of mass, we first need to check whether the eHEX algorithm itself has any influence on
-the center of mass motion. This is done by applying the algorithm without the thermostat and calculating the center of mass velocity. This turns out
-to be the case, i.e. the center of mass motion stands still.\\
-To check the integrity of the algortihm, a simple case needs to be tested. Such a simple case is $\Delta Q = 0$, i.e. no energy begin pumped into the
-system. The expected outcome of this case is a thermal equilibration between the incoming gas particles, the outgoing gas particles and the center of
-mass motion of the nano particle with the incoming gas temperature being the reference value. 
-The measurement yields this result and backs up the claim. Furthermore, the internal temperature of the nano particle also equilibrates to the 
-temperature of the surrounding gas. This was checked on a system with pressure $P=0.8$, temperature of the incoming particles $T_\text{imp} = 0.05$, 
-internal temperature of the nano particle (set with the velocity verlet) $T_\text{intern} = 0.2$ and a system size of $N=32$ atoms in the nano
-particle. The resulting temperatures are:
-\begin{eqnarray*}
-    T_\text{em} = 0.0494 \pm 0.0009\\
-    T_\text{COM} = 0.051 \pm 0.108\\
-    T_\text{int} = 0.048 \pm 0.007\\
-\end{eqnarray*}
-The calculation of the temperature is done by using the following equations:
-\begin{eqnarray}
-    \label{eq:ekin}E_\text{Kin} &=& \frac{m}{2 M} \sum_i \vec{v}_i^{\, 2} \\
-    \label{eq:temp}E &=& \frac{f}{2} k_B T
-\end{eqnarray}
-In the first line, $m$ denotes the mass, which is $m=0.1$ for the gas particles, $m=1$ for the atoms in the nano particle and $m=32$ for the center
-of mass. $M$ denotes the number of particles that are being summed over, which is $M=1$ for the center of mass, $M=32$ for the internal temperature
-and fluctuates for the outgoing gas particles. The number of the outgoing particles is not constant and depends heavily on the pressure and the
-temperature of the impinging gas particles, as can be seen in \eqref{eq:numberofparticles}. In the second line, $f$ denotes the number of degrees of
-freedom of the system, which is $f=3$ for all three systems. Putting these two equations together, we get an equation for the temperature:
-\begin{equation}
-    T = \frac{m}{3 M} \sum_i \vec{v}_i^{\, 2} \\
-\end{equation}
-Since this describes an instantaneous value, the mean value of the velocities over the course of the simulation has to be used to get the mean value
-of the temperature of each subsystem. The simulation results compared with the mean values can be seen in
-Figs.~\ref{fig:outgoingequib},~\ref{fig:vcomsqdequib} and \ref{fig:internalequib}.
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/gastemp.pdf}
-        \caption{Instantaneous and mean temperature of the outgoing gas particles of the surrounding gas}
-        \label{fig:outgoingequib}
-    \end{center}
-\end{figure}
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/vcomsqd.pdf}
-        \caption{Instantaneous and mean temperature of the center of mass of the nano particle in the laser trap}
-        \label{fig:vcomsqdequib}
-    \end{center}
-\end{figure}
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/t_int.pdf}
-        \caption{Instantaneous and mean internal temperature of the nano particle in the laser trap}
-        \label{fig:internalequib}
-    \end{center}
-\end{figure}
-\\With the calculated mean temperature of the center of mass, we can take a look at the distribution of the components of the center of mass velocity.
-The expected distribution is a gaussian with $\sigma=\sqrt{\frac{k_B T}{m}}$, i.e.
-\begin{equation}
-    \label{eq:gaussianvel}
-    p(v_i) = \frac{m}{2\pi k_B T} e^{-\frac{x^2 m}{2 k_B T}}
-\end{equation}
-where $i$ denotes the chosen component of the center of mass velocity vector. A histogram of the x-component of the center of mass velocity  
-with a curve representing \eqref{eq:gaussianvel} can be seen in Fig.~\ref{fig:comvel}. Since the y and z axis yield very similar histograms, the
-x-component has been chosen to represent the velocity components. The graph in Fig.~\ref{fig:comvel} clearly shows, that the velocities indeed follow
-the gaussian distribution.
-\begin{figure}[h]
-    \begin{center}
-        \includegraphics[scale=0.4]{images/vcomtemp.pdf}
-        \caption{A histogram of the x-component of the center of mass velocities with the expected gaussian distribution}
-        \label{fig:comvel}
-    \end{center}
-\end{figure}
-\\The case of $\Delta Q=0$ is a good way to start and to check the integrity of the algorithms, but the more interesting cases are the ones where
-$\Delta Q > 0$. Besides $\Delta Q$, there are two values that determine the dynamics of the system: the pressure $P$ and the temperature of the
-impinging gas particles $T_\text{imp}$.\\
-Since the surrounding gas acts as a thermostat, the pressure of the impinging gas at a given gas temperature determines the maximum amount of heat that
-can be pumped into the system without particles of the nano particle leaving the simulation box -- the higher the pressure, the more energy can be
-pumped into the system. The algorithm is applied at two surrounding gas temperatures of $T_\text{imp} = 0.05$ and $T_\text{imp} = 0.1$, 
-which are one quarter and one half the temperature of the nano particle before applying the eEHX and the thermostat algorithm. The pressure varied
-from $P=0.1$ to $P=1.2$ in steps of $\Delta P=0.1$ and the heat is generally varied from $\Delta Q=0$ to $\Delta Q=0.5$ in steps of 0.05. The term
-''generally'' is used because there are cases where the stepsize has to be decreased or one value for $\Delta Q$ is skipped over. The stepsize has to
-be decresed for low pressures, such as $P=0.1$ or $P=0.2$ because there would otherwise not be enough data points. Some values of $\Delta Q$ need to
-be skipped because it can happen that atoms of the nano particle leave the system due to statistical fluctuations of velocities, which leads to
-unusable results.\\
-To see the comparison between the different temperatures of the impinging gas temperatures, three values for the pressure are chosen: $P=0.1$, $P=0.5$
-and $P=1.0$. The simulation results for the center of mass temperature, the internal temperature and the temperature of the outgoing gas particles are
-plotted for each of the pressure values and both temperatures, which can be seen in~\ref{fig:comintout}.
-\begin{figure}[h]
-    \begin{center}
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p01_com.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p05_com.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p1_com.pdf}
-        \end{subfigure} 
-        \\
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p01_int.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p05_int.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p1_int.pdf}
-        \end{subfigure} 
-        \\
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p01_out.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p05_out.pdf}
-        \end{subfigure} 
-        \
-        \begin{subfigure}[t]{0.3\textwidth}
-            \includegraphics[scale=0.13]{images/p1_out.pdf}
-        \end{subfigure} 
-        \caption{The temperature of the center of mass, the internal temperature and the outgoing temperature of the surrounding gas as functions of
-            heat being pumped into the system, for pressures
-        $P=0.1$, $P=0.5$ and $P=1$ at impinging gas temperatures of $T_\text{imp}=0.05$ (red) and $T_\text{imp}=0.1$. The three columns each show 
-        a fixed pressure value and the rows show fixed measured quantities. As the pressure relates to the maximum amount of heat that can be injected
-    into the system, the ranges on the abscissa in one row is different for every plot.}
-        \label{fig:comintout}
-    \end{center}
-\end{figure}
-The three columns of the plot correspond to fixed pressure values and different measured quantities, whereas every row corresponds to a fixed measured
-quantity at different pressures. As previously mentioned, the amount of pressure relates to the number of surrounding particles and thus limits the
-amount of heat that can be pumped into the system. Therefore, the range of values in the abscissa differs for different pressures -- at $P=0.1$ the
-values of $\Delta Q$ range from $0$ to $0.02$ and at $P=1.0$ from $0$ to $0.5$.\\
-The first thing that is noticable, is the increase of the temperature of the center of mass, the internal temperature and the outgoing gas temperature
-with increasing $\Delta Q$. This makes sense, as the amount of heat that is put into the system is also transfered in the interaction between the nano
-particle and the surrounding gas.\\
-One observation that is not so obvious due to the different ranges of $\Delta Q$ values: all temperature values increase faster for low pressures and
-increase slower for higher pressures. Take the center of mass temperature for $P=0.1$ and $P=1$ at $T=0.05$ (red line) as an example: while it reaches
-a value of a $T_\text{com} = 0.1$ at about $\Delta Q = 0.015$ at $P=0.1$, it takes a $\Delta Q$ of a little under $0.1$ to achieve the same value.
-This can be observed for every one of the three quantities. This is to be expected, as the surrounding gas acts as a thermostat. \\
-The development of the different temperatures seem to be depending of the temperature of the surrounding gas as well. The blue curves (which
-correspont do a surrounding gas temperature of $T=0.1$) are growing slightly faster than the red curves ($T=0.05$).\\
-In their paper~\cite{MillenJ.2014} Millen et al. give a formula for the center of mass temperature:
-\begin{equation}
-    T_\text{COM} = \frac{T_\text{imp}^{3/2}+\tfrac\pi 8 \ T_\text{em}^{3/2}}{T_\text{imp}^{1/2}+\tfrac\pi8 T_\text{em}^{1/2}}
-\end{equation}
-This formula together with the measured values of $T_\text{imp}$ and $T_\text{em}$ can be used to compare the results. The comparison between the
-calculated and the measured value can be seen for a representative choice of parameters in Fig.~\ref{fig:measurevscalc}.
-
-%\begin{figure}[h]
-    %\begin{center}
-        %\begin{subfigure}[t]{0.3\textwidth}
-            %\includegraphics[scale=0.3]{images/fcc-2.pdf}
-            %\caption{Schematic figure of a face centered cubic (FCC) lattice}
-            %\label{fig:fcc}
-        %\end{subfigure} 
-        %\
-        %\begin{subfigure}[t]{0.3\textwidth}
-            %\includegraphics[scale=0.15]{images/fcc_unit.png}
-            %\caption{Schematic of the number of particles contained in one FCC unit cell. The atoms on the edges are shared by 8 cells and the ones 
-                %on the faces by 2 cells\cite{unitcell}}
-            %\label{fig:fcccut}
-        %\end{subfigure} 
-        %\
-        %%\hspace{60pt}
-        %\begin{subfigure}[t]{0.3\textwidth}
-            %\includegraphics[scale=0.3]{images/unit_cell.pdf}
-            %\caption{Setup for successively building a FCC lattice}
-            %\label{fig:unitcell}
-        %\end{subfigure}
-    %\end{center}
-%\end{figure}
-
-\newpage
-\section{Conclusion and Further Remarks}
-% what does this all tell us? do we need to study more of this? 
-In this thesis I have modeled an optical tweezer experiment with methods from the field of computational physics. The experiment was broken down into
-different sections, that have been modeled separately. With the use of molecular dynamics I simulated the behaviour of a nano particle in a laser trap
-and investigated the effects of the laser power and temperature of the surrounding gas on the center of mass movement of the nano particle.\\
-The results show, that the center of mass temperature (and thus its movement) does indeed depend on the energy of the laser, as well as on the
-pressure and temperature of the surrounding gas.\\
-The development of the center of mass temperature shows a curve that seems to have a convex form, whereas the internal temperature shows a more
-concave behaviour. So while the internal temperature initially grows faster, it could be possible for the center of mass temperature to pass the
-internal temperature.\\
-There are several ways to explore this subject further. The first idea would be to decrease the pressure even more and investiate the case of very low
-energies, since this is not easily feasible in an experimental setup. Another idea would be to increase the number of atoms in the nano particle to
-see if the surface of the particle plays a role in the development of the temperatures. Furthermore, the interaction between the gas particles and the
-nano particle in the trap could be varied to include an attractive part in the potential. Generally, these simulations could be expanded by running
-them on faster computers to investigate long term behaviour of the systems and expand the ranges of the different values like pressure or laser
-energy.
-
-\newpage
-\bibliography{references}
-\bibliographystyle{unsrt}
+An erster Stelle möchte ich mich bei meinem Betreuer Prof. Christoph Dellago bedanken, der es mir nicht nur ermöglicht hat
+meine Masterarbeit zu schreiben, sondern mir auch die Möglichkeit gegeben hat ein Teil der Gruppe für Copmputergestützte Physik
+zu sein. Durch das wöchtentliche Gruppenseminar und den Kontakt mit anderen Mitgliedern der Gruppe war es mir möglich neben meiner
+Masterarbeit auch Einblick in die aktuelle Forschung zu bekommen.\\
+Das bringt mich zu den nächsten Personen bei denen ich mich bedanken möchte: die Mitglieder der Gruppe für Computergestützte Physik. Sie sind mir
+bei vielen Problemen und Fragen mit Rat und Tat zur Seite gestanden und waren ausgezeichnete Gesprächspartner in den vielen Diskussionen, die 
+geführt wurden.\\
+Weiters möchte ich mich bei all meinen Studienkollegen und Freunden bedanken, die mich auf meinem Weg begleitet haben. Allen voran
+Manuel Reisenbauer, Stefanie Belohlavy, Michael Partener und Veronika Baumann, die ich seit meinem ersten Studienjahr zu meinen besten Freunden 
+zählen darf und ohne deren Hilfe nie so weit gekommen wäre.\\
+Zu guter Letzt möchte ich meinen allergrößten Dank an meine Eltern und Familie aussprechen. Meine Eltern haben mich nicht nur Unterstützt und
+ermutigt, sondern mir auch den Raum gegeben, den ich brauchte, um mich zu entfalten. Ohne sie wäre das alles nicht möglich gewesen und ich 
+bin ihnen dafür auf Ewig dankbar.
 

### 2017-03-06 doc/final_ver1.tex
final version of the thesis
stuff

--- 2017-03-03 doc/draft4.tex
@@
 \usepackage{graphicx}
+\usepackage{pdfpages}
 
@@
 \begin{document}
-\title{Simulation of Nano Particles in a Laser Trap}
-\author{Mathias H\"old, BSc.}
-\date{2016}
-\maketitle
+\includepdf[pages={1}]{deckblatt_besser_neu.pdf}
+\newpage\thispagestyle{empty}~\newpage\thispagestyle{empty} 
+\section*{Danksagung}
+\thispagestyle{empty}
+An erster Stelle möchte ich mich bei meinem Betreuer Prof. Christoph Dellago bedanken, der es mir nicht nur ermöglicht hat
+meine Masterarbeit zu schreiben, sondern mir auch die Möglichkeit gegeben hat ein Teil der Gruppe für Copmputergestützte Physik
+zu sein. Durch das wöchtentliche Gruppenseminar und den Kontakt mit anderen Mitgliedern der Gruppe war es mir möglich neben meiner
+Masterarbeit auch Einblick in die aktuelle Forschung zu bekommen.\\
+Das bringt mich zu den nächsten Personen bei denen ich mich bedanken möchte: die Mitglieder der Gruppe für Computergestützte Physik. Sie sind mir
+bei vielen Problemen und Fragen mit Rat und Tat zur Seite gestanden und waren ausgezeichnete Gesprächspartner in den vielen Diskussionen, die 
+geführt wurden.\\
+Weiters möchte ich mich bei all meinen Studienkollegen und Freunden bedanken, die mich auf meinem Weg begleitet haben. Allen voran
+Manuel Reisenbauer, Stefanie Belohlavy, Michael Partener und Veronika Baumann, die ich seit meinem ersten Studienjahr zu meinen besten Freunden 
+zählen darf und ohne deren Hilfe ich nie so weit gekommen wäre.\\
+Zu guter Letzt möchte ich meinen allergrößten Dank an meine Eltern und Familie aussprechen. Meine Eltern haben mich nicht nur Unterstützt und
+ermutigt, sondern mir auch den Raum gegeben, den ich brauchte, um mich zu entfalten. Ohne sie wäre das alles nicht möglich gewesen und ich 
+bin ihnen dafür auf Ewig dankbar. 
+\newpage\thispagestyle{empty}~\newpage\thispagestyle{empty} 
+\newpage
+%\title{Simulation of Nano Particles in a Laser Trap}
+%\author{Mathias H\"old, BSc.}
+%\date{2016}
+%\maketitle
 \thispagestyle{empty}
 \newpage
+\setcounter{page}{1}
 \tableofcontents
@@
         contains a heat source (red), $\Gamma_1$, which is moving with velocity $v_{\Gamma_1}$ and a heat sink (blue), $\Gamma_2$, which is moving 
-        with velocity $v_{\Gamma_2}$. The regions that are neither heat sources or heat sinks are denoted by $\Gamma_0$.}
+        with velocity $v_{\Gamma_2}$. The regions that are neither heat sources or heat sinks are denoted by $\Gamma_0$. Image tagen from 
+    \cite{Wirnsberger2015}.}
         \label{fig:ehex} 
@@
 and $P=1.0$. The simulation results for the center of mass temperature, the internal temperature and the temperature of the outgoing gas particles are
-plotted for each of the pressure values and both temperatures, which can be seen in~\ref{fig:comintout}.
+plotted for each of the pressure values and both temperatures, which can be seen in Fig.~\ref{fig:comintout}.
 \begin{figure}[h]
@@
 \bibliographystyle{unsrt}
+\newpage
+\section*{Abstract}
+\thispagestyle{empty}
+Optical trapping is an experimental method, in which objects with a size ranging from a subatomic to micrometer  scale are 
+localized with a laser due to a gradient force. In this thesis, we use methods from the field of computational physics
+to simulate the behaviour of a nano particle in a laser trap and its interactions with the laser and the surrounding gas.\\
+Gieseler et al. performed such an optical tweezer experiment to investigate the fluctiation theorem. In their experiment
+they use a Langevin equation to model the dynamics of a silica nanosphere in a laser trap, where the temperature of the heat bath 
+is used to describe the stochastic force acting on the nano particle. Millen et al. conducted a similar experiment and raised
+the point, that there are four different temperatures present during this experiment: the temperature of the impinging gas particles, the 
+emerging gas particles, the center of mass and the surface of the nano particle. With these temperature regimes in mind, the experiment
+is simulated using molecular dynamics in order to determine the influence of the laser power and the temperature of the 
+surrounding gas on the center of mass temperature.\\
+To simulate the experiment, each part needs to be modeled appropriately. The nano sphere is represented by a set of atoms that interact
+with each other via a Lennard-Jones potential. The gradient force of the laser, which localizes the nano particle, is approximated by 
+a harmonic potential. Since the laser energy is absorbed by the nano particle, the internal temperature of the nano particle increses, which 
+is achieved by applying the eHEX algorithm to the nano particle. The surrounding gas acts as a barostat and a thermostat and will be represented
+by an algorithm that uses an ideal gas as pressure medium.\\
+The simulation shows that the center of mass temperature depends on the choice of laser power and temperature of the surrounding gas.
+\newpage
+\section*{Zusammenfassung}
+\thispagestyle{empty}
+Optical trapping bezeichnet eine experimentelle Methode, bei der Objekte mit einer Größe, die von einer subatomaren bis zur Mikrometer-Skala
+reicht, durch die Gradientenkraft eines Lasers lokalisiert werden. In dieser Arbeit benutzen wir Methoden aus dem Feld der Computational Physics, 
+um das Verhalten eines Nanoteilchens in so einer Laserfalle und seine Interaktionen mit dem Laser und dem umgebenden Gas zu simulieren.\\
+Gieseler et al. führten ein sogenanntes ,,Optical Tweezer''-Experiment durch, um das Fluktuationstheorem zu untersuchen. In diesem Experiment
+verwenden sie die Lagevin-Gleichung, um das Verhalten einer Nanokugel aus Siliciumdioxid in einer Laserfalle zu modellieren, wobei die 
+Temperatur des Umgebungsgases als Wärmereservoir in dem Modell verwendet wird, um die stochastische Kraft auf das Nanoteilchen zu beschreiben. 
+Millen et al. führen ein ähnliches Experiment durch und zeigten auf, dass zwischen vier Temperaturen unterschieden werden muss: die Temperatur 
+des eingehenden Gases, des ausgehenden Gases, der Schwerpunktsbewegung und der Oberfläche des Nanoteilchens. Mit Augenmerk auf diese verschiedenen
+Temperaturen wird das Experiment mit Molecular Dynamics Methoden simuliert, um den Einfluss der Stärke des Lasers und des Umgebungsgases auf 
+die Schwerpunktstemperatur des Nanoteilchens zu untersuchen.\\
+Um das Experiment zu simulieren, müssen die einzelnen Bestandteile modelliert werden. Das Nanoteilchen wird durch ein System aus Atomen 
+dargestellt, die über ein Lennard-Jones Potential miteinander wechselwirken. Die Gradientenkraft des Lasers, die das Nanoteilchen lokalisiert, 
+wird durch ein harmonisches Potential approximiert. Da die Energie des Lasers vom Nanoteilchen absorbiert wird, erhöht sich die interne 
+Temperatur des Nanoteilchens, was durch Anwendung des eHEX Algorithmus erreicht wird. Das umgebende Gas hat die Funktion eines Barostaten und die 
+eines Thermostaten und wird in der Simulation durch einen Algorithmus repräsentiert, der ein ideales Gas als Druckmedium verwendet.\\
+Die Simulation zeigt, dass die Schwerpunktstemperatur von der Lasertemperatur und der Temperatur des einströmenden Gases abhängt. 
+\newpage\thispagestyle{empty}~\newpage\thispagestyle{empty}
 

### 2017-03-06 doc/presentation_ver1.tex
added pres

--- 2017-01-27 doc/expose.tex
@@
-\documentclass[12pt]{article}
+\documentclass{beamer}
 \usepackage{geometry}
@@
 \begin{document}
-\title{Expos\'e - Simulation of Nano Particles in a Laser Trap}
-\author{Mathias H\"old, BSc.}
-\maketitle
-\thispagestyle{empty}
-\section{Konzept}
-In dieser Masterarbeit wird mit Hilfe von Methoden der Computational Physics ein Experiment simuliert, in dem Nanopartikel in einer Laserfalle
-gefangen werden. Das Hauptaugenmerk dabei liegt auf der Interaktion des Nanopartikels mit dem umgebenden Gas und dem Einfluss des Lasers auf die
-Bewegung des Nanopartikels in der Falle. Die benutzte Technik in dieser Simulation nennt sich ,,molecular dynamics'' 
-und wird dazu verwendet, um die zeitliche Entwicklung des Zustandes des Nanopartikels und des Gases zu berechnen.\\
-Das Experiment besteht aus einem Nanopartikel, in diesem Fall eine Siliziumdioxid-Kugel mit einem Durchmesser von $r \approx 75 nm$, das in von einem
-Laser durch Gradientenkraft lokalisiert wird. Das Nanopartikel befindet sich in einer Vakuumkammer mit variablem Druck.\\
-Die Gasteilchen, die das Nanopartikel umgeben, interagieren nicht miteinander, wenn die mittlere freie Wegl\"ange der Gasteilchen viel gr\"o\ss er ist
-als der Radius des Nanopartikels. Das bedeutet, dass die Interaktion nur zwischen den Gas Teilchen und dem Nanopartikel in der Falle stattfindet. 
-Dadurch entstehen zwei W\"armereservoirs des umgebenden -- ein ,,kaltes'', bestehend aus Teilchen vor der Interaktion mit dem Nanopartikel und ein
-,,hei\ss es'', bestehend aus Teilchen nach der Interaktion mit dem Nanopartikel.\\
-Dem Nanopartikel selbst sind auch zwei Temperaturen zuzuweisen. Da es sich in einer Laserfalle befindet, absorbiert das Nanopartikel die Energie des
-Lasers in Form von W\"arme, die an die innere Energie der Atome, aus dem das Nanopartikel besteht, abgegeben wird. Die Temperatur, die aus dieser
-inneren Energie abzuleiten ist, wird ,,Oberfl\"achentemperatur'' genannt. Die zweite Temperatur, die dem Nanopartikel zuzuordnen ist, ist seine
-Schwerpunktstemperatur -- jene Temperatur, die aus der Geschwindigkeit der Schwerpunktsbewegung berechnet wird. Durch die Lokalisierung in der
-Laserfalle ist wird diese Bewegung minimiert. \\
-Die Simulation dieses Experiments wird durch Modellierung der Einzelnen Bestandteile durchgef\"uhrt. Das Nanopartikel wird durch einen FCC Kristall
-approximiert, in dem die Teilchen miteinander \"uber ein Lennard-Jones Potential miteinander wechselwirken. Der Laser wird durch zwei unterschiedliche
-Teile modelliert. Zum einen wird die durch den Laser abgegebene und das Nanopartikel aufgenommene Energie durch den so genannten ,,eHEX'' -- enhanced
-heat exchange algorithm -- modelliert. Die Lokalisierung des Nanopartikels durch den Laser wird durch ein harmonisches Potential modelliert, das auf
-den Schwerpunkt wirkt. Das umgebende Gas wird durch einen Barostat modelliert, der in diesem Fall eher eine Funktion als Thermostat einnimmt. Die
-Bewegung der einzelnen Atome wird durch den velocity Verlet Algorithmus berechnet.\\
-Die Fragestellung, der in dieser Arbeit nachgegangen wird ist: Hat die erh\"ohte Temperatur der inneren Freiheitsgrade des Nanopartikels einen
-Einfluss auf dessen Schwerpunktsbewegung?
-
-\section{Vorl\"aufiges Inhaltsverzeichnis}
-Die Arbeit wird in folgende Kapitel unterteilt sein:
-\begin{itemize}
-    \item{Introduction}
-    \item{Motivation}
-    \item{Simulation}
-    \item{Results}
-    \item{Conclusion}
-\end{itemize}
-
-\section{Auswahlbibliographie}
-\nocite{*} 
-\bibliography{references}
-\bibliographystyle{unsrt}
-
+\begin{frame}
+    test!
+\end{frame}
 \end{document}

### 2017-03-21 doc/presentation/pres_long_ver1.tex
added presentation

\documentclass[12pt]{article}
\usepackage{geometry}
\geometry{a4paper}
\usepackage[utf8]{inputenc}
\usepackage{amsmath,amssymb}
\usepackage{hyperref} % verlinkungen
\usepackage{textcomp}
\usepackage{flafter}
\usepackage{booktabs}
\usepackage{array}
\usepackage{paralist}
\usepackage{dsfont} \usepackage{color}
\usepackage{bbold}
\usepackage{float}
\usepackage[font=scriptsize,labelfont=bf]{caption}
%\usepackage[font=footnotesize,labelfont=bf]{caption}
\usepackage[font=scriptsize]{subcaption}
\usepackage{fancyhdr}
\setlength{\headheight}{15.2pt}
\pagestyle{fancy}
\usepackage{listings} 
\usepackage{pict2e}
\usepackage{xfrac}
\usepackage[english]{babel}
\usepackage{mathtools}
\usepackage{graphicx}
\usepackage{pdfpages}

%%%%%%%%         EIGENEBEFEHLE  %%%%%%%%%
\newcommand{\ddt}{\frac{\partial}{\partial{t}}}
\newcommand{\dnach}[1]{\frac{\partial}{\partial{#1}}}
\newcommand{\ddnach}[2]{\frac{\partial{#1}}{\partial{#2}}}
\newcommand{\dddnach}[3]{\frac{\partial^2{#1}}{\partial{#2} \partial{#3}}}
\newcommand{\dznach}[2]{\frac{\partial^2{#1}}{\partial{#2}^2}}
\newcommand{\vnabla}{\mathbf{\nabla}}
\newcommand{\emathbf}[1]{\mathbf{\hat{#1}}}
\newcommand{\bra}[1]{\langle{#1}|}
\newcommand{\ket}[1]{|{#1}\rangle}
\newcommand{\bracket}[2]{\langle{#1}|{#2}\rangle}
\newcommand{\up}{\uparrow}
\newcommand{\down}{\downarrow}
\newcommand{\updown}{\uparrow\downarrow}
\newcommand{\downup}{\downarrow\uparrow}
\newcommand{\upup}{\uparrow\uparrow}
\newcommand{\downdown}{\downarrow\downarrow}
\newcommand{\sandwich}[2]{\bra{#1} #2 \ket{#1}}
\newcommand{\fsqrt}[2]{\sqrt{\frac{#1}{#2}}}
\newcommand{\GammaK}{{\Gamma_k}}

\definecolor{mygreen}{rgb}{0,0.6,0}
\definecolor{mygray}{rgb}{0.5,0.5,0.5}
\definecolor{mymauve}{rgb}{0.58,0,0.82}

\lstset{%
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  basicstyle=\footnotesize,        % the size of the fonts that are used for the code
  breakatwhitespace=false,         % sets if automatic breaks should only happen at whitespace
  breaklines=true,                 % sets automatic line breaking
  captionpos=b,                    % sets the caption-position to bottom
  commentstyle=\color{mygreen},    % comment style
  %deletekeywords={...},            % if you want to delete keywords from the given language
  %escapeinside={\%*}{*)},          % if you want to add LaTeX within your code
  extendedchars=true,              % lets you use non-ASCII characters; for 8-bits encodings only, does not work with UTF-8
  frame=single,                    % adds a frame around the code
  keepspaces=true,                 % keeps spaces in text, useful for keeping indentation of code (possibly needs columns=flexible)
  keywordstyle=\color{blue},       % keyword style
  language=C,                 % the language of the code
  %morekeywords={*,...},            % if you want to add more keywords to the set
  numbers=left,                    % where to put the line-numbers; possible values are (none, left, right)
  numbersep=5pt,                   % how far the line-numbers are from the code
  numberstyle=\tiny\color{mygray}, % the style that is used for the line-numbers
  rulecolor=\color{black},         % if not set, the frame-color may be changed on line-breaks within not-black text (e.g. comments (green here))
  showspaces=false,                % show spaces everywhere adding particular underscores; it overrides 'showstringspaces'
  showstringspaces=false,          % underline spaces within strings only
  showtabs=false,                  % show tabs within strings adding particular underscores
  stepnumber=2,                    % the step between two line-numbers. If it's 1, each line will be numbered
  stringstyle=\color{mymauve},     % string literal style
  tabsize=2,                       % sets default tabsize to 2 spaces
  %title=\lstname                   % show the filename of files included with \lstinputlisting; also try caption instead of title
}

\lhead[\ ]{\leftmark }
\chead[\ ]{\ }
\rhead[\ ]{\rightmark }

\lfoot[\ ]{\ }
\cfoot[\ ]{\thepage }
\rfoot[\ ]{\ }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{document}

\section{Einleitung}
In meiner Masterarbeit beschäftigte ich mich mit der Simulation eines Nano Teilchens in einer Laserfalle. Diese Arbeit verbindet damit 
zwei Teilbereiche der Physik miteinander: Computational Physics und Optische Fallen, die ich beide kurz vorstellen möchte.\\
Die Computational Physics (oder auch Computergestützte Physik) verwendet numerische Methoden, um physikalische Prozesse auf dem Computer
zu simulieren. Ein Meilenstein, der genannt werden sollte, ist die Arbeit von Metropolis et al., die im Jahre 1953 mit Hilfe des Computers
MANIAC (Mathematical Analyzer Numerical Integrator And Computer) in Los Alamos die Zustandsgleichung eines Fluids berechneten. Die dort 
benutzte Methode, der Metropolis Monte Carlo Algorithmus, ist bis heute ein sehr wichtiger Bestandteil der Compuational Physics. Seit
1953 hat sich nicht nur die Rechenleistung von Computern enorm verbessert, sondern es wurden auch Zahlreiche neue Methoden zur Simulation 
von physikalischen Prozessen entwickelt, wie zum Beispiel Transition Path Sampling, Finite-Elemente-Methode und Molekulardynamik-Simulation. 
Letztere wurde in dieser Arbeit verwendet und wird später etwas detaillierter beschrieben.\\
Laserfallen, auch Optische Pinzette (englisch \textit{optical tweezers} genannt) ist eine experimentelle Methode, bei der Objekte durch die 
Gradientenkraft eines fokussierten Laserstrahls lokalisiert werden. Die Größe der Objekte reicht von der subatomaren (Kühlung einzelner Atome) bis zur
Mikrometer-Skala (Zellen). Die Idee für diese Methode entstand schon zu Beginn des 20. Jahrhunderts, als Lebedev und Nichols und Hull die 
Existenz von Strahlungsdruck experimentell nachweisen konnten. Da die Technik zu der Zeit noch nicht so ausgereift war, wurde die Idee erst 
1970 wieder durch Ashkin aufgegriffen. Dieser verwendete Laser, um die Bewegung von Atomen und Mikrometergroßen Objekten zu beeinflussen.\\

\section{Motivation}
Die Motivation für diese Arbeit ist ein Optical Tweezer Experiment, das von Jan Gieseler, Romain Quidant, Christoph Dellago und Lukas Novotny
durchgeführt wurde. Ziel des Experiments war der Nachweis des Fluktuationstheorems.\\
Im Experiment befindet sich eine Nanokugel mit einem Durchmesser von etwa 75 Nanometer und einer Masse von $3 \times 10^{-18}$ kg in einer
Laserfalle in einer Vakuumkammer. Die Kugel wird durch die Gradientenkraft des Lasers lokalisiert. In der Laserfalle fluktuiert die Position 
der Kugel in alle 3 Raumrichtungen. Die Fluktuation in jede Richtung ist von den anderen Richtungen entkoppelt und kann daher in einer 
eindimensionalen Bewegungsgleichung, der Langevin-Gleichung, beschrieben werden:
\begin{equation}
    \label{eq:langevin}
    \ddot{x} + \Gamma_0 \dot{x} + \Omega^2_0x = \frac 1 m \left(F_\text{fluct} + F_\text{ext}\right)
\end{equation}
Auf der linken Seite: 
$x$ bezeichnet den Ort des Teilchens, $\dot{x}$ und $\ddot{x}$ die Geschwindigkeit und Beschleunigung, $\Gamma_0$ ist der Reibungskoeffizient
und $\Omega_0$ ist die Winkelfrequenz, die die Fluktuation entlang der gewählten Raumrichtung beschreibt. Auf der rechten Seite befinden sich 
zwei Kräfte. Die erste, $F_\text{fluct}$ beschreibt die stochastische Kraft, die durch die Stöße mit dem umgebenden Gas verursacht wird. 
Diese Kraft ist gegeben durch:
\begin{equation}
    \label{eq:ffluct}
    F_\text{fluct} = \sqrt{2m\Gamma_0k_BT_0} \ \xi\left(t\right)
\end{equation}
Hier ist $T_0$ die Temperatur des Wärmereservoirs, also des umgebenden gases, $k_B$ ist die Boltzmann-Konstante und $\xi(t)$ ist weißes 
Rauschen für das gilt $\left\langle\xi(t)\right\rangle= 0$ und $\left\langle\xi(t)\xi(t')\right\rangle= \delta(t-t')$, d.h. es handelt sich um 
eine randomisierte Kraft. Der Term $\Gamma_0$ taucht in der Formel wegen des Fluktuations-Dissipations-Theorems auf, das die Dämpfungsrate
und die stochastische Kraft verbindet. \\
Die andere Kraft, $F_\text{ext}$ ist eine externe, zeitabhängige Kraft, die das Teilchen in einen stationären Nichtgleichgewichtszustand zu bringen. 
Die Kraft wird zum Zeitpunkt $t = t_\text{off}$ deaktiviert und die Relaxation des Teilchens wird gemessen. Diese Messung wird 
durchgeführt, um das Fluktuationstheorem zu untersuchen, das geschrieben werden kann als:
\begin{equation}
    \label{eq:fluctuationtheorem}
    \frac{p(-\Delta S)}{\Delta S} = e^{-\Delta S}
\end{equation}
wobei $\Delta S$ die relative Entropieänderung ist, gegeben durch:
\begin{equation}
    \Delta S = \beta_0 Q + \Delta \phi
\end{equation}
$Q$ bezeichnet die vom Wärmereservoir absorbierte Wärme, $\beta$ ist die reziproke Temperatur bei der die Wärme absorbiert wird
und $\Delta \phi$ ist die Differenz der trajektoriebasierten Entropie. Das Fluktuationstheorem beschreibt die 
Wahrscheinlichkeit für Verletzung des 2. Hauptsatzes für finite (kleine) Zeitintervalle. Das ist ein statistischer Effekt und steht nicht
im Widerspruch zum 2. Hauptsatz.\\
Die Untersuchung des Fluktuationstheorems wird bei Relaxationsprozessen von zwei unterschiedlichen stationären Nichtgleichgewichtszuständen
durchgeführt. Der erste wird durch einen Feedback-Mechanismus erzeugt und der zweite durch eine Kombination des Feedback-Mechanismus und einem
zusätzlichen Feedback-Drives. Die Ergebnisse zeigen, dass beide Zustände beim Relaxationsprozess dem Fluktuationstheorem genügen, wodurch 
das Theorem für diesen Fall experimentell bestätigt wird.\\
J. Millen, T. Deesuwan, P. Barker und J. Anders führten ein ähnliches Experiment durch (ohne den Feedback Mechanismus), wo sie die Temperatur
des umgebenden Gases untersuchten.\\
Während in dem Modell von Gieseler et al. nur eine Temperatur $T_0$ verwendet wird, stellen Millen et al. ein Modell mit vier unterschiedlichen
Temperaturen auf: die Temperatur des einströmenden und ausströmenden Gases, der Oberfläche und des Schwerpunkts des Teilchens.\\
Das Nanoteilchen in der Falle absorbiert Wärme vom Laser, was zu einer erhöhten Oberflächentemperatur führt, die im Allgmeinen höher ist 
als die Temperatur des einströmenden Gases. Die eiströmenden Gasteilchen wechselwirken mit dem Teilchen in der Falle, nehmen dabei Energie 
auf und verlassen das Teilchen mit einer höheren Temperatur. Im Modell wird angenommen, dass die Gasteilchen nur mit dem Nanopartikel in der
Falle und nicht untereinander interagieren, was zu der Entstehung von zwei Wärmereservoirs mit unterschiedlicher Temperatur führt.\\
Diese Unterschiedlichen Temperaturen führen zu einer Modifikation der eindimensionalen Langevin-Gleichung:
\begin{equation}
    M\ddot{x}(t) + M\left(\Gamma^\text{imp}+\Gamma^\text{em}\right)\dot{x}(t) + M\omega^2x(t) =F^\text{imp}+F^\text{em}
\end{equation}
x, $\dot{x}$ und $\ddot{x}$ sind Position, Geschwindigkeit und Beschleunigung des Nanoteilchens, M ist die Masse des Nanopartikels, 
$\omega$ ist die Frequenz der Falle, $\Gamma^\text{Imp}$ und $\Gamma^\text{Em}$ sind die Dämpfungsterme für das ein- und ausgehende Gas
und $F^\text{Imp}$ und $F^\text{Em}$ sind die zugehörigen Noise-Terme.\\
Im Experiment zeigen Millen et al., dass die Laserintensität und Teilchengröße einen Einfluss auf die Temperatur der ausgehenden Gastteilchen hat.\\
Man sieht also, dass im Modell für die Bewegung des Teilchens in der Falle mehr Temperaturen als nur die Temperatur des umgebenden Gases
berücksichtigt werden sollen. In dieser Masterarbeit beschäftigte ich mich daher mit der Frage: Welchen Einfluss haben Laserintensität
und die Temperatur des eingehenden Gases auf die Bewegung des Nanoteilchens in der Falle?

\section{Simulation}
Um das Experiment zu simulieren, müssen die einzelnen Bestandteile des Experiments modelliert werden. Das Nanoteilchen in der Falle
wird durch ein System aus Atomen in einem FCC Gitter approximiert, die Gradientenkraft des Lasers durch ein harmonisches Potential, 
die Absorption der Laserenergie im Teilchen durch den eHEX Algorithmus und das umgebende Gas durch einen Thermostaten mit einem idealen Gas
als Druckmedium. Die Dynamik des Systems wird mit Hilfe der Methode der Molekulardynamik-Simulation berechnet.\\
\subsection{Molekulardynamik - Die Idee}
Molekulardynamik Simulationen werden verwendet um die Dynamik klassischer atomarer oder molekularer Vielteilchensysteme zu berechnen. 
Klassisch
in diesem Sinne bedeutet, dass die Trajektorien mit Hilfe von Modellen der klassischen Physik und nicht der Quantenmechanik 
berechnet werden. Für große Systeme liefert das sehr gute Ergebnisse. Für kleine Systeme wie Wasserstoff oder Helium sind quantenmechanische
Effekte nicht mehr vernachlässigbar und daher sollten andere Methoden dafür verwendet werden.

\subsection{Velocity-Verlet Algorithmus}
Mikroskopisch gesehen beschreitet das System eine Trajektorie im Phasenraum, 
wo jeder Punkt einem Set von Positionen und Impulsen entspricht. Die Verbindung zwischen
zwei Punkten entspricht der Entwicklung von einem Zustand zum anderen. Diese Entwicklung gilt es zu simulieren. Dazu wird 
die Methode der finiten Differenzen gewählt. Die Trajektorie im Phasenraum wird in Abschnitte mit Länge $\Delta t$ geteilt und die 
Bewegungsgleichungen für jedes segment werden berechnet.\\
Der Algorithmus, der hier verwendet wird und eine sehr zuverlässige Wahl ist, ist der Velocity Verlet Algorithmus. Er basiert auf der 
Taylorreihenentwicklung der Koordinaten und Geschwindigkeiten. Der Algorithmus kann durch 2 simple Formeln dargestellt werden:
\begin{equation}
    \label{eq:velocityverlet}
    \begin{aligned}
        \mathbf{r}_i(t+\Delta t) = \mathbf{r}_i(t) + {\mathbf{v}}_i(t) \Delta t + \frac1{2m} {\mathbf{F}}_i(t) \Delta t^2\\
        \mathbf{v}_i(t+\Delta t) = \mathbf{v}_i(t) + \frac1{2m} \Big[\mathbf{F}_i(t) + \mathbf{F}_i(t+\Delta t)\Big] \Delta t
    \end{aligned}
\end{equation}
Wo $r_i$ die Position des Teilchens $i$ ist, $v_i$ ist seine Geschwindigkeit und $F_i$ ist die auf das Teilchen wirkende Kraft, 
$\Delta t$ ist der benützte Zeitschritt. 
Umgesetzt wird der Algorithmus durch folgende Schritte
\begin{enumerate}
    \item Berechne die Kräfte zwischen den Teilchen (nur erster Schritt)
    \item Berechne neue Positionen
    \item Berechne neue Geschwindigkeiten 
    \item Berechne Kräfte mit neuen Positionen 
    \item Berechne neue Geschwindigkeit mit neuen Kräften -> benütze diese für Schritt 2 und setze dort fort
\end{enumerate}

\subsection{Das Nanoteilchen}
Um das Nanoteilchen in der Falle zu modellieren wird ein System aus N Teilchen verwendet, die durch das Lennard-Jones Potential 
wechselwirken:
\begin{equation}
    \label{eq:lj}
    U(r) = 4\varepsilon\left[\left(\frac\sigma r\right)^{12} - \left(\frac\sigma r\right)^6\right]
\end{equation}
hier ist $\varepsilon$ die Potentialtiefe, $\sigma$ die Distanz bei der das Potential gleich Null ist und r ist der intermolekulare Abstand.\\
Um Rechnungen zu vereinfachen werden so genannte reduzierte Einheiten verwendet: $\sigma$ als Einheit der Länge, $\varepsilon$ als Einheit 
der Energie und die molekulare Masse $m$ als Masse. Üblicherweise werden reduzierte Variablen mit einem Stern versehen. Da aber hier reduzierte 
Einheiten ständig verwendet werden, wird der Stern weggelassen.
%Daraus können andere Einheiten abgeleitet werden. Zum Beispiel
%\begin{itemize}
    %\item {distance:} $r^* = r/\sigma$
    %\item {potential energy:} $U^* = U/\varepsilon$
    %\item {temperature:} $T^* = k_B T/\varepsilon$
    %\item {time:} $t^* = t\sqrt{\varepsilon/(m\sigma^2)}$
    %\item {pressure:} $P^* = P\sigma^3/\varepsilon$
    %\item {density:} $\rho^* = \rho \sigma^3$
%\end{itemize}
Das Lennard Jones Potential kann damit geschrieben werden als:
\begin{equation}
    U(r^*) = 4\left[{r}^{-12} - {r}^{-6}\right].
\end{equation}
Die zugehörige Kraft, die für den Velocity Verlet Algorithmus gebraucht wird, wird durch den negativen Gradienten des
Potentials berechnet. Als Beispiel die Kraft in x-Richtung:
\begin{eqnarray}
    F_{x} &=& -\frac{\partial}{\partial x} U(r) \nonumber\\
                &=& -\frac{\partial}{\partial x} 4\left[{r}^{-12} - {r}^{-6}\right] \nonumber\\
    \label{eq:ljforce} &=& 48 \left[r^{-14} - 0.5 \ r^{-8}\right] x
\end{eqnarray}
(y,z analog).\\
Ein wichtiger Schritt in der Simulation ist die Initialisierung der Teilchen. Für diese Arbeit wurde als 
Startkonfiguration ein kubisch flächenzentriertes Gitter (FCC, face centered cubic) gewählt. 
Die Anzahl der Teilchen in einer 
Elementarzelle eines FCC Gitters ist 4 (1/8 an jeder Ecke und 1/2 auf jeder der 6 Seitenflächen). Um einen Würfel aus solchen
Einheitszelle zu erzeugen, werden $M^3$ solcher Zellen benötigt und damit $4M^3$ Teilchen. Die Bedingung 
$N = 4M^4 = 4,32,108,256,\ldots$ führt zu den so genannten magic numbers. Um Zeit und Rechenaufwand zu sparen wurden in dieser
Arbeit $N=32$ Teilchen verwendet.\\
Die Erzeugung des Gitters wird mit Hilfe von 4 eindeutigen Punkten durchgeführt:
\begin{eqnarray*}
    p_1 &=& \{0,0,0\}\\
    p_2 &=& \{0.5 \ a,0.5 \ a,0\}\\
    p_3 &=& \{0.5 \ a,0,0.5 \ a\}\\
    p_4 &=& \{0,0.5 \ a,0.5 \ a\}
\end{eqnarray*}
wobei $a$ die Gitterkonstante des FCC Gitters ist, gegeben durch 
From the particle number $N$ and the number of FCC unit cells per edge $M$ the \textit{lattice constant} $a$ can be calculated
\begin{equation}
    a = \frac{L}{M}
\end{equation}
wobei $M$ die Anzahl der FCC Einheitszellen entlang einer Kante ist und $L$ die Seitenlänge des gesamten Systems, gegeben durch 
\begin{equation}
    L = \sqrt[3]{\frac{N}{\rho}}
\end{equation}
$\rho$ ist die Dichte des Systems.

\subsection{Laserstrahl - Wärmequelle}
Das Teilchen in der Falle wird durch den Laserstrahl gefangen. Während die Schwerpunktsbewegung dabei lokalisert ist, absorbieren
die einzelnen Atome des Nanoteilchens Wärme vom Laser und bekommen damit eine hohe interne Temperatur.\\
Um dieses Verhalten zu simulieren wird anstatt eines Thermostaten ein Algorithmus verwendet, der Wärmeaustausch als Grundlage hat.
Dieser Algorithmus heißt eHEX.\\
Der grundsätzliche Aufbau sieht folgendermaßen aus: Im Gesamtsystem werden Regionen festgelegt, die entweder als Wärmequelle oder 
Wärmesenken dienen. Diese Regionen werden mit $\Gamma_k$ bezeichnet und haben korrespondierende Mengen von ausgetauschter Wärme $\Delta Q_k$. 
Negatives $\Delta Q_k$ bedeutet, dass die Region Wärme von der Umgebung aufnimmt und positives $\Delta Q_k$ bedeutet, dass die Region 
Wärme an die Umgebung abgibt. Die Teilchen in der Simulationsbox $\Gamma$ und in den ,,Wärme''-Regionen besitzen 
Schwerpunktsgeschwindigkeiten $v_\Omega$ und $v_{\Gamma_k}$. Die Änderung der Energie in den Regionen wird vorgenommen durch das Reskalieren 
der Geschwindigkeiten mit einem Faktor $\xi_k$ und einem Shift:
\begin{equation}
    \mathbf{v}_i \rightarrow \mathbf{\bar{v}}_i = \xi_k \mathbf{v}_i + (1-\xi_k)\mathbf{v}_{\Gamma_k}
\end{equation}
\begin{equation}
    \xi_k = \sqrt{1+\frac{\Delta Q_{\Gamma_k}}{\mathcal{K}_{\Gamma_k}}}
\end{equation}
wobei $\Delta Q_{\Gamma_k}$ die ausgetauschte Wärme in der Region ist und $\mathcal{K}_{\Gamma_k}$ die kinetische (Ruhe)-Energie der Region:
\begin{equation}
    \label{eq:kineticK}
    \mathcal{K}_{\Gamma_k} = \sum_{i \in \gamma_k} \frac{m_i v_i^2}{2} - \frac{m_{\Gamma_k}v_{\Gamma_k}^2}{2}
\end{equation}
Um den Algorithmus durchzuführen werden noch einige Variablen benötigt, auf die ich hier nicht im Detail eingehen möchte.\\
Die Situation für die Masterarbeit ist etwas einfacher: es gibt nur eine Region, die als Wärmequelle agiert und die
Schwerpunktsgeschwindigkeit gleich Null.\\
Der Algorithmus basiert auf einem Velocity-Verlet Algorithmus mit einem Halbschritt in der Geschwindigkeit, ist jedoch etwas komplizierter. 
Deswegen werde
ich nicht alle Formeln aufschreiben, sondern die Einzelnen Schritte beschreiben:
\begin{enumerate}
\item Berechnung aller intermolekularen Kräfte (Initialisierung)
\item Reskalierung der Geschwindigkeiten -> $\bar{v}^n$
\item Halbschritt der skalierten Geschwindigkeit durch addieren der Kraft -> $\bar{v}^{n+1/2}$
\item Berechnung der skalierten Positionen mit dem Halbschritt der Geschwindigkeiten -> $\bar{r}^{n+1}$
\item Berechnung der Kräfte basierend auf den neuen Positionen
\item Halbschritt der skalierten Geschwindigkeit mit vorigem Halbschritt und neuen Kräften -> $\bar{v}^{n+1}$
\item Berechnung der Geschwindigkeit durch erneute skalierung mit aktualisiertem $\xi$
\item Berechnung der neuen Positionen und Korrektur
\end{enumerate}
Der Korrekturterm im letzten Schritt ist das Herzstück des eHEX Algorithmus. Er sorgt dafür, dass der Algorithmus keinen 
Drift in der Energie im Langzeitverhalten aufweist.

\subsection{Laserstrahl - Lokalisierung}
Die Lokalisierung des Teilchens wird durch ein harmonisches Potential simuliert. Kraft und Potential können geschrieben werden als
\begin{eqnarray}
    \mathbf{F} &=& -k \Big[\mathbf{r}_\text{COM}-\mathbf{x}_0\Big]\\
    U &=& \frac12 k \Big[\mathbf{r}_\text{COM}-\mathbf{x}_0\Big]^2
\end{eqnarray}
wobei $r_\text{COM}$ die Position des Schwerpunkts, $x_0$ die Position des Minimums des Potentials und k die Federkonstante ist. Die 
Schwerpunktsposition wird berechnet durch  
\begin{equation}
    \mathbf{r}_\text{COM} = \frac1N\sum_{i=1}^N \mathbf{r}_i
\end{equation}
Um das Potential in die Bewegungsgleichung einfließen zu lassen wird die Kraft auf den Schwerpunkt berechnet und zu den intermolekularen 
Kräften hinzugefügt.

\subsection{Umgebendes Gas - Thermostat}
Im Experiment ist Teilchen von einem Gas in einer Vakuumkammer umgeben. Dieses Gas wird durch einen Thermostaten/Barostaten 
simuliert, der ein ideales Gas als Druckmedium verwendet.\\
Die Gasteilchen strömen dabei in die Simulationsbox, interagieren mit dem Teilchen in der Falle und verlassen das System wieder (mit einer höheren
Temperatur). Die Gasteilchen interagieren also nur mit dem Teilchen in der Falle, nicht untereinander. Dadurch ist dieses Modell der 
Situation im Experiment sehr ähnlich.\\
Dazu wird das simulierte Teilchen mit einem Volumen umgeben, das auf der Geometrie des Systems basiert. Der Algorithmus ist so 
gestaltet, dass sich dieses Volumen im Laufe der Simulation der Geometrie des Systems anpasst. Ich habe mich entschlossen, eine simple, sich nicht
ändernde Geometrie zu wählen: einen Würfel mit Seitenlänge L, der eine Seitenlänge des LJ-Systems vom Teilchen in der Falle entfernt ist. Damit gilt
$L = 3s$. Das verwendete Potential für die Interaktion zwischen Gas und Nanoteilchen ist ein Soft-Sphere Potential
\begin{equation}
    \label{eq:softsphere}
    U(r) = \varepsilon \left(\frac{\sigma}{r}\right)^{12}
\end{equation}
wo $\varepsilon$ die Interaktionsstärke ist, $\sigma$ die Interaktionslänge und $r$ die Distanz zwischen Gasteilchen und Atom im Nanoteilchen.\\
Der Algorithmus kann für die vorliegende Situation folgendermaßen zusammengefasst werden:
\begin{enumerate}
\item Erzeuge Zufallszahl an Teilchen, die an einer Seite des umgebenden Volumens einströmen soll aus einer Poisson-Verteilung mit 
Erwartungswert 
        \begin{equation}
            %\langle N_\text{fac}\rangle = \Delta t L^2 P \left(\frac{1}{2\pi m k_B T}\right)^\frac12
            \label{eq:numberofparticles}
            \langle N_\text{fac}\rangle = \Delta t L^2 P \sqrt{\frac{1}{2\pi m k_B T}}
        \end{equation}
wobei $\Delta t$ der Zeitschritt, $L$ Seitenlänge des Umgebenden Volumens, $P$ eingestellter Druck, $m$ Masse, $k_B$ Boltzmann Konstante und 
$T$ die Temperatur.
Die Positionen der Teilchen werden über entsprechende Fläche gleichmäßig veerteilt. Die Geschwindigkeiten werden aus zwei unterschiedlichen 
Verteilungen gewürfelt. Die Komponente, die normal auf die gewählte Fläche des Volumens ist, wird aus einer Rayleigh Verteilung gewählt
        \begin{equation}
            p(v_i) = \frac{m}{k_B T}v_i \ e^{-\frac{mv_i^2}{2k_BT}}
        \end{equation}
Die restlichen Komponenten werden aus einer Maxwell-Boltzmann Verteilung gezogen.
\item Führe den ersten Velocity Verlet Schritt aus, um neue Positionen zu berechnen
\item Wenn Gasteilchen das umgebende Volumen verlassen habe, entferne sie
\item Berechne die Kräfte
\item Führe den zweiten Schritt des Velocity Verlet Algorithmus aus
\end{enumerate}

\section{Ergebnisse}
Da ich alle Modelle selbst umgesetzt habe, sollten alle Bestandteile auf Korrektheit untersucht werden.
\subsection{Nanoteilchen}
Um zu verfizieren, dass es sich um einen FCC Kristall handelt wird die Darstellung so gewählt, dass man das für den FCC Kristall typische
ABC-Muster der close packed structure sieht.

\subsection{Velocity Verlet}
Simulationen mit der Molekulardynamik Methode finden im mikrokanonischen Ensemble statt, d.h. Volumen, Teilchenzahl und Gesamtenergie sind konstant.
Die Geschwindigkeiten der Teilchen werden bei der Initialisierung aus einer Maxwell-Boltzmann Verteilung gewürfelt. Danach müssen zwei Anpassungen 
vorgenommen werden.\\
Um zu verhindern, dass das System sich bewegt, wird die Schwerpunktsgeschwindigkeit berechnet und von allen Teilchen subtrahiert.
\begin{eqnarray}
    \mathbf{v}_\text{COM} &=& \frac1N\sum_{i=1}^N \mathbf{v}_i\\
    %\mathbf{v}_\text{cut} &=& \frac{\mathbf{v}_\text{tot}}N\\
    \mathbf{v}_i^\text{new} &=& \mathbf{v}_i - \mathbf{v}_\text{COM}
\end{eqnarray}
Da es für die Teilchen kein sehr natürlicher Zustand ist sich genau auf den Punkten in einem FCC Gitter zu befinden benötigt das System 
einige Zeitschritte zur Equilibrierung, was dazu führen kann, dass sich eine andere Temperatur einstellt, als die, die gewünscht ist. Um zu 
versichern, dass die Temperatur der gewünschten entspricht, müssen die Geschwindigkeiten der Teilchen reskaliert werden:
\begin{eqnarray}
\label{eq:rescale1}   E &=& \sum_{i=1}^N \frac{\mathbf{v}_i^2}{2} \\
 \label{eq:rescale2}   \lambda &=& \sqrt{\frac{3(N-1)T}{2E}} \\
 \label{eq:rescale3}   \mathbf{v}_i^\text{new} &=& \lambda \mathbf{v}_i  
\end{eqnarray}
Diese Reskalierung wird einige Male vorgenommen, bis das System die gewünschte Temperatur erreicht hat. Da diese Reskalierung keinem 
natürlichen internen Prozess entspricht, kann man einen Sprung in der Kurve der Gesamtenergie sehen.

\subsection{eHEX}
Der eHEX Algorithmus besteht aus vielen Schritten und vielen Variablen, was ihn sehr fehleranfällig macht. Weiters sollte der Effekt von 
verschiedenen Werten von $\Delta Q$ untersucht werden.

\end{document}

### 2017-03-21 doc/presentation/pres_long_ver1.tex
presentation presentation presentation

@@
 Der eHEX Algorithmus besteht aus vielen Schritten und vielen Variablen, was ihn sehr fehleranfällig macht. Weiters sollte der Effekt von 
-verschiedenen Werten von $\Delta Q$ untersucht werden.
+verschiedenen Werten von $\Delta Q$ untersucht werden. Was wir erwarten ist eine kontinuierliche Erhöhung der
+internen Temperatur des Systems. Die Rate der Erhöhung sollte mit steigendem $\Delta Q$ steigen und die Temperatur sollte sich 
+für $\Delta Q = 0$ nicht verändern. Weiters sollte mit steigender interner Temperatur auch die Gesamtenergie des Systems
+steigen. Diese zwei Erwartungen werden erfüllt, wie die beiden Bilder zeigen.
+
+\subsection{Thermostat}
+Der Thermostat hält das System davon ab sich zu überhitzen. Durch Interaktion von Gasteilchen mit dem Nanoteilchen in der Falle
+wird Energie auf die Gasteilchen übertragen und somit stellt sich eine neue, höhere konstante Temperatur ein. Diese höhrere
+Temperatur ist die Basis für die Messungen.
+
+\subsection{Simulation des Experiments}
+Da wir jetzt festgestellt haben, dass alle Teile individuell funktionieren, ist es Zeit sie zusammenzuführen, um das Experiment zu 
+simulieren, um die Frage zu beantworten: Welchen Einfluss hat die Laserintensität auf die Schwerpunktsbewegung des Gases? Und welche 
+Rolle spielt die Temperatur des Umgebenden Gases?\\
+Zunächst müssen wir feststellen, ob die Anwendung des eHEX Algorithmus die Schwerpunktsbewegung des Nanoteilchens beeinflusst. Diese Untersuchung
+resultiert in einer langweiligen Grafik mit einer konstanten, waagrechten Gerade, die zeigt, dass der eHEX Algorithmus alleine keinen 
+Einfluss auf die Schwerpunktsbewegung hat.\\
+Der nächste Schritt ist eine Art Kalibrierung. Bei Laserintensität $\Delta Q = 0$ und unter Einwirkung des Thermostaten sollte sich das
+gesamte System auf die selbe Temperatur, nämlich die Temperatur des umgebenden Gases einstellen. Dies ist der Fall, wie in 
+folgenden Grafiken zu sehen ist. Die gewählten Parameter sind: $P=0.8$, $T_\text{imp} = 0.05$, $T = 0.2$, $N=32$.\\
+Da nun der Algorithmus kalibriert ist, werden die interessanteren Fälle von $\Delta Q > 0$ untersucht. Der eingestellte Druck 
+ist proportional zur Anzahl der erzeugten Teilchen auf jeder Fläche. Höherer Druck bedeutet also mehr Teilchen, die Wärme vom 
+Nanoteilchen abführen können, was zu einem höheren Maximalwert von $\Delta Q$ führt. Die Temperatur des Umgebenden Gases fließt ebenfalls 
+in die Anzahl der erzeugten Teilchen ein, deshalb wurden zwei Temperaturen zur untersuchung gewählt: 0.05 und 0.1 - ein Viertel und die 
+Hälfte der ursprünglichen internen Temperatur des Gasteilchens. Der Druck wird von 0.1 bis 1.2 in Schritten von 0.1 variiert und die 
+zugeführte Wärmemenge von 0 bis 0.5 in Schritten von 0.05 im Allgemeinen (nicht konstant, sondern hängt von den Umständen ab, vor 
+allem bei $P=0.1$ und $P=0.2$). Für die Darstellung werden 3 Werte für den Druck gewählt: $P=0.1$, $P=0.5$ und $P=1.0$. Für jeden 
+dieser Druckwerte werden die Graphen für die Schwerpunktstemperatur, die interne Temperatur und die Temperatur des ausgehenden Gases 
+geplottet. Dabei ist auf die Unterschiedliche Skalierung der Achsen zu achten.\\
+Die Erste Beobachtung ist, dass die Schwerpunktsbewegung und die Temperatur des ausgehenden Gases mit steigendem $\Delta Q$ ansteigen.
+Das ist zu erwarten, da bei höherer ins System gempumpter Wärme mehr Energie an die umgebenden Teilchen abgegeben wird. Weiters wird 
+beobachtet, dass die Temperatur schneller für niedrige Drücke ansteigt (was durch die Unterschiedliche Skalierung der 
+Ordinate nicht sofort auffällt). Dieses Verhalten ist ebenfalls zu erwarten, weil das Umgebende Gas als Thermostat wirkt und 
+höherer Druck eine höhere Anzahl an Gasteilchen zur Folge hat, die mehr Wärme vom System abtransportieren.\\
+Die Entwicklung der Temperatur hängt auch von der Temperatur des Umgebenden Gases ab. Die zur Temperatur $T=0.1$ korrespondieren (blau) 
+steigen etwas schneller an, als die roten Kurven.\\
+Millen et al. geben eine Formel für die Entwicklung der Schwerpunktstemperatur an. Diese Formel kann mit den Ergebnissen verglichen werden. 
+Man sieht in der Grafik (die als Repräsentation gewählt wurde), 
+dass die gemessene Temperatur schneller ansteigt als die berechnete. Dies kann einige Gründe haben. Ein Grund dafür könnte
+ein Unterschiedlicher thermischer Akkomodationskoeffizient sein. Dieser ist gegeben durch
+\begin{equation}
+    \alpha = \frac{T_\text{em} - T_\text{imp}}{T_\text{sur} - T_\text{imp}}
+\end{equation}
+Der Akkomodationskoeffizient ist ein Maß für die übertragung der Energie vom Nanoteilchen auf das Gas. Der Wert für $\alpha$ im Paper von 
+Millen et al. ist $\alpha = 0.777$, während der berechnete Wert für diese Simulation bei $\alpha = 0.016$. Das bedeutet, dass 
+die Übertragung der Energie von Nanoteilchen auf Gasteilchen nicht auf die gleiche Art und Weise stattfindet.
+
+\subsection{Conclusio}
+Um zusammenzufassen: Für diese Arbeit habe ich ein Optical Tweezer Experiment simuliert, um den Einfluss der Laserintensität 
+auf die Schwerpunktsbewegung zu untersuchen. Die Ergebnisse zeigen, dass die Laserintensität, die Temperatur
+des eingehenden Gases und der umgebende Druck die Schwerpunktsbewegung beeinflussen.
 
